---
title: "Exploration of native taxa in the LC"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(htmltools)
library(sf)
library(here)
library(flextable)
library(ggrepel)
```

```{r, imported parameters}

# collection = 'CUBG'
# load('/Users/jakepowell/Cambridge/Enriched_reports_Jake/CUBG_June2023_enriched_report.rda')
# coordinates = c(52.19378853629289,0.1277234065606053)


# 
# collection = 'LC'
# load('/Users/jakepowell/Cambridge/Enriched_reports_Jake/Vilnius_enriched_report.rda')
# coordinates = c(54.737851403492456, 25.413512725031616)
# load('/Users/jakepowell/Cambridge/WCVP/Version 11/wcvp_with_redlistcategory.rda')
# load('/Users/jakepowell/Cambridge/Geography Module/wgsrpd3.rda')
# 
# native = 'Naturally occurring only'
# extinct = TRUE
# doubtful_locations = FALSE
# min_year = 1970
# export_data = F
# table_font_size = 14
# report_kind = 'static'
# ggtheme = NULL
# separate_figure_folder = FALSE
# value_on_fig = TRUE
# report_kind = 'static'
```

```{r theme_ggplot2, echo = FALSE}
library(ggplot2)

# Changing the default theme
if(is.null(ggtheme)){
   ggtheme <- function(base_size = 16) {
      ggplot2::theme_bw(base_size = base_size) %+replace%
        ggplot2::theme(
          plot.title = ggplot2::element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
          panel.grid.minor = ggplot2::element_blank(),
          panel.border = ggplot2::element_blank(),
          axis.title = ggplot2::element_text(size = rel(0.85), face = "bold"),
          axis.text = ggplot2::element_text(size = rel(0.70), face = "bold"),
          axis.line = ggplot2::element_line(color = "black"),
          legend.title = ggplot2::element_text(size = rel(0.85), face = "bold"),
          legend.text = ggplot2::element_text(size = rel(0.70), face = "bold"),
          legend.key = ggplot2::element_rect(fill = "transparent", colour = NA),
          legend.key.size = ggplot2::unit(1.5, "lines"),
          legend.background = ggplot2::element_rect(fill = "transparent", colour = NA),
          strip.background = ggplot2::element_rect(fill = "#17252D", color = "#17252D"),
          strip.text = ggplot2::element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
        )
    }
}
theme_set(ggtheme())
update_geom_defaults("line", list(size = 2))

# Set the colour scale for the plots
scale_colour_continuous <- scale_colour_continuous
scale_colour_discrete   <- scale_colour_discrete
scale_colour_binned     <- scale_colour_binned
scale_fill_continuous <- scale_fill_continuous
scale_fill_discrete <- scale_fill_discrete
scale_fill_binned <- scale_fill_binned
```

```{r, Add combined geography column to enriched report}
geography_data = enriched_report[,grepl('_area_code_l3', names(enriched_report))]
want = c('000','010','001','011','100','110','101')

#Depending on the values of the options reduce `want` to only the satisfactory values.
##A) introduced / naturally occurring only.
if(native == 'Introduced only'){
 want = want[grepl('^1',want)]
}else if(native == 'Naturally occurring only'){
 want = want[grepl('^0',want)]
}
## B) Extinct
if(!extinct){
 want = want[!grepl('010|011|110',want)]
}
## C) Doubtful
if(!doubtful_locations){
 want = want[!grepl('001|011|101',want)]
}
# Get the columns in wanted info that contain the geography information we want.
geography_data = geography_data[,grepl(paste0(want,collapse='|'), names(geography_data))]

if(length(want) > 1){
 level3codes =do.call("paste", c(geography_data, sep = ", "))
 level3codes = stringr::str_remove(level3codes, ', NA$')
 level3codes = stringr::str_remove(level3codes, 'NA, ')
}else{
 level3codes = geography_data
}
level3codes[level3codes == "NA"] = NA

enriched_report$geography_codes = level3codes
```

```{r, Get location_type_text}
location_type_text = ''
if(native == 'Naturally occurring only'){
 location_type_text = paste0(location_type_text, 'naturally occurringonly,', collapse = ' ')  
}else if(native == 'Introduced only'){
  location_type_text = paste0(location_type_text, 'introduced only,', collapse = ' ')  
 
}else{
  location_type_text = paste0(location_type_text, 'Either naturally occurring or introduced,', collapse = ' ')  
}
if(extinct){
   location_type_text = paste0(location_type_text, ' allowing extinct regions and', collapse = ' ')  
}else{
  location_type_text = paste0(location_type_text, ' not including extinct regions and', collapse = ' ')  
}
if(doubtful_locations){
   location_type_text = paste0(location_type_text, ' allowing doubtful regions.', collapse = ' ')  
 
}else{
  location_type_text = paste0(location_type_text, ' not allowing doubtful regions.', collapse = ' ')  
}
```

`r if(any(! c('AccNoFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType') %in% names(enriched_report))){"*** \n **Warning!**\n\n This enriched report contains missing 'original' columns needed to perform the analyses in this report. Default values have been created. In particular, the following were missing "}`

`r if(!'AccNoFull' %in% names(enriched_report)){"- AccNoFull: assume each item is its own accession.\n"}`
`r if(!'AccYear' %in% names(enriched_report)){"- AccYear: Cannot do analyses over time.\n"}`
`r if(!'ItemStatusDate' %in% names(enriched_report)){"- AccNoFull: Cannot do analyses over time.\n"}`
`r if(!'ItemStatusType' %in% names(enriched_report)){"- AccNoFull: Cannot do analyses over time.\n"}`

`r if(any(! c('AccNoFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType') %in% names(enriched_report))){"***"}`


```{r, Extracting information from enriched report, echo = F}
# Add required columns if missing.
do_over_time = TRUE
needed_columns = c('AccNoFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType')
if(!'AccNoFull' %in% names(enriched_report)){
  enriched_report$AccNoFull = 1:nrow(enriched_report)
}
if(!'AccYear' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$AccYear = rep(0, nrow(enriched_report))
}
if(!'ItemStatusDate' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$ItemStatusDate = rep(NA, nrow(enriched_report))
}
if(!'ItemStatusType' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$ItemStatusType = rep('Existing', nrow(enriched_report))
}
if(all(is.na(enriched_report$ItemStatusDate))){
  do_over_time = FALSE
}
if(all(is.na(enriched_report$ItemStatusType))){
  do_over_time = FALSE
}
if(all(is.na(enriched_report$AccYear))){
  do_over_time = FALSE
}
#Extract endemic information from enriched_report
endemic = rep('Not Endemic', nrow(enriched_report))
endemic_index = which(stringr::str_length(enriched_report$geography_codes) == 3)
endemic[endemic_index] = 'Endemic'
enriched_report$endemic = endemic
rm(endemic)

# Extract threatened. 
threat_cat = c('VU','EN','CR','EW','EX')
threatened = rep('Not Threatened', nrow(enriched_report))
threatened[which(enriched_report$redList_category %in% threat_cat)] = 'Threatened'
threatened[which(enriched_report$POWO_Red_category %in% threat_cat)] = 'Threatened'
enriched_report$threatened = threatened
rm(threatened)

# Extract threatened category. 
threatened_category = rep(NA, nrow(enriched_report))
for(i in 1:length(threat_cat)){
 threatened_category[which(enriched_report$redList_category == threat_cat[i])] = threat_cat[i]
 threatened_category[which(enriched_report$POWO_Red_category == threat_cat[i])] = threat_cat[i]
}
enriched_report$threatened_category = threatened_category
rm(threatened_category)

# Convert Provenance code to text.
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'G'] = 'Garden'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'U'] = 'Unknown'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'W'] = 'Wild'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'Z'] = 'Wild-derived'

# Create a sanitised taxonomic name.
enriched_report$good_name = paste0(enriched_report$sanitised_taxon, ' ', enriched_report$extracted_author)

best_name = paste0(enriched_report$POWO_taxon_name, ' ', enriched_report$POWO_taxon_authors)
best_name[which(best_name == 'NA NA')] = enriched_report$good_name[which(best_name == 'NA NA')]
enriched_report$good_name = paste0(enriched_report$sanitised_taxon, ' ', enriched_report$extracted_author)

report_original = enriched_report

if(separate_figure_folder){
  figures_dir = paste0(output_dir,'/',collection,'_Figures')
  dir.create(figures_dir,showWarnings = FALSE)
}

exporting_data_store = list()
```

```{r, Find LC location, echo = FALSE}
coord_contained = TRUE
# Extract the location of the collection.
pnts <- data.frame( x = coordinates[2], y = coordinates[1])

# create a points collection
pnts_sf <- do.call("st_sfc",c(lapply(1:nrow(pnts), 
                                     function(i) {sf::st_point(as.numeric(pnts[i, ]))}), list("crs" = 4326))) 

pnts_trans <- st_transform(pnts_sf, 2163) # apply transformation to pnts sf
tt1_trans <- st_transform(wgsrpd3, 2163)      # apply transformation to polygons sf

intersect = as.vector(st_intersects(tt1_trans, pnts_trans, sparse = FALSE))
collection_geog_details = wgsrpd3[which(intersect),-5]
location_name = collection_geog_details$LEVEL3_NAM
location_code = collection_geog_details$LEVEL3_COD

if(length(location_name) == 0){
  coord_contained = FALSE
  # Try and find the nearest polygon.
  collection_geog_details  =  wgsrpd3[which.min(as.vector(sf::st_distance(tt1_trans, pnts_trans))),-5]
  location_name = collection_geog_details$LEVEL3_NAM
  location_code = collection_geog_details$LEVEL3_COD
}
```

This report explores the taxa existing in `r collection` that are native to `r location_name`, and how the representation of native species has varied over time.

 To determine the geographic distribution of items in the LC we enrich with information from [Plants of the World Online](https://powo.science.kew.org), by matching taxonomic names. POWO uses [TDWG geographical codes](https://www.tdwg.org/standards/wgsrpd/) (Brummitt, 2001) expressed to that system's third level. This splits the world into 369 regions. Thus, each accepted name in POWO is associated with a list of geographic regions. POWO splits the geographic regions associated with a taxa by three conditions:

- Native / introduced,
- Extinct / Not extinct,
- Location is doubtful / otherwise.

This document considers locations that are: `r location_type_text` 
 
`r if(coord_contained){paste0("Within TDWG geographical codes ",collection," is  found within the region ",location_name,".")}`
`r if(!coord_contained){paste0("Within TDWG geographical codes ",collection," is not found within any regions, however the nearest region is ",location_name,".")}`
Therefore, any taxa whose geographic distribution includes `r location_name` is classed as native to the LC. Otherwise, taxa are classed as non-native.

The location of the LC and the native region are shown below.

```{r, LC location and native region plot, fig.fullwidth=TRUE, fig.dim = c(10, 6), results='hide',fig.keep='all', message=FALSE}
collection_geog_details$occurrence_type = 'native'
lims <- st_bbox(collection_geog_details)
p <- rWCVP::wcvp_distribution_map(collection_geog_details, crop_map=TRUE) + 
  theme(legend.position="none") +
  geom_sf(data=pnts_sf, fill="purple",col="purple", shape=20, size = 3) +
  coord_sf(xlim=lims[c(1,3)], ylim=lims[c(2,4)]) +
  theme(legend.position = "none",
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(hjust = 0.5))

if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Map_LC_location_and_native region.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

p
```

***

### How many native species are in the LC?

```{r, How many native species are in the LC - data}
# Find all the native taxa to the location of the LC in WCVP
geography_data = wcvp$geography[,grepl('_area_code_l3', names(wcvp$geography))]
want = c('000','010','001','011','100','110','101')
#Depending on the values of the options reduce `want` to only the satisfactory values.
##A) introduced / naturally occurring only.
if(native == 'Introduced only'){
 want = want[grepl('^1',want)]
}else if(native == 'Naturally occurring only'){
 want = want[grepl('^0',want)]
}
## B) Extinct
if(!extinct){
 want = want[!grepl('010|011|110',want)]
}
## C) Doubtful
if(!doubtful_locations){
 want = want[!grepl('001|011|101',want)]
}
# Get the columns in wanted info that contain the geography information we want.
geography_data = geography_data[,grepl(paste0(want,collapse='|'), names(geography_data))]

if(length(want) > 1){
 level3codes =do.call("paste", c(geography_data, sep = ", "))
 level3codes = stringr::str_remove(level3codes, ', NA$')
 level3codes = stringr::str_remove(level3codes, 'NA, ')
}else{
 level3codes = geography_data
}
level3codes_wcvp = level3codes

native_index = grepl(location_code,level3codes_wcvp)
native_plant_ids =  wcvp$geography$plant_name_id[native_index]
native_plant_ids = native_plant_ids[!is.na(native_plant_ids)]

native_wcvp = wcvp$wcvp_names[wcvp$wcvp_names$plant_name_id %in% native_plant_ids,]

# native_wcvp currently also contains genus only (records) so we need to reduce to species.
native_wcvp = native_wcvp[-which(native_wcvp$taxon_rank == 'Genus'),]

no_native_taxa = nrow(native_wcvp)
no_native_families = length(unique(native_wcvp$family))
no_native_genus = length(unique(native_wcvp$genus))
no_native_species = length(unique(paste0(native_wcvp$genus, ' ', native_wcvp$species)))


# Get the existing native plants from the LC to the location. 
LC_native = enriched_report[enriched_report$POWO_plant_name_id %in% native_plant_ids,]
LC_native_exisiting = LC_native[LC_native$ItemStatusType == 'Existing',]

no_native_items_LC = nrow(LC_native_exisiting)
no_native_accessions_LC = length(unique(LC_native_exisiting$AccNoFull))
no_native_families_LC = length(unique(LC_native_exisiting$POWO_family))
no_native_genus_LC = length(unique(LC_native_exisiting$POWO_genus))
no_native_species_LC = length(unique(paste0(LC_native_exisiting$POWO_genus, ' ', LC_native_exisiting$POWO_species)))
no_native_taxa_LC = length(unique(paste0(LC_native_exisiting$POWO_taxon_name, ' ', LC_native_exisiting$POWO_taxon_authors)))

category = c('Family', 'Genus', 'Species', 'Taxa')
WCVP = c(no_native_families, no_native_genus, no_native_species, no_native_taxa)
LC  = c(no_native_families_LC, no_native_genus_LC, no_native_species_LC, no_native_taxa_LC)
proportion = paste0(round(LC/WCVP*100, digits = 2),'%')

df = data.frame(category, WCVP,LC, proportion)
names(df) = c('Category', 'Number found in WCVP', ' Number found in LC', 'Proportion of WCVP found in LC')

if(export_data){
  exporting_data_store$Representative_of_native_species_in_LC_current = df
}

# MANAGEMENT SHEET: Export the management sheet of all native accessions.
native_accessions = LC_native_exisiting
native_accessions = native_accessions[match(unique(native_accessions$AccNoFull),native_accessions$AccNoFull),]
powo_name = paste0(native_accessions$POWO_taxon_name, ' ', native_accessions$POWO_taxon_authors)
powo_name[powo_name == 'NA NA'] = NA
native_accessions$powo_name = powo_name
native_accessions = native_accessions[,match(c('AccNoFull', 'TaxonNameFull', 'powo_name', 'threatened', 'endemic'), names(native_accessions))]
names(native_accessions) = c('Accession Number', 'Original Taxonomic Name and Author', 'POWO Taxonomic Name and Author', 'Threatened', 'Endemic')

writexl::write_xlsx(native_accessions, path =paste0(output_dir,'/',collection,'_native_accessions_list.xlsx'))

```

In the LC there are `r format(no_native_items_LC, big.mark=',')` items corresponding to  `r format(no_native_accessions_LC, big.mark=',')` accessions of existing native plants. In the following table we show how many families/genus/species/taxa of native plants are represented in the LC (compared to all plants from `r location_name` in WCVP). Note that here species is defined as the genus species pair and taxa is the taxonomic name and author (thus including subspecies, variety, forma).

```{r, How many native species are in the LC - table}
if(report_kind == 'static'){
  df |>
  flextable() |>
  set_caption(caption = "Native plants in the collection compared to WCVP") |> 
  font(fontname = "Calibri (Body)", part = "all") |> 
  fontsize(size = table_font_size, part = "body") |> 
  # add footer if you want
  # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
  #                colwidths = 4) |> 
  theme_booktabs() |> # default theme
  autofit()
}
if(report_kind == 'interactive'){
  DT::datatable(df)
}

```

`r if(do_over_time){"Below we produce figures detailing how the representation of native families/genus/species/taxa has changed over time in the LC."}`



```{r, How many native species are in the LC over time - data}
# LC_native contains, The LC reduced to native plants.
LC_native_over_time = LC_native
# The current year.
year_cur = as.numeric(format(Sys.Date(),'%Y'))

# Get which items are existing each year in the collection.
years = min_year:year_cur
date = paste0(years, '-01-01')

# Add Genus species columns from WCVP.
GenusSpecies = paste0(LC_native_over_time$POWO_genus, ' ', LC_native_over_time$POWO_species)
LC_native_over_time$GenusSpecies_new = GenusSpecies

# Remove those with missing item status date.
LC_native_over_time = LC_native_over_time[!is.na(LC_native_over_time$ItemStatusDate),]
LC_native_over_time = LC_native_over_time[LC_native_over_time$AccYear > 1650 &
                                            LC_native_over_time$AccYear <= year_cur,]
LC_native_over_time = LC_native_over_time[!is.na(LC_native_over_time$AccYear),]
LC_native_over_time = LC_native_over_time[!is.na(LC_native_over_time$ItemStatusDate),]
LC_native_over_time = LC_native_over_time[!is.na(LC_native_over_time$ItemStatusType),]
# Get the species/genera that are existing
plant_existing_LC_native = BGSmartR::exist_at_date(date, AccessionYear = as.character(LC_native_over_time$AccYear),
                       ItemStatusDate = as.character(LC_native_over_time$ItemStatusDate),
                       ItemStatusType = as.character(LC_native_over_time$ItemStatusType))

# For each year get the number in each group (species and genera)
over_time_info = lapply(plant_existing_LC_native, function(x){
    garden_current = LC_native_over_time[x,]

    # Get the values of interest.
    no_families_LC = length(unique(garden_current$POWO_family))
    no_genus_LC = length(unique(garden_current$POWO_genus))
    no_species_LC = length(unique(paste0(garden_current$POWO_genus, ' ', garden_current$POWO_species)))
    no_taxa_LC = length(unique(paste0(garden_current$POWO_taxon_name, ' ', garden_current$POWO_taxon_authors)))
   
    return(c(no_families_LC, no_genus_LC, no_species_LC, no_taxa_LC))
}) |>
                  data.frame() |>
                  t() |> 
                  data.frame()
names(over_time_info) = c('Family', 'Genus', 'Species', 'Taxa')
over_time_info = data.frame(Date = date, over_time_info)
row.names(over_time_info) = 1:nrow(over_time_info)

if(export_data){
  exporting_data_store$Representative_of_native_species_in_LC_over_time = over_time_info
}

```

```{r  How many native species are in the LC - static figures, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'static' & do_over_time}
to_do = c('Family', 'Genus', 'Species', 'Taxa')
titles  = c('Families', 'Genera', 'Species', 'Taxa')
# Loop over to_do.
for(i in 1:length(to_do)){
  count = as.numeric(unlist(over_time_info[match(to_do[i], names(over_time_info))]))
  percent = round(count/df$`Number found in WCVP`[which(df$Category == to_do[i])]*100, digits = 3)
  
  plot_dataA = data.frame(date =years, 
                               percent = percent,
                               represented = rep('Represented', length(date)))
  plot_dataB = data.frame(date =years, 
                             percent = 100-percent,
                              represented = rep('Not Represented', length(date)))
  
  plot_data = rbind(plot_dataA, plot_dataB)
  
  
  names(plot_data) = c('dates', 'percentage', 'represented')

  p = ggplot(plot_data, aes(fill=represented, y=percentage, x=dates)) + 
    geom_bar(position="stack", stat="identity", width = 1) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(x ="Year",
        y = "Percentage (%)") +
    guides(fill=guide_legend(title=paste0("Native ", titles[i]))) +
    scale_fill_discrete() #+
    # ggtitle(paste0("Representation of ", titles[i] , " over time"))
  
  if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','How_many_native_species_are_in_the_LC', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}
  print(p)
}


 
```

`r if(report_kind == 'interactive' & do_over_time){"##### Representation over time graphs {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Families"}`
```{r  How many native species are in the LC - interactive Family plot, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
to_do = c('Family', 'Genus', 'Species', 'Taxa')
titles  = c('Families', 'Genera', 'Species', 'Taxa')
# Loop over to_do.
i=1
  count = as.numeric(unlist(over_time_info[match(to_do[i], names(over_time_info))]))
  percent = round(count/df$`Number found in WCVP`[which(df$Category == to_do[i])]*100, digits = 3)
  
  plot_dataA = data.frame(date =years, 
                               percent = percent,
                               represented = rep('Represented', length(date)))
  plot_dataB = data.frame(date =years, 
                             percent = 100-percent,
                              represented = rep('Not Represented', length(date)))
  
  plot_data = rbind(plot_dataA, plot_dataB)
  
  
  names(plot_data) = c('dates', 'percentage', 'represented')

  # plot_data$text = paste0(round(plot_data$percent,digits = 2), '% of accessions have ', plot_data$total_items, ' items')
  # hoverinfo = "text",  hovertext = ~text
fig <- plot_ly(plot_data, x = ~dates, y = ~percentage, color = ~represented, type = 'bar')
fig <- fig |> layout(yaxis = list(title = paste0('Percentage of ',titles[i])),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text=paste0("Native ", titles[i]))),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig
  
 
  if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-percentage_native_',titles[i],'_represented_over_time.html'))
}

 
```

`r if(report_kind == 'interactive' & do_over_time){"###### Genera"}`
```{r  How many native species are in the LC - interactive Genera plot, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
to_do = c('Family', 'Genus', 'Species', 'Taxa')
titles  = c('Families', 'Genera', 'Species', 'Taxa')
# Loop over to_do.
i=2
  count = as.numeric(unlist(over_time_info[match(to_do[i], names(over_time_info))]))
  percent = round(count/df$`Number found in WCVP`[which(df$Category == to_do[i])]*100, digits = 3)
  
  plot_dataA = data.frame(date =years, 
                               percent = percent,
                               represented = rep('Represented', length(date)))
  plot_dataB = data.frame(date =years, 
                             percent = 100-percent,
                              represented = rep('Not Represented', length(date)))
  
  plot_data = rbind(plot_dataA, plot_dataB)
  
  
  names(plot_data) = c('dates', 'percentage', 'represented')

  # plot_data$text = paste0(round(plot_data$percent,digits = 2), '% of accessions have ', plot_data$total_items, ' items')
  # hoverinfo = "text",  hovertext = ~text
fig <- plot_ly(plot_data, x = ~dates, y = ~percentage, color = ~represented, type = 'bar')
fig <- fig |> layout(yaxis = list(title = paste0('Percentage of ',titles[i])),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text=paste0("Native ", titles[i]))),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig
  
 
  if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-percentage_native_',titles[i],'_represented_over_time.html'))
}

 
```

`r if(report_kind == 'interactive' & do_over_time){"###### Species"}`
```{r  How many native species are in the LC - interactive Species plot, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
to_do = c('Family', 'Genus', 'Species', 'Taxa')
titles  = c('Families', 'Genera', 'Species', 'Taxa')
# Loop over to_do.
i=3
  count = as.numeric(unlist(over_time_info[match(to_do[i], names(over_time_info))]))
  percent = round(count/df$`Number found in WCVP`[which(df$Category == to_do[i])]*100, digits = 3)
  
  plot_dataA = data.frame(date =years, 
                               percent = percent,
                               represented = rep('Represented', length(date)))
  plot_dataB = data.frame(date =years, 
                             percent = 100-percent,
                              represented = rep('Not Represented', length(date)))
  
  plot_data = rbind(plot_dataA, plot_dataB)
  
  
  names(plot_data) = c('dates', 'percentage', 'represented')

  # plot_data$text = paste0(round(plot_data$percent,digits = 2), '% of accessions have ', plot_data$total_items, ' items')
  # hoverinfo = "text",  hovertext = ~text
fig <- plot_ly(plot_data, x = ~dates, y = ~percentage, color = ~represented, type = 'bar')
fig <- fig |> layout(yaxis = list(title = paste0('Percentage of ',titles[i])),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text=paste0("Native ", titles[i]))),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig
  
 
  if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-percentage_native_',titles[i],'_represented_over_time.html'))
}

 
```

`r if(report_kind == 'interactive' & do_over_time){"###### Taxa"}`
```{r  How many native species are in the LC - interactive Taxa plot, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=4
to_do = c('Family', 'Genus', 'Species', 'Taxa')
titles  = c('Families', 'Genera', 'Species', 'Taxa')
# Loop over to_do.
  count = as.numeric(unlist(over_time_info[match(to_do[i], names(over_time_info))]))
  percent = round(count/df$`Number found in WCVP`[which(df$Category == to_do[i])]*100, digits = 3)
  
  plot_dataA = data.frame(date =years, 
                               percent = percent,
                               represented = rep('Represented', length(date)))
  plot_dataB = data.frame(date =years, 
                             percent = 100-percent,
                              represented = rep('Not Represented', length(date)))
  
  plot_data = rbind(plot_dataA, plot_dataB)
  
  
  names(plot_data) = c('dates', 'percentage', 'represented')

  # plot_data$text = paste0(round(plot_data$percent,digits = 2), '% of accessions have ', plot_data$total_items, ' items')
  # hoverinfo = "text",  hovertext = ~text
fig <- plot_ly(plot_data, x = ~dates, y = ~percentage, color = ~represented, type = 'bar')
fig <- fig |> layout(yaxis = list(title = paste0('Percentage of ',titles[i])),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text=paste0("Native ", titles[i]))),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig
  
 
  if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-percentage_native_',titles[i],'_represented_over_time.html'))
}

 
```


### What proportion of the LC at `r collection` is native and how has this varied over time?

In this section we explore the number of native plants in the LC compared to all plants in the LC. In particular, we will determine if the number of native plants in the LC is stable, increasing or decreasing. 

We will examine the proportion of the LC that is native, this is shown in the table below.

```{r, What proportion of the LC at `r collection` is native and how has this varied over time}
LC = enriched_report
LC_existing = LC[LC$ItemStatusType == 'Existing',]

is_native = rep(FALSE, nrow(LC_existing))
is_native[which(LC_existing$POWO_plant_name_id %in% native_plant_ids)] = TRUE
LC_existing$is_native = is_native


no_items = nrow(LC_existing)
no_items_native = sum(LC_existing$is_native)

accessions_data = LC_existing[match(unique(LC_existing$AccNoFull), LC_existing$AccNoFull),]
no_accessions = nrow(accessions_data)
no_accessions_native = sum(accessions_data$is_native)

taxa_data = LC_existing[match(unique(LC_existing$TaxonNameFull), LC_existing$TaxonNameFull),]
no_taxa = nrow(taxa_data)
no_taxa_native = sum(taxa_data$is_native)

species_data = LC_existing[!is.na(LC_existing$POWO_plant_name_id),]
species_data = LC_existing[match(unique(paste0(species_data$POWO_taxon_name, species_data$POWO_taxon_authors)),
                            paste0(species_data$POWO_taxon_name, species_data$POWO_taxon_authors)),]
no_species = nrow(species_data)
no_species_native = sum(species_data$is_native)

Kind = c('Items', 'Accessions', 'Taxa', 'Species')
All = c(no_items, no_accessions, no_taxa, no_species)
Native = c(no_items_native, no_accessions_native, no_taxa_native, no_species_native)
Proportion = paste0(round(Native/All*100,digits=2),'%')

to_show = data.frame(Kind = Kind, All = All, Native = Native, Proportion = Proportion)

if(export_data){
  exporting_data_store$Native_species_in_LC_current = to_show
}
```

```{r, What proportion of the LC at `r collection` is native and how has this varied over time - table}
if(report_kind == 'static'){
  to_show |>
    flextable() |>
    set_caption(caption = paste0('Native / non native plants (', location_name, ') in the LC')) |> 
    font(fontname = "Calibri (Body)", part = "all") |> 
    fontsize(size = table_font_size, part = "body") |> 
    # add footer if you want
    # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
    #                colwidths = 4) |> 
    theme_booktabs() |> # default theme
    autofit()
}
if(report_kind == 'interactive'){
  DT::datatable(to_show)
}
```

`r if(do_over_time){"Next we show how the number of native plants in the LC has varied over time. Note that the numbers provided in 'Native / non native plants (Great Britain) in the LC' table are calculated using plants that are currently existing (irrespective of whether they have an accession year). Whereas, in the over time analysis we remove records without an accession year (since we do not know which years such records are in the LC). Therefore, the numbers shown for the current year in the figures below might be margially smaller than those show in the previous table."}`

`r if(!do_over_time){"We do not have the information to perform how the number of native plants in the LC has varied over time."}`


```{r, What proportion of the LC at `r collection` is native and how has this varied over time - data}
# Get the number of items accession taxa species for native plants in the LC.
over_time_info = lapply(plant_existing_LC_native, function(x){
    garden_current = LC_native_over_time[x,]

    # Get the values of interest.
    no_items = nrow(garden_current)
    no_accessions = length(unique(garden_current$AccNoFull))
    no_taxa = length(unique(garden_current$TaxonNameFull))
    no_species = length(unique(paste0(garden_current$POWO_taxon_name, garden_current$POWO_taxon_authors)[!is.na(garden_current$POWO_plant_name_id)]))
   
    return(c(no_items, no_accessions, no_taxa, no_species))
}) |>
                  data.frame() |>
                  t() |> 
                  data.frame()
names(over_time_info) = c('Items', 'Accessions', 'Taxa', 'Species')
over_time_info = data.frame(Date = date, over_time_info)
row.names(over_time_info) = 1:nrow(over_time_info)
native_data = over_time_info

if(export_data){
  exporting_data_store$Number_of_native_species_in_LC_over_time = native_data
}
```

```{r, What proportion of the LC at `r collection` is native and how has this varied over time - static number figures, eval = report_kind == 'static' & do_over_time,  fig.fullwidth=TRUE, fig.dim = c(10, 6)}
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')
for(i in 1:length(to_do)){
  counts = native_data[,match(to_do[i], names(native_data))]
  plot_data = data.frame(date = years, counts = counts)
  
  p = ggplot(data=plot_data, aes(x=date, y=counts)) +
  geom_line()+
  geom_point() +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = paste0("Number of ",titles[i])) #+
    # guides(fill=guide_legend(title=paste0("Native ", titles[i]))) +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("Number of native ", titles[i] , " over time")
  
    if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','number_native_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
  
  print(p)
}
```

`r if(report_kind == 'interactive' & do_over_time){"##### Representation of number of native plants over time graphs {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Items"}`
```{r  Number of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=1
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

counts = native_data[,match(to_do[i], names(native_data))]
plot_data = data.frame(date = years, counts = counts)

fig <- plot_ly(plot_data, x = ~date, y = ~counts, type = 'scatter', mode = 'lines+markers')

fig <- fig |> layout(yaxis = list(title = paste0("Number of ",titles[i])),
                   xaxis = list(title = "Year"))
fig <- fig |> layout(hovermode = 'x unified')
fig
  
  
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','line-number_of_native_',titles[i],'__over_time.html'))
}
  
```

`r if(report_kind == 'interactive' & do_over_time){"###### Accessions"}`
```{r  Number of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=2
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

counts = native_data[,match(to_do[i], names(native_data))]
plot_data = data.frame(date = years, counts = counts)

fig <- plot_ly(plot_data, x = ~date, y = ~counts, type = 'scatter', mode = 'lines+markers')

fig <- fig |> layout(yaxis = list(title = paste0("Number of ",titles[i])),
                   xaxis = list(title = "Year"))
fig <- fig |> layout(hovermode = 'x unified')
fig
  
  
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','line-number_of_native_',titles[i],'__over_time.html'))
}
  
```

`r if(report_kind == 'interactive' & do_over_time){"###### Taxa"}`
```{r  Number of taxa accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=3
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

counts = native_data[,match(to_do[i], names(native_data))]
plot_data = data.frame(date = years, counts = counts)

fig <- plot_ly(plot_data, x = ~date, y = ~counts, type = 'scatter', mode = 'lines+markers')

fig <- fig |> layout(yaxis = list(title = paste0("Number of ",titles[i])),
                   xaxis = list(title = "Year"))
fig <- fig |> layout(hovermode = 'x unified')
fig
  
  
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','line-number_of_native_',titles[i],'__over_time.html'))
}
  
```

`r if(report_kind == 'interactive' & do_over_time){"###### Species"}`
```{r  Number of native species over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=4
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

counts = native_data[,match(to_do[i], names(native_data))]
plot_data = data.frame(date = years, counts = counts)

fig <- plot_ly(plot_data, x = ~date, y = ~counts, type = 'scatter', mode = 'lines+markers')

fig <- fig |> layout(yaxis = list(title = paste0("Number of ",titles[i])),
                   xaxis = list(title = "Year"))
fig <- fig |> layout(hovermode = 'x unified')
fig
  
  
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','line-number_of_native_',titles[i],'__over_time.html'))
}
  
```

`r if(report_kind == 'interactive' & do_over_time){"##### {-}"}`

`r if(do_over_time){"Next we view native plants in the LC as a percentage of all plants in the LC over time."}`

```{r, What proportion of the LC at `r collection` is native and how has this varied over time - percentage data, eval = do_over_time}
# Do the same for the whole collection.
report = enriched_report
report = report[report$AccYear > 1650 & report$AccYear<=year_cur,]
report = report[!is.na(report$ItemStatusDate),]
plant_existing = BGSmartR::exist_at_date(date, AccessionYear = as.character(report$AccYear),
                       ItemStatusDate = as.character(report$ItemStatusDate),
                       ItemStatusType = as.character(report$ItemStatusType))

# For each year get the number in each group (species and genera)
over_time_info = lapply(plant_existing, function(x){
    garden_current = report[x,]
    
    # Get the values of interest.

    no_items = nrow(garden_current)
    no_accessions = length(unique(garden_current$AccNoFull))
    no_taxa = length(unique(garden_current$TaxonNameFull))
    no_species = length(unique(paste0(garden_current$POWO_taxon_name, garden_current$POWO_taxon_authors)[!is.na(garden_current$POWO_plant_name_id)]))

    return(c(no_items, no_accessions, no_taxa, no_species))
}) |>
                  data.frame() |>
                  t() |> 
                  data.frame()
names(over_time_info) = c('Items', 'Accessions', 'Taxa', 'Species')
over_time_info = data.frame(Date = date, over_time_info)
row.names(over_time_info) = 1:nrow(over_time_info)
LC_data = over_time_info

if(export_data){
  exporting_data_store$Total_species_in_LC_over_time = LC_data
}
```

```{r, What proportion of the LC at `r collection` is native and how has this varied over time - static percentage figures, eval = report_kind == 'static' & do_over_time, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
to_do =  c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

# Loop over to_do.
for(i in 1:length(to_do)){
  native_counts = native_data[,match(to_do[i], names(native_data))]
  total_counts = LC_data[,match(paste0(to_do[i]), names(LC_data))]
  
  percent = round(native_counts/total_counts*100, digits = 3)

  plot_dataA = data.frame(date =years, 
                               percent = percent,
                               represented = rep('Native', length(date)))
  plot_dataB = data.frame(date =years, 
                             percent = 100-percent,
                              represented = rep('Non Native', length(date)))
  
  plot_data = rbind(plot_dataA, plot_dataB)
  plot_data$represented = factor(plot_data$represented,levels = c('Non Native','Native'))
  
  names(plot_data) = c('dates', 'percentage', 'represented')

  p = ggplot(plot_data, aes(fill=represented, y=percentage, x=dates)) + 
    geom_bar(position="stack", stat="identity", width = 1) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = "Percentage (%)") +
    guides(fill=guide_legend(title=paste0("Native ", titles[i]))) +
   scale_fill_discrete() # +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("Proportion of ", titles[i] , " that are native in the LC over time"))
    
      if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','proportion_native_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
    
  print(p)
}
 

```

`r if(report_kind == 'interactive' & do_over_time){"#####  Proportion of native plants over time graphs {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Items"}`
```{r  Proportion of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=1
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

native_counts = native_data[,match(to_do[i], names(native_data))]
total_counts = LC_data[,match(paste0(to_do[i]), names(LC_data))]

percent = round(native_counts/total_counts*100, digits = 3)

plot_dataA = data.frame(date =years, 
                             percent = percent,
                             represented = rep('Native', length(date)))
plot_dataB = data.frame(date =years, 
                           percent = 100-percent,
                            represented = rep('Non Native', length(date)))

plot_data = rbind(plot_dataA, plot_dataB)
plot_data$represented = factor(plot_data$represented,levels = c('Non Native','Native'))

names(plot_data) = c('dates', 'percentage', 'represented')

fig <- plot_ly(plot_data, x = ~dates, y = ~percentage, color = ~represented, type = 'bar')
fig <- fig |> layout(yaxis = list(title = paste0('Percentage of ',titles[i])),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text=paste0("Native ", titles[i]))),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_bar-percentage_of_native',titles[i],'__over_time.html'))
}
  
```

`r if(report_kind == 'interactive' & do_over_time){"###### Accessions"}`
```{r  Proportion of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=2
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

native_counts = native_data[,match(to_do[i], names(native_data))]
total_counts = LC_data[,match(paste0(to_do[i]), names(LC_data))]

percent = round(native_counts/total_counts*100, digits = 3)

plot_dataA = data.frame(date =years, 
                             percent = percent,
                             represented = rep('Native', length(date)))
plot_dataB = data.frame(date =years, 
                           percent = 100-percent,
                            represented = rep('Non Native', length(date)))

plot_data = rbind(plot_dataA, plot_dataB)
plot_data$represented = factor(plot_data$represented,levels = c('Non Native','Native'))

names(plot_data) = c('dates', 'percentage', 'represented')

fig <- plot_ly(plot_data, x = ~dates, y = ~percentage, color = ~represented, type = 'bar')
fig <- fig |> layout(yaxis = list(title = paste0('Percentage of ',titles[i])),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text=paste0("Native ", titles[i]))),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig
  
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_bar-percentage_of_native',titles[i],'__over_time.html'))
}
  
```

`r if(report_kind == 'interactive' & do_over_time){"###### Taxa"}`
```{r  Proportion of native taxa over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=3
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

native_counts = native_data[,match(to_do[i], names(native_data))]
total_counts = LC_data[,match(paste0(to_do[i]), names(LC_data))]

percent = round(native_counts/total_counts*100, digits = 3)

plot_dataA = data.frame(date =years, 
                             percent = percent,
                             represented = rep('Native', length(date)))
plot_dataB = data.frame(date =years, 
                           percent = 100-percent,
                            represented = rep('Non Native', length(date)))

plot_data = rbind(plot_dataA, plot_dataB)
plot_data$represented = factor(plot_data$represented,levels = c('Non Native','Native'))

names(plot_data) = c('dates', 'percentage', 'represented')

fig <- plot_ly(plot_data, x = ~dates, y = ~percentage, color = ~represented, type = 'bar')
fig <- fig |> layout(yaxis = list(title = paste0('Percentage of ',titles[i])),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text=paste0("Native ", titles[i]))),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_bar-percentage_of_native',titles[i],'__over_time.html'))
}
  
```

`r if(report_kind == 'interactive' & do_over_time){"###### Species"}`
```{r  Proportion of native species over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=4
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

native_counts = native_data[,match(to_do[i], names(native_data))]
total_counts = LC_data[,match(paste0(to_do[i]), names(LC_data))]

percent = round(native_counts/total_counts*100, digits = 3)

plot_dataA = data.frame(date =years, 
                             percent = percent,
                             represented = rep('Native', length(date)))
plot_dataB = data.frame(date =years, 
                           percent = 100-percent,
                            represented = rep('Non Native', length(date)))

plot_data = rbind(plot_dataA, plot_dataB)
plot_data$represented = factor(plot_data$represented,levels = c('Non Native','Native'))

names(plot_data) = c('dates', 'percentage', 'represented')

fig <- plot_ly(plot_data, x = ~dates, y = ~percentage, color = ~represented, type = 'bar')
fig <- fig |> layout(yaxis = list(title = paste0('Percentage of ',titles[i])),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text=paste0("Native ", titles[i]))),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_bar-percentage_of_native',titles[i],'__over_time.html'))
}
  
```

`r if(report_kind == 'interactive' & do_over_time){"##### {-}"}`

***

### Among the native plants in a LC, how many of them are endemic or redlisted?

```{r, Among the native plants in a LC, how many of them are endemic or redlisted - Data}
# LC_native_exisiting, native and existing part of LC database.

endemic = rep(FALSE, nrow(LC_native_exisiting))
endemic[which(stringr::str_length(LC_native_exisiting$POWO_Dist_110_area_code_l3) == 3)] = TRUE
total_endemic = sum(endemic)

threatened = rep(FALSE, nrow(LC_native_exisiting))
threatened[which(LC_native_exisiting$POWO_Red_category %in% c('EN', 'VU', 'CR', 'EW', 'EX'))] = TRUE
threatened[which(LC_native_exisiting$redList_category %in% c('EN', 'VU', 'CR', 'EW', 'EX'))] = TRUE
total_threatened = sum(threatened)

total_threat_endemic = sum(endemic & threatened)
need_venn = total_endemic > 0 | total_threatened > 0

```

`r if(need_venn){paste0("In this section, we concentrate on native plants in the LC that are either endemic to ", location_name ," or redlisted by the IUCN.")}`

`r if(need_venn){"Below we produce a Venn diagram of endemic and threatened plants that are native in the LC."}`

`r if(!need_venn){"There are no endemic or threatened plants that are native in the LC."}`


```{r, Among the native plants in a LC, how many of them are endemic or redlisted - Venn Diagram, echo = FALSE, results='hide',fig.keep='all', message=FALSE}
if(need_venn){
  library("VennDiagram") 

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
# draw.pairwise.venn(area1=total_endemic, area2=total_threatened,cross.area=total_threat_endemic, 
#                    category=c("Endemic","Threatened"),fill=c("Yellow", "Red"))

draw.pairwise.venn(
  area1 = total_endemic,
  area2 = total_threatened,
  cross.area = total_threat_endemic,
  category = c("Endemic","Threatened"),
  col=c("#440154ff", '#21908dff'),
  fill = c(alpha("#440154ff",0.3), alpha('#21908dff',0.3)),
  cex = 1.1,
  cat.cex = 1.3,
  cat.default.pos = "outer",
  cat.pos = c(0, 0),
  cat.dist = 0.055,
  # cat.just = list(c(-1, -1), c(1, 1)), ext.pos = 30,
  # ext.dist = -0.05,
  # ext.length = 0.85,
  # ext.line.lwd = 2,
  # ext.line.lty = "dashed",
  scaled = TRUE,
  lwd = c(1,1),
  lty = c(1,1)
)

      if(separate_figure_folder){
ggplot2::ggsave(filename = paste0(figures_dir, '/','Native_threatened_endemic_Venn_diagram.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
}

```


`r if(need_venn){"For these native and endemic/threatened plants we produce a table below detailing how many of each species is in the collection together with the accession number and when each item was last checked."}`

```{r, Among the native plants in a LC, how many of them are endemic or redlisted - get important taxa, ft.keepnext = FALSE, eval = need_venn}
accessions = LC_native_exisiting$AccNoFull
taxa = paste0(LC_native_exisiting$POWO_taxon_name, ' ', LC_native_exisiting$POWO_taxon_authors)
ItemStatusDate = LC_native_exisiting$ItemStatusDate

data = data.frame(accessions =accessions, 
                  taxa = taxa,
                  ItemStatusDate = ItemStatusDate, 
                  endemic = endemic, 
                  threatened = threatened)

data_compressed <- data |>
  dplyr::group_by(.data$taxa) |>
  dplyr::summarise(endemic = .data$endemic[1],
                   threatened = .data$threatened[1],
                   ItemStatusDates = toString(.data$ItemStatusDate),
                   Accessions = toString(unique(.data$accessions)),
                   Items = length(.data$ItemStatusDate)
                   
  ) |>
  dplyr::ungroup()

data_compressed = data_compressed[data_compressed$endemic | data_compressed$threatened,]

if(export_data){
  exporting_data_store$Detail_threatened_and_endemic_native_plants_in_LC = data.frame(data_compressed)
}

# MANAGEMENT SHEET: native plants that are native and either endemic or threatened.
writexl::write_xlsx(data_compressed, path =paste0(output_dir,'/',collection,'_native_and_either_endemic_or_threatened_list.xlsx'))


```

```{r, Among the native plants in a LC, how many of them are endemic or redlisted - table, ft.keepnext = FALSE, eval = need_venn}
if(report_kind == 'static'){
  data_compressed |>
  flextable() |>
  set_caption(caption = paste0('Native plants in the LC that are endemic or threatened')) |> 
  font(fontname = "Calibri (Body)", part = "all") |> 
  fontsize(size = table_font_size, part = "body") |> 
  # add footer if you want
  # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
  #                colwidths = 4) |> 
  theme_booktabs() |> # default theme
  autofit() |>
  set_table_properties(width = 1, layout = "autofit")
}
if(report_kind == 'interactive'){
    DT::datatable(data_compressed)

}

```

***

### Threatened, native plants in the LC

In this section we are exploring the threatened, native plants in the collection. We shall do this in two parts: existing plants in the LC and how this has changed over time.

```{r,  Threatened, native plants in the LC - data}
# LC_native_exisiting
# 
redlistcatA = LC_native_exisiting$redList_category
redlistcatB = LC_native_exisiting$POWO_Red_category

redlist_used = rep(NA, nrow(LC_native_exisiting))
redlist_used[redlistcatA == 'VU' | redlistcatB == 'VU'] = 'VU'
redlist_used[redlistcatA == 'EN' | redlistcatB == 'EN'] = 'EN'
redlist_used[redlistcatA == 'CR' | redlistcatB == 'CR'] = 'CR'
redlist_used[redlistcatA == 'EW' | redlistcatB == 'EW'] = 'EW'
redlist_used[redlistcatA == 'EX' | redlistcatB == 'EX'] = 'EX'

LC_native_exisiting$redlist_used = redlist_used

redlistcat_items = data.frame(table(LC_native_exisiting$redlist_used))
if(nrow(redlistcat_items) == 0){
  redlistcat_items = data.frame(matrix(ncol = 2, nrow = 0))
}
names(redlistcat_items) = c('RedListCategory', 'Items')
redlistcat_items$RedListCategory = as.character(redlistcat_items$RedListCategory)
redlistcat_items$RedListCategory  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(redlistcat_items$RedListCategory, c('VU', 'EN', 'CR', 'EW', 'EX'))] 

if(nrow(redlistcat_items) > 0){
redlistcat_items$percent = paste0(round(redlistcat_items$Items / sum(redlistcat_items$Items)*100,digits = 1),'%')
redlistcat_items <- redlistcat_items %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Items))), 
         pos = Items/2 + dplyr::lead(csum, 1),
         pos = dplyr::if_else(is.na(pos), Items/2, pos))
}

LC_native_exisiting_redlist = LC_native_exisiting[LC_native_exisiting$redlist_used %in% c('VU', 'EN', 'CR', 'EW', 'EX'),]

redlistcat_accessions = data.frame(table(LC_native_exisiting_redlist$redlist_used[match(unique(LC_native_exisiting_redlist$AccNoFull), LC_native_exisiting_redlist$AccNoFull)]))
if(nrow(redlistcat_accessions) == 0){
  redlistcat_accessions = data.frame(matrix(ncol = 2, nrow = 0))
}
names(redlistcat_accessions) = c('RedListCategory', 'Accessions')
redlistcat_items$RedListCategory = as.character(redlistcat_items$RedListCategory)
redlistcat_accessions$RedListCategory  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(redlistcat_accessions$RedListCategory, c('VU', 'EN', 'CR', 'EW', 'EX'))]  

if(nrow(redlistcat_accessions) > 0){
redlistcat_accessions$percent = paste0(round(redlistcat_accessions$Accessions / sum(redlistcat_accessions$Accessions)*100,digits = 1),'%')
redlistcat_accessions <- redlistcat_accessions %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Accessions))), 
         pos = Accessions/2 + dplyr::lead(csum, 1),
         pos = dplyr::if_else(is.na(pos), Accessions/2, pos))
}

if(nrow(redlistcat_items) > 0){
  flex_data = data.frame(redlistcat_items$RedListCategory,
                       redlistcat_items$Items, 
                       paste0(round(redlistcat_items$Items/sum(redlistcat_items$Items)*100,digits=2),'%'),
                       redlistcat_accessions$Accessions,
                       paste0(round(redlistcat_accessions$Accessions/sum(redlistcat_accessions$Accessions)*100,digits=2),'%'))
}else{
  flex_data = data.frame(matrix(ncol = 5, nrow = 0))
}
names(flex_data) = c('Red List Category', 'Number of items', 'Proportion of items', 'Number of accessions', 'Proportion of accessions')

have_redlist = eval = nrow(flex_data)>0
if(export_data){
  exporting_data_store$Threatened_category_of_existing_native_plants_in_the_LC = flex_data
}
```

`r if(!have_redlist){"The LC does not contain any theatened native plants."}`


```{r, Threatened, native plants in the LC - table, eval = have_redlist}
if(report_kind == 'static'){
  flex_data |>
  flextable() |>
  set_caption(caption = paste0('Threatened category of existing native plants in the LC')) |> 
  font(fontname = "Calibri (Body)", part = "all") |> 
  fontsize(size = table_font_size, part = "body") |> 
  # add footer if you want
  # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
  #                colwidths = 4) |> 
  theme_booktabs() |> # default theme
  separate_header() |> 
  align(align = "center", part = "all") |>
  autofit()
}
if(report_kind == 'interactive'){
    DT::datatable(flex_data)

}


```

`r if(have_redlist){"Putting the data into a pie charts yields"}`

```{r,  Threatened, native, existing plants in the LC - static pie charts, echo = FALSE, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'static' & have_redlist}
if(nrow(flex_data) > 0){
  # Basic piechart
pie1 = ggplot(redlistcat_items, aes(x="", y=Items, fill=RedListCategory)) +
  geom_bar(stat="identity", width=1, color="white") +
  ggtheme(base_size = 18) +
  coord_polar("y", start=0) +
  ggtitle('Items') +
  theme(legend.position = "none",
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5))

pie2 = ggplot(redlistcat_accessions, aes(x="", y=Accessions, fill=RedListCategory)) +
  geom_bar(stat="identity", width=1, color="white") +
  ggtheme(base_size = 18) +
  coord_polar("y", start=0) +
  ggtitle('Accessions') +
  theme(legend.position = "none",
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5))

    if(value_on_fig){
      pie1 = pie1 + geom_label_repel(aes(y = pos,label = percent),
                   size = 4.5, nudge_x = 1, show.legend = FALSE, fill = "white")   
      
      pie2 = pie2 + geom_label_repel(aes(y = pos,label = percent),
                   size = 4.5, nudge_x = 1, show.legend = FALSE, fill = "white") 
  }

p = ggpubr::ggarrange(pie1, pie2, ncol=2, nrow=1, common.legend = TRUE, legend="bottom",
                  font.label = list(size = table_font_size))

    if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','native_threatened.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}
p
}

```

`r if(report_kind == 'interactive' & have_redlist){"#####  Threatened category of native species in the LC {.tabset}"}`

`r if(report_kind == 'interactive' & have_redlist){"###### Items"}`
```{r  Threatened Category of existing native items in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'& have_redlist}
text = paste0(redlistcat_items$Items, ' native items with red list category "', redlistcat_items$RedListCategory, '"')
             
fig <- plot_ly(redlistcat_items, labels = ~RedListCategory, values = ~Items, hoverinfo = "text",  hovertext = text, sort = FALSE)
fig <- fig %>% add_pie(hole = 0.6)
fig <- fig %>% layout(title = "",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Pie-red_list_category_native_items.html'))
}

  
```

`r if(report_kind == 'interactive' & have_redlist){"###### Accessions"}`
```{r  Threatened category of existing native accessions in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & have_redlist}
text = paste0(redlistcat_accessions$Accessions, ' native accessions with red list category "', redlistcat_accessions$RedListCategory, '"')
             
fig <- plot_ly(redlistcat_accessions, labels = ~RedListCategory, values = ~Accessions, hoverinfo = "text",  hovertext = text, sort = FALSE)
fig <- fig %>% add_pie(hole = 0.6)
fig <- fig %>% layout(title = "",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Pie-red_list_category_native_accessions.html'))
}

  
```

`r if(report_kind == 'interactive'){"##### {-}"}`



`r if(have_redlist & do_over_time){"We will now look at how the threatened categories of native plants in the LC has changed over time."}`

`r if(!do_over_time){"We do not have the information to explore how hreatened categories of native plants in the LC has changed over time."}`

```{r, Threatened, native plants in the LC - data over time, fig.fullwidth=FALSE, fig.dim = c(10, 6), eval = do_over_time}
# LC_native_over_time contains the over time native plants.
redlistcatA = LC_native_over_time$redList_category
redlistcatB = LC_native_over_time$POWO_Red_category

redlist_used = rep(NA, nrow(LC_native_over_time))
redlist_used[redlistcatA == 'VU' | redlistcatB == 'VU'] = 'VU'
redlist_used[redlistcatA == 'EN' | redlistcatB == 'EN'] = 'EN'
redlist_used[redlistcatA == 'CR' | redlistcatB == 'CR'] = 'CR'
redlist_used[redlistcatA == 'EW' | redlistcatB == 'EW'] = 'EW'
redlist_used[redlistcatA == 'EX' | redlistcatB == 'EX'] = 'EX'

LC_native_over_time$redlist_used = redlist_used

# For each year get the number in each group (species and genera)
over_time_info = lapply(plant_existing_LC_native, function(x){
  garden_current = LC_native_over_time[x,]
  redlistcat_items = data.frame(table(garden_current$redlist_used))
  if(ncol(redlistcat_items) != 2){redlistcat_items = data.frame(matrix(ncol = 2, nrow = 0))}
  names(redlistcat_items) = c('RedListCategory', 'Items')
  
  redlistcat_accessions = data.frame(table(garden_current$redlist_used[match(unique(garden_current$AccNoFull), garden_current$AccNoFull)]))
  if(ncol(redlistcat_accessions) != 2){redlistcat_accessions = data.frame(matrix(ncol = 2, nrow = 0))}
  
  names(redlistcat_accessions) = c('RedListCategory', 'Accessions')
  
  taxa = paste0(garden_current$POWO_taxon_name, ' ', garden_current$POWO_taxon_authors)
  redlistcat_taxa = data.frame(table(garden_current$redlist_used[match(unique(taxa), taxa)]))
  if(ncol(redlistcat_taxa) != 2){redlistcat_taxa = data.frame(matrix(ncol = 2, nrow = 0))}
  
  names(redlistcat_taxa) = c('RedListCategory', 'Accessions')
  
  
  return(list(redlistcat_items = redlistcat_items,
              redlistcat_accessions = redlistcat_accessions,
              redlistcat_taxa = redlistcat_taxa))
})


```

```{r, Threatened, native plants in the LC - static proportion over time, fig.fullwidth=FALSE, fig.dim = c(10, 6), eval = report_kind == 'static' & have_redlist & do_over_time}
to_do = c('items', 'accessions')
if(nrow(flex_data) > 0){
for(i in 1:length(to_do)){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  if(export_data){
    item_name = paste0('Threatened_native_plants_in_the_LC_proportion_over_time_',to_do[i])
  exporting_data_store$AA = data_wanted
  names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
}
  
    p = ggplot(data_wanted, aes(fill=RedListCategory, y=prop, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = "Percentage (%)") +
    guides(fill=guide_legend(title=paste0('RedList Category')))# +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("RedList Category of ", to_do[i] , " that are native in the LC over time"))
    
    if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_threatened_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}
  print(p)

}
}
```

`r if(report_kind == 'static' &  have_redlist & do_over_time){"Note that there might not be many threatened native plants. Therefore, we also include line graphs of the number of each threatened category in the LC."}`

```{r, Threatened, native plants in the LC - number over time, fig.fullwidth=FALSE, fig.dim = c(10, 6), eval = report_kind == 'static' & have_redlist & do_over_time}
to_do = c('items', 'accessions', 'taxa')
if(nrow(flex_data) > 0){
for(i in 1:length(to_do)){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  
  # Create plot.
  p = ggplot(data=data_wanted, aes(x=year, y=data_wanted[,3], group=RedListCategory)) +
  geom_line(aes(color=RedListCategory)) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = paste0('Number of ',to_do[i])) +
    guides(color=guide_legend(title=paste0('RedList Category'))) #+
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("RedList Category of ", to_do[i] , " that are native in the LC over time"))
  
      if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Number_of_threatened_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
      }
  
  print(p)
}
}
```


`r if(report_kind == 'static' & have_redlist & do_over_time){"We do this in two ways, looking at the number of plants in each threatened category ove time and the proportion if each threatened cateogory over all threatened taxa."}`

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"#####  Threatened plants over time graphs {.tabset}"}`

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"###### Items (Number)"}`
```{r  Number of threatened native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
i=1
to_do = c('items', 'accessions', 'taxa')
if(nrow(flex_data) > 0){

  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  
  fig <- plot_ly(data_wanted, x = ~year, y = ~Items, color = ~RedListCategory, type = 'scatter', mode = 'lines+markers')

  fig <- fig |> layout(yaxis = list(title = paste0("Number of ",to_do[i])),
                   xaxis = list(title = "Year"))
  fig <- fig |> layout(hovermode = 'x unified')

 
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line-number_of_threatened_',to_do[i],'_over_time.html'))
}
  
 fig
}

```

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"###### Accessions (Number)"}`
```{r  Number of threatened native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & have_redlist & do_over_time}
i=2
to_do = c('items', 'accessions', 'taxa')
if(nrow(flex_data) > 0){

  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  
  fig <- plot_ly(data_wanted, x = ~year, y = ~Accessions, color = ~RedListCategory, type = 'scatter', mode = 'lines+markers')

  fig <- fig |> layout(yaxis = list(title = paste0("Number of ",to_do[i])),
                   xaxis = list(title = "Year"))
  fig <- fig |> layout(hovermode = 'x unified')

 
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line-number_of_threatened_',to_do[i],'_over_time.html'))
}
  fig
}


```

`r if(report_kind == 'interactive'& have_redlist & do_over_time){"###### Taxa (Number)"}`
```{r  Number of threatened native taxa over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & have_redlist & do_over_time}
i=3
to_do = c('items', 'accessions', 'taxa')
if(nrow(flex_data) > 0){

  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  
  fig <- plot_ly(data_wanted, x = ~year, y = ~Accessions, color = ~RedListCategory, type = 'scatter', mode = 'lines+markers')

  fig <- fig |> layout(yaxis = list(title = paste0("Number of ",to_do[i])),
                   xaxis = list(title = "Year"))
  fig <- fig |> layout(hovermode = 'x unified')

 
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line-number_of_threatened_',to_do[i],'_over_time.html'))
}
  fig
}


```

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"###### Items (Proportion)"}`
```{r  Proportion of threatened native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & have_redlist & do_over_time}
i=1
to_do = c('items', 'accessions')
if(nrow(flex_data) > 0){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  if(export_data){
    item_name = paste0('Threatened_native_plants_in_the_LC_proportion_over_time_',to_do[i])
  exporting_data_store$AA = data_wanted
  names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
}
  
   text = paste0(round(data_wanted$prop,digits=2), '% (',data_wanted$Items,' items) ')
data_wanted$text = text
fig <- plot_ly(data_wanted, x = ~year, y = ~prop, color = ~RedListCategory, type = 'bar', text = ~text,
               hovertemplate = '%{text}',
               sort = FALSE)
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text='Red List Category')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)



if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-native_threatened_over_time_by_item.html'))
}

fig
}

```

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"###### Accessions (Proportion)"}`
```{r  Proportion of threatened native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & have_redlist & do_over_time}
i=2
to_do = c('items', 'accessions')
if(nrow(flex_data) > 0){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  if(export_data){
    item_name = paste0('Threatened_native_plants_in_the_LC_proportion_over_time_',to_do[i])
  exporting_data_store$AA = data_wanted
  names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
}
  
   text = paste0(round(data_wanted$prop,digits=2), '% (',data_wanted$Accessions,' accessions) ')
data_wanted$text = text
fig <- plot_ly(data_wanted, x = ~year, y = ~prop, color = ~RedListCategory, type = 'bar', text = ~text,
               hovertemplate = '%{text}',
               sort = FALSE)
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text='Red List Category')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)



if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-native_threatened_over_time_by_accessions.html'))
}

fig
}

```


`r if(report_kind == 'interactive' & have_redlist & do_over_time){"##### {-}"}`


***

### Provenance of native plants in the LC

In this section we are exploring the provenance of native plants in the collection. We shall do this in two parts: existing plants in the LC and how this has changed over time.

Below we produce a table of provenance for items and accessions.

```{r, Provenance of native plants in the LC}
# LC_native_exisiting the data of existing native plants in the LC. 

prov_items = data.frame(table(LC_native_exisiting$ProvenanceCode))
names(prov_items) = c('Provenance', 'Items')
# prov_items$Provenance  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(prov_items$Provenance, c('G', 'U', 'W', 'Z'))]  
prov_items$percent = paste0(round(prov_items$Items / sum(prov_items$Items)*100,digits = 1),'%')
prov_items <- prov_items %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Items))), 
         pos = Items/2 + dplyr::lead(csum, 1),
         pos = dplyr::if_else(is.na(pos), Items/2, pos))


prov_accessions = data.frame(table(LC_native_exisiting$ProvenanceCode[match(unique(LC_native_exisiting$AccNoFull), LC_native_exisiting$AccNoFull)]))
names(prov_accessions) = c('Provenance', 'Accessions')
# prov_accessions$Provenance  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(prov_accessions$Provenance, c('G', 'U', 'W', 'Z'))]
prov_accessions$percent = paste0(round(prov_accessions$Accessions / sum(prov_accessions$Accessions)*100,digits = 1),'%')
prov_accessions <- prov_accessions %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Accessions))), 
         pos = Accessions/2 + dplyr::lead(csum, 1),
         pos = dplyr::if_else(is.na(pos), Accessions/2, pos))

flex_data = data.frame(prov_items$Provenance,
                       prov_items$Items, 
                       paste0(round(prov_items$Items/sum(prov_items$Items)*100,digits=2),'%'),
                       prov_accessions$Accessions,
                       paste0(round(prov_accessions$Accessions/sum(prov_accessions$Accessions)*100,digits=2),'%'))
names(flex_data) = c('Provenance', 'Number of items', 'Proportion of items', 'Number of accessions', 'Proportion of accessions')

if(export_data){
  exporting_data_store$Provenance_of_existing_native_plants_in_the_LC = flex_data
}

### MANAGEMENT SHEET: Wild/Wild-derived accessions of native plants
```

```{r Provenance of native plants in the LC - Table}
if(report_kind == 'static'){
flex_data |>
  flextable() |>
  set_caption(caption = paste0('Provenance of existing native plants in the LC')) |> 
  font(fontname = "Calibri (Body)", part = "all") |> 
  fontsize(size = table_font_size, part = "body") |> 
  # add footer if you want
  # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
  #                colwidths = 4) |> 
  theme_booktabs() |> # default theme
  separate_header() |> 
  align(align = "center", part = "all") |>
  autofit()
}
if(report_kind == 'interactive'){
    DT::datatable(flex_data)
}

```

Putting the data into a pie charts yields

```{r, Provenance of native plants in the LC pie charts, echo = FALSE, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'static'}
# Basic piechart
pie1 = ggplot(prov_items, aes(x="", y=Items, fill=Provenance)) +
  geom_bar(stat="identity", width=1, color="white") +
  ggtheme(base_size = 18) +
  coord_polar("y", start=0) +
  ggtheme(base_size = 18) +
  ggtitle('Items') +
  scale_fill_discrete() +
  theme(legend.position = "none",
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5))

pie2 = ggplot(prov_accessions, aes(x="", y=Accessions, fill=Provenance)) +
  geom_bar(stat="identity", width=1, color="white") +
  ggtheme(base_size = 18) +
  coord_polar("y", start=0) +
  ggtitle('Accessions') + 
  scale_fill_discrete() +
  theme(legend.position = "none",
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5))

    if(value_on_fig){
      pie1 = pie1 + geom_label_repel(aes(y = pos,label = percent),
                   size = 4.5, nudge_x = 1, show.legend = FALSE, fill = "white")   
      
      pie2 = pie2 + geom_label_repel(aes(y = pos,label = percent),
                   size = 4.5, nudge_x = 1, show.legend = FALSE, fill = "white") 
  }

p = ggpubr::ggarrange(pie1, pie2, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")

if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Provenance_of_native_plants.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}


p

```

`r if(report_kind == 'interactive'){"#####  Provenance of native species in the LC {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Items"}`
```{r  Provenance of existing native items in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
text = paste0(prov_items$Items, ' native items with provenance "', prov_items$Provenance, '"')
             
fig <- plot_ly(prov_items, labels = ~Provenance, values = ~Items, hoverinfo = "text",  hovertext = text, sort = FALSE)
fig <- fig %>% add_pie(hole = 0.6)
fig <- fig %>% layout(title = "",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Pie-provenance_native_items.html'))
}

  
```

`r if(report_kind == 'interactive'){"###### Accessions"}`
```{r  Provenance of existing native accessions in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
text = paste0(prov_accessions$Accessions, ' native accessions with provenance "', prov_accessions$Provenance, '"')
             
fig <- plot_ly(prov_accessions, labels = ~Provenance, values = ~Accessions, hoverinfo = "text",  hovertext = text, sort = FALSE)
fig <- fig %>% add_pie(hole = 0.6)
fig <- fig %>% layout(title = "",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Pie-provenance_native_accessions.html'))
}

  
```

`r if(report_kind == 'interactive'){"##### {-}"}`

`r if(do_over_time){"We will now look at how the provenance of native plants in the LC has changed over time."}`

`r if(!do_over_time){"We do not have the information to explore how provenance of native plants in the LC has changed over time."}`

```{r, Provenance of native plants in the LC- change over time data, eval = do_over_time}
# LC_native_over_time contains the over time native plants.
redlistcatA = LC_native_over_time$redList_category
redlistcatB = LC_native_over_time$POWO_Red_category

redlist_used = rep(NA, nrow(LC_native_over_time))
redlist_used[redlistcatA == 'VU' | redlistcatB == 'VU'] = 'VU'
redlist_used[redlistcatA == 'EN' | redlistcatB == 'EN'] = 'EN'
redlist_used[redlistcatA == 'CR' | redlistcatB == 'CR'] = 'CR'
redlist_used[redlistcatA == 'EW' | redlistcatB == 'EW'] = 'EW'
redlist_used[redlistcatA == 'EX' | redlistcatB == 'EX'] = 'EX'

LC_native_over_time$redlist_used = redlist_used

# For each year get the number in each group (species and genera)
over_time_info = lapply(plant_existing_LC_native, function(x){
  garden_current = LC_native_over_time[x,]
  
  prov_items =  data.frame(table(garden_current$ProvenanceCode))
  if(nrow(prov_items) == 0){
    return(list(prov_items = prov_items,
              prov_accessions = prov_items,
              redlistcat_items = prov_items,
              redlistcat_accessions = prov_items,
              redlistcat_taxa = prov_items))
  }
  names(prov_items) = c('Provenance', 'Items')
  
  prov_accessions = data.frame(table(garden_current$ProvenanceCode[match(unique(garden_current$AccNoFull), garden_current$AccNoFull)]))
  names(prov_accessions) = c('Provenance', 'Accessions')
  
  redlistcat_items = data.frame(table(garden_current$redlist_used))
  if(ncol(redlistcat_items) != 2){redlistcat_items = data.frame(matrix(ncol = 2, nrow = 0))}
  names(redlistcat_items) = c('RedListCategory', 'Items')
  
  redlistcat_accessions = data.frame(table(garden_current$redlist_used[match(unique(garden_current$AccNoFull), garden_current$AccNoFull)]))
  if(ncol(redlistcat_accessions) != 2){redlistcat_accessions = data.frame(matrix(ncol = 2, nrow = 0))}
  
  names(redlistcat_accessions) = c('RedListCategory', 'Accessions')
  
  taxa = paste0(garden_current$POWO_taxon_name, ' ', garden_current$POWO_taxon_authors)
  redlistcat_taxa = data.frame(table(garden_current$redlist_used[match(unique(taxa), taxa)]))
  if(ncol(redlistcat_taxa) != 2){redlistcat_taxa = data.frame(matrix(ncol = 2, nrow = 0))}
  
  names(redlistcat_taxa) = c('RedListCategory', 'Accessions')
  
  
  return(list(prov_items = prov_items,
              prov_accessions = prov_accessions,
              redlistcat_items = redlistcat_items,
              redlistcat_accessions = redlistcat_accessions,
              redlistcat_taxa = redlistcat_taxa))
})
prov_redlist_native_over_time = over_time_info

```

```{r, Provenance of native plants in the LC- change over time figures, fig.fullwidth=FALSE, fig.dim = c(10, 6), , eval = report_kind == 'static' & do_over_time}
to_do = c('items', 'accessions')
for(i in 1:length(to_do)){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('prov',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
    if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)),
                           info_year = info_year,
                           prop = prop)
    data_wanted = rbind(data_wanted, info_year)

  }
  names(data_wanted) = c('year', 'Provenance', 'Items', 'prop')
  data_wanted[,2] = as.character(data_wanted[,2])
  # data_wanted[,2]  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(data_wanted[,2], c('G', 'U', 'W', 'Z') )]
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  
    if(export_data){
    item_name = paste0('Provenance_of_native_plants_in_the_LC_proportion_over_time_',to_do[i])
  exporting_data_store$AA = data_wanted
  names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
    }
  
    p = ggplot(data_wanted, aes(fill=Provenance, y=prop, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = "Percentage (%)") +
    guides(fill=guide_legend(title=paste0('Provenance')))# +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("Provenance of ", to_do[i] , " that are native in the LC over time"))
    
    if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Provenance_of_native_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}
  print(p)

}
```

`r if(report_kind == 'interactive'& do_over_time){"#####  Provenance of native species in the LC over time {.tabset}"}`

`r if(report_kind == 'interactive'& do_over_time){"###### Items"}`
```{r  Provenance of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'& do_over_time}
to_do = c('items', 'accessions')
i=1

# Get the data from over_time_info.
data_wanted = NULL
wanted_item = which(grepl('prov',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
for(jj in 1:length(over_time_info)){
  info_year = over_time_info[[jj]][[wanted_item]]
  if(nrow(info_year) == 0){
    next
  }
  prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
  info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)),
                         info_year = info_year,
                         prop = prop)
  data_wanted = rbind(data_wanted, info_year)

}
names(data_wanted) = c('year', 'Provenance', 'Items', 'prop')
data_wanted[,2] = as.character(data_wanted[,2])
# data_wanted[,2]  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(data_wanted[,2], c('G', 'U', 'W', 'Z') )]
data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))


  if(export_data){
  item_name = paste0('Provenance_of_native_plants_in_the_LC_proportion_over_time_',to_do[i])
exporting_data_store$AA = data_wanted
names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
  }


text = paste0(round(data_wanted$prop,digits=2), '% (',data_wanted$Items,' items) ')
data_wanted$text = text
fig <- plot_ly(data_wanted, x = ~year, y = ~prop, color = ~Provenance, type = 'bar', text = ~text,
               hovertemplate = '%{text}',
               sort = FALSE)
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text='Provenance Type')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-native_provenance_over_time_by_item.html'))
}
```

`r if(report_kind == 'interactive'& do_over_time){"###### Accessions"}`
```{r  Provenance of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'& do_over_time}
to_do = c('items', 'accessions')
i=2

# Get the data from over_time_info.
data_wanted = NULL
wanted_item = which(grepl('prov',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
for(jj in 1:length(over_time_info)){
  info_year = over_time_info[[jj]][[wanted_item]]
  if(nrow(info_year) == 0){
    next
  }
  prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
  info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)),
                         info_year = info_year,
                         prop = prop)
  data_wanted = rbind(data_wanted, info_year)

}
names(data_wanted) = c('year', 'Provenance', 'Items', 'prop')
data_wanted[,2] = as.character(data_wanted[,2])
# data_wanted[,2]  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(data_wanted[,2], c('G', 'U', 'W', 'Z') )]
data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))


  if(export_data){
  item_name = paste0('Provenance_of_native_plants_in_the_LC_proportion_over_time_',to_do[i])
exporting_data_store$AA = data_wanted
names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
  }


text = paste0(round(data_wanted$prop,digits=2), '% (',data_wanted$Items,' accessions) ')
data_wanted$text = text
fig <- plot_ly(data_wanted, x = ~year, y = ~prop, color = ~Provenance, type = 'bar', text = ~text,
               hovertemplate = '%{text}',
               sort = FALSE)
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text='Provenance Type')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)
fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-native_provenance_over_time_by_accession.html'))
}
```

`r if(report_kind == 'interactive'& do_over_time){"##### {-}"}`


***

### Turnover of native plants

This section looks at the turnover (gain/loss) of items and accessions of native plants in the LC.

`r if(!do_over_time){"As we do not have information to explore the data over time we cannot perform turnover analysis."}`

`r if(do_over_time){"We can also view the gain/loss of native items/accessions as a proportion of all gained/lost items each year. This will show if the incoming proportion of native plants was more or less than the proportion of native plants in the collection."}` 

`r if(do_over_time){"##### Items"}`

```{r, Native Turnover - Analyse items, eval = do_over_time}
#Lost items
lost_items = data.frame(t(data.frame(lapply(years, function(year){
  # Accessions for the year
  have_status_year = LC_native_over_time[which(LC_native_over_time$status_year == year),]
  # Number of items lost.
  lost_items = have_status_year[which(have_status_year$ItemStatusType == 'NotExisting'),]
  no_lost_items = nrow(lost_items)
  
  return(c(year,no_lost_items))
}))))
names(lost_items) = c('Year', 'Number of items')
row.names(lost_items) = 1:nrow(lost_items)

# Gained items
gain_items = data.frame(t(data.frame(lapply(years, function(year){
  # Accessions for the year
  have_accYear = LC_native_over_time[which(LC_native_over_time$AccYear == year),]

  no_gain_items = nrow(have_accYear)
  
  return(c(year,no_gain_items))
}))))
names(gain_items) = c('Year', 'Number of items')
row.names(gain_items) = 1:nrow(gain_items)

# items
items = data.frame(lost_items, gain_items[,2], gain_items[,2] - lost_items[,2])
names(items) = c('Year', 'Number of lost items', 'Number of gained items', 'Net number of items')


if(export_data){
  exporting_data_store$Native_turnover_items = items
}
# DT::datatable(items, rownames = FALSE, caption = paste0('Native turnover (items)'),
#               options = list(scrollX =  TRUE,
#                              scrollY = '70vh',
#                              pageLength =  min(nrow(items),200),
#                              columnDefs = list(list(className = 'dt-center', targets = 0:3))
#                              ),
#               escape = FALSE)
```


```{r, Native Turnover - Gain/Loss items figures,  fig.fullwidth=FALSE, fig.dim = c(10, 4), results='hide',fig.keep='all', message=FALSE, eval = report_kind == 'static' & do_over_time}
dataA = data.frame(items[,c(1,2)], rep('Lost', nrow(items)))
names(dataA) = LETTERS[1:3]
dataB = data.frame(items[,c(1,3)], rep('Gain', nrow(items)))
names(dataB) = LETTERS[1:3]
items_data = rbind(dataA, dataB)

p = ggplot(data=items_data, aes(x=A, y=B, group=C)) +
  geom_line(aes(color=C)) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = paste0('Number of items')) +
    guides(color=guide_legend(title=paste0('Gain/Loss'))) +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("Gain/Loss of native items in the LC over time"))
  scale_color_manual(values=c("blue", "red"))
  print(p)
  
  if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_turnover_items.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

  
# Color based on value
color <- ifelse(items$`Net number of items` < 0, "pink", "lightblue")

p = ggplot(items, aes(x = Year, y = `Net number of items`)) +
  geom_bar(stat = "identity",
           show.legend = FALSE,
           fill = color,      # Background color
           color = "white") + # Border color
  # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(title="",
      x ="Year",
      y = paste0('Number of items')) #+
  # theme(text = element_text(size=table_font_size)) +
  # ggtitle(paste0("Net turnover of native items in the LC over time"))
print(p)  

  if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_net_turnover_items.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}
```

`r if(report_kind == 'interactive' & do_over_time){"#####  Turnover of native plants over time figures {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Gain/Loss"}`
```{r  Gain/Loss of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
dataA = data.frame(items[,c(1,2)], rep('Lost', nrow(items)))
names(dataA) = LETTERS[1:3]
dataB = data.frame(items[,c(1,3)], rep('Gain', nrow(items)))
names(dataB) = LETTERS[1:3]
items_data = rbind(dataA, dataB)

colors = c("lightblue", "pink")
fig <- plot_ly(items_data, x = ~A, y = ~B, color = ~C, colors = colors, type = 'scatter', mode = 'lines+markers')
fig <- fig |> layout(yaxis = list(title = paste0("Number of items")),
                   xaxis = list(title = "Year"))
fig <- fig |> layout(hovermode = 'x unified')

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','line-Gain_Loss_of_native_items.html'))
}

fig
```

`r if(report_kind == 'interactive' & do_over_time){"###### Net"}`
```{r  Net turnover of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
colors = c("pink", "lightblue")
color <- ifelse(items$`Net number of items` < 0, "pink", "lightblue")
text = paste0('Net: ', items$`Net number of items`, '<br>',
              'Gain: ', items$`Number of gained items`, '<br>',
              'Loss: ', items$`Number of lost items`, '<br>'
              )
fig = plot_ly(items, x = ~Year, y = ~`Net number of items`, type = 'bar', hoverinfo = "text",  hovertext = text, color = color, colors = rev(colors), showlegend = FALSE)
fig <- fig |> layout(yaxis = list(title = paste0("Net number of items")),
                   xaxis = list(title = "Year"))

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Bar-Net_turnover_of_native_items.html'))
}

fig

```

`r if(report_kind == 'interactive' & do_over_time){"##### {-}"}`

`r if(do_over_time){"As proportion of loss/gained items each year"}`

```{r, Native Turnover - Gain/Loss items proportion data, message=FALSE, eval = do_over_time}
lost_items_all = data.frame(t(data.frame(lapply(years, function(year){
  # Accessions for the year
  have_status_year = report[report$status_year == year,]
  
  # Number of items lost.
  lost_items = have_status_year[have_status_year$ItemStatusType == 'NotExisting',]
  no_lost_items = nrow(lost_items)
  
  return(c(year,no_lost_items))
}))))
names(lost_items_all) = c('Year', 'Number of items')
row.names(lost_items_all) = 1:nrow(lost_items_all)

# Gained items
gain_items_all = data.frame(t(data.frame(lapply(years, function(year){
  # Accessions for the year
  have_accYear = report[report$AccYear == year,]

  no_gain_items = nrow(have_accYear)
  
  return(c(year,no_gain_items))
}))))
names(gain_items_all) = c('Year', 'Number of items')
row.names(gain_items_all) = 1:nrow(gain_items_all)

# items
items_all = data.frame(lost_items_all, gain_items_all[,2], gain_items_all[,2] - lost_items_all[,2])
names(items_all) = c('Year', 'Number of lost items', 'Number of gained items', 'Net number of items')

if(export_data){
  exporting_data_store$Whole_LC_turnover_items = items_all
}
```

```{r, Native Turnover - Gain/Loss items proportion figures, fig.fullwidth=TRUE, fig.dim = c(10, 4), results='hide',fig.keep='all', message=FALSE,  eval = report_kind == 'static' & do_over_time}
### Plot of proportion of gained items being native.
plot_data = data.frame(year = gain_items$Year,
                       native = gain_items$`Number of items`, 
                       all = gain_items_all$`Number of items`)
plot_data$prop_native = round(plot_data$native / plot_data$all *100,digits=2)
native_counts = native_data$Items
total_counts = LC_data$Items
plot_data$native_LC = round(native_counts / total_counts *100,digits=2)

 p = ggplot(plot_data, aes(y=prop_native, x=year)) + 
    geom_bar(stat="identity", width = 1, fill = 'lightblue', col = 'black') +
   geom_point(aes(x = year, y = native_LC))+
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = "Native Percentage (%)") +
    guides(fill=guide_legend(title=paste0('Provenance'))) +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("Proportion of gained items that are native in the LC over time"))
      labs(caption = "Black points are proportion of items in the LC that are native each year.")
  print(p)
  
      if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_gain_items_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
  
### Plot of proportion of lost items being native.

plot_data = data.frame(year = lost_items$Year,
                       native = lost_items$`Number of items`, 
                       all = lost_items_all$`Number of items`)
plot_data$prop_native = round(plot_data$native / plot_data$all *100,digits=2)
native_counts = native_data$Items
total_counts = LC_data$Items
plot_data$native_LC = round(native_counts / total_counts *100,digits=2)

 p = ggplot(plot_data, aes(y=prop_native, x=year)) + 
    geom_bar(stat="identity", width = 1, fill = 'pink', col = 'black') +
   geom_point(aes(x = year, y = native_LC))+
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = "Native Percentage (%)") +
    guides(fill=guide_legend(title=paste0('Provenance'))) +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("Proportion of lost items that are native in the LC over time"))
      labs(caption = "Black points are proportion of items in the LC that are native each year.")
  print(p)

        if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_loss_items_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
```

`r if(report_kind == 'interactive' & do_over_time){"#####  Turnover of native plants over time figures (proportion) {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Gain"}`
```{r  Gain proportion of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
### Plot of proportion of gained items being native.
plot_data = data.frame(year = gain_items$Year,
                       native = gain_items$`Number of items`, 
                       all = gain_items_all$`Number of items`)
plot_data$prop_native = round(plot_data$native / plot_data$all *100,digits=2)
native_counts = native_data$Items
total_counts = LC_data$Items
plot_data$native_LC = round(native_counts / total_counts *100,digits=2)

fig = plot_ly(plot_data, x = ~year, y = ~prop_native, type = 'bar', name = '% Native items gained', marker = list(color = 'lightblue'))
fig <- fig |> layout(yaxis = list(title = paste0("Native percentage")),
                   xaxis = list(title = "Year"))
fig <- fig |> add_trace(x = ~year, y = ~native_LC, type = 'scatter', mode = 'markers', name = '% Native items in LC', marker = list(color = 'black'))
fig <- fig |> layout(hovermode = 'x unified')

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_gained_native_items.html'))
}

fig
```

`r if(report_kind == 'interactive' & do_over_time){"###### Loss"}`
```{r  Loss proportion of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
plot_data = data.frame(year = lost_items$Year,
                       native = lost_items$`Number of items`, 
                       all = lost_items_all$`Number of items`)
plot_data$prop_native = round(plot_data$native / plot_data$all *100,digits=2)
native_counts = native_data$Items
total_counts = LC_data$Items
plot_data$native_LC = round(native_counts / total_counts *100,digits=2)

fig = plot_ly(plot_data, x = ~year, y = ~prop_native, type = 'bar', name = '% Native items lost',  marker = list(color = 'pink'))
fig <- fig |> layout(yaxis = list(title = paste0("Native percentage")),
                   xaxis = list(title = "Year"))
fig <- fig |> add_trace(x = ~year, y = ~native_LC, type = 'scatter', mode = 'markers', name = '% Native items in LC',  marker = list(color = 'black'))
fig <- fig |> layout(hovermode = 'x unified')

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_lost_native_items.html'))
}

fig
```

`r if(report_kind == 'interactive'& do_over_time){"##### {-}"}`


`r if(do_over_time){"##### Accessions"}`

`r if(do_over_time){"We now create the same figures using accessions rather than items, where an accession is lost only when all the items from the accession are not existing. "}`

```{r, Native Turnover - Analyse accessions, eval = do_over_time}

accessions_alive = unique(LC_native_over_time$AccNoFull[LC_native_over_time$ItemStatusType == 'Existing'])
accessions = LC_native_over_time[!LC_native_over_time$AccNoFull %in% accessions_alive,]

# To reduce from items to accessions only use the most recent status_year
accessions = accessions[rev(order(accessions$status_year)),]
unique_accessions = unique(accessions$AccNoFull)
accessions = accessions[match(unique_accessions, accessions$AccNoFull),]  


#Lost accessions
lost_accessions = data.frame(t(data.frame(lapply(years, function(year){
  # Accessions for the year
  have_status_year = accessions[which(accessions$status_year == year),]
  
  # Number of accessions lost.
  lost_accessions = have_status_year[which(have_status_year$ItemStatusType == 'NotExisting'),]
  no_lost_accessions = nrow(lost_accessions)
  
  return(c(year,no_lost_accessions))
}))))
names(lost_accessions) = c('Year', 'Number of accessions')
row.names(lost_accessions) = 1:nrow(lost_accessions)

# Gained accessions
gain_accessions = data.frame(t(data.frame(lapply(years, function(year){
  # Accessions for the year
  have_accYear = accessions[which(accessions$AccYear == year),]

  no_gain_accessions = nrow(have_accYear)
  
  return(c(year,no_gain_accessions))
}))))
names(gain_accessions) = c('Year', 'Number of accessions')
row.names(gain_accessions) = 1:nrow(gain_accessions)

# accessions
accessions = data.frame(lost_accessions, gain_accessions[,2], gain_accessions[,2] - lost_accessions[,2])
names(accessions) = c('Year', 'Number of lost accessions', 'Number of gained accessions', 'Net number of accessions')


if(export_data){
  exporting_data_store$Native_turnover_accessions = accessions
}

# DT::datatable(accessions, rownames = FALSE, caption = paste0('Native turnover (accessions)'),
#               options = list(scrollX =  TRUE,
#                              scrollY = '70vh',
#                              pageLength =  min(nrow(accessions),200),
#                              columnDefs = list(list(className = 'dt-center', targets = 0:3))
#                              ),
#               escape = FALSE)
```

```{r, Native Turnover - Gain/Loss accessions figures,  fig.fullwidth=FALSE, fig.dim = c(10, 4), results='hide',fig.keep='all', message=FALSE, eval = report_kind == 'static' & do_over_time}
dataA = data.frame(accessions[,c(1,2)], rep('Lost', nrow(accessions)))
names(dataA) = LETTERS[1:3]
dataB = data.frame(accessions[,c(1,3)], rep('Gain', nrow(accessions)))
names(dataB) = LETTERS[1:3]
accessions_data = rbind(dataA, dataB)

p = ggplot(data=accessions_data, aes(x=A, y=B, group=C)) +
  geom_line(aes(color=C)) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = paste0('Number of accessions')) +
    guides(color=guide_legend(title=paste0('Gain/Loss'))) +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("Gain/Loss of native accessions in the LC over time"))
  scale_color_manual(values=c("blue", "red"))
  print(p)
  
  
    if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_turnover_accessions.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
  
# Color based on value
color <- ifelse(accessions$`Net number of accessions` < 0, "pink", "lightblue")

p = ggplot(accessions, aes(x = Year, y = `Net number of accessions`)) +
  geom_bar(stat = "identity",
           show.legend = FALSE,
           fill = color,      # Background color
           color = "white") + # Border color
  # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(title="",
      x ="Year",
      y = paste0('Number of accessions')) #+
  # theme(text = element_text(size=table_font_size))
  # ggtitle(paste0("Net turnover of native accessions in the LC over time"))
print(p)  

    if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_net_turnover_accessions.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
```


`r if(report_kind == 'interactive' & do_over_time){"#####  Turnover of native plants over time figures {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Gain/Loss"}`
```{r  Gain/Loss of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
dataA = data.frame(accessions[,c(1,2)], rep('Lost', nrow(accessions)))
names(dataA) = LETTERS[1:3]
dataB = data.frame(accessions[,c(1,3)], rep('Gain', nrow(accessions)))
names(dataB) = LETTERS[1:3]
accessions_data = rbind(dataA, dataB)

colors = c("lightblue", "pink")
fig <- plot_ly(accessions_data, x = ~A, y = ~B, color = ~C, colors = colors, type = 'scatter', mode = 'lines+markers')
fig <- fig |> layout(yaxis = list(title = paste0("Number of accessions")),
                   xaxis = list(title = "Year"))
fig <- fig |> layout(hovermode = 'x unified')

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','line-Gain_Loss_of_native_accessions.html'))
}

fig
```

`r if(report_kind == 'interactive' & do_over_time){"###### Net"}`
```{r  Net turnover of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
colors = c("pink", "lightblue")
color <- ifelse(accessions$`Net number of accessions` < 0, "pink", "lightblue")
text = paste0('Net: ', accessions$`Net number of accessions`, '<br>',
              'Gain: ', accessions$`Number of gained accessions`, '<br>',
              'Loss: ', accessions$`Number of lost accessions`, '<br>'
              )
fig = plot_ly(accessions, x = ~Year, y = ~`Net number of accessions`, type = 'bar', hoverinfo = "text",  hovertext = text, color = color, colors = rev(colors), showlegend = FALSE)
fig <- fig |> layout(yaxis = list(title = paste0("Net number of accessions")),
                   xaxis = list(title = "Year"))

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Bar-Net_turnover_of_native_accessions.html'))
}

fig

```

`r if(report_kind == 'interactive' & do_over_time){"##### {-}"}`

`r if(do_over_time){"As proportion of loss/gained items each year"}`

```{r, Native Turnover - Gain/Loss accessions proportion data, eval = do_over_time}
accessions_all = report[rev(order(report$status_year)),]
unique_accessions_all = unique(accessions_all$AccNoFull)
accessions_all = accessions_all[match(unique_accessions_all, accessions_all$AccNoFull),]  


lost_accessions_all = data.frame(t(data.frame(lapply(years, function(year){
  # Accessions for the year
  have_status_year = accessions_all[accessions_all$status_year == year,]
  
  # Number of accessions lost.
  lost_accessions = have_status_year[have_status_year$ItemStatusType == 'NotExisting',]
  no_lost_accessions = nrow(lost_accessions)
  
  return(c(year,no_lost_accessions))
}))))
names(lost_accessions_all) = c('Year', 'Number of accessions')
row.names(lost_accessions_all) = 1:nrow(lost_accessions_all)

# Gained accessions
gain_accessions_all = data.frame(t(data.frame(lapply(years, function(year){
  # Accessions for the year
  have_accYear = accessions_all[accessions_all$AccYear == year,]

  no_gain_accessions = nrow(have_accYear)
  
  return(c(year,no_gain_accessions))
}))))
names(gain_accessions_all) = c('Year', 'Number of accessions')
row.names(gain_accessions_all) = 1:nrow(gain_accessions_all)

# accessions
accessions_all = data.frame(lost_accessions_all, gain_accessions_all[,2], gain_accessions_all[,2] - lost_accessions_all[,2])
names(accessions_all) = c('Year', 'Number of lost accessions', 'Number of gained accessions', 'Net number of accessions')


if(export_data){
  exporting_data_store$Whole_LC_turnover_accessions = accessions_all
}
```

```{r, Native Turnover - Gain/Loss accessions proportion figures, fig.fullwidth=TRUE, fig.dim = c(10, 4), results='hide',fig.keep='all', message=FALSE, eval = report_kind == 'static' & do_over_time}
### Plot of proportion of gained accessions being native.

plot_data = data.frame(year = gain_accessions$Year,
                       native = gain_accessions$`Number of accessions`, 
                       all = gain_accessions_all$`Number of accessions`)
plot_data$prop_native = round(plot_data$native / plot_data$all *100,digits=2)
native_counts = native_data$Accessions
total_counts = LC_data$Accessions
plot_data$native_LC = round(native_counts / total_counts *100,digits=2)

 p = ggplot(plot_data, aes(y=prop_native, x=year)) + 
    geom_bar(stat="identity", width = 1, fill = 'lightblue', col = 'black') +
   geom_point(aes(x = year, y = native_LC))+
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = "Native Percentage (%)") +
    guides(fill=guide_legend(title=paste0('Provenance'))) +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("Proportion of gained accessions that are native in the LC over time"))
      labs(caption = "Black points are proportion of accessions in the LC that are native each year.")
  print(p)
  
          if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_gain_accessions_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
  
### Plot of proportion of lost accessions being native.

plot_data = data.frame(year = lost_accessions$Year,
                       native = lost_accessions$`Number of accessions`, 
                       all = lost_accessions_all$`Number of accessions`)
plot_data$prop_native = round(plot_data$native / plot_data$all *100,digits=2)
native_counts = native_data$Accessions
total_counts = LC_data$Accessions
plot_data$native_LC = round(native_counts / total_counts *100,digits=2)

 p = ggplot(plot_data, aes(y=prop_native, x=year)) + 
    geom_bar(stat="identity", width = 1, fill = 'pink', col = 'black') +
   geom_point(aes(x = year, y = native_LC))+
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = "Native Percentage (%)") +
    guides(fill=guide_legend(title=paste0('Provenance'))) +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("Proportion of lost accessions that are native in the LC over time"))
      labs(caption = "Black points are proportion of accessions in the LC that are native each year.")
  print(p)

          if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_loss_accessions_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
```


`r if(report_kind == 'interactive' & do_over_time){"#####  Turnover of native plants over time figures (proportion) {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Gain"}`
```{r  Gain proportion of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
### Plot of proportion of gained items being native.
plot_data = data.frame(year = gain_accessions$Year,
                       native = gain_accessions$`Number of accessions`, 
                       all = gain_accessions_all$`Number of accessions`)
plot_data$prop_native = round(plot_data$native / plot_data$all *100,digits=2)
native_counts = native_data$Accessions
total_counts = LC_data$Accessions
plot_data$native_LC = round(native_counts / total_counts *100,digits=2)

fig = plot_ly(plot_data, x = ~year, y = ~prop_native, type = 'bar', name = '% Native accessions gained', marker = list(color = 'lightblue'))
fig <- fig |> layout(yaxis = list(title = paste0("Native percentage")),
                   xaxis = list(title = "Year"))
fig <- fig |> add_trace(x = ~year, y = ~native_LC, type = 'scatter', mode = 'markers', name = '% Native accessions in LC', marker = list(color = 'black'))
fig <- fig |> layout(hovermode = 'x unified')

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_gained_native_accessions.html'))
}

fig
```

`r if(report_kind == 'interactive' & do_over_time){"###### Loss"}`
```{r  Loss proportion of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_over_time}
plot_data = data.frame(year = lost_accessions$Year,
                       native = lost_accessions$`Number of accessions`, 
                       all = lost_accessions_all$`Number of accessions`)
plot_data$prop_native = round(plot_data$native / plot_data$all *100,digits=2)
native_counts = native_data$Accessions
total_counts = LC_data$Accessions
plot_data$native_LC = round(native_counts / total_counts *100,digits=2)

fig = plot_ly(plot_data, x = ~year, y = ~prop_native, type = 'bar', name = '% Native accessions lost',  marker = list(color = 'pink'))
fig <- fig |> layout(yaxis = list(title = paste0("Native percentage")),
                   xaxis = list(title = "Year"))
fig <- fig |> add_trace(x = ~year, y = ~native_LC, type = 'scatter', mode = 'markers', name = '% Native accessions in LC',  marker = list(color = 'black'))
fig <- fig |> layout(hovermode = 'x unified')

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_lost_native_accessions.html'))
}

fig
```

`r if(report_kind == 'interactive' & do_over_time){"##### {-}"}`


***

### Survival of native plants in the LC 

```{r, survival analysis, echo = FALSE}
# Remove those without accession year.
do_survival = TRUE
if(!do_over_time){
  do_survival = FALSE
}
enriched_report_wanted = enriched_report[as.numeric(enriched_report$AccYear) > 1650 & as.numeric(enriched_report$AccYear) < as.numeric(format(Sys.Date(),'%Y')),]

# Create a data-frame of the information we want.
wanted_data = data.frame(enriched_report_wanted$TaxonNameFull, enriched_report_wanted$AccYear, enriched_report_wanted$ItemStatusDate, enriched_report_wanted$ItemStatusType)
names(wanted_data) = c('TaxonNameFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType')

wanted_data = wanted_data[!is.na(wanted_data$ItemStatusDate),]
wanted_data = wanted_data[!is.na(wanted_data$AccYear),]
if(nrow(wanted_data) == 0){
  do_survival = FALSE
}
if(do_survival){
  #Set accession date (just chose the first of June)
pre_date = rep(NA, length(wanted_data$AccYear))
pre_date = paste(wanted_data$AccYear, "06", "01", sep = "-")

# Convert ItemStatusDate into all dates (i.e if month or day missing add one)
post_date = rep(as.character(Sys.Date()), length(wanted_data$AccYear))
post_date[which(wanted_data$ItemStatusType == "NotExisting")] = as.character(wanted_data$ItemStatusDate[which(wanted_data$ItemStatusType == 
  "NotExisting")])

post_date[which(stringr::str_length(post_date) == 4)] = paste(post_date[which(stringr::str_length(post_date) == 
  4)], "12", "28", sep = "-")
post_date[which(stringr::str_length(post_date) == 7)] = paste(post_date[which(stringr::str_length(post_date) == 
  7)], "28", sep = "-")

# Create the time difference between accession and ItemStatusDate.
time = as.numeric(as.Date(post_date) - as.Date(pre_date)) /365.25

# Create status to tell whether the last item status date was item death or just a check.
status = rep(0,length(wanted_data$AccYear))
status[wanted_data$ItemStatusType == 'NotExisting'] = 1

wanted_data = data.frame(wanted_data, AccessionDate = pre_date,
                         StatusDate = post_date,
                         time = time,
                         Alive_dead = status)


# We haven't removed those with missing ItemStatusDate, these will most likely have negative time so remove these.
missing_itemstatusdate_index = which(wanted_data$time > 0)
wanted_data = wanted_data[missing_itemstatusdate_index,]
enriched_report = enriched_report[missing_itemstatusdate_index,]


## 1) Native or non-native (or doesn't have geography info.)
location = location_code
#Get which records are native.
native_non_native = rep('Not Native',nrow(enriched_report))
native_non_native[which(enriched_report$POWO_plant_name_id %in% native_plant_ids)] = 'Native'
native_non_native[is.na(enriched_report$POWO_Dist_000_area_code_l3)] = NA

data_to_analyse = data.frame(wanted_data$TaxonNameFull,
                             wanted_data$time,
                             wanted_data$Alive_dead,
                             native_non_native)
names(data_to_analyse) = c("TaxonNameFull", "time", "Alive_dead", "native_status")

if(export_data){
  exporting_data_store$Data_for_survival = data_to_analyse
}
}

```

In this section, we look at whether the survival period of native plants are longer than non-native plants. Note that plants can be long-lived and most collections will have plants that are older than the collection itself. Therefore, we need to lean into tools from statistics to analyse the lifespan of plants in a collection. Herein, we use survival analysis to calculate the probability that plants will survive for any number of years. 

`r if(FALSE){"Note that \"lifespan\" may be a preferable way to describe this section, however we use survival because accessioned plants may arrive in the garden already a few years old. As we will be using the accession date and item status date to determine the length of time plants have been in the collection this might not align perfectly with the age of the plant. Moreover, it is important to be aware that the data used in the analysis is not \"perfect\" in the sense that plants might have already died a large period of time prior to an audit, moreover this analysis doesn't account for circumstances of plant death which could be natural, or for renovation of a section of the garden, etc. "}`

In the following analysis, we assume that existing items are still existing to the present day. Therefore, even if the last item status date was a year ago if the plant was found to be existing then we assume it is still existing now. (I could change this to just use alive at Item status date, it won't make much difference. )

Since we only have the accession year we set the accession date to be the middle of the year (1st June).

Native plants are any that are from `r location_name`, non-native plants are plants that have geographic information (matched to WCVP) but are not from `r location_name`. This means that we exclude plants not matched to WCVP for the comparison (cultivars, etc). 
 
`r if(do_survival){"Below we show the survival probability (Kaplan-Meier plots) for native and non-native plants."}`

`r if(!do_survival){"Survival analysis cannot be performed. (This could be due to lack of accession year or item status date."}`

```{r, survival analysis - figure, echo  = FALSE, fig.fullwidth=TRUE, fig.dim = c(10, 6), results='hide',fig.keep='all', message=FALSE, eval = do_survival}
# Create Kaplan-Meier plots comparing native and not native
p = ggsurvfit::survfit2(survival::Surv(time, Alive_dead) ~ native_status, data = data_to_analyse) |> 
  ggsurvfit::ggsurvfit() +
  ggplot2::labs(
    x = "Years",
    y = "Overall survival probability"
  ) + 
  ggsurvfit::add_confidence_interval()+
    # theme(text = element_text(size=table_font_size))+
  # labs(title=paste0("Survival probability of native and non native plants in the LC")) +
  labs(caption = "Shaded region represents 95% confidence interval.") +
  ggtheme() + scale_colour_discrete()

if(report_kind == 'static'){
  if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','native_notnative_survival.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
  }
  p
} 
```

```{r survival_analysis - interactive figure, eval = do_survival}
if(report_kind == 'interactive'){
  fig = ggplotly(p)
  fig$x$layout$title$text =''
  fig$x$layout$legend$title$text = 'Type of Taxa'
  fig$x$layout$hovermode = 'x unified'
  fig$x$data[[2]]$name = 'Not Native'
  fig$x$data[[2]]$legendgroup = 'Not Native'
  fig$x$data[[1]]$name = 'Native'
  fig$x$data[[1]]$legendgroup = 'Native'
  fig$x$data[[4]]$hoverinfo = 'none'
  fig$x$data[[3]]$hoverinfo = 'none'
  
  time_region = round(fig$x$data[[4]]$text |>
    stringr::str_extract(pattern = 'time:[ ]+[0-9\\.]+') |>
    stringr::str_extract(pattern = '[0-9]+\\.[0-9]+') |>
    as.numeric(), digits = 3) 
  lower_region = round(fig$x$data[[4]]$text |>
    stringr::str_extract(pattern = 'conf.low: [0-9\\.]+') |>
    stringr::str_extract(pattern = '[01]\\.[0-9]+') |>
    as.numeric()*100,digits = 2)
  upper_region = round(fig$x$data[[4]]$text |>
    stringr::str_extract(pattern = 'conf.high: [0-9\\.]+') |>
    stringr::str_extract(pattern = '[01]\\.[0-9]+') |>
    as.numeric()*100, digits = 2)
  time = round(fig$x$data[[2]]$x,digits = 3)
  lower = lower_region[match(time, time_region)]
  upper = upper_region[match(time, time_region)]
  
  fig$x$data[[2]]$text = paste0(#'Age (years): ',round(fig$x$data[[2]]$x,digits = ), '<br>',
                                'Survival probability: ', round(fig$x$data[[2]]$y*100,digits = 2),'% <br>',
                                '95% confidence Interval: (', lower,'%, ', upper, '%)')
  
    time_region = round(fig$x$data[[3]]$text |>
    stringr::str_extract(pattern = 'time:[ ]+[0-9\\.]+') |>
    stringr::str_extract(pattern = '[0-9]+\\.[0-9]+') |>
    as.numeric(), digits = 3) 
  lower_region = round(fig$x$data[[3]]$text |>
    stringr::str_extract(pattern = 'conf.low: [0-9\\.]+') |>
    stringr::str_extract(pattern = '[01]\\.[0-9]+') |>
    as.numeric()*100,digits = 2)
  upper_region = round(fig$x$data[[3]]$text |>
    stringr::str_extract(pattern = 'conf.high: [0-9\\.]+') |>
    stringr::str_extract(pattern = '[01]\\.[0-9]+') |>
    as.numeric()*100, digits = 2)
  time = round(fig$x$data[[1]]$x,digits = 3)
  lower = lower_region[match(time, time_region)]
  upper = upper_region[match(time, time_region)]
  
  fig$x$data[[1]]$text = paste0('Age (years): ',round(fig$x$data[[1]]$x,digits = ), '<br>',
                                'Survival probability: ', round(fig$x$data[[1]]$y*100,digits = 2),'% <br>',
                                '95% confidence Interval: (', lower,'%, ', upper, '%)')
  
  if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Native_survival_analysis.html'))
}

fig
}
```

`r if(do_survival){"Next we give the survival probability for 10-year intervals."}`

```{r, survival analysis - tables, echo = FALSE, message=FALSE, eval = do_survival}
# We can also get survival probabilities in a table
# 
ggsurvfit::survfit2(survival::Surv(time, Alive_dead) ~ native_status, data = data_to_analyse) |> 
  gtsummary::tbl_survfit(
    times = c(2,5,10),
    label_header = "{time}-year survival (95% CI)")

ggsurvfit::survfit2(survival::Surv(time, Alive_dead) ~ native_status, data = data_to_analyse) |> 
  gtsummary::tbl_survfit(
    times = c(25,50,100),
    label_header = "{time}-year survival (95% CI)")
# ggsurvfit::survfit2(survival::Surv(time, Alive_dead) ~ native_status, data = data_to_analyse) |> 
#   gtsummary::tbl_survfit(
#     times = c(70,80,90),
#     label_header = "{time}-year survival (95% CI)")
```


```{r, export data, echo = FALSE}
if(export_data){
  #remove all data except wanted to be exported.
  # rm(list=setdiff(ls(), c("exporting_data_store", "output_dir", "collection")))
  # Do some changes (which we do in smaller chunks in the actual code.)
  
  exporting_data_store$Number_of_native_species_in_LC_over_time
  Proportion_species_in_LC_over_time = exporting_data_store$Total_species_in_LC_over_time
  for(i in 2:5){
    Proportion_species_in_LC_over_time[[i]] = round(exporting_data_store$Number_of_native_species_in_LC_over_time[[i]] / exporting_data_store$Total_species_in_LC_over_time[[i]] *100, digits = 2)
  }
  exporting_data_store$Total_species_in_LC_over_time = Proportion_species_in_LC_over_time
  names(exporting_data_store)[ names(exporting_data_store) == 'Total_species_in_LC_over_time'] = 'Proportion_species_in_LC_over_time'
  
  # save as .rda.
  save(exporting_data_store, file = paste0(output_dir,'/',collection, '_native_data.rda'))
  
}
```

```{r, save excel version of data, eval = save_excel & export_data}
library(xlsx)
file = paste0(output_dir,'/',collection, '_native_data.xlsx')
wb <- createWorkbook()
datas <- exporting_data_store
data_titles = names(datas) |> stringr::str_replace_all('_', ' ')
sheetnames <- paste0('Sheet ', 1:length(datas))
sheets <- lapply(sheetnames,  createSheet, wb = wb)
for(i in 1:length(datas)){
  # First we'll create a new row, let's put the title on row 2
  titleRow <-createRow(sheets[[i]], rowIndex=1)
  # We'll add a cell to that row (in the first column) for the title
  sheetTitle <-createCell(titleRow, colIndex=1)
  # We'll add the sheet title to the cell
  setCellValue(sheetTitle[[1,1]], data_titles[i])
  # We'll choose the formatting for the cell
  xlsx::addDataFrame(datas[[i]], sheet = sheets[[i]],row.names = FALSE, startRow = 3)
}
saveWorkbook(wb, file = file)

```
***
