---
title: "Retrospective Overview of the Collection - New Accessions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(ggplot2)
library(plotly)
library(rjson)
library(xlsx)
library(sf)
```


```{r, echo = F}
# Inputs
# As an input we get 
# - Enriched report
# - min_year for retrospective reports.
# - collection the name to be printed in the report.

# load('/Users/jakepowell/Cambridge/Enriched reports/CUBG_enriched_report.rda')
# report = enriched_report
# collection = 'CUBG'
# create_excel = TRUE
# save_widgets = TRUE
# min_year = 1980
# collection_coords = c(0.128305,52.193403)
# do_coord = !is.na(collection_coords[1])
# output_dir = getwd()

report = enriched_report

if(save_widgets){
  widget_dir = paste0(output_dir,'/',collection,'_Widgets')
  dir.create(widget_dir,showWarnings = FALSE)
  
  whole_collection_widget_dir = paste0(widget_dir,'/Whole_collection')
  dir.create(whole_collection_widget_dir,showWarnings = FALSE)
    
  new_accessions_widget_dir = paste0(widget_dir,'/New_accessions')
  dir.create(new_accessions_widget_dir,showWarnings = FALSE)
  
  lost_accessions_widget_dir = paste0(widget_dir,'/Lost_accessions')
  dir.create(lost_accessions_widget_dir,showWarnings = FALSE)
  
  net_accessions_widget_dir = paste0(widget_dir,'/Net_accessions')
  dir.create(net_accessions_widget_dir,showWarnings = FALSE)
}

# Add endemic/notendemic split.
report$endemic = rep(NA, nrow(report))
report$endemic[!is.na(report$POWO_Dist_area_code_l3)] = 'Not Endemic'
endemic_index = which(stringr::str_length(report$POWO_Dist_area_code_l3) ==3)
report$endemic[endemic_index] = 'Endemic'

# Add threatened/notthreatened split.
report$threatened = rep('Not Threatened', nrow(report))
threat_index = which(report$redList_category %in% c('VU','EN','CR', 'EW', 'EX'))
report$threatened[threat_index] = 'Threatened'

# Update the original family to only have first letter capitalised.
report$Family = stringr::str_to_title(report$Family)

# Set the min year of the report
if(is.null(min_year)){
  min_year = min(enriched_report$AccYear[enriched_report$AccYear>1750],na.rm = T) + 1
}

# Load the geographic information.
wgsrpd3 = BGSmartR::wgsrpd3

# Extract the location of the collection (if we have coordinates)
if(do_coord){
pnts <- data.frame( x = collection_coords[1], y = collection_coords[2])

# create a points collection
pnts_sf <- do.call("st_sfc",c(lapply(1:nrow(pnts), 
                                     function(i) {sf::st_point(as.numeric(pnts[i, ]))}), list("crs" = 4326))) 

pnts_trans <- st_transform(pnts_sf, 2163) # apply transformation to pnts sf
tt1_trans <- st_transform(wgsrpd3, 2163)      # apply transformation to polygons sf

intersect = as.vector(st_intersects(tt1_trans, pnts_trans, sparse = FALSE))
collection_geog_details = wgsrpd3[which(intersect),-5]
level3_name = collection_geog_details$LEVEL3_NAM
level3_code = collection_geog_details$LEVEL3_COD
level2_name = collection_geog_details$LEVEL2_NAM
level2_code = collection_geog_details$LEVEL2_COD
level1_name = collection_geog_details$LEVEL1_NAM
level1_code = collection_geog_details$LEVEL1_COD

text = paste0(level3_name, ', ', level2_name, ', ', level1_name,'.')

}

# Get the current year.
year_cur = as.numeric(format(Sys.Date(),'%Y'))

# extract and calculate the number of bad accessions years.
bad_accessions_index = which(report$AccYear < 1750 | report$AccYear> year_cur | is.na(report$AccYear))
if(length(bad_accessions_index)>0){
  bad_accessions = report[which(report$AccYear < 1750 | report$AccYear> year_cur | is.na(report$AccYear)),]
  no_bad_accessions = nrow(bad_accessions)
  #Export the bad accessions.
  # write.csv(bad_accessions[,1:(which(names(bad_accessions) == 'sanitised_taxon')-1)], file = paste0(output_dir,'/', collection,'_issue_accession_year.csv'), row.names = FALSE)
  # Reduce report to not include those with bad accessions.
  report = report[-which(report$AccYear < 1750 | report$AccYear> year_cur |  is.na(report$AccYear)),]
} else{
  no_bad_accessions = 0 
}
```

This report uses data from **`r collection`** and analyses the collection from **`r min_year`** to **`r year_cur`**.

 A retrospective overview uses a combination of the accession year and item status date to calculate which plants are existing each year. This approach assumes that plants only die when their record is updated to reflect this. 

The collection contains **`r no_bad_accessions`** items that have either a missing/incorrect accession year, since we cannot accurately describe when these plants are existing we remove these records for the following analysis. These records can be found in the file **`r collection`_issue_accession_year.csv**.

This document outlines the changes in the number of accessions, items, taxa, species, families, genera, provenance, endemic species and threatened species. 

<hr style="border:4px solid gray">

## New Accessions

In this section we look at how the composition of new accessions varies. 

```{r, new accessions, echo = F}
years = min_year:year_cur
#########
# Reduce data frame to only accessions.
#########
# Remove entries without an AccYear.
accessions = report[!is.na(report$AccYear),]
# Reduce report to only the earliest year of accession.
accessions = accessions[order(accessions$AccYear),]
unique_accessions = unique(accessions$AccNoFull)
accessions = report[match(unique_accessions, accessions$AccNoFull),]  

# For each year strip the quantities of interest into a list.
new_accessions = lapply(years, function(year){
  # Accessions for the year
  garden_current = accessions[accessions$AccYear == year,]
  
   # Number of families.
  families = data.frame(original = garden_current$Family, POWO = garden_current$POWO_family)
  use_family = families$POWO
  use_family[is.na(use_family)] = families$original[is.na(use_family)]
  family = unique(use_family)
  no_families <- length(family)

  # Number of genera.
  genus = data.frame(original = garden_current$Genus, POWO = garden_current$POWO_genus)
  use_genus = genus$POWO
  use_genus[is.na(use_genus)] = genus$original[is.na(use_genus)]
  no_genus <- length(unique(use_genus))

  # Number of Accessions.
  no_accessions = nrow(garden_current)
  
  #No species (#remove cultivars, indeterminants, hybrids and those without infrageneric level)
  POWO_GenusSpecies = rep(NA, nrow(garden_current))
  POWO_GenusSpecies[!is.na(garden_current$POWO_plant_name_id)] = paste0(garden_current$POWO_genus[!is.na(garden_current$POWO_plant_name_id)], ' ', garden_current$POWO_species[!is.na(garden_current$POWO_plant_name_id)])
  species_data = data.frame(original = garden_current$GenusSpecies, POWO = POWO_GenusSpecies, infra = garden_current$infrageneric_level)
  species_only = species_data[!grepl('5|6|0',species_data$infra),]
  species_only = species_only[!is.na(species_only$infra),]
  use_species = species_only$POWO
  use_species[is.na(use_species)] = species_only$original[is.na(use_species)]
  no_species = length(unique(use_species))
  
  # Breakdown of endemic.
  endemic = table(garden_current$endemic)
  
  # Number of endemic (local)
  if(do_coord){
    no_endemic_local = sum(garden_current$POWO_Dist_area_code_l3[garden_current$endemic == 'Endemic'] == level3_code, na.rm = T)
  }else{
    no_endemic_local = NA
  }
  breakdown_endemic = table(garden_current$endemic)
  
  # Breakdown of provenance.
  provenance = table(garden_current$ProvenanceCode)
  
  # Breakdown of threatened.
  breakdown_threatened = table(garden_current$threatened)
  
  # Breakdown of threatened categories
  breakdown_threatened_cateogory = table(garden_current$redList_category)
  breakdown_threatened_cateogory = breakdown_threatened_cateogory[match(c('VU','EN','CR', 'EW', 'EX'), names(breakdown_threatened_cateogory))]
  
  # Breakdown of infrageneric diversity.
  breakdown_infrageneric = table(garden_current$infrageneric_level)
  
  
  return(list(no_accessions = no_accessions, no_families = no_families, no_genus = no_genus, no_species = no_species,
              breakdown_endemic = endemic, no_endemic_local = no_endemic_local,
              breakdown_provenance = provenance,
              breakdown_threatened = breakdown_threatened, breakdown_threatened_cateogory = breakdown_threatened_cateogory,
              breakdown_infrageneric = breakdown_infrageneric))
})
names(new_accessions) = paste0('year-', years)

time_series_info = new_accessions
```

#### Number of Accessions each Year

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 4)}
no_accessions = unlist(lapply(time_series_info, function(x){x$no_accessions}))
accessions_data = data.frame(year = years, no_accessions = no_accessions)
accessions_data$text = paste0('Date: ',years, ' <br>', 'Accessions: ', accessions_data$no_accessions)


fig <- plot_ly(accessions_data, x = ~year, y = ~no_accessions, type = 'bar', text = ~text, hoverinfo = 'text')
fig <- fig %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Accessions")) 
fig <- fig %>% layout(hovermode = 'x')

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/Bar_no_accessions.html'),selfcontained = T)
}

fig
```

***

#### Number of Species Accessed Each Year

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 4)}
no_species = unlist(lapply(time_series_info, function(x){x$no_species}))
species_data = data.frame(year = years, no_species = no_species)
species_data$text = paste0('Date: ',years, ' <br>', 'Species: ', species_data$no_species)

fig <- plot_ly(species_data, x = ~year, y = ~no_species, type = 'bar', text = ~text, hoverinfo = 'text')
fig <- fig %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Species")) 
fig <- fig %>% layout(hovermode = 'x')

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/Bar_no_species.html'),selfcontained = T)
}

fig
```

We ignore records that are indeterminants, cultivars and hybrids. We then use `GenusSpecies` from matched POWO records where available, otherwise we use `GenusSpecies` from the original report. Note that here we condense subspecies, varieties and forma into their corresponding species. For example `Tephrosia longipes var. drummondii` would be condensed to the species `Tephrosia longipes`. 

***
#### Number of Families Accessed Each Year

```{r, echo = F, fig.dim = c(10, 4)}
families = unlist(lapply(time_series_info, function(x){x$no_families}))
families_data = data.frame(year = years, no_families = families)
families_data$text = paste0('Date: ',names(families), ' <br>', 'Families: ', families_data$no_families)

fig <- plot_ly(families_data, x = ~year, y = ~no_families, type = 'bar', text = ~text, hoverinfo = 'text')
fig <- fig %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Families")) 
fig <- fig %>% layout(hovermode = 'x')

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/Bar_no_families.html'),selfcontained = T)
}

fig
```

***

#### Number of Genera Accessed Each Year

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 4)}
no_genus = unlist(lapply(time_series_info, function(x){x$no_genus}))
genus_data = data.frame(year = years, no_genus = no_genus)
genus_data$text = paste0('Date: ',names(no_genus), ' <br>', 'Genus: ', genus_data$no_genus)

fig <- plot_ly(genus_data, x = ~year, y = ~no_genus, type = 'bar', text = ~text, hoverinfo = 'text')
fig <- fig %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Genera")) 
fig <- fig %>% layout(hovermode = 'x')

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/Bar_no_genera.html'),selfcontained = T)
}

fig
```

We use the genus taken from matched POWO record where available, otherwise we use the genus from the original report.

***


### Number of species, infraspecific, and cultivars accessed each year

We group the plants into:

- Indeterminant: consists of indeterminant plants.
- Species: consists of species, where subspecies, varieties and forma have not been included.
- Infraspecific: includes subspecies (subsp.), varieties (var.) and forma (f.).
- Horticultural: consists of hybrids and cultivars.


```{r, echo = F}
no_items = data.frame(t(data.frame(lapply(time_series_info, function(x){
  details = x$breakdown_infrageneric
  
  indet = sum(as.numeric(details[grepl('0',names(details))]))
  species = sum(as.numeric(details[grepl('1',names(details))]))
  infra = sum(as.numeric(details[grepl('2|3|4',names(details))]))
  hort = sum(as.numeric(details[grepl('5|6',names(details))]))
  
  return(c(indet, species, infra, hort))
  }))))
names(no_items) = c('Indeterminat', 'Species', 'Infraspecific', 'Horticultral')
prop = round(no_items/rowSums(no_items)*100,digits=2)
names(prop) =paste0(names(prop),'_prop')
infra_data = data.frame(year = years, no_items, prop)
infra_data$text = paste0('Date: ', names(time_series_info), ' <br>', 'Indeterminat: ', infra_data$indet,
                   ' <br>', 'Species: ', infra_data$species,
                   ' <br>', 'Infraspecific: ', infra_data$infra,
                   ' <br>', 'Horticultral: ', infra_data$hort)
```


#### Change in Total Number of Plants over time

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## Grouped species, hort etc RAW NUMBERS FOUR GRAPH OPTIONS.
infra_data1 = infra_data[,1:(length(no_items)+1)]
infra_columns =2:(length(no_items)+1)
colours = viridis::inferno(length(infra_columns))
infra_labels = names(infra_data1)[infra_columns]

fig1 <- plot_ly(infra_data1, x = ~year) 
for(i in 1:length(infra_columns)){
  fig1 <- fig1 %>% add_trace(y = infra_data1[,infra_columns[i]], name = infra_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Accessions")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(infra_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(infra_columns)){
  fig2 <- fig2 %>% add_trace(y = infra_data1[,infra_columns[i]], name = infra_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Accessions")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(infra_data1, x = ~year)
for(i in 1:length(infra_columns)){
  fig3 <- fig3 %>% add_trace(y = infra_data1[,infra_columns[i]],
                             name = infra_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Accessions"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(infra_data1, x = ~year)
for(i in 1:length(infra_columns)){
  fig4 <- fig4 %>% add_trace(y = infra_data1[,infra_columns[i]],
                             type='bar',
                             name = infra_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Accessions"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/number_species_infraspecific_and_cultivars.html'),selfcontained = T)
}

fig
```


#### Change in Proportion of Accessed Plants each year

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## Grouped species, hort etc RAW NUMBERS FOUR GRAPH OPTIONS.
infra_data1 = infra_data[,c(1, which(grepl('prop', names(infra_data))))]
infra_columns =2:(length(no_items)+1)
colours = viridis::inferno(length(infra_columns))
infra_labels = names(no_items)

fig1 <- plot_ly(infra_data1, x = ~year) 
for(i in 1:length(infra_columns)){
  fig1 <- fig1 %>% add_trace(y = infra_data1[,infra_columns[i]], name = infra_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Accessions")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(infra_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(infra_columns)){
  fig2 <- fig2 %>% add_trace(y = infra_data1[,infra_columns[i]], name = infra_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Accessions")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(infra_data1, x = ~year)
for(i in 1:length(infra_columns)){
  fig3 <- fig3 %>% add_trace(y = infra_data1[,infra_columns[i]],
                             name = infra_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Accessions"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(infra_data1, x = ~year)
for(i in 1:length(infra_columns)){
  fig4 <- fig4 %>% add_trace(y = infra_data1[,infra_columns[i]],
                             type='bar',
                             name = infra_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Accessions"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/proportion_species_infraspecific_and_cultivars.html'),selfcontained = T)
}

fig
```

<hr style="border:4px solid gray">

### Distribution and Number of Accessions by Provenance

```{r, echo = F}
#Get the unique Provenance codes for the collection
unique_provenance_values = unique(report$ProvenanceCode)

# Get the number of plants with provenance code for each year.
no_items = data.frame(t(data.frame(lapply(time_series_info, function(x){
  details = x$breakdown_provenance
  
  no_prov = as.numeric(details[match(unique_provenance_values, names(details))])
  no_prov[is.na(no_prov)] = 0
  # garden = sum(as.numeric(details[grepl('G',names(details))]))
  # unknown = sum(as.numeric(details[grepl('U',names(details))]))
  # wild = sum(as.numeric(details[grepl('W',names(details))]))
  return(no_prov)
  # return(c(garden, unknown, wild))
  }))))
names(no_items) = unique_provenance_values
# Change the labels for known provenances.
names(no_items)[match(c('G','W','U'), names(no_items))[!is.na(match(c('G','W','U'), names(no_items)))]] = c('Garden', 'Wild', 'Unknown')[!is.na(match(c('G','W','U'), names(no_items)))]

prop = round(no_items/rowSums(no_items)*100,digits=2)
names(prop) =paste0(names(prop),'_prop')
prov_data = data.frame(year = years, no_items, prop)
```

To extract provenance information we use the data field `ProvenanceCode`.
The collection has provenance codes: **`r paste0(unique_provenance_values, collapse=', ')`**.

Where we have:

- G = Garden: plants of garden origin.
- W = Wild: plants of wild origin.
- U = Unknown: plants whose origin is unknown.

#### Total Number of Accessions each year by provenance

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## PROVENANCE RAW NUMBERS FOUR GRAPH OPTIONS.
prov_data1 = prov_data[,1:(length(no_items)+1)]
prov_columns =2:(length(no_items)+1)
colours = viridis::inferno(length(prov_columns))
prov_labels = names(prov_data1)[prov_columns]

fig1 <- plot_ly(prov_data1, x = ~year) 
for(i in 1:length(prov_columns)){
  fig1 <- fig1 %>% add_trace(y = prov_data1[,prov_columns[i]], name = prov_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(prov_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(prov_columns)){
  fig2 <- fig2 %>% add_trace(y = prov_data1[,prov_columns[i]], name = prov_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(prov_data1, x = ~year)
for(i in 1:length(prov_columns)){
  fig3 <- fig3 %>% add_trace(y = prov_data1[,prov_columns[i]],
                             name = prov_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(prov_data1, x = ~year)
for(i in 1:length(prov_columns)){
  fig4 <- fig4 %>% add_trace(y = prov_data1[,prov_columns[i]],
                             type='bar',
                             name = prov_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/number_provenance.html'),selfcontained = T)
}

fig
```


#### Proportion of Accessions each year by provenance

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## PROVENANCE Proportion FOUR GRAPH OPTIONS.
prov_data1 = prov_data[,c(1,which(grepl('prop',names(prov_data))))]
prov_columns = 2:(length(no_items)+1)
colours = viridis::inferno(length(prov_columns))
prov_labels = names(no_items)

fig1 <- plot_ly(prov_data1, x = ~year) 
for(i in 1:length(prov_columns)){
  fig1 <- fig1 %>% add_trace(y = prov_data1[,prov_columns[i]], name = prov_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(prov_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(prov_columns)){
  fig2 <- fig2 %>% add_trace(y = prov_data1[,prov_columns[i]], name = prov_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(prov_data1, x = ~year)
for(i in 1:length(prov_columns)){
  fig3 <- fig3 %>% add_trace(y = prov_data1[,prov_columns[i]],
                             name = prov_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(prov_data1, x = ~year)
for(i in 1:length(prov_columns)){
  fig4 <- fig4 %>% add_trace(y = prov_data1[,prov_columns[i]],
                             type='bar',
                             name = prov_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/proportion_provenance.html'),selfcontained = T)
}

fig
```

<hr style="border:4px solid gray">

### Distribution and Number of Accessions by Endemic Species

```{r, echo = F}
unique_no_endemic_values = unique(report$endemic)[!is.na(unique(report$endemic))]

# Get the number of plants with provenance code for each year.
no_items = data.frame(t(data.frame(lapply(time_series_info, function(x){
  details = x$breakdown_endemic
  
  no_endemic = as.numeric(details[match(unique_no_endemic_values, names(details))])
  no_endemic[is.na(no_endemic)] = 0

  return(no_endemic)
  }))))
names(no_items) = unique_no_endemic_values

prop = round(no_items/rowSums(no_items)*100,digits=2)
names(prop) =paste0(names(prop),'_prop')
endemic_data = data.frame(year = years, no_items, prop)
```


#### Total Number of Accessions each year by endemic status

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## PROVENANCE RAW NUMBERS FOUR GRAPH OPTIONS.
endemic_data1 = endemic_data[,1:(length(no_items)+1)]
endemic_columns =2:(length(no_items)+1)
colours = viridis::inferno(length(endemic_columns))
endemic_labels = names(endemic_data1)[endemic_columns]
endemic_labels = stringr::str_replace_all(endemic_labels, '\\.', ' ')

fig1 <- plot_ly(endemic_data1, x = ~year) 
for(i in 1:length(endemic_columns)){
  fig1 <- fig1 %>% add_trace(y = endemic_data1[,endemic_columns[i]], name = endemic_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(endemic_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(endemic_columns)){
  fig2 <- fig2 %>% add_trace(y = endemic_data1[,endemic_columns[i]], name = endemic_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(endemic_data1, x = ~year)
for(i in 1:length(endemic_columns)){
  fig3 <- fig3 %>% add_trace(y = endemic_data1[,endemic_columns[i]],
                             name = endemic_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(endemic_data1, x = ~year)
for(i in 1:length(endemic_columns)){
  fig4 <- fig4 %>% add_trace(y = endemic_data1[,endemic_columns[i]],
                             type='bar',
                             name = endemic_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/number_endemic.html'),selfcontained = T)
}

fig
```


#### Proportion of Accessions each year by endemic status

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## PROVENANCE Proportion FOUR GRAPH OPTIONS.
endemic_data1 = endemic_data[,c(1,which(grepl('prop',names(endemic_data))))]
endemic_columns = 2:(length(no_items)+1)
colours = viridis::inferno(length(endemic_columns))
endemic_labels = names(no_items)

fig1 <- plot_ly(endemic_data1, x = ~year) 
for(i in 1:length(endemic_columns)){
  fig1 <- fig1 %>% add_trace(y = endemic_data1[,endemic_columns[i]], name = endemic_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(endemic_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(endemic_columns)){
  fig2 <- fig2 %>% add_trace(y = endemic_data1[,endemic_columns[i]], name = endemic_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(endemic_data1, x = ~year)
for(i in 1:length(endemic_columns)){
  fig3 <- fig3 %>% add_trace(y = endemic_data1[,endemic_columns[i]],
                             name = endemic_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(endemic_data1, x = ~year)
for(i in 1:length(endemic_columns)){
  fig4 <- fig4 %>% add_trace(y = endemic_data1[,endemic_columns[i]],
                             type='bar',
                             name = endemic_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/proportion_endemic.html'),selfcontained = T)
}

fig
```


The collection is located in `r level3_name`. Below we show how the number of accessions endemic to `r level3_name` each year.

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 4)}
no_endemic_local = unlist(lapply(time_series_info, function(x){x$no_endemic_local}))
endemic_local_data = data.frame(year = as.numeric(stringr::str_extract(names(time_series_info),'[0-9]{4}')), no_endemic_local = no_endemic_local)
endemic_local_data$text = paste0('Date: ',names(time_series_info), ' <br>', 'Local Endemic: ', endemic_local_data$no_endemic_local)

fig <- plot_ly(endemic_local_data, x = ~year, y = ~no_endemic_local, type = 'bar', text = ~text, hoverinfo = 'text')
fig <- fig %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items")) 
fig <- fig %>% layout(hovermode = 'x')

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/Bar_endemic_local.html'),selfcontained = T)
}

fig
```


<hr style="border:4px solid gray">

### Distribution and Number of Accessions by Threatened Species

```{r, echo = F}
unique_no_threatened_values = unique(report$threatened)[!is.na(unique(report$threatened))]

# Get the number of plants with provenance code for each year.
no_items = data.frame(t(data.frame(lapply(time_series_info, function(x){
  details = x$breakdown_threatened
  
  no_threatened = as.numeric(details[match(unique_no_threatened_values, names(details))])
  no_threatened[is.na(no_threatened)] = 0

  return(no_threatened)
  }))))
names(no_items) = unique_no_threatened_values

prop = round(no_items/rowSums(no_items)*100,digits=2)
names(prop) =paste0(names(prop),'_prop')
threatened_data = data.frame(year = years, no_items, prop)
```

#### Total Number of Accessions each year by threatened status

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## PROVENANCE RAW NUMBERS FOUR GRAPH OPTIONS.
threatened_data1 = threatened_data[,1:(length(no_items)+1)]
threatened_columns =2:(length(no_items)+1)
colours = viridis::inferno(length(threatened_columns))
threatened_labels = names(threatened_data1)[threatened_columns]
threatened_labels = stringr::str_replace_all(threatened_labels, '\\.', ' ')

fig1 <- plot_ly(threatened_data1, x = ~year) 
for(i in 1:length(threatened_columns)){
  fig1 <- fig1 %>% add_trace(y = threatened_data1[,threatened_columns[i]], name = threatened_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(threatened_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(threatened_columns)){
  fig2 <- fig2 %>% add_trace(y = threatened_data1[,threatened_columns[i]], name = threatened_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(threatened_data1, x = ~year)
for(i in 1:length(threatened_columns)){
  fig3 <- fig3 %>% add_trace(y = threatened_data1[,threatened_columns[i]],
                             name = threatened_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(threatened_data1, x = ~year)
for(i in 1:length(threatened_columns)){
  fig4 <- fig4 %>% add_trace(y = threatened_data1[,threatened_columns[i]],
                             type='bar',
                             name = threatened_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/number_threatened.html'),selfcontained = T)
}

fig
```


#### Proportion of Accessions each year by threatened status

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## PROVENANCE Proportion FOUR GRAPH OPTIONS.
threatened_data1 = threatened_data[,c(1,which(grepl('prop',names(threatened_data))))]
threatened_columns = 2:(length(no_items)+1)
colours = viridis::inferno(length(threatened_columns))
threatened_labels = names(no_items)

fig1 <- plot_ly(threatened_data1, x = ~year) 
for(i in 1:length(threatened_columns)){
  fig1 <- fig1 %>% add_trace(y = threatened_data1[,threatened_columns[i]], name = threatened_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(threatened_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(threatened_columns)){
  fig2 <- fig2 %>% add_trace(y = threatened_data1[,threatened_columns[i]], name = threatened_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(threatened_data1, x = ~year)
for(i in 1:length(threatened_columns)){
  fig3 <- fig3 %>% add_trace(y = threatened_data1[,threatened_columns[i]],
                             name = threatened_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(threatened_data1, x = ~year)
for(i in 1:length(threatened_columns)){
  fig4 <- fig4 %>% add_trace(y = threatened_data1[,threatened_columns[i]],
                             type='bar',
                             name = threatened_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/proportion_threatened.html'),selfcontained = T)
}

fig
```


Concentrating only on the threatened species the following figures outline the categories of the threatened taxa over the history of the collection. 

```{r, echo = F}
unique_no_threatened_cat_values =  c('VU','EN','CR', 'EW', 'EX')

# Get the number of plants with provenance code for each year.
no_items = data.frame(t(data.frame(lapply(time_series_info, function(x){
  details = x$breakdown_threatened_cateogory
  
  no_threatened = as.numeric(details[match(unique_no_threatened_cat_values, names(details))])
  no_threatened[is.na(no_threatened)] = 0

  return(no_threatened)
  }))))
names(no_items) = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')

prop = round(no_items/rowSums(no_items)*100,digits=2)
names(prop) =paste0(names(prop),'_prop')
threatened_cat_data = data.frame(year = years, no_items, prop)
```

#### Total Number of Accessions each year by threatened category

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## PROVENANCE RAW NUMBERS FOUR GRAPH OPTIONS.
threatened_data1 = threatened_cat_data[,1:(length(no_items)+1)]
threatened_columns =2:(length(no_items)+1)
colours = viridis::inferno(length(threatened_columns))
threatened_labels = names(threatened_cat_data)[threatened_columns]
threatened_labels = stringr::str_replace_all(threatened_labels, '\\.', ' ')

fig1 <- plot_ly(threatened_data1, x = ~year) 
for(i in 1:length(threatened_columns)){
  fig1 <- fig1 %>% add_trace(y = threatened_data1[,threatened_columns[i]], name = threatened_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(threatened_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(threatened_columns)){
  fig2 <- fig2 %>% add_trace(y = threatened_data1[,threatened_columns[i]], name = threatened_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(threatened_data1, x = ~year)
for(i in 1:length(threatened_columns)){
  fig3 <- fig3 %>% add_trace(y = threatened_data1[,threatened_columns[i]],
                             name = threatened_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(threatened_data1, x = ~year)
for(i in 1:length(threatened_columns)){
  fig4 <- fig4 %>% add_trace(y = threatened_data1[,threatened_columns[i]],
                             type='bar',
                             name = threatened_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Number of Items"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/number_threatened_category.html'),selfcontained = T)
}

fig
```


#### Proportion of Accessions each year by threatened category

```{r, echo = F, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
## PROVENANCE Proportion FOUR GRAPH OPTIONS.
threatened_data1 = threatened_cat_data[,c(1,which(grepl('prop',names(threatened_cat_data))))]
threatened_columns = 2:(length(no_items)+1)
colours = viridis::inferno(length(threatened_columns))
threatened_labels = names(no_items)

fig1 <- plot_ly(threatened_data1, x = ~year) 
for(i in 1:length(threatened_columns)){
  fig1 <- fig1 %>% add_trace(y = threatened_data1[,threatened_columns[i]], name = threatened_labels[i], mode = 'lines',type = 'scatter', line = list(color = colours[i])) 
}
fig1 <- fig1 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items")) 
fig1 <- fig1 %>% layout(hovermode = 'x unified')


#Stacked Line graph
fig2 <- plot_ly(threatened_data1, x = ~year, mode = 'none', stackgroup = 'one')
for(i in 1:length(threatened_columns)){
  fig2 <- fig2 %>% add_trace(y = threatened_data1[,threatened_columns[i]], name = threatened_labels[i],type = 'scatter', fillcolor = colours[i]) 
}
fig2 <- fig2 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items")) 
fig2 <- fig2 %>% layout(hovermode = 'x unified')

# Bar graph
fig3 <- plot_ly(threatened_data1, x = ~year)
for(i in 1:length(threatened_columns)){
  fig3 <- fig3 %>% add_trace(y = threatened_data1[,threatened_columns[i]],
                             name = threatened_labels[i],
                             marker = list(color = colours[i],line = list(color = 'rgb(8,48,107)', width = 0)),
                             type = 'bar')
}
fig3 <- fig3 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items"),
          barmode = 'group')
fig3 <- fig3 %>% layout(hovermode = 'x unified')

# Stacked bar graph
fig4 <- plot_ly(threatened_data1, x = ~year)
for(i in 1:length(threatened_columns)){
  fig4 <- fig4 %>% add_trace(y = threatened_data1[,threatened_columns[i]],
                             type='bar',
                             name = threatened_labels[i],
                             marker = list(color = colours[i], line = list(color = colours[i], width = 0)))
}
fig4 <- fig4 %>% layout(title = "",
         xaxis = list(title = "Year"),
         yaxis = list (title = "Proportion of Items"),
          barmode = 'stack')
fig4 <- fig4 %>% layout(hovermode = 'x unified', bargap =0)

fig = manipulateWidget::combineWidgets(fig1, fig2, fig3, fig4, nrow = 2)

if(save_widgets){
  htmlwidgets::saveWidget(fig, file = paste0(new_accessions_widget_dir,'/proportion_threatened_category.html'),selfcontained = T)
}

fig
```



<hr style="border:4px solid gray">


```{r,echo=F}
# Return excel workbook with data for each figure and table

library(xlsx)
wb = xlsx::createWorkbook()
sh1 = xlsx::createSheet(wb, 'Accessions')
xlsx::addDataFrame(accessions_data[,-3], sh1, startRow = 1,row.names = F)
sh1A = xlsx::createSheet(wb, 'Families')
xlsx::addDataFrame(families_data[,-3], sh1A, startRow = 1,row.names = F)
sh1B = xlsx::createSheet(wb, 'Genera')
xlsx::addDataFrame(genus_data[,-3], sh1B, startRow = 1,row.names = F)
sh1C = xlsx::createSheet(wb, 'Species')
xlsx::addDataFrame(species_data[,-3], sh1C, startRow = 1,row.names = F)
sh4 = xlsx::createSheet(wb, 'Species, infraspecific, and cultivars')
xlsx::addDataFrame(infra_data[,-ncol(infra_data)], sh4, startRow = 1,row.names = F)
sh5 = xlsx::createSheet(wb, 'Provanence')
xlsx::addDataFrame(prov_data, sh5, startRow = 1,row.names = F)
sh6 = xlsx::createSheet(wb, 'Endemic')
xlsx::addDataFrame(endemic_data, sh6, startRow = 1,row.names = F)
if(do_coord){
  sh6B = xlsx::createSheet(wb, paste0('Endemic to ',level3_name))
  xlsx::addDataFrame(endemic_local_data[,-ncol(endemic_local_data)], sh6B, startRow = 1,row.names = F)  
}
sh7 = xlsx::createSheet(wb, 'Threatened')
xlsx::addDataFrame(threatened_data, sh7, startRow = 1,row.names = F)
sh8 = xlsx::createSheet(wb, 'Red list category')
xlsx::addDataFrame(threatened_cat_data, sh8, startRow = 1,row.names = F)
xlsx::saveWorkbook(wb, file = paste0(output_dir,'/',collection,'_data_for_figures_reterospective_new_accessions.xlsx'))
```


