---
title: "Threatened species in the living collection"
output: html_document
always_allow_html: true
---
  
###### {-} 
<!-- This section is for loading/setting up the data and writing the intro -->
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(htmltools)
library(sf)
library(here)
library(flextable)
library(ggrepel)
library(downloadthis)
library(VennDiagram)

interactive_colour <- function(n){
  color = c('#f46d43', "#FDAE6B", '#848484', '#e6e6e6')
  if(n==2){
    return(color[c(1,4)])
  }
  color[1:n]
}
color_binary <- c('#f46d43', '#e6e6e6')
palette = 'Oranges'
```

```{css, echo = FALSE}
/* from https://ianlunn.github.io/Hover/ */
.hvr-sweep-to-left {
  display: inline-block;
  vertical-align: middle;
  -webkit-transform: perspective(1px) translateZ(0);
  transform: perspective(1px) translateZ(0);
  box-shadow: 0 0 1px rgba(0, 0, 0, 0);
  position: relative;
  -webkit-transition-property: color;
  transition-property: color;
  -webkit-transition-duration: 0.3s;
  transition-duration: 0.3s;
}
.hvr-sweep-to-left:before {
  content: "";
  position: absolute;
  z-index: -1;
  top: 0;
  left: 0;
  right: 0;
  bottom: 0;
  background: #f46d43;
  -webkit-transform: scaleX(0);
  transform: scaleX(0);
  -webkit-transform-origin: 100% 50%;
  transform-origin: 100% 50%;
  -webkit-transition-property: transform;
  transition-property: transform;
  -webkit-transition-duration: 0.3s;
  transition-duration: 0.3s;
  -webkit-transition-timing-function: ease-out;
  transition-timing-function: ease-out;
}
.hvr-sweep-to-left:hover, .hvr-sweep-to-left:focus, .hvr-sweep-to-left:active {
  color: white;
}
.hvr-sweep-to-left:hover:before, .hvr-sweep-to-left:focus:before, .hvr-sweep-to-left:active:before {
  -webkit-transform: scaleX(1);
  transform: scaleX(1);
}

.blackbox {
  padding: 1em;
  background: #FDAE6B;
  border: 2px solid #e6e6e6;
  border-radius: 10px;
}
.center {
  text-align: center;
}
```

```{r, imported parameters}
# 
# collection = 'CUBG'
# load('/Users/jakepowell/Cambridge/Enriched_reports_Jake/CUBG_June2023_enriched_report.rda')
# coordinates = c(52.19378853629289,0.1277234065606053)
# #
# #
# load('/Users/jakepowell/Cambridge/Geography Module/wgsrpd3.rda')
# load('/Users/jakepowell/Cambridge/WCVP/Version 11/wcvp_with_redlistcategory.rda')
# 
# native = 'Naturally occurring only'
# extinct = TRUE
# doubtful_locations = FALSE
# min_year = 1970
# export_data = F
# table_font_size = 14
# ggtheme = NULL
# separate_figure_folder = FALSE
# value_on_fig = TRUE
# report_kind = 'static'
# color_binary = c('darkgray','darkgreen')
# palette = 'Greens'
# endemic_species_per_region = NULL
# output_dir =getwd()
```

```{r, add logo, eval = report_kind == 'interactive'}
htmltools::img(src = knitr::image_uri(paste0(system.file(package = "BGSmartR"), "/logo.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width:87.8px; height:100px')
```

```{r, functions for turnover}
turnover_items <- function(report){
  report = report[(report$LossYear >= min_year & report$LossYear <= max_year) | is.na(report$LossYear),]
  combined = data.frame(Year = years)
  
  # Items
  gain = data.frame(table(report$AccYear))
  if(nrow(gain) == 0){
    gained = rep(0,nrow(combined))
    combined$gain_items = gained
  }else{
    names(gain) = c('Year', 'Items')
    gain$Year = as.numeric(as.character(gain$Year))
    gain = gain[gain$Year >=min_year & gain$Year <= max_year,]
    gained = rep(0,nrow(combined))
    gained[match(gain$Year, combined$Year)] = gain$Items
    combined$gain_items = gained
  }
  
  
  loss = data.frame(table(report$LossYear))
  if(nrow(loss) == 0){
    lost = rep(0,nrow(combined))
    combined$loss_items = lost
  }else{
    names(loss) = c('Year', 'Items')
    loss$Year = as.numeric(as.character(loss$Year))
    loss = loss[loss$Year >=min_year & loss$Year <= max_year,]
    lost = rep(0,nrow(combined))
    lost[match(loss$Year, combined$Year)] = loss$Items
    combined$loss_items = lost
  }
  
  
  combined$net_items = combined$gain_items - combined$loss_items
  return(combined)
}

plots_turnover <- function(turnover, trend, report_kind, text = '', separate_figure_folder, data_type = 'items'){
    dataA = data.frame(turnover[,c(1,3)], rep('Lost', nrow(turnover)))
    names(dataA) = LETTERS[1:3]
    dataB = data.frame(turnover[,c(1,2)], rep('Gain', nrow(turnover)))
    names(dataB) = LETTERS[1:3]
    items_data = rbind(dataA, dataB)
  if(report_kind == 'static'){

    p = ggplot(data=items_data, aes(x=A, y=B, group=C)) +
      geom_line(aes(color=C)) +
      # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
      labs(title=paste0("Turnover of ",text," in the LC (By ",data_type,")"),
           x ="Year",
           y = paste0('Number of ',data_type,'')) +
      guides(color=guide_legend(title=paste0('Gain/Loss'))) +
      # theme(text = element_text(size=table_font_size)) +
      scale_color_manual(values=c("blue", "red"))+
      theme(legend.direction = "horizontal", legend.position = "top", legend.justification = "right")
    
    
    if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','turnover_',stringr::str_replace_all(text, ' ', '_'),'_gain_loss_',data_type,'.pdf'),device = 'pdf', scale = 1, width = 20, height = 12, limitsize = FALSE)
    }
    p1=p
    
    # Color based on value
    color <- ifelse(turnover$net_items < 0, "pink", "lightblue")
    p = ggplot(turnover, aes(x = Year, y = net_items)) +
      geom_bar(stat = "identity",
               show.legend = FALSE,
               fill = color,      # Background color
               color = "white") + # Border color
      # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
      labs(title=paste0("Net turnover of ",text," in the LC (By ",data_type,")"),
           x ="Year",
           y = paste0('Number of ',data_type,''))
    # theme(text = element_text(size=table_font_size))
    
    if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Net_turnover_',stringr::str_replace_all(text, ' ', '_'),'_',data_type,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
    p2=p
    
    
    # Trend plot.
    trend_data = data.frame(year = turnover$Year, trend = trend)
    p = ggplot(trend_data, aes(x = year, y = trend)) +
      geom_line(width = 1.1) +
      labs(title="",
           x ="Year",
           y = paste0('Number of ',data_type,''))
    p3 = p
    return(list(gain_loss = p1, net = p2, trend = p3))
  }
  if(report_kind == 'interactive'){
    # gain_loss plot
    colors = color_binary |> rev()
    fig <- plot_ly(items_data, x = ~A, y = ~B, color = ~C, colors = colors, type = 'scatter', mode = 'lines+markers')
    fig <- fig |> layout(yaxis = list(title = paste0("Number of ",data_type |> tolower(),"")),
                         xaxis = list(title = "Year"))
    fig <- fig |> layout(hovermode = 'x unified')
    
    if(separate_figure_folder){
      htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','line-turnover_',stringr::str_replace_all(text, ' ', '_'),'_gain_loss_',data_type,'.html'))
    }
    fig1 = fig
    
    # net plot
    colors = color_binary |> rev()
    color <- ifelse(turnover$net_items < 0, colors[1], colors[2])
    texto = paste0('Net: ', turnover$net_items, '<br>',
                   'Gain: ', turnover$gain_items, '<br>',
                   'Loss: ', turnover$loss_items, '<br>'
    )
    fig = plot_ly(turnover, x = ~Year, y = ~net_items, type = 'bar', hoverinfo = "x+text",  hovertext = texto, color = color, colors = rev(colors), showlegend = FALSE)
    fig <- fig |> layout(yaxis = list(title = paste0("Net number of ",data_type |> tolower(),"")),
                         xaxis = list(title = "Year"))
    fig <- fig |> layout(hovermode = 'x unified')
    
    if(separate_figure_folder){
      htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Bar-Net_',stringr::str_replace_all(text, ' ', '_'),'turnover_of_',data_type,'.html'))
    }
    fig2 = fig
    
    # Trend plot.
    trend_data = data.frame(year = turnover$Year, trend = trend)
    fig = plot_ly(trend_data, x = ~year, y = ~trend, type = 'scatter', mode = 'line', hoverinfo = "x+text", hovertext = trend)
    fig <- fig |> layout(yaxis = list(title = paste0("Number of ",data_type |> tolower(),"")),
                         xaxis = list(title = "Year"))
    fig <- fig |> layout(hovermode = 'x unified')
    fig3= fig
    
    # Combined trend and net.
    fig <- plot_ly()
    # Add traces
    ay <- list(
      tickfont = list(color = "black"),
      overlaying = "y",
      side = "right",
      title = paste0("Total ", tolower(data_type)))
    
      fig <- fig %>% add_trace(x = years, y = turnover$net_items, name = "Net", type = 'bar', hoverinfo = "x+text",  hovertext = texto, color = color, colors = rev(colors), showlegend = FALSE)
     
    fig <- fig %>% add_trace(x = years, y = trend, name = "Total", mode = "lines+markers", type = "scatter", line = list(color = 'black'), marker = list(color = 'black'),yaxis = "y2")
    
  
    # Set figure title, x and y-axes titles
    fig <- fig |> layout(
      title = "", yaxis2 = ay,
      xaxis = list(title="Year"),
      yaxis = list(title = paste0("Net turnover of ", tolower(data_type)))
    ) |>
      layout(plot_bgcolor='white',
             xaxis = list(
               zerolinecolor = '#ffff',
               zerolinewidth = 2,
               gridcolor = 'ffff'),
             yaxis2 = list(
               zeroline = TRUE,
               zerolinecolor = 'ffff',
               zerolinewidth = 2,
               gridcolor = 'ffff',
               showgrid = FALSE),
              yaxis = list(
                zeroline = TRUE,
                zerolinecolor = '#f0f0f0',
                zerolinewidth = 1,
                gridcolor = '#f0f0f0',
                griddash = 'solid',
                gridwidth = 1,showgrid = T),
             margin = list(t = 10, l = 20, r = 70, b = 20, pad = 4),
             autosize = T
      )
    fig <- fig |> layout(hovermode = 'x unified')
    fig4 = fig
    
    return(list(gain_loss = fig1, net = fig2, trend = fig3, trend_and_net = fig4))
    
  }
}

proportional_plots_turnover <- function(turnover,
                                        overall_turnover,
                                        collection_proportion,
                                        separate_figure_folder, 
                                        report_kind,
                                        text = '',
                                        data_type = 'items',
                                        quantity = ''){
  if(report_kind == 'static'){
    # Gain
    plot_data = data.frame(year = turnover[,1],
                           specific = turnover[,2], 
                           all = overall_turnover[,2])
    plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
    plot_data$prop_specific_collection = collection_proportion*100
    plot_data_gain = plot_data
    p = ggplot(plot_data, aes(y=prop_specific, x=year)) + 
      geom_bar(stat="identity", width = 1, fill = 'lightblue', col = 'black') +
      geom_point(aes(x = year, y = prop_specific_collection))+
      labs(title=paste0("Proportion of gained ",data_type," that are ",text," in the LC"),
           x ="Year",
           y = "Percentage (%)") +
      # theme(text = element_text(size=table_font_size)) +
      labs(caption = paste0("Black points are proportion of ",data_type," in the LC that are ",text," each year."))
    p1 = p
    
    if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/',stringr::str_replace_all(text,' ','_'),'_gain_',data_type,'_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
    
    ### Plot of proportion of lost items being native.
    
    plot_data = data.frame(year =  turnover[,1],
                           specific = turnover[,3], 
                           all = overall_turnover[,3])
    plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
    plot_data$prop_specific_collection = collection_proportion*100
    plot_data_loss = plot_data
    
    p = ggplot(plot_data, aes(y=prop_specific, x=year)) + 
      geom_bar(stat="identity", width = 1, fill = 'pink', col = 'black') +
      geom_point(aes(x = year, y = prop_specific_collection))+
      # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
      labs(title=paste0("Proportion of lost ",data_type," that are ",text," in the LC"),
           x ="Year",
           y = "Percentage (%)") +
      # theme(text = element_text(size=table_font_size)) +
      labs(caption = paste0("Black points are proportion of ",data_type," in the LC that are ",text," each year."))
    
    if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/',stringr::str_replace_all(text,' ','_'),'_loss_',data_type,'_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
    
    return(list(gain = p1, loss = p, data_gain = plot_data_gain, data_loss = plot_data_loss))
  }
  if(report_kind =='interactive'){
    # Gain
    plot_data = data.frame(year = turnover[,1],
                           specific = turnover[,2], 
                           all = overall_turnover[,2])
    plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
    plot_data$prop_specific_collection = collection_proportion*100
    plot_data_gain = plot_data
    fig = plot_ly(plot_data, x = ~year, y = ~prop_specific, type = 'bar',
                  name = paste0('New accessions'),
                  marker = list(color = color_binary[2]),
                  hovertemplate = paste("%{y:.2f}% of new accessions <extra></extra>")
                  )
                  
    fig <- fig |> layout(yaxis = list(title = paste0("Percentage")),
                         xaxis = list(title = "Year"))
    fig <- fig |> add_trace(x = ~year, y = ~prop_specific_collection, type = 'scatter', mode = 'lines+markers',
                            name = paste0('Collection'),
                            marker = list(color = 'black'),
                            line = list(color = 'black'),
                             hovertemplate = paste("%{y:.2f}% of the collection <extra></extra>")
                            )
    fig <- fig |> layout(hovermode = 'x unified',
                         yaxis = list(ticksuffix = '%'),
                         xaxis = list(hoverformat = paste0('fvf')),
                         legend=list(title=list(text=paste0('<b>',text, '</b>'))))
    
    if(separate_figure_folder){
      htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_gained_',stringr::str_replace_all(text, ' ', '_'),'_',data_type,'.html'))
    }
    
    fig1 = fig
    
    ### Plot of proportion of lost items being native.
    
    plot_data = data.frame(year =  turnover[,1],
                           specific = turnover[,3], 
                           all = overall_turnover[,3])
    plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
    plot_data$prop_specific_collection = collection_proportion*100
    plot_data_loss = plot_data
    
    fig = plot_ly(plot_data, x = ~year,
                  y = ~prop_specific,
                  type = 'bar',
                  name = paste0('Lost accessions'),
                  marker = list(color = color_binary[1]),
                  hovertemplate = paste("%{y:.2f}% of lost accessions <extra></extra>"))
    fig <- fig |> layout(yaxis = list(title = paste0("Percentage")),
                         xaxis = list(title = "Year"))
    fig <- fig |> add_trace(x = ~year,
                            y = ~prop_specific_collection,
                            type = 'scatter',
                            mode = 'lines+markers',
                            name = paste0('Collection'),
                            marker = list(color = 'black'),
                            line = list(color = 'black'),
                            hovertemplate = paste("%{y:.2f}% of the collection<extra></extra>"))
    fig <- fig |> layout(hovermode = 'x unified',
                         yaxis = list(ticksuffix = '%'),
                         legend=list(title=list(text=paste0('<b>',text, '</b>'))))
    
    if(separate_figure_folder){
      htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_lost_',stringr::str_replace_all(text, ' ', '_'),'_',data_type,'.html'))
    }
    
    return(list(gain = fig1, loss = fig,  data_gain = plot_data_gain, data_loss = plot_data_loss))
    
  }
}
```

```{r, region names}

level_2 = c("10, 11, 12, 13, 14, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 40, 41, 42, 43, 50, 51, 60, 61, 62, 63, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 90, 91, Northern Europe, Middle Europe, Southwestern Europe, Southeastern Europe, Eastern Europe, Northern Africa, Macaronesia, West Tropical Africa, West-Central Tropical Africa, Northeast Tropical Africa, East Tropical Africa, South Tropical Africa, Southern Africa, Middle Atlantic Ocean, Western Indian Ocean, Siberia, Russian Far East, Middle Asia, Caucasus, Western Asia, Arabian Peninsula, China, Mongolia, Eastern Asia, Indian Subcontinent, Indo-China, Malesia, Papuasia, Australia, New Zealand, Southwestern Pacific, South-Central Pacific, Northwestern Pacific, North-Central Pacific, Subarctic America, Western Canada, Eastern Canada, Northwestern U.S.A., North-Central U.S.A., Northeastern U.S.A., Southwestern U.S.A., South-Central U.S.A., Southeastern U.S.A., Mexico, Central America, Caribbean, Northern South America, Western South America, Brazil, Southern South America, Subantarctic Islands, Antarctic Continent")
level_1 = c("1, 2, 3, 4, 5, 6, 7, 8, 9, Europe, Africa, Asia-Temperate, Asia-Tropical, Australasia, Pacific, Northern America, Southern America, Antarctic")

level_2 = data.frame(matrix(unlist(stringr::str_split(level_2, ', ')), ncol = 2))
names(level_2) = c('code', 'name') ; level_2$code = as.numeric(level_2$code)

level_1 = data.frame(matrix(unlist(stringr::str_split(level_1, ', ')), ncol = 2))
names(level_1) = c('code', 'name') ; level_1$code = as.numeric(level_1$code)

```

```{r theme_ggplot2}
library(ggplot2)

# Changing the default theme
if(is.null(ggtheme)){
  ggtheme <- function(base_size = 16) {
    ggplot2::theme_bw(base_size = base_size) %+replace%
      ggplot2::theme(
        plot.title = ggplot2::element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
        panel.grid.minor = ggplot2::element_blank(),
        panel.border = ggplot2::element_blank(),
        axis.title = ggplot2::element_text(size = rel(0.85), face = "bold"),
        axis.text = ggplot2::element_text(size = rel(0.70), face = "bold"),
        axis.line = ggplot2::element_line(color = "black"),
        legend.title = ggplot2::element_text(size = rel(0.85), face = "bold"),
        legend.text = ggplot2::element_text(size = rel(0.70), face = "bold"),
        legend.key = ggplot2::element_rect(fill = "transparent", colour = NA),
        legend.key.size = ggplot2::unit(1.5, "lines"),
        legend.background = ggplot2::element_rect(fill = "transparent", colour = NA),
        strip.background = ggplot2::element_rect(fill = "#17252D", color = "#17252D"),
        strip.text = ggplot2::element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
      )
  }
}
theme_set(ggtheme())
update_geom_defaults("line", list(size = 2))

# Set the colour scale for the plots
scale_colour_continuous <- scale_colour_continuous
scale_colour_discrete   <- scale_colour_discrete
scale_colour_binned     <- scale_colour_binned
scale_fill_continuous <- scale_fill_continuous
scale_fill_discrete <- scale_fill_discrete
scale_fill_binned <- scale_fill_binned
```

```{r, Add combined geography column to enriched report}
geography_data = enriched_report[,grepl('_area_code_l3', names(enriched_report))]
want = c('000','010','001','011','100','110','101')

#Depending on the values of the options reduce `want` to only the satisfactory values.
##A) introduced / naturally occurring only.
if(native == 'Introduced only'){
  want = want[grepl('^1',want)]
}else if(native == 'Naturally occurring only'){
  want = want[grepl('^0',want)]
}
## B) Extinct
if(!extinct){
  want = want[!grepl('010|011|110',want)]
}
## C) Doubtful
if(!doubtful_locations){
  want = want[!grepl('001|011|101',want)]
}
geog_want_values = want
# Get the columns in wanted info that contain the geography information we want.
geography_data = geography_data[,grepl(paste0(want,collapse='|'), names(geography_data))]

if(length(want) > 1){
  level3codes =do.call("paste", c(geography_data, sep = ", "))
  level3codes = stringr::str_remove(level3codes, ', NA$')
  level3codes = stringr::str_remove(level3codes, 'NA, ')
}else{
  level3codes = geography_data
}
level3codes[level3codes == "NA"] = NA

enriched_report$geography_codes = level3codes
```

```{r, Get location_type_text}
location_type_text = ''
if(native == 'Naturally occurring only'){
  location_type_text = paste0(location_type_text, 'naturally occurring only,', collapse = ' ')  
}else if(native == 'Introduced only'){
  location_type_text = paste0(location_type_text, 'introduced only,', collapse = ' ')  
  
}else{
  location_type_text = paste0(location_type_text, 'Either naturally occurring or introduced,', collapse = ' ')  
}
if(extinct){
  location_type_text = paste0(location_type_text, ' allowing extinct regions and', collapse = ' ')  
}else{
  location_type_text = paste0(location_type_text, ' not including extinct regions and', collapse = ' ')  
}
if(doubtful_locations){
  location_type_text = paste0(location_type_text, ' allowing doubtful regions.', collapse = ' ')  
  
}else{
  location_type_text = paste0(location_type_text, ' not allowing doubtful regions.', collapse = ' ')  
}
```

`r if(any(! c('AccNoFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType', 'TaxonNameFull') %in% names(enriched_report))){"*** \n **Warning!**\n\n This enriched report contains missing 'original' columns needed to perform the analyses in this report. Default values have been created. In particular, the following were missing "}`

`r if(!'AccNoFull' %in% names(enriched_report)){"- AccNoFull: assume each item is its own accession.\n"}`
`r if(!'TaxonNameFull' %in% names(enriched_report)){"- TaxonNameFull: set TaxonNameFull to TaxonName.\n"}`
`r if(!'AccYear' %in% names(enriched_report)){"- AccYear: Cannot do analyses over time.\n"}`
`r if(!'ItemStatusDate' %in% names(enriched_report)){"- AccNoFull: Cannot do analyses over time.\n"}`
`r if(!'ItemStatusType' %in% names(enriched_report)){"- AccNoFull: Cannot do analyses over time.\n"}`

`r if(any(! c('AccNoFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType') %in% names(enriched_report))){"***"}`

```{r, Find LC location}
coord_contained = TRUE
# Extract the location of the collection.
pnts <- data.frame( x = coordinates[2], y = coordinates[1])

# create a points collection
pnts_sf <- do.call("st_sfc",c(lapply(1:nrow(pnts), 
                                     function(i) {sf::st_point(as.numeric(pnts[i, ]))}), list("crs" = 4326))) 

pnts_trans <- st_transform(pnts_sf, 2163) # apply transformation to pnts sf
tt1_trans <- st_transform(wgsrpd3, 2163)      # apply transformation to polygons sf

intersect = as.vector(st_intersects(tt1_trans, pnts_trans, sparse = FALSE))
collection_geog_details = wgsrpd3[which(intersect),-5]
location_name = collection_geog_details$LEVEL3_NAM
location_code = collection_geog_details$LEVEL3_COD

if(length(location_name) == 0){
  coord_contained = FALSE
  # Try and find the nearest polygon.
  collection_geog_details  =  wgsrpd3[which.min(as.vector(sf::st_distance(tt1_trans, pnts_trans))),-5]
  location_name = collection_geog_details$LEVEL3_NAM
  location_code = collection_geog_details$LEVEL3_COD
}
```


```{r, Extracting information from enriched report, echo = F}
# Add required columns if missing.
do_over_time = TRUE
needed_columns = c('AccNoFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType')
if(!'AccNoFull' %in% names(enriched_report)){
  enriched_report$AccNoFull = 1:nrow(enriched_report)
}
if(!'AccYear' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$AccYear = rep(0, nrow(enriched_report))
}
if(!'ItemStatusDate' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$ItemStatusDate = rep(NA, nrow(enriched_report))
}
if(!'ItemStatusType' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$ItemStatusType = rep('Existing', nrow(enriched_report))
}
if(all(is.na(enriched_report$ItemStatusDate))){
  do_over_time = FALSE
}
if(all(is.na(enriched_report$ItemStatusType))){
  do_over_time = FALSE
}
if(all(is.na(enriched_report$AccYear))){
  do_over_time = FALSE
}

#Extract endemic information from enriched_report
endemic = rep('Widespread', nrow(enriched_report))
endemic_index = which(stringr::str_length(enriched_report$geography_codes) == 3)
endemic[endemic_index] = 'Endemic'
enriched_report$endemic = endemic
rm(endemic)

# Extract threatened. 
threat_cat = c('VU','EN','CR','EW','EX')
threatened = rep('Not Threatened', nrow(enriched_report))
threatened[which(enriched_report$redList_category %in% threat_cat)] = 'Threatened'
threatened[which(enriched_report$POWO_Red_category %in% threat_cat)] = 'Threatened'
enriched_report$threatened = threatened
rm(threatened)

# Extract threatened category. 
threatened_category = rep(NA, nrow(enriched_report))
for(i in 1:length(threat_cat)){
  threatened_category[which(enriched_report$redList_category == threat_cat[i])] = threat_cat[i]
  threatened_category[which(enriched_report$POWO_Red_category == threat_cat[i])] = threat_cat[i]
}
enriched_report$threatened_category = threatened_category
rm(threatened_category)

# Convert Provenance code to text.
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'G'] = 'Garden'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'U'] = 'Unknown'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'W'] = 'Wild'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'Z'] = 'Wild-derived'

# Extract tree.
tree =rep('Not Tree', nrow(enriched_report))
tree[enriched_report$Enrich_is_tree] = 'Tree'
enriched_report$tree = tree
rm(tree)

# Add native column.
nativeo = rep('Non-native', nrow(enriched_report))
native_index = which(grepl(location_code, enriched_report$geography_codes))
nativeo[native_index] = 'Native'
enriched_report$native = nativeo
rm(nativeo)

LossYear = rep(NA,nrow(enriched_report))
ItemStatusYear = as.numeric(stringr::str_extract(enriched_report$ItemStatusDate,pattern = '[0-9]{4}'))
LossYear[enriched_report$ItemStatusType == 'NotExisting'] = ItemStatusYear[enriched_report$ItemStatusType == 'NotExisting']
enriched_report$LossYear = LossYear

# Create a sanitised taxonomic name.
enriched_report$good_name = paste0(enriched_report$sanitised_taxon, ' ', enriched_report$extracted_author)

best_name = paste0(enriched_report$POWO_taxon_name, ' ', enriched_report$POWO_taxon_authors)
best_name[which(best_name == 'NA NA')] = enriched_report$good_name[which(best_name == 'NA NA')]
enriched_report$best_name = best_name


if(!is.null(old_accession_year_codes)){
  has_old_accessions_code = which(enriched_report$AccYear %in% old_accession_year_codes)
  no_old_accessions = length(has_old_accessions_code)
  enriched_report$AccYear[has_old_accessions_code] = earliest_allowable_record
}
report = enriched_report
report_original = enriched_report

if(separate_figure_folder){
  figures_dir = paste0(output_dir,'/',collection,'_Figures')
  dir.create(figures_dir,showWarnings = FALSE)
}

exporting_data_store = list()
```

###### {-}

This report explores threatened species in **`r collection`**, and how the number of threatened species has varied over time.

A taxa in the collection is classed as **threatened** if it is classified as either extinct, extinct in the wild, critically endangered, endangered or vulnerable by [IUCN's red list](https://www.iucnredlist.org).

<!-- Since there is sometimes disagreement between accepted taxonomic names in POWO and IUCN's red list taxonomic names we use two methods to extract the IUCN's red list category. First, we match the taxonomic name in the report directly to IUCN's red list. Second, we first match to the taxonomic name in the report to POWO and use the accepted taxonomic name in POWO to match to IUCN's red list. If the two methods result in different red list categories we use the most severe category.  -->

<!-- Note that in this document we will show the geographic area that threatened taxa in the living collection belong to. To determine the geographic distribution of items in the living collection we enrich with information from [Plants of the World Online](https://powo.science.kew.org), by matching taxonomic names. POWO uses [TDWG geographical codes](https://www.tdwg.org/standards/wgsrpd/) (Brummitt, 2001) expressed to that system's third level. This splits the world into 369 regions. Thus, each accepted name in POWO is associated with a list of geographic regions. POWO splits the geographic regions associated with a taxa by three conditions: -->

<!-- - Native / introduced, -->
<!-- - Extinct / Not extinct, -->
<!-- - Location is doubtful / otherwise. -->

<!-- This document considers locations that are: `r location_type_text`  -->

In analyses over time we use the current version of IUCN's red list retrospectively on the historic data. 
 
```{r get threatened only information}
threatened = enriched_report[which(enriched_report$threatened == 'Threatened'),]
max_year = format(Sys.Date(), '%Y') |> as.numeric()

# threatened must match to POWO so no naming issues.
threatened_existing = threatened[which(threatened$ItemStatusType == 'Existing'),] 
no_threatened_items_existing = nrow(threatened_existing)
no_threatened_accessions_existing = length(unique(threatened_existing$AccNoFull))
no_threatened_species_existing = length(unique(paste0(threatened_existing$POWO_genus, ' ', threatened_existing$POWO_species)))
no_threatened_taxa_existing = length(unique(paste0(threatened_existing$POWO_taxon_name, ' ', threatened_existing$POWO_taxon_authors)))

to_keep = which(as.numeric(threatened$AccYear) >= earliest_allowable_record  & as.numeric(threatened$AccYear) <= max_year)
still_issue = nrow(threatened) - length(to_keep)
threatened = threatened[to_keep,]



# For all items we need to be more careful. Items and accessions are fine. 
# For species and taxa only consider those that have matched to POWO. 
LC_existing = enriched_report[enriched_report$ItemStatusType == 'Existing',]
no_LC_items_existing = nrow(LC_existing)
no_LC_accessions_existing = length(unique(LC_existing$AccNoFull))
no_LC_species_existing = length(unique(paste0(LC_existing$POWO_genus, ' ', LC_existing$POWO_species)))
no_LC_species_existing = no_LC_species_existing - sum(any(is.na(LC_existing$POWO_genus)))
no_LC_taxa_existing = length(unique(paste0(LC_existing$POWO_taxon_name, ' ', LC_existing$POWO_taxon_authors)))
no_LC_taxa_existing = no_LC_taxa_existing - sum(any(is.na(LC_existing$POWO_genus)))


category = c('Items', 'Accessions', 'Species', 'Taxa')
threatened_values = c(no_threatened_items_existing, no_threatened_accessions_existing, no_threatened_species_existing, no_threatened_taxa_existing)
LC_values  = c(no_LC_items_existing, no_LC_accessions_existing, no_LC_species_existing, no_LC_taxa_existing)
proportion = paste0(round(threatened_values/LC_values*100, digits = 2),'%')

threatened_count_table = data.frame(category, threatened_values,LC_values, proportion)
names(threatened_count_table) = c('Category', 'Number of threatened', ' Total in living collection', 'Proportion of threatened')

if(export_data){
  exporting_data_store$threatened_count_table = threatened_count_table
}

### Over time information
### A) threatened
year_cur = as.numeric(format(Sys.Date(), '%Y'))
years = max(min_year, min(enriched_report$AccYear[enriched_report$AccYear > 1650],na.rm = T)) :year_cur
dates = paste0(years,'-12-31')
threatened_existing_each_year = BGSmartR::exist_at_date(date = dates,
                                                     AccessionYear = as.character(threatened$AccYear),
                                                     ItemStatusDate = as.character(threatened$ItemStatusDate),
                                                     ItemStatusType = as.character(threatened$ItemStatusType),
                                                     post_date = '3000-01-01')

# Get the number of items accession taxa species for threatened plants in the LC.
over_time_info = lapply(threatened_existing_each_year, function(x){
  garden_current = threatened[x,]
  
  # Get the values of interest.
  no_items = nrow(garden_current)
  no_accessions = length(unique(garden_current$AccNoFull))
  no_taxa = length(unique(garden_current$TaxonNameFull))
  no_species = length(unique(paste0(garden_current$POWO_taxon_name, garden_current$POWO_taxon_authors)[!is.na(garden_current$POWO_plant_name_id)]))
  
  return(c(no_items, no_accessions, no_taxa, no_species))
}) |>
  data.frame() |>
  t() |> 
  data.frame()
names(over_time_info) = c('Items', 'Accessions', 'Taxa', 'Species')
over_time_info = data.frame(Date = dates, over_time_info)
row.names(over_time_info) = 1:nrow(over_time_info)
threatened_over_time = over_time_info

if(export_data){
  exporting_data_store$Number_of_threatened_species_in_LC_over_time = threatened_over_time
}

### B) Whole LC.
# Do the same for the whole collection.
report = enriched_report
report = report[report$AccYear > 1650 & report$AccYear<=year_cur,]
report = report[!is.na(report$ItemStatusDate),]
plant_existing = BGSmartR::exist_at_date(dates, AccessionYear = as.character(report$AccYear),
                                         ItemStatusDate = as.character(report$ItemStatusDate),
                                         ItemStatusType = as.character(report$ItemStatusType))

# For each year get the number in each group (species and genera)
over_time_info = lapply(plant_existing, function(x){
  garden_current = report[x,]
  
  # Get the values of interest.
  no_items = nrow(garden_current)
  no_accessions = length(unique(garden_current$AccNoFull))
  no_taxa = length(unique(garden_current$TaxonNameFull))
  no_species = length(unique(paste0(garden_current$POWO_taxon_name, garden_current$POWO_taxon_authors)[!is.na(garden_current$POWO_plant_name_id)]))
  
  return(c(no_items, no_accessions, no_taxa, no_species))
}) |>
  data.frame() |>
  t() |> 
  data.frame()
names(over_time_info) = c('Items', 'Accessions', 'Taxa', 'Species')
over_time_info = data.frame(Date = dates, over_time_info)
row.names(over_time_info) = 1:nrow(over_time_info)
LC_data = over_time_info

if(export_data){
  exporting_data_store$Total_counts_of_LC_over_time = LC_data
}

### MANAGEMENT SHEET: All existing threatened accessions.
threatened_accessions = threatened_existing
threatened_accessions = threatened_accessions[match(unique(threatened_accessions$AccNoFull),threatened_accessions$AccNoFull),]
powo_name = paste0(threatened_accessions$POWO_taxon_name, ' ', threatened_accessions$POWO_taxon_authors)
powo_name[powo_name == 'NA NA'] = NA
threatened_accessions$powo_name = powo_name
threatened_accessions = threatened_accessions[,match(c('AccNoFull', 'TaxonNameFull', 'powo_name', 'threatened', 'threatened_category', 'endemic', 'geography_codes', 'ProvenanceCode', 'no_gardens'), names(threatened_accessions))]
names(threatened_accessions) = c('Accession Number', 'Original Taxonomic Name and Author', 'POWO Taxonomic Name and Author', 'Threatened', 'Red list category', 'endemic', 'Geographic codes..', 'Provenance', 'Number of Global Collections Held in')
dir.create(output_dir,showWarnings = F,recursive = T)
writexl::write_xlsx(threatened_accessions, path =paste0(output_dir,'/',collection,'_threatened_accessions_list.xlsx'))


```

### Threatened collection size

In the living collection there are `r format(no_threatened_items_existing, big.mark=',')` items corresponding to  `r format(no_threatened_accessions_existing, big.mark=',')` accessions of existing threatened plants. The number of threatened species existing in the living collection is summarised below. 

<!-- Note that here species is defined as the genus species pair and taxa is the taxonomic name and author (thus including subspecies, variety, forma). -->

```{r, How many threatened species are in the living collection - table}
if(report_kind == 'static'){
  threatened_count_table |>
    flextable() |>
    set_caption(caption = "threatened plants in the living collection") |> 
    font(fontname = "Calibri (Body)", part = "all") |> 
    fontsize(size = table_font_size, part = "body") |> 
    # add footer if you want
    # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
    #                colwidths = 4) |> 
    theme_booktabs() |> # default theme
    autofit()
}
if(report_kind == 'interactive'){
  
   threatened_count_table |>  DT::datatable(extensions = 'Buttons',
                rownames = FALSE,
                options = list(scrollX = TRUE,
                               pageLength = nrow(threatened_count_table),
                               dom = 'Bt',
                               buttons = c('copy', 'csv', 'excel', 'pdf'),
                               paging=FALSE,
                               ordering=F))
}

```

Next we split the threatened plants in the living collection by their IUCN red list category.


```{r,  Threatened, native plants in the LC - data}
# threatened_existing
# 
redlistcatA = threatened_existing$redList_category
redlistcatB = threatened_existing$POWO_Red_category

redlist_used = rep(NA, nrow(threatened_existing))
redlist_used[redlistcatA == 'VU' | redlistcatB == 'VU'] = 'VU'
redlist_used[redlistcatA == 'EN' | redlistcatB == 'EN'] = 'EN'
redlist_used[redlistcatA == 'CR' | redlistcatB == 'CR'] = 'CR'
redlist_used[redlistcatA == 'EW' | redlistcatB == 'EW'] = 'EW'
redlist_used[redlistcatA == 'EX' | redlistcatB == 'EX'] = 'EX'

threatened_existing$redlist_used = redlist_used

redlistcat_items = data.frame(table(threatened_existing$redlist_used))
if(nrow(redlistcat_items) == 0){
  redlistcat_items = data.frame(matrix(ncol = 2, nrow = 0))
}
names(redlistcat_items) = c('RedListCategory', 'Items')
redlistcat_items$RedListCategory = as.character(redlistcat_items$RedListCategory)
redlistcat_items$RedListCategory  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(redlistcat_items$RedListCategory, c('VU', 'EN', 'CR', 'EW', 'EX'))] 

if(nrow(redlistcat_items) > 0){
redlistcat_items$percent = paste0(round(redlistcat_items$Items / sum(redlistcat_items$Items)*100,digits = 1),'%')
redlistcat_items <- redlistcat_items %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Items))), 
         pos = Items/2 + dplyr::lead(csum, 1),
         pos = dplyr::if_else(is.na(pos), Items/2, pos))
}

threatened_existing_redlist = threatened_existing[threatened_existing$redlist_used %in% c('VU', 'EN', 'CR', 'EW', 'EX'),]

redlistcat_accessions = data.frame(table(threatened_existing_redlist$redlist_used[match(unique(threatened_existing_redlist$AccNoFull), threatened_existing_redlist$AccNoFull)]))
if(nrow(redlistcat_accessions) == 0){
  redlistcat_accessions = data.frame(matrix(ncol = 2, nrow = 0))
}
names(redlistcat_accessions) = c('RedListCategory', 'Accessions')
redlistcat_items$RedListCategory = as.character(redlistcat_items$RedListCategory)
redlistcat_accessions$RedListCategory  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(redlistcat_accessions$RedListCategory, c('VU', 'EN', 'CR', 'EW', 'EX'))]  

if(nrow(redlistcat_accessions) > 0){
redlistcat_accessions$percent = paste0(round(redlistcat_accessions$Accessions / sum(redlistcat_accessions$Accessions)*100,digits = 1),'%')
redlistcat_accessions <- redlistcat_accessions %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Accessions))), 
         pos = Accessions/2 + dplyr::lead(csum, 1),
         pos = dplyr::if_else(is.na(pos), Accessions/2, pos))
}

if(nrow(redlistcat_items) > 0){
  flex_data = data.frame(redlistcat_items$RedListCategory,
                       redlistcat_items$Items, 
                       paste0(round(redlistcat_items$Items/sum(redlistcat_items$Items)*100,digits=2),'%'),
                       redlistcat_accessions$Accessions,
                       paste0(round(redlistcat_accessions$Accessions/sum(redlistcat_accessions$Accessions)*100,digits=2),'%'))
}else{
  flex_data = data.frame(matrix(ncol = 5, nrow = 0))
}
names(flex_data) = c('Red List Category', 'Number of items', 'Proportion of items', 'Number of accessions', 'Proportion of accessions')

have_redlist = eval = nrow(flex_data)>0
if(export_data){
  exporting_data_store$Threatened_category_of_existing_native_plants_in_the_LC = flex_data
}
```

```{r, Threatened, native plants in the LC - table, eval = have_redlist}
if(report_kind == 'static'){
  flex_data |>
  flextable() |>
  set_caption(caption = paste0('Threatened category of existing plants in the living collection')) |> 
  font(fontname = "Calibri (Body)", part = "all") |> 
  fontsize(size = table_font_size, part = "body") |> 
  # add footer if you want
  # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
  #                colwidths = 4) |> 
  theme_booktabs() |> # default theme
  separate_header() |> 
  align(align = "center", part = "all") |>
  autofit()
}
if(report_kind == 'interactive'){
  flex_data_table = flex_data |>  DT::datatable(extensions = 'Buttons',
                  rownames = FALSE,
                  options = list(scrollX = TRUE,
                                 pageLength = nrow(flex_data),
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))
    # DT::datatable(flex_data, options = list(dom = 't'))

}

```

```{r  duel pie charts}
### Create duel pie chart
   endangered_cat =  c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')
  endangered_cat_order = endangered_cat[which(!is.na(match(endangered_cat, redlistcat_accessions$RedListCategory)))]
  redlistcat_accessions$RedListCategory = factor(redlistcat_accessions$RedListCategory, levels = endangered_cat_order)
  redlistcat_accessions = redlistcat_accessions[order(redlistcat_accessions$RedListCategory),]

  endangered_cat_order = endangered_cat[which(!is.na(match(endangered_cat, redlistcat_items$RedListCategory)))]
  redlistcat_items$RedListCategory = factor(redlistcat_items$RedListCategory, levels = endangered_cat_order)
  redlistcat_items = redlistcat_items[order(redlistcat_items$RedListCategory),]
 
figs = list()
# Items
text = paste0(redlistcat_items$Items, ' native items with red list category "', redlistcat_items$RedListCategory, '"')
fig1 <- plot_ly(redlistcat_items, labels = ~RedListCategory, values = ~Items, hoverinfo = "text",  hovertext = text, sort = FALSE)
fig1 <- fig1 %>% add_pie(hole = 0.6)
fig1 <- fig1 %>% layout(title = "Items",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

figs$items = fig1
#Accessions
text = paste0(redlistcat_accessions$Accessions, ' native accessions with red list category "', redlistcat_accessions$RedListCategory, '"')
             
fig2 <- plot_ly(redlistcat_accessions, labels = ~RedListCategory, values = ~Accessions, hoverinfo = "text",  hovertext = text, sort = FALSE)
fig2 <- fig2 %>% add_pie(hole = 0.6)
fig2 <- fig2 %>% layout(title = "Accessions",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

figs$accessions = fig2
colo = scales::brewer_pal(palette = palette, direction = 1)(5)
names(colo) = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')
# colo = c("Vulnerable" = rgb(242,141,60, max = 255),
#          "Endangered" = rgb(238,81,53, max = 255),
#          "Critically Endangered" = rgb(129,27,39, max = 255),
#          "Extinct in the Wild" = rgb(88,88,88, max = 255),
#          "Extinct" = rgb(0,0,0, max = 255)
#          )
colo = colo[match(redlistcat_items$RedListCategory, names(colo))]


# Together
p <- plot_ly() |>
  add_pie(data = redlistcat_items, labels = ~RedListCategory, values = ~Items, hole = 0.6,
          title = "Items", name ='', sort = FALSE, marker = list(colors = colo),
          domain = list(row = 0, column = 0)) |>
  add_pie(data = redlistcat_accessions, labels = ~RedListCategory, values = ~Accessions, hole = 0.6,
          title = "Accessions", name ='', sort = FALSE, marker = list(colors = colo),
         domain = list(row = 0, column = 1))  |>
  layout(title = "", showlegend = TRUE,
          grid=list(rows=1, columns=2),
         plot_bgcolor="#f0f0f0",
         legend = list(orientation = 'h', xanchor = "center", x = 0.5),
         xaxis = list(title=paste0('Items', 'Accessions'), showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         margin = c(0,0,0,0))


if(separate_figure_folder){
  htmlwidgets::saveWidget(figs$items, file = paste0(figures_dir, '/','native_threatened_cat_existing_items.html'))
  htmlwidgets::saveWidget(figs$accessions, file = paste0(figures_dir, '/','native_threatened_cat_existing_accessions.html'))
}

duel_pie_chart = p

### (END) duel pie chart
```

```{r, How many threatened cateogories in the living collection - Table, eval = T}
#items
threatened_categories =c('VU', 'EN', 'CR', 'EW', 'EX')
cat_items = table(threatened_existing$threatened_category)
cat_items = cat_items[match(threatened_categories, names(cat_items))]

threatened_existing_acc = threatened_existing[match(unique(threatened_existing$AccNoFull),
                                                               threatened_existing$AccNoFull),]
cat_acc = table(threatened_existing_acc$threatened_category)
cat_acc = cat_acc[match(threatened_categories, names(cat_acc))]

best_GenusSpecies = threatened_existing$GenusSpecies
best_GenusSpecies[!is.na(threatened_existing$POWO_taxon_name)] = paste0(threatened_existing$POWO_genus[!is.na(threatened_existing$POWO_taxon_name)],
                                                                        ' ',                                                                      threatened_existing$POWO_species[!is.na(threatened_existing$POWO_taxon_name)] )
threatened_existing$best_GenusSpecies = best_GenusSpecies

threatened_existing_species = threatened_existing[match(unique(threatened_existing$best_GenusSpecies),
                                                               threatened_existing$best_GenusSpecies),]
cat_species = table(threatened_existing_species$threatened_category)
cat_species = cat_species[match(threatened_categories, names(cat_species))]

threatened_existing_taxa = threatened_existing[match(unique(threatened_existing$best_name),
                                                               threatened_existing$best_name),]
cat_taxa = table(threatened_existing_acc$threatened_category)
cat_taxa = cat_taxa[match(threatened_categories, names(cat_taxa))]

cat_items = as.numeric(cat_items) ; cat_items[is.na(cat_items)] = 0
cat_acc= as.numeric(cat_acc) ; cat_acc[is.na(cat_acc)] = 0
cat_species= as.numeric(cat_species) ; cat_species[is.na(cat_species)] = 0
cat_taxa= as.numeric(cat_taxa) ; cat_taxa[is.na(cat_taxa)] = 0

cat_table = data.frame(c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct'),
                       cat_items,
                       cat_acc,
                       cat_species,
                       cat_taxa)
names(cat_table) = c('Category', 'Items','Accessions','Species','Taxa')

if(report_kind == 'static'){
  cat_table |>
    flextable() |>
    set_caption(caption = "") |> 
    font(fontname = "Calibri (Body)", part = "all") |> 
    fontsize(size = table_font_size, part = "body") |> 
    # add footer if you want
    # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
    #                colwidths = 4) |> 
    theme_booktabs() |> # default theme
    autofit()
}
if(report_kind == 'interactive'){
  cat_table = cat_table |>  DT::datatable(extensions = 'Buttons',
                  rownames = FALSE,
                  options = list(scrollX = TRUE,
                                 pageLength = nrow(cat_table),
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))
}
```

`r if(report_kind == 'interactive' & have_redlist){"#####  {.tabset}"}`


`r if(report_kind == 'interactive' & have_redlist){"###### Table"}`

`r if(report_kind == 'interactive'){cat_table}`

`r if(report_kind == 'interactive' & have_redlist){"###### Chart"}`

```{r,  Threatened existing plants in the LC - static pie charts, echo = FALSE, fig.fullwidth=TRUE, fig.dim = c(10, 5), eval = report_kind == 'interactive' & have_redlist}
duel_pie_chart
```

`r if(report_kind == 'interactive'){"##### {-}"}`

***

### Change of threatened species over time

In this section we explore how the number of threatened species in the living collection has varied over time.

<!-- For each year we use the accession year and item status date to determine which items in the collection are existing on the 31st of December each year. -->

```{r, What proportion of the living collection at `r collection` is threatened and how has this varied over time - static number figures, eval = report_kind == 'static' & do_over_time,  fig.fullwidth=TRUE, fig.dim = c(10, 6)}
to_do = c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')
for(i in 1:length(to_do)){
  counts = threatened_over_time[,match(to_do[i], names(threatened_over_time))]
  plot_data = data.frame(date = years, counts = counts)
  
  p = ggplot(data=plot_data, aes(x=date, y=counts)) +
    geom_line()+
    geom_point() +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title=paste0("Number of threatened ", titles[i] , " over time"),
         x ="Year",
         y = paste0("Number of ",titles[i])) #+
  # guides(fill=guide_legend(title=paste0("", titles[i]))) +
  # theme(text = element_text(size=table_font_size))
  
  if(separate_figure_folder){
    ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','number_threatened_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
  }
  
  print(p)
}
```

`r if(report_kind == 'interactive' & do_over_time){"#####  {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Items"}`
```{r  Number of threatened items over time in the living collection - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_over_time}
number_over_time <- function(data, type, title){
counts = data[,match(type, names(data))]
plot_data = data.frame(date = years, counts = counts)

fig <- plot_ly(plot_data, x = ~date, y = ~counts, type = 'scatter', mode = 'lines+markers', line = list(color = interactive_colour(1)), marker = list(color = interactive_colour(1)))

fig <- fig |> layout(yaxis = list(title = paste0("Number of ",title)),
                   xaxis = list(title = "Year"))
fig <- fig |> layout(hovermode = 'x unified')
return(fig)
}

fig = number_over_time(data = threatened_over_time, type = 'Items', title = 'items')

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','number_of_threatened_items_over_time.html'))
}

fig
```

`r if(report_kind == 'interactive' & do_over_time){"###### Accessions"}`
```{r  Number of threatened accessions over time in the living collection - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_over_time}
fig = number_over_time(data = threatened_over_time, type = 'Accessions', title = 'accessions')

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','number_of_threatened_accessions_over_time.html'))
}

fig
```

`r if(report_kind == 'interactive' & do_over_time){"###### Taxa"}`
```{r  Number of threatened taxa over time in the living collection - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_over_time}
fig = number_over_time(data = threatened_over_time, type = 'Taxa', title = 'taxa')

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','number_of_threatened_taxa_over_time.html'))
}

fig

```

`r if(report_kind == 'interactive' & do_over_time){"###### Species"}`
```{r  Number of threatened species over time in the living collection - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_over_time}
fig = number_over_time(data = threatened_over_time, type = 'Species', title = 'species')

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','number_of_threatened_species_over_time.html'))
}

fig
```

`r if(report_kind == 'interactive' & do_over_time){"##### {-}"}`

`r if(do_over_time){"Next we view threatened plants as a percentage of all plants in the living collection over time."}`

```{r, What proportion of the living collection at `r collection` is threatened and how has this varied over time - percentage data, eval = do_over_time}
# Do the same for the whole collection.
report = report_original
report = report[report$AccYear > 1650 & report$AccYear<=year_cur,]
report = report[!is.na(report$ItemStatusDate),]
date = paste0(years,'-01-01')
plant_existing = BGSmartR::exist_at_date(date, AccessionYear = as.character(report$AccYear),
                       ItemStatusDate = as.character(report$ItemStatusDate),
                       ItemStatusType = as.character(report$ItemStatusType))

# For each year get the number in each group (species and genera)
over_time_info = lapply(plant_existing, function(x){
    garden_current = report[x,]
    
    # Get the values of interest.

    no_items = nrow(garden_current)
    no_accessions = length(unique(garden_current$AccNoFull))
    no_taxa = length(unique(garden_current$TaxonNameFull))
    no_species = length(unique(paste0(garden_current$POWO_taxon_name, garden_current$POWO_taxon_authors)[!is.na(garden_current$POWO_plant_name_id)]))

    return(c(no_items, no_accessions, no_taxa, no_species))
}) |>
                  data.frame() |>
                  t() |> 
                  data.frame()
names(over_time_info) = c('Items', 'Accessions', 'Taxa', 'Species')
over_time_info = data.frame(Date = date, over_time_info)
row.names(over_time_info) = 1:nrow(over_time_info)
LC_data = over_time_info

if(export_data){
  exporting_data_store$Total_species_in_LC_over_time = LC_data
}
```

```{r, What proportion of the LC at `r collection` is threatened and how has this varied over time - static percentage figures, eval = report_kind == 'static' & do_over_time, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
to_do =  c('Items', 'Accessions', 'Taxa', 'Species')
titles = c('items', 'accessions', 'taxa', 'species')

# Loop over to_do.
for(i in 1:length(to_do)){
  threatened_counts = threatened_over_time[,match(to_do[i], names(threatened_over_time))]
  total_counts = LC_data[,match(paste0(to_do[i]), names(LC_data))]
  
  percent = round(threatened_counts/total_counts*100, digits = 3)
  
  plot_dataA = data.frame(date =years, 
                          percent = percent,
                          represented = rep('Threatened', length(date)))
  plot_dataB = data.frame(date =years, 
                          percent = 100-percent,
                          represented = rep('Widespread', length(date)))
  
  plot_data = rbind(plot_dataA, plot_dataB)
  plot_data$represented = factor(plot_data$represented,levels = c('Widespread','Threatened'))
  
  names(plot_data) = c('dates', 'percentage', 'represented')
  
  p = ggplot(plot_data, aes(fill=represented, y=percentage, x=dates)) + 
    geom_bar(position="stack", stat="identity", width = 1) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title=paste0("Proportion of ", titles[i] , " that are threatened in the LC over time"),
         x ="Year",
         y = "Percentage (%)") +
    guides(fill=guide_legend(title=paste0("", titles[i]))) +
    scale_fill_discrete() # +
  # theme(text = element_text(size=table_font_size))
  
  if(separate_figure_folder){
    ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','proportion_threatened_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
  }
  
  print(p)
}


```

`r if(report_kind == 'interactive' & do_over_time){"#####  {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Items"}`
```{r  Proportion of threatened items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_over_time}
proportion_plot <- function(data, LC_data, type, title){
  native_counts = data[,match(type, names(data))]
  total_counts = LC_data[,match(paste0(type), names(LC_data))]

  percent = round(native_counts/total_counts*100, digits = 3)
  

  # Line graph of only the native plants. 
  plot_data = data.frame(date =years, 
                               percent = percent
                               )
  fig <- plot_ly(plot_data, x = ~date, y = ~percent, type = 'scatter', mode = 'lines+markers', line = list(color = interactive_colour(1)), marker = list(color = interactive_colour(1)))

fig <- fig |> layout(yaxis = list(title = paste0("Percentage of ",title)),
                   xaxis = list(title = "Year"))
fig <- fig |> layout(hovermode = 'x unified',
                     yaxis = list(ticksuffix  ='%', hoverformat = '.1f'))
  

  # Stacked Bar version. 
   # plot_dataA = data.frame(date =years, 
   #                             percent = percent,
   #                             represented = rep('Native', length(date)))
   # 
  # plot_dataB = data.frame(date =years, 
  #                            percent = 100-percent,
  #                             represented = rep('Non-Native', length(date)))
  
  # plot_data = rbind(plot_dataA, plot_dataB)
  # plot_data$represented = factor(plot_data$represented, levels = c('Native', 'Non-Native'))
  # fig <- plot_ly(plot_data, x = ~date, y = ~percent, color = ~represented, type = 'bar', colors = interactive_colour(2))
  # fig <- fig |> layout(yaxis = list(title = paste0('Percentage of ',title)),
  #                      xaxis = list(title = "Year"),
  #                      legend=list(title=list(text=paste0("Native ", title))),
  #                      barmode = 'stack',
  #                      hoverlabel = list(namelength = -1))
  # fig <- fig |> layout(hovermode = 'x unified',
  #                      bargap =0,
  #                      yaxis = list(ticksuffix  ='%', hoverformat = '.1f')
                       
return(fig)
}

fig = proportion_plot(data = threatened_over_time, LC_data = LC_data, type = 'Items', title = 'items')

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','percentage_of_threatened_items_over_time.html'))
}
  
fig
```

`r if(report_kind == 'interactive' & do_over_time){"###### Accessions"}`
```{r  Proportion of threatened accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_over_time}
fig = proportion_plot(data = threatened_over_time, LC_data = LC_data, type = 'Accessions', title = 'accessions')

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','percentage_of_threatened_accessions_over_time.html'))
}
  
fig
```

`r if(report_kind == 'interactive' & do_over_time){"###### Taxa"}`
```{r  Proportion of threatened taxa over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_over_time}
fig = proportion_plot(data = threatened_over_time, LC_data = LC_data, type = 'Taxa', title = 'taxa')

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','percentage_of_threatened_taxa_over_time.html'))
}
  
fig
```

`r if(report_kind == 'interactive' & do_over_time){"###### Species"}`
```{r  Proportion of threatened species over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_over_time}
fig = proportion_plot(data = threatened_over_time, LC_data = LC_data, type = 'Species', title = 'species')

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','percentage_of_threatened_species_over_time.html'))
}
  
fig
```

`r if(report_kind == 'interactive' & do_over_time){"##### {-}"}`


`r if(have_redlist & do_over_time){"Next we split the threatened plants by their IUCN designation and explore how this has changed over time."}`

```{r, Threatened cat plants in the LC - data over time, fig.fullwidth=FALSE, fig.dim = c(10, 6), eval = do_over_time}
# threatenedcontains all theatened plants over the LC's history.
redlistcatA = threatened$redList_category
redlistcatB = threatened$POWO_Red_category

redlist_used = rep(NA, nrow(threatened))
redlist_used[redlistcatA == 'VU' | redlistcatB == 'VU'] = 'VU'
redlist_used[redlistcatA == 'EN' | redlistcatB == 'EN'] = 'EN'
redlist_used[redlistcatA == 'CR' | redlistcatB == 'CR'] = 'CR'
redlist_used[redlistcatA == 'EW' | redlistcatB == 'EW'] = 'EW'
redlist_used[redlistcatA == 'EX' | redlistcatB == 'EX'] = 'EX'

threatened$redlist_used = threatened$threatened_category

# For each year get the number in each group (species and genera)
over_time_info = lapply(threatened_existing_each_year, function(x){
  garden_current = threatened[x,]
  redlistcat_items = data.frame(table(garden_current$redlist_used))
  if(ncol(redlistcat_items) != 2){redlistcat_items = data.frame(matrix(ncol = 2, nrow = 0))}
  names(redlistcat_items) = c('RedListCategory', 'Items')
  
  redlistcat_accessions = data.frame(table(garden_current$redlist_used[match(unique(garden_current$AccNoFull), garden_current$AccNoFull)]))
  if(ncol(redlistcat_accessions) != 2){redlistcat_accessions = data.frame(matrix(ncol = 2, nrow = 0))}
  
  names(redlistcat_accessions) = c('RedListCategory', 'Accessions')
  
  taxa = paste0(garden_current$POWO_taxon_name, ' ', garden_current$POWO_taxon_authors)
  redlistcat_taxa = data.frame(table(garden_current$redlist_used[match(unique(taxa), taxa)]))
  if(ncol(redlistcat_taxa) != 2){redlistcat_taxa = data.frame(matrix(ncol = 2, nrow = 0))}
  
  names(redlistcat_taxa) = c('RedListCategory', 'Accessions')
  
  
  return(list(redlistcat_items = redlistcat_items,
              redlistcat_accessions = redlistcat_accessions,
              redlistcat_taxa = redlistcat_taxa))
})


```

```{r, Threatened, native plants in the LC - static proportion over time, fig.fullwidth=FALSE, fig.dim = c(10, 6), eval = report_kind == 'static' & have_redlist & do_over_time}
to_do = c('items', 'accessions')
if(nrow(flex_data) > 0){
for(i in 1:length(to_do)){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  if(export_data){
    item_name = paste0('Threatened_native_plants_in_the_LC_proportion_over_time_',to_do[i])
  exporting_data_store$AA = data_wanted
  names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
}
  
    p = ggplot(data_wanted, aes(fill=RedListCategory, y=prop, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = "Percentage (%)") +
    guides(fill=guide_legend(title=paste0('RedList Category')))# +
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("RedList Category of ", to_do[i] , " that are native in the LC over time"))
    
    if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Native_threatened_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}
  print(p)

}
}
```

`r if(report_kind == 'static' &  have_redlist & do_over_time){"Note that there might not be many threatened native plants. Therefore, we also include line graphs of the number of each threatened category in the living collection."}`

```{r, Threatened, native plants in the LC - number over time, fig.fullwidth=FALSE, fig.dim = c(10, 6), eval = report_kind == 'static' & have_redlist & do_over_time}
to_do = c('items', 'accessions', 'taxa')
if(nrow(flex_data) > 0){
for(i in 1:length(to_do)){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  
  # Create plot.
  p = ggplot(data=data_wanted, aes(x=year, y=data_wanted[,3], group=RedListCategory)) +
  geom_line(aes(color=RedListCategory)) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title="",
        x ="Year",
        y = paste0('Number of ',to_do[i])) +
    guides(color=guide_legend(title=paste0('RedList Category'))) #+
    # theme(text = element_text(size=table_font_size)) +
    # ggtitle(paste0("RedList Category of ", to_do[i] , " that are native in the LC over time"))
  
      if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Number_of_threatened_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
      }
  
  print(p)
}
}
```


`r if(report_kind == 'static' & have_redlist & do_over_time){"We do this in two ways, looking at the number of plants in each threatened category ove time and the proportion if each threatened cateogory over all threatened taxa."}`

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"##### {.tabset}"}`

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"###### Items (Number)"}`
```{r  Number of threatened native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_over_time}

i=1
to_do = c('items', 'accessions', 'taxa')
if(nrow(flex_data) > 0){

  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))

colo = scales::brewer_pal(palette = palette, direction = 1)(5)
names(colo) = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')

# colo = c("Vulnerable" = rgb(242,141,60, max = 255),
#          "Endangered" = rgb(238,81,53, max = 255),
#          "Critically Endangered" = rgb(129,27,39, max = 255),
#          "Extinct in the Wild" = rgb(88,88,88, max = 255),
#          "Extinct" = rgb(0,0,0, max = 255)
#          )
  endangered_cat =  c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')
  endangered_cat_order = endangered_cat[which(!is.na(match(endangered_cat, data_wanted$RedListCategory)))]
  data_wanted$RedListCategory = factor(data_wanted$RedListCategory, levels = endangered_cat_order)
colo = colo[match(levels(data_wanted$RedListCategory), names(colo))]
  
  fig <- plot_ly(data_wanted, x = ~year, y = ~Items, color = ~RedListCategory, type = 'scatter', mode = 'lines+markers', colors =  colo)

  fig <- fig |> layout(yaxis = list(title = paste0("Number of ",to_do[i])),
                   xaxis = list(title = "Year"))
  fig <- fig |> layout(hovermode = 'x unified')

 
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line-number_of_threatened_',to_do[i],'_over_time.html'))
}
  
 fig
}

```

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"###### Accessions (Number)"}`
```{r  Number of threatened native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & have_redlist & do_over_time}
i=2
to_do = c('items', 'accessions', 'taxa')
if(nrow(flex_data) > 0){

  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
   endangered_cat =  c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')
  endangered_cat_order = endangered_cat[which(!is.na(match(endangered_cat, data_wanted$RedListCategory)))]
  data_wanted$RedListCategory = factor(data_wanted$RedListCategory, levels = endangered_cat_order)
colo = colo[match(levels(data_wanted$RedListCategory), names(colo))]
  
  fig <- plot_ly(data_wanted, x = ~year, y = ~Accessions, color = ~RedListCategory, type = 'scatter', mode = 'lines+markers',colors =  colo)

  fig <- fig |> layout(yaxis = list(title = paste0("Number of ",to_do[i])),
                   xaxis = list(title = "Year"))
  fig <- fig |> layout(hovermode = 'x unified')

 
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line-number_of_threatened_',to_do[i],'_over_time.html'))
}
  fig
}


```

`r if(report_kind == 'interactive'& have_redlist & do_over_time){"###### Taxa (Number)"}`
```{r  Number of threatened native taxa over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & have_redlist & do_over_time}
i=3
to_do = c('items', 'accessions', 'taxa')
if(nrow(flex_data) > 0){

  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
   endangered_cat =  c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')
  endangered_cat_order = endangered_cat[which(!is.na(match(endangered_cat, data_wanted$RedListCategory)))]
  data_wanted$RedListCategory = factor(data_wanted$RedListCategory, levels = endangered_cat_order)
colo = colo[match(levels(data_wanted$RedListCategory), names(colo))]
  fig <- plot_ly(data_wanted, x = ~year, y = ~Accessions, color = ~RedListCategory, type = 'scatter', mode = 'lines+markers', colors =  colo)

  fig <- fig |> layout(yaxis = list(title = paste0("Number of ",to_do[i])),
                   xaxis = list(title = "Year"))
  fig <- fig |> layout(hovermode = 'x unified')

 
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line-number_of_threatened_',to_do[i],'_over_time.html'))
}
  fig
}


```

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"###### Items (Proportion)"}`
```{r  Proportion of threatened native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & have_redlist & do_over_time}
i=1
to_do = c('items', 'accessions')
if(nrow(flex_data) > 0){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
   endangered_cat =  rev(c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct'))
  endangered_cat_order = endangered_cat[which(!is.na(match(endangered_cat, data_wanted$RedListCategory)))]
  data_wanted$RedListCategory = factor(data_wanted$RedListCategory, levels = endangered_cat_order)
colo = colo[match(levels(data_wanted$RedListCategory), names(colo))]
  if(export_data){
    item_name = paste0('Threatened_native_plants_in_the_LC_proportion_over_time_',to_do[i])
  exporting_data_store$AA = data_wanted
  names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
}
  
   text = paste0(round(data_wanted$prop,digits=2), '% (',data_wanted$Items,' items) ')
data_wanted$text = text
fig <- plot_ly(data_wanted, x = ~year, y = ~prop, color = ~RedListCategory, type = 'bar', text = ~text,
               hovertemplate = '%{text}',
               sort = FALSE,colors =  colo)
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text=''),  traceorder= 'reversed'),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)



if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-native_threatened_over_time_by_item.html'))
}

fig
}

```

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"###### Accessions (Proportion)"}`
```{r  Proportion of threatened native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & have_redlist & do_over_time}
i=2
to_do = c('items', 'accessions')
if(nrow(flex_data) > 0){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('redlistcat',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
  for(jj in 1:length(over_time_info)){
    info_year = over_time_info[[jj]][[wanted_item]]
     if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)), info_year, prop = prop)
    data_wanted = rbind(data_wanted, info_year)
  }
 data_wanted[,2]  = c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct')[match(data_wanted[,2], c('VU', 'EN', 'CR', 'EW', 'EX'))]  
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
   endangered_cat =  rev(c('Vulnerable', 'Endangered', 'Critically Endangered', 'Extinct in the Wild', 'Extinct'))
  endangered_cat_order = endangered_cat[which(!is.na(match(endangered_cat, data_wanted$RedListCategory)))]
  data_wanted$RedListCategory = factor(data_wanted$RedListCategory, levels = endangered_cat_order)
colo = colo[match(levels(data_wanted$RedListCategory), names(colo))]
  if(export_data){
    item_name = paste0('Threatened_native_plants_in_the_LC_proportion_over_time_',to_do[i])
  exporting_data_store$AA = data_wanted
  names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
}
  
   text = paste0(round(data_wanted$prop,digits=2), '% (',data_wanted$Accessions,' accessions) ')
data_wanted$text = text
fig <- plot_ly(data_wanted, x = ~year, y = ~prop, color = ~RedListCategory, type = 'bar', text = ~text,
               hovertemplate = '%{text}',
                colors =  colo)
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text=''),  traceorder= 'reversed'),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)



if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-native_threatened_over_time_by_accessions.html'))
}

fig
}

```

`r if(report_kind == 'interactive' & have_redlist & do_over_time){"##### {-}"}`


```{r, get WCVP plants that were in the LC but no more}

df = report[which(!is.na(report$POWO_plant_name_id)),]
df = df[which(df$threatened == "Threatened"),]
df = df[which(!is.na(df$LossYear)),]


all_ids =  df$POWO_plant_name_id |> unique()
existing_ids = df$POWO_plant_name_id[df$ItemStatusType == 'Existing'] |> unique()
no_more_ids = all_ids[!all_ids %in% existing_ids]
no_more_info = data.frame(1,2,3,4,5,6,7,8,0)
names(no_more_info) = c('Taxa', 'Last year in collection', 'Total years in collection', 'Presence history', 'Threatened', 'Threatened category', 'Endemic', 'Naive', 'Tree')
no_more_info = no_more_info[FALSE,]
if(length(no_more_ids) > 0){
  column_interest = match(c('AccYear', 'LossYear', 'threatened', 'threatened_category', 'endemic', 'tree', 'native'), names(df))
no_more_info = pbapply::pblapply(no_more_ids, function(x){
  
  plant = df[df$POWO_plant_name_id == x, match(c("POWO_taxon_name", "POWO_taxon_authors"), names(df))][1,] |> paste0(collapse = ' ')
  info = df[df$POWO_plant_name_id == x, column_interest]
  last_seen = max(info$LossYear) ; first_seen = min(info$AccYear)
  all_years = first_seen:last_seen
  seen_years = apply(info, 1, function(x){x[1]:x[2]}) |> unlist() |> unique() |> sort()
  has_year = all_years %in% seen_years
  total_years = sum(has_year)
  consec = split(which(has_year),cumsum(c(1,diff(which(has_year))!=1)))
  
   mess = ''
  for(ii in 1:length(consec)){
    mess = 
    paste0(mess,all_years[min(consec[[ii]])], '-', all_years[max(max(consec[[ii]]), 1)], ', ' )
  }
  mess <- stringr::str_remove(mess,', $')
  
  # characteristics.
  threat = info$threatened[1] ; threat_cat = info$threatened_category[1]
  endemic = info$endemic[1] ; tree = info$tree[1] ; nativeo = info$native[1]
  return(c(plant, last_seen, total_years, mess, threat, threat_cat, endemic, nativeo, tree))
}) |> data.frame() |> t() |> data.frame()
names(no_more_info) = c('Taxa', 'Last year in collection', 'Total years in collection', 'Presence history', 'Threatened', 'Threatened category', 'Endemic', 'Naive', 'Tree')
rownames(no_more_info) = 1:nrow(no_more_info)
no_more_info$`Total years in collection` = no_more_info$`Total years in collection` |> as.numeric()
no_more_info$`Last year in collection` = no_more_info$`Last year in collection` |> as.numeric()

}

```


In `r collection`'s history there are `r nrow(no_more_info)` threatened taxa which are no longer present. These are outlined below, ordered by how recently the taxa was present at the living collection.

```{r,  no_more_info table}

no_more_info_table = no_more_info[order(no_more_info$`Last year in collection`, no_more_info$`Total years in collection`, decreasing = T),]
no_more_info_table = no_more_info_table[,c(1,2,3,4,6,7,8)]
no_more_info_table |>  DT::datatable(extensions = 'Buttons',
                  rownames = FALSE,
                  options = list(scrollX = TRUE,
                                 pageLength = min(nrow(no_more_info_table), 10),
                                 dom = 'Btp',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=T,
                                 ordering=T),
                  filter = list(position = "top"))
```


***
  
### Geographic coverage of threatened species

```{r check if we are given the number of accepted species or endemic species}
# if(!is.null(wcvp) & (is.null(accepted_species_per_region) | is.null(endemic_species_per_region))){
   geography_data = wcvp$geography[,grepl('_area_code_l3', names(wcvp$geography))]
  want = c('000','010','001','011','100','110','101')

  #Depending on the values of the options reduce `want` to only the satisfactory values.
  ##A) introduced / naturally occurring only.
  native = 'Introduced only'
  if(native == 'Introduced only'){
    want = want[grepl('^1',want)]
  }else if(native == 'Naturally occurring only'){
    want = want[grepl('^0',want)]
  }
  ## B) Extinct
  if(!extinct){
    want = want[!grepl('010|011|110',want)]
  }
  ## C) Doubtful
  if(!doubtful_locations){
    want = want[!grepl('001|011|101',want)]
  }
  geog_want_values = want
  # Get the columns in wanted info that contain the geography information we want.
  geography_data = geography_data[,grepl(paste0(geog_want_values,collapse='|'), names(geography_data))]

  # If multiple columns join the strings such that each contains only a single list of regions.
  if(length(geog_want_values) > 1){
    level3codes =do.call("paste", c(geography_data, sep = ", "))
    level3codes = stringr::str_remove(level3codes, ', NA$')
    level3codes = stringr::str_remove(level3codes, 'NA, ')
  }else{
    level3codes = geography_data
  }
  level3codes[level3codes == "NA"] = NA

  wcvp_geography = data.frame(id = wcvp$geography$plant_name_id,
                              codes = level3codes)

  #Next we need to extract all the plant name ids from wcvp that are  accepted
  plant_ids = wcvp$wcvp_names$plant_name_id[wcvp$wcvp_names$taxon_status %in% c('Accepted')& !is.na(wcvp$wcvp_names$species)]

  # reduce wcvp_geography to only accepted plant name ids. (Note that usually only accepted plants have geography so this won't actually remove many records)
  wcvp_geography = wcvp_geography[wcvp_geography$id %in% plant_ids,]
  if(!exists('accepted_species_per_region')){
    no_accepted_species_wcvp = unlist(lapply(All_locations, function(x){
    sum(grepl(x, wcvp_geography$codes))}))
    
    accepted_species_per_region = data.frame(code = All_locations, no_accepted_taxa_wcvp = no_accepted_species_wcvp)
  }
  

  # Reduce to only endemic plants.
  wcvp_geography_endemic = wcvp_geography[stringr::str_length(wcvp_geography$codes) == 3,]

    if(!exists('endemic_species_per_region')){
      no_endemic_accepted_species_wcvp = unlist(lapply(All_locations, function(x){
    sum(grepl(x, wcvp_geography_endemic$codes))
  }))
  endemic_species_per_region = data.frame(code = All_locations, no_endemic_accepted_species_wcvp = no_endemic_accepted_species_wcvp)
  }

# }
```

```{r threatened Species map -  data}
# Get all BRU level 3 location codes.
All_locations = wgsrpd3$LEVEL3_COD

wanted_info = threatened_existing
wanted_info_species = wanted_info[match(unique(wanted_info$best_name), wanted_info$best_name),]



#Get the number of speices for each location.
no_threatened_items = unlist(lapply(All_locations, function(x){
  sum(grepl(x, wanted_info$geography_codes))
}))
#Get the number of speices for each location.
no_threatened_species = unlist(lapply(All_locations, function(x){
  sum(grepl(x, wanted_info_species$geography_codes))
}))


# Combine the wanted information into location data.
location_data = data.frame(code =  wgsrpd3$LEVEL3_COD,
                           name = wgsrpd3$LEVEL3_NAM,
                           rep_threatened = no_threatened_items > 0,
                           no_threatened_items = no_threatened_items,
                           no_threatened_species = no_threatened_species
)

location_data$no_endemic_accepted_wcvp_species = endemic_species_per_region[,2]
location_data$no_accepted_wcvp_species = accepted_species_per_region[,2]

  

location_data$name = stringr::str_replace(location_data$name, 'Is\\.', 'Islands')
location_data$name = stringr::str_replace(location_data$name, 'I\\.', 'Island')

threat_match_IUCN = sum(!is.na(detailied_IUCN$POWO_plant_id))
prop_threat_match_IUCN = round(threat_match_IUCN/nrow(detailied_IUCN)*100,digits = 2)
# number of threatened per geographic area. 
threatened_per_region = unlist(lapply(location_data$code, function(code){
  sum(grepl(code, detailied_IUCN$geography))
}))
location_data$threatened_per_region = threatened_per_region
  prop = round(location_data$no_threatened_species / threatened_per_region*100,digits=2)
  prop[which(prop |> as.numeric() > 100)] = 100 # Set max to 100% even if we actually have more
  prop_true = prop
  prop_true[is.nan(prop_true)] = 'NA'
  prop[prop ==  0] = -0.01

without_threatened_regions =   location_data$name[threatened_per_region ==0]
if(!is.null(wcvp)){
  
  ## Which regions do we have 0 threatened items (when species exist in wcvp)
  with_threatened = sum(location_data$no_threatened_items > 0)
  missing_regions = location_data[which(location_data$no_threatened_items == 0 & location_data$threatened_per_region !=0), match(c('name', 'code', 'threatened_per_region'), names(location_data))]
  missing_codes = missing_regions$code
  missing_names = missing_regions$name
  number_no_threatened_regions= length(missing_codes)
  wgsrpd3$LEVEL1_COD[match(missing_regions$code,wgsrpd3$LEVEL3_COD)]
  level_1_names = level_1$name[match(wgsrpd3$LEVEL1_COD[match(missing_regions$code,wgsrpd3$LEVEL3_COD)], level_1$code)]
  level_2_names = level_2$name[match(wgsrpd3$LEVEL2_COD[match(missing_regions$code,wgsrpd3$LEVEL3_COD)], level_2$code)]
  missing_regions = data.frame(missing_regions$name, level_2_names, level_1_names, missing_regions$threatened_per_region)
  
  
  names(missing_regions) = c('Region', 'Level 2 region','Level 1 region', 'Number of Threatened Species')
  missing_regions = missing_regions[order(as.numeric(missing_regions$`Number of Threatened Species`), decreasing = T),]
   info <- wcvp_geography[grep(paste0(missing_codes, collapse = '|'),wcvp_geography$codes),]
   info = info[info$id %in% detailied_IUCN$POWO_plant_id,]
   ids = info$id
  region_wanted = rep('',nrow(info))
  for(i in 1:length(missing_codes)){
    region_wanted[grep(missing_codes[i],info$codes)] = paste0(region_wanted[grep(missing_codes[i],info$codes)], ', ',missing_names[i])
  }
  region_wanted = stringr::str_remove(region_wanted, pattern = '^, ')
  
  wcvp_wantedo = wcvp$wcvp_names[match(ids,wcvp$wcvp_names$plant_name_id) ,]
  wcvp_wantedo$web_link = paste0("https://powo.science.kew.org/taxon/urn:lsid:ipni.org:names:",wcvp_wantedo$powo_id)
  wcvp_wantedo$plant_native_to =region_wanted
  IUCN_indo = detailied_IUCN[match(wcvp_wantedo$plant_name_id, detailied_IUCN$POWO_plant_id),]
  wcvp_wantedo = data.frame(wcvp_wantedo, IUCN_indo)
  
    wcvp_wantedo = wcvp_wantedo[c('taxon_name', 'plant_native_to', 'web_link', "category", "population_trend", "threat_hist", "first_threat", 'powo_id', 'ipni_id', "family", "genus", "species_hybrid", "species",              "infraspecific_rank", "infraspecies", "parenthetical_author", "primary_author", "publication_author", "nomenclatural_remarks", "geographic_area", "lifeform_description", "climate_description")]
  
  wcvp_wantedo = wcvp_wantedo[order(wcvp_wantedo$taxon_name),]
}



plot_info <- wgsrpd3 |>
  #add the location data to the geometry data
  dplyr::left_join(location_data, by=c("LEVEL3_COD"="code"))

# Do we want to export the data into an excel spreadsheet and r data file.
if(export_data){
  l = location_data
  names(l) = c('BRU level 3 code',
                 'BRU level 3 name',
                 'Has Threatened Species in living collection',
                 'Number of Threatened Items',
                 'Number of Threatened Species')
  
  exporting_data_store$map_of_threatened_species = l
  rm(l)
}

percent_rep = round(sum(plot_info$rep_threatened) / length(plot_info$rep_threatened) *100,digits = 2)
no_rep = plot_info$LEVEL3_NAM[!plot_info$rep_threatened]

without_geog = unique(threatened_existing$best_name[which(is.na(threatened_existing$geography_codes))])
```

The maps in this section show which botanical countries are represented by threatened species in the collection, how many threatened species from each botanical country are in the collection, and the proportion of threatened species in the collection relative to the total number of threatened species in a botanical country.

<!-- Note that a threatened taxa may belong to multiple geographic regions. Therefore, a single taxa/item can contribute to the count of multiple regions on the map. As such if one was to add the number of items from all the regions this will be larger than the number of threatened items in the living collection.  -->

 In the living collection there are `r length(unique(threatened_existing$best_name[which(is.na(threatened_existing$geography_codes))]))` taxa that are classed as threatened without geographic information provided by WCVP, these can be viewed below.

<details>
<summary>Threatened plants in the living collection without geography</summary>
`r paste0(without_geog, collapse = ' <br>')`
</details>

```{r Get threatened without geography table, eval = FALSE}

threat_without_geography = threatened_existing[which(is.na(threatened_existing$geography_codes)),]
threat_without_geography$POWO_web_address[8]
if(report_kind == 'static'){
  threatened_count_table |>
    flextable() |>
    set_caption(caption = "Threatened plants in the living collection") |> 
    font(fontname = "Calibri (Body)", part = "all") |> 
    fontsize(size = table_font_size, part = "body") |> 
    # add footer if you want
    # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
    #                colwidths = 4) |> 
    theme_booktabs() |> # default theme
    autofit()
}
if(report_kind == 'interactive'){
  threatened_count_table|>  DT::datatable(extensions = 'Buttons',
                  rownames = FALSE,
                  options = list(scrollX = TRUE,
                                 pageLength = nrow(threatened_count_table),
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))
  # DT::datatable(threatened_count_table, rownames = FALSE)
}
```

```{r, Number threatened Species in the living collection - figure, fig.fullwidth=FALSE, fig.dim = c(10, 6), eval = report_kind == 'static'}
top10_name = plot_info$LEVEL3_NAM[order(plot_info$no_threatened_species, decreasing = TRUE)][1:10]
top10_value = plot_info$no_threatened_species[order(plot_info$no_threatened_species, decreasing = TRUE)][1:10]
top10 = data.frame(region = top10_name, no_threatened = top10_value)

###################################################################
# Plot of the representation of threatened species in each geographic region.
# Regions that have no threatened plants are set to NA.
plot_info$rep_threatened[plot_info$rep_threatened == TRUE] = 'Yes'
plot_info$rep_threatened[plot_info$rep_threatened == FALSE] = 'No'

plot_info$`Have Threatened` = plot_info$rep_threatened

p = ggplot(plot_info)+
  geom_sf(aes(fill=`Have Threatened`),  col="transparent")+
  stat_sf_coordinates(aes(col=`Have Threatened`), show.legend = FALSE)+
  coord_sf(expand=FALSE) + 
  ggtitle("Geography of Threatened Species in the living collection")

if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Map_representation_of_threatened_species.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

p
###################################################################
###################################################################
# Plot of the number of threatened species in each geographic region.
plot_info$no_threatened_species[plot_info$no_threatened_species == 0] = NA
p = ggplot(plot_info)+
  geom_sf(aes(fill=no_threatened_species),  col="transparent")+
  stat_sf_coordinates(aes(col=no_threatened_species))+
  scale_colour_distiller(direction=1, 
                         name="Number\nof Species", na.value = 'black') +
  scale_fill_distiller(direction=1,
                       name="Number\nof Species", na.value = 'black')+
  coord_sf(expand=FALSE) + 
  ggtitle("Number of Threatened Species")

if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Map_number_of_threatened_species.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

p

```

```{r, interactive map data setup, eval = report_kind == 'interactive'}
# Due to the high level of detail in wgsrpd3 we need to go to a simplified geometry for the interactive geometry plots. 
# load all geo areas
wgsrpd3_level3_simp = BGSmartR::wgsrpd3_level3_simp

# Reduce to only regions that have threatened species.
plot_int_info <- wgsrpd3_level3_simp |>
  dplyr::left_join(location_data, by=c("code"="code"))
plot_int_info$name = plot_int_info$name.x

data_g = st_geometry(plot_int_info)
centers = st_centroid(data_g)
AA = unlist(centers)
center_df = data.frame(long = AA[seq(1,length(AA),2)], lat = AA[seq(2,length(AA),2)])

# Convert geometry into a format accepted by plotly
mapp = sf::st_cast(wgsrpd3_level3_simp, "MULTIPOLYGON")
geo_sf = geojsonsf::sf_geojson(mapp)
data = rjson::fromJSON(geo_sf)
feat = data$features
for(i in 1:length(feat)){
  feat[[i]]$id = feat[[i]]$properties$code
}
data$features = feat

index = seq(0,1,0.01)
col_rep = rev(color_binary)
colours = c(rep(col_rep[1],51), rep(col_rep[2],50))
counter = 1:length(index)
col_scale_rep = lapply(counter, function(i){return(c(index[i], colours[i]))})

colours_cont = scales::brewer_pal(palette = palette, direction = 1)(7)
ramp <- scales::colour_ramp(colours_cont)
colos = ramp(seq(0, 1, length = 1001))


index = seq(0,1,0.001)
colours_cont = c( "#e6e6e6","#e6e6e6",colos)
counter = 1:length(index)
col_scale = lapply(counter, function(i){return(c(index[i], colours_cont[i]))})

plot_int_info$name = stringr::str_replace(plot_int_info$name, 'Is\\.', 'Islands')
plot_int_info$name = stringr::str_replace(plot_int_info$name, 'I\\.', 'Island')
```

`r if(report_kind == 'interactive'){"#####  {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Represented regions"}`
```{r  Map of the representation of threatened species - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
rep_threatened = rep('No', nrow(plot_int_info))
rep_threatened[plot_int_info$rep_threatened] = 'Yes'
plot_int_info$rep_threatened_yes_no = rep_threatened
# text = paste0('Region: ', plot_int_info$name, '<br />', 
#               'Represented: ', rep_threatened, '<br />',
#               'Species: ', plot_int_info$no_threatened_species, '<br />',
#               'Items: ', plot_int_info$no_threatened_items)
text = paste0('Region: ', plot_int_info$name, '<br>', 
              'Represented: ', rep_threatened, '<br>',
              'Taxa in LC: ', plot_int_info$no_threatened_species, '/',threatened_per_region, '<br>',
              'Items in LC: ', plot_int_info$no_threatened_items, '<br>',
              'Threatened from region in LC: ', prop_true, '%')
plot_int_info$hover = text

fig = plot_ly(plot_int_info,  text = ~hover, hoverinfo = 'text') |>
  add_trace(type="choroplethmapbox",
            geojson=data,
            locations=plot_int_info$code,
            z=as.numeric(plot_int_info$rep_threatened),
            colorscale=col_scale_rep,
            reversescale = FALSE,
            marker=list(line=list(
              width=0.01,color ='black'),
              text = plot_int_info$hover,
              hoverinfo = 'text'
            )) |>
  layout(
    mapbox=list(
      style="white-bg",
      center = list(lon = 0 ,lat= 60),
      zoom =0.46, 
      lataxis = list(range = c(-59, 90)))
  ) |>
  hide_colorbar()
# colorbar(title = "Species <br> Represented <br> in LC")  
fig = fig |>  add_trace(type="scattermapbox",
                        mode = 'markers',
                        data = center_df,
                        lon = center_df$long,
                        lat = center_df$lat,
                        text = plot_int_info$hover,
                        color = I(col_rep[as.numeric(plot_int_info$rep_threatened)+1]), 
                        hoverinfo = 'text',
                        inherit = FALSE, showlegend = FALSE,
                        marker = list(color = I(col_rep[as.numeric(plot_int_info$rep_threatened)+1]), line = list(width = 0)),
                        hoverlabel = list(bgcolor = I(col_rep[as.numeric(plot_int_info$rep_threatened)+1]))) 

fig <- fig   |> config(modeBarButtonsToRemove = c("select2d", "lasso2d"))


if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Geography-Representation_of_threatened_Species.html'))
}
suppressWarnings(fig)

```

`r if(report_kind == 'interactive'){"###### Number of threatened species held/region"}`
```{r  Map of the number of threatened species - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
rep_threatened = rep('No', nrow(plot_int_info))
rep_threatened[plot_int_info$rep_threatened] = 'Yes'

  plot_int_info$no_threatened_species_use = plot_int_info$no_threatened_species
  plot_int_info$no_threatened_species_use[plot_int_info$no_threatened_species == 0] = -1
  
  colscale_0_1 = as.numeric(unlist(lapply(col_scale, function(x){x[[1]]})))
  point_value = (plot_int_info$no_threatened_species_use+1)/max(plot_int_info$no_threatened_species_use+1)
  color_index = unlist(lapply(point_value, function(x){which.min(abs(x - colscale_0_1))}))
  fillcolor = unlist(lapply(color_index, function(x){col_scale[[x]][2]}))
  
  
  # fig = fig |>  add_trace(type="scattermapbox",
  #                         mode = 'markers',
  #                         data = center_df,
  #                         lon = center_df$long,
  #                         lat = center_df$lat,
  #                         text = plot_int_info$hover,
  #                         # fill = plot_int_info$no_threatened_species+1,
  #                         # fillcolor=col_scale,
  #                         hoverinfo = 'skip',
  #                         inherit = FALSE,
  #                         showlegend = FALSE,
  #                         marker = list(color = fillcolor, line = list(width = 0)))


  
  fig = plot_ly(plot_int_info,  text = plot_int_info$hover, hoverinfo = 'text') |>
      add_trace(type="scattermapbox",
                          mode = 'markers',
                          data = center_df,
                          lon = center_df$long,
                          lat = center_df$lat,
                          text = plot_int_info$hover,
                          # fill = plot_int_info$no_threatened_species+1,
                          # fillcolor=col_scale,
                          hoverinfo = 'text',
                          inherit = FALSE,
                          showlegend = FALSE,
                          marker = list(color = fillcolor, line = list(width = 0)),
                          hoverlabel = list(bgcolor = fillcolor)) |>
    add_trace(type="choroplethmapbox",
     geojson=data,
     locations=plot_int_info$code,
     z=plot_int_info$no_threatened_species_use+1,
     zmin = 0, zmax = max(plot_int_info$no_threatened_species_use),
     colorscale=col_scale,
     reversescale = FALSE,
     marker=list(line=list(
     width=0.01,color ='black')),
     text = plot_int_info$hover,
     hoverinfo = 'text',
     hoverlabel = list(bgcolor = fillcolor)
    ) |>
    layout(
    mapbox=list(
      style="white-bg",
      center = list(lon = 0 ,lat= 60),
      zoom =0.46,
      lataxis = list(range = c(-59, 90)))
  ) |>
    colorbar(title = "Number of <br> species held")


  fig <- fig   |> config(modeBarButtonsToRemove = c("select2d", "lasso2d")
                 
                         )

  if(separate_figure_folder){
    htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Geography-Number_of_threatened_Species.html'))
  }
  suppressWarnings(fig)

```

`r if(report_kind == 'interactive'){"###### % held relative to native threatened species"}`
```{r  Map of the proportion of threatened species - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}

colscale_0_1 = as.numeric(unlist(lapply(col_scale, function(x){x[[1]]})))
point_value = (prop+0.01)/max(prop+0.01,na.rm = T)
point_value[is.nan(point_value)] = 0
color_index = unlist(lapply(point_value, function(x){which.min(abs(x - colscale_0_1))}))
fillcolor = unlist(lapply(color_index, function(x){col_scale[[x]][2]}))
fillcolor[is.nan(prop)] = 'white'
  
  fig = plot_ly(plot_int_info,  text = ~hover, hoverinfo = 'text') |>
    add_trace(type="choroplethmapbox",
     geojson=data,
     locations=plot_int_info$code,
     z= prop+ 0.01,
     colorscale=col_scale,
     reversescale = FALSE,
     marker=list(line=list(
     width=0.01,color ='black')),
     text = plot_int_info$hover,
     hoverinfo = 'text',
    hoverlabel = list(bgcolor = fillcolor)
    ) |>
    layout(
    mapbox=list(
      style="white-bg",
      center = list(lon = 0 ,lat= 60),
      zoom =0.46,
      lataxis = list(range = c(-59, 90)))
  ) |>
    # hide_colorbar()
    colorbar(title = "Proportion of <br> threatened species <br> held")

  
  fig = fig |>  add_trace(type="scattermapbox",
                          mode = 'markers',
                          data = center_df,
                          lon = center_df$long,
                          lat = center_df$lat,
                          text = plot_int_info$hover,
                          hoverinfo = 'text',
                          inherit = FALSE,
                          showlegend = FALSE,
                          marker = list(color = fillcolor, line = list(width = 0)),
                           hoverlabel = list(bgcolor = fillcolor))

  fig <- fig   |> config(modeBarButtonsToRemove = c("select2d", "lasso2d")
                     )

  if(separate_figure_folder){
    htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Geography-Number_of_threatened_Species.html'))
  }
  suppressWarnings(fig)

```
 
`r if(report_kind == 'interactive'){"##### {-}"}`

***

### Duplication of threatened species

```{r, duplication of threatened species}
### How many items/accessions for each threatened taxa in the LC.
threatened_items_per_taxa = table(paste0(threatened_existing$POWO_taxon_name, ' ', threatened_existing$POWO_taxon_authors))
threatened_items_per_taxa_table = table(as.numeric(threatened_items_per_taxa))

threatened_items_per_taxa_simp = as.numeric(threatened_items_per_taxa)
threatened_items_per_taxa_simp[threatened_items_per_taxa_simp >= 5] = '>4'
threatened_items_per_taxa_table_simp = table(threatened_items_per_taxa_simp)

threatened_accessions = threatened_existing[match(unique(threatened_existing$AccNoFull), threatened_existing$AccNoFull),]
threatened_accessions_per_taxa = table(paste0(threatened_accessions$POWO_taxon_name, ' ', threatened_accessions$POWO_taxon_authors))
threatened_accessions_per_taxa_table = table(as.numeric(threatened_accessions_per_taxa))

threatened_accessions_per_taxa_simp = as.numeric(threatened_accessions_per_taxa)
threatened_accessions_per_taxa_simp[threatened_accessions_per_taxa_simp >= 5] = '>4'
threatened_accessions_per_taxa_table_simp = table(threatened_accessions_per_taxa_simp)

duplication_items = data.frame(threatened_items_per_taxa_table_simp)
names(duplication_items) = c('Duplication', 'Items')

duplication_items$percent = paste0(round(duplication_items$Items / sum(duplication_items$Items)*100,digits = 1),'%')
duplication_items <- duplication_items %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Items))), 
                pos = Items/2 + dplyr::lead(csum, 1),
                pos = dplyr::if_else(is.na(pos), Items/2, pos))


duplication_accessions = data.frame(threatened_accessions_per_taxa_table_simp)
names(duplication_accessions) = c('Duplication', 'Accessions')
# duplication_accessions$Provenance  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(duplication_accessions$Provenance, c('G', 'U', 'W', 'Z'))]
duplication_accessions$percent = paste0(round(duplication_accessions$Accessions / sum(duplication_accessions$Accessions)*100,digits = 1),'%')
duplication_accessions <- duplication_accessions %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Accessions))), 
                pos = Accessions/2 + dplyr::lead(csum, 1),
                pos = dplyr::if_else(is.na(pos), Accessions/2, pos))

flex_data = data.frame(duplication_items$Duplication,
                       duplication_items$Items, 
                       paste0(round(duplication_items$Items/sum(duplication_items$Items)*100,digits=2),'%'),
                       duplication_accessions$Accessions,
                       paste0(round(duplication_accessions$Accessions/sum(duplication_accessions$Accessions)*100,digits=2),'%'))
names(flex_data) = c('Quantity', 'Number of Threatened Species (by items)', 'Proportion of Threatened Species (by items)', 'Number of Threatened Species (by accessions)', 'Proportion of Threatened Species (by accessions)')

if(export_data){
  exporting_data_store$Duplication_of_existing_threatened_plants_in_the_LC = flex_data
}

```

In this section, we explore the duplication of existing threatened species in the living collection. 

Below we produce pie charts and a table of the number of threatened species that have 1, 2, 3, 4 or >4 items or accessions in the living collection. 

```{r Duplication of threatened plants in the living collection - Table}
flex_data  = flex_data[c(2:5,1),]
if(report_kind == 'static'){
  flex_data |>
    flextable() |>
    set_caption(caption = paste0('Duplication of existing threatened plants in the living collection')) |> 
    font(fontname = "Calibri (Body)", part = "all") |> 
    fontsize(size = table_font_size, part = "body") |> 
    # add footer if you want
    # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
    #                colwidths = 4) |> 
    theme_booktabs() |> # default theme
    separate_header() |> 
    align(align = "center", part = "all") |>
    autofit()
}


```


`r if(report_kind == 'static'){"Converting the table to pie charts yields "}`

```{r, Duplication of threatened plants in the living collection pie charts, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'static'}
# Basic piechart
pie1 = ggplot(duplication_items, aes(x="", y=Items, fill=Duplication)) +
  geom_bar(stat="identity", width=1, color="white") +
  ggtheme(base_size = 18) +
  coord_polar("y", start=0) +
  ggtheme(base_size = 18) +
  ggtitle('Items') +
  scale_fill_discrete() +
  theme(legend.position = "none",
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5))

pie2 = ggplot(duplication_accessions, aes(x="", y=Accessions, fill=Duplication)) +
  geom_bar(stat="identity", width=1, color="white") +
  ggtheme(base_size = 18) +
  coord_polar("y", start=0) +
  ggtitle('Accessions') + 
  scale_fill_discrete() +
  theme(legend.position = "none",
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5))

if(value_on_fig){
  pie1 = pie1 + geom_label_repel(aes(y = pos,label = percent),
                                 size = 4.5, nudge_x = 1, show.legend = FALSE, fill = "white")   
  
  pie2 = pie2 + geom_label_repel(aes(y = pos,label = percent),
                                 size = 4.5, nudge_x = 1, show.legend = FALSE, fill = "white") 
}

p = ggpubr::ggarrange(pie1, pie2, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")

if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Duplication_of_threatened_plants.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}


p

```

`r if(report_kind == 'interactive'){"#####  Duplication of threatened species in the living collection {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Chart"}`
```{r  Duplication of existing threatened items in the living collection - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 5), eval = report_kind == 'interactive'}
figs = list()

# Items
text = paste0(duplication_items$Items, ' Threatened taxa has ', duplication_items$Duplication, ' items')
fig1 <- plot_ly(duplication_items, labels = ~Duplication, values = ~Items, hoverinfo = "text",  hovertext = text, sort = FALSE, marker = list(colors = scales::brewer_pal(palette = palette, direction = 1)(5)))
fig1 <- fig1 |> add_pie(hole = 0.6)
fig1 <- fig1 |> layout(title = "Items",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
figs$items = fig1
#Accessions
text = paste0(duplication_accessions$Accessions, ' threatened taxa has', duplication_accessions$Duplication, ' accessions')
fig2 <- plot_ly(duplication_accessions, labels = ~Duplication, values = ~Accessions, hoverinfo = "text",  hovertext = text, sort = FALSE, marker = list(colors = scales::brewer_pal(palette = palette, direction = 1)(5)))
fig2 <- fig2 |> add_pie(hole = 0.6)
fig2 <- fig2 |> layout(title = "Accessions",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
figs$accessions = fig2
# duplication_items <<- duplication_items
# duplication_accessions <<- duplication_accessions

duplication_items = duplication_items[c(2:5,1),]
duplication_accessions = duplication_accessions[c(2:5,1),]
# Together
p <- plot_ly() |>
  add_pie(data = duplication_items, labels = ~Duplication, values = ~Items, hole = 0.6,
          title = "Items", name ='', sort = FALSE, marker = list(
            colors =  scales::brewer_pal(palette = palette, direction = 1)(5)),
          domain = list(row = 0, column = 0)) |>
  add_pie(data = duplication_accessions, labels = ~Duplication, values = ~Accessions, hole = 0.6,
          title = "Accessions", name ='', sort = FALSE, marker = list(
            colors = scales::brewer_pal(palette = palette, direction = 1)(5)),
         domain = list(row = 0, column = 1))  |>
  layout(title = "", showlegend = TRUE,
          grid=list(rows=1, columns=2),
         plot_bgcolor="#f0f0f0",
         legend = list(orientation = 'h', xanchor = "center", x = 0.5),
         xaxis = list(title=paste0('Items', 'Accessions'), showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         margin = c(0,0,0,0))


if(separate_figure_folder){
  htmlwidgets::saveWidget(figs$items, file = paste0(figures_dir, '/','threatened_duplication_existing_items.html'))
  htmlwidgets::saveWidget(figs$accessions, file = paste0(figures_dir, '/','threatened_duplication_existing_accessions.html'))
}

p
  

# 
# text = paste0(duplication_items$Items, ' threatened Taxa has ', duplication_items$Duplication, ' items')
# 
# fig <- plot_ly(duplication_items, labels = ~Duplication, values = ~Items, hoverinfo = "text",  hovertext = text, sort = FALSE)
# fig <- fig %>% add_pie(hole = 0.6)
# fig <- fig %>% layout(title = "",
#                       xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
#                       yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
# 
# fig
# 
# if(separate_figure_folder){
#   htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Pie-duplication_threatened_items.html'))
# }
# 

```

`r if(report_kind == 'interactive' & do_over_time){"###### Table"}`
```{r Duplication of existing threatened items in the living collection - table, , eval = report_kind == 'interactive'}
flex_data |>  DT::datatable(extensions = 'Buttons',
                  rownames = FALSE,
                  options = list(scrollX = TRUE,
                                 pageLength = nrow(flex_data),
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))
```

`r if(report_kind == 'interactive'){"##### {-}"}`

Next we look at how duplication has changed over time. 

```{r, duplication of threatened species over time}
over_time_info = lapply(threatened_existing_each_year, function(x){
  garden_current = threatened[x,]

  threatened_items_per_taxa = table(paste0(garden_current$POWO_taxon_name, ' ', garden_current$POWO_taxon_authors))
threatened_items_per_taxa_table = table(as.numeric(threatened_items_per_taxa))

threatened_items_per_taxa_simp = as.numeric(threatened_items_per_taxa)
threatened_items_per_taxa_simp[threatened_items_per_taxa_simp >= 5] = '>4'
threatened_items_per_taxa_table_simp = table(threatened_items_per_taxa_simp)
formatted = threatened_items_per_taxa_table_simp[match( c('1', '2', '3','4', '>4'), names(threatened_items_per_taxa_table_simp))]
formatted = as.numeric(formatted) ; formatted[is.na(formatted)] = 0

return(formatted)
}) |>
  data.frame() |>
  t() |>
  data.frame()
over_time_info = data.frame(Date = years, over_time_info)
names(over_time_info) = c('year','1', '2', '3', '4', '>4')
row.names(over_time_info) = 1:nrow(over_time_info)
duplication_items_over_time = over_time_info


over_time_info = lapply(threatened_existing_each_year, function(x){
  garden_current = threatened[x,]

threatened_accessions = garden_current[match(unique(garden_current$AccNoFull), garden_current$AccNoFull),]
threatened_accessions_per_taxa = table(paste0(threatened_accessions$POWO_taxon_name, ' ', threatened_accessions$POWO_taxon_authors))
threatened_accessions_per_taxa_table = table(as.numeric(threatened_accessions_per_taxa))

threatened_accessions_per_taxa_simp = as.numeric(threatened_accessions_per_taxa)
threatened_accessions_per_taxa_simp[threatened_accessions_per_taxa_simp >= 5] = '>4'
threatened_accessions_per_taxa_table_simp = table(threatened_accessions_per_taxa_simp)
formatted = threatened_accessions_per_taxa_table_simp[match( c('1', '2', '3','4', '>4'), names(threatened_accessions_per_taxa_table_simp))]
formatted = as.numeric(formatted) ; formatted[is.na(formatted)] = 0

return(formatted)

}) |>
  data.frame() |>
  t() |>
  data.frame()
over_time_info = data.frame(Date = years, over_time_info)
names(over_time_info) = c('year','1', '2', '3', '4', '>4')
row.names(over_time_info) = 1:nrow(over_time_info)
duplication_accessions_over_time = over_time_info


```

`r if(report_kind == 'interactive'& do_over_time){"#####  Duplication of threatened species over time {.tabset}"}`

`r if(report_kind == 'interactive'& do_over_time){"###### Items"}`
```{r  duplication of threatened items over time in the living collection - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive'& do_over_time}
prop_data = duplication_items_over_time
prop_data[,-1] = prop_data[,-1]/rowSums(prop_data[,-1])*100

dd = reshape(prop_data, idvar = "year", varying = list(2:ncol(prop_data)),
                   v.names = "count", timevar = "group", times = names(prop_data)[-1], direction = "long")
dd$group = factor(dd$group, levels = c('1', '2', '3', '4', '>4'))
prop_data = dd
text = paste0(round(prop_data$count,digits=2), '%')
prop_data$text = text
fig <- plot_ly(prop_data, x = ~year, y = ~count, color = ~group, type = 'bar', customdata = ~text,
               hovertemplate = '%{customdata}',  colors = scales::brewer_pal(palette = palette, direction = 1)(5))
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text='')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-threatened_duplication_over_time_by_item.html'))
}
```

`r if(report_kind == 'interactive'& do_over_time){"###### Accessions"}`
```{r  duplication of threatened accessions over time in the living collection - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive'& do_over_time}
prop_data = duplication_accessions_over_time
prop_data[,-1] = prop_data[,-1]/rowSums(prop_data[,-1])*100

dd = reshape(prop_data, idvar = "year", varying = list(2:ncol(prop_data)),
                   v.names = "count", timevar = "group", times = names(prop_data)[-1], direction = "long")
dd$group = factor(dd$group, levels = c('1', '2', '3', '4', '>4'))
prop_data = dd
text = paste0(round(prop_data$count,digits=2), '%')
prop_data$text = text
fig <- plot_ly(prop_data, x = ~year, y = ~count, color = ~group, type = 'bar', customdata = ~text,
               hovertemplate = '%{customdata}',  colors = scales::brewer_pal(palette = palette, direction = 1)(5))
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text='')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-threatened_duplication_over_time_by_accessions.html'))
}
```
`r if(report_kind == 'interactive'& do_over_time){"##### {-}"}`


***

### Rarity of threatened species 

```{r, rarity of threatened species}
### How many items/accessions for each threatened taxa in the living collection.
threatened_wanted_info = threatened_existing[c("AccNoFull","best_name", "no_gardens")]
threatened_wanted_info_og = threatened_wanted_info
# rarity_table = threatened_wanted_info |>
#   dplyr::group_by(AccNoFull)


threatened_wanted_info <- threatened_wanted_info |> 
        dplyr::group_by(best_name) |> 
        dplyr::summarise(global = no_gardens[1], 
                         count = length(no_gardens)) |>
  dplyr::ungroup() 

not_found = sum(is.na(threatened_wanted_info$global))
threatened_wanted_info$global[is.na(threatened_wanted_info$global)] = 0

rarity_plot_info <- threatened_wanted_info |> 
        dplyr::group_by(global) |> 
        dplyr::summarise(global = global[1], 
                         count = length(best_name)) |>
  dplyr::ungroup() 

# QQ = rarity_plot_info[rarity_plot_info$global %in% 11:50,]
unique_ = sum(rarity_plot_info$count[rarity_plot_info$global %in% 0:1])  
v_rare = sum(rarity_plot_info$count[rarity_plot_info$global %in% 2:10])  
rare = sum(rarity_plot_info$count[rarity_plot_info$global %in% 11:50]) 
common = sum(rarity_plot_info$count[rarity_plot_info$global %in% 51:100]) 
v_common = sum(rarity_plot_info$count[rarity_plot_info$global %in% 100:1500]) 

doughnut_data = data.frame(rarity = c('Unique','Very Rare', 'Rare', 'Common', 'Very Common'),
                           no_items = c(unique_,v_rare, rare, common, v_common),
                           # color = c('#f46d43','#41ab5d', '#78c679', '#addd8e', "#d9f0a3")),
                           color = scales::brewer_pal(palette = palette, direction = 1)(5) |> rev())
                           doughnut_data$percent = paste0(round(doughnut_data$no_items / sum(doughnut_data$no_items)*100,digits = 1),'%')


# data formatted if we want to export/save it.
globally_held_taxa_grouped = doughnut_data[,c(1,2,4,3)]
names(globally_held_taxa_grouped) = c("Rarity in global living collections",   "Number of Taxa in living collection", "Percentage",  "Colour" )

# For downloaded table
threatened_wanted_info_og$no_gardens[is.na(threatened_wanted_info_og$no_gardens)] = 0
to_download = threatened_wanted_info_og[threatened_wanted_info_og$no_gardens < 10,]

to_download = to_download |>
  dplyr::group_by(best_name) |> 
  summarise(acc = toString(unique(AccNoFull)),
            no_gardens = no_gardens[1]
            )

names(to_download) = c('Taxa', 'Accessions', 'Number of global collections')
```

In this section, we explore the rarity of existing threatened species in the living collection on the global scale. To do this we used BGCI's PlantSearch to extract the number of global collections plants are held within.

We group taxa by the number of collections they are found in globally. These are:

- **Unique**: plants found nowhere else (0 or 1 collection). These are names that do not match any name or match a name reported in only one collection in BGCI’s PlantSearch tool. Since it is not possible to know to which collection it belongs, BGsmartR assumes it belongs to yours and the name is in PlantSearch thanks to your input to that database. 
- **Very Rare**: plants found in 2-10 collections.
- **Rare**: plants found in 11-50 collections.
- **Common**: plants found in 51-100 collections.
- **Very Common**: plants found in more than 100 collections.

`r if(report_kind == 'interactive'){"#####  Rarity of threatened species {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Summary"}`
```{r, grouped rarity - interactive figure, fig.dim = c(10, 4), eval = report_kind == 'interactive'}
text = paste0(doughnut_data$no_items, ' records classed as ', doughnut_data$rarity )
doughnut_data$rarity = factor(doughnut_data$rarity, levels = doughnut_data$rarity)

fig <- plot_ly(doughnut_data, labels = ~rarity, values = ~no_items, hoverinfo = "text",  hovertext = text, marker = list(colors = ~color),  sort = FALSE )
fig <- fig %>% add_pie(hole = 0.6)
fig <- fig %>% layout(title = "",  showlegend = T,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Pie-grouped_global_rarity.html'))
}

fig

```

`r if(report_kind == 'interactive'& do_over_time){"###### Detailed"}`
```{r, taxa in collection against how many sites globally - interactive figure, fig.dim = c(10, 4), eval = report_kind == 'interactive'}
data_plot = rarity_plot_info
text = paste0(data_plot$count, ' threatened species in the garden are found in ', data_plot$global, ' collections globally.')
colo_rare = rev(scales::brewer_pal(palette = palette, direction = 1)(5))
fig <- plot_ly(data_plot, x = ~global, y = ~count, 
         type = 'scatter',  mode = 'markers', hoverinfo = "text",  hovertext = text, colors = 'black', color = 'color')
fig <- fig |> layout(title = "",
  xaxis = list(title = "Number of living collections <br> holding the taxa worldwide"),
  yaxis = list (title = "Number of threatened taxa <br> found in the collection")) 
fig <- layout(fig,
             shapes = list(
                list(type = "rect",
                    fillcolor = colo_rare[1], line = list(color = colo_rare[1]), opacity = 0.3,
                    x0 = 0, x1 = 1.5, xref = "x",
                    y0 = 0, y1 = max(data_plot$count), yref = "y"),
               list(type = "rect",
                    fillcolor = colo_rare[2], line = list(color = colo_rare[2]), opacity = 0.3,
                    x0 = 1.5, x1 = 10.5, xref = "x",
                    y0 = 0, y1 = max(data_plot$count), yref = "y"),
                list(type = "rect",
                    fillcolor = colo_rare[3], line = list(color = colo_rare[3]), opacity = 0.3,
                    x0 = 10.5, x1 = 50.5, xref = "x",
                    y0 = 0, y1 = max(data_plot$count), yref = "y"),
               list(type = "rect",
                    fillcolor = colo_rare[4], line = list(color = colo_rare[4]), opacity = 0.3,
                    x0 = 50.5, x1 = 100.5, xref = "x",
                    y0 = 0, y1 = max(data_plot$count), yref = "y"),
                list(type = "rect",
                    fillcolor = colo_rare[5], line = list(color = colo_rare[5]), opacity = 0.3,
                    x0 = 100.5, x1 = max(data_plot$global), xref = "x",
                    y0 = 0, y1 = max(data_plot$count), yref = "y")               
              ))

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','grown_in_LC_against_global.html'))
}

fig

```

`r if(report_kind == 'interactive'& do_over_time){"##### {-}"}`

:::: {.blackbox data-latex=""}

Note that unique/very rare taxa may suffer from inaccurate naming. These taxa can be downloaded using the button below

```{r}
to_download |>
  download_this(
    output_name = "unique_and_rare_threatened_taxa",
    output_extension = ".xlsx",
    button_label = "Download unique and rare threatened taxa (xlsx)",
    button_type = "default",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "hvr-sweep-to-left",
    self_contained =  T
  )
```
::::

***

### Provenance of threatened species
  
In this section we are exploring the provenance of native plants in the collection. Each accession in the garden has one of the following provenance types:

- Wild: plants that were collected directly from the wild,
- Wild-derived: cultivated plants descended from plants of known wild origin,
- Garden: cultivated plants descended from plants not of known wild origin,
- Unknown: plants with unknown origin.

Below we produce a chart and table of provenance for items and accessions.

```{r, Provenance of threatened plants in the living collection,  results='hide',fig.keep='all', message=FALSE}
# threatened_existing the data of existing threatened plants in the living collection. 
wanted_order = c('Wild', 'Wild-derived', 'Garden', 'Unknown')
prov_items = data.frame(table(threatened_existing$ProvenanceCode))
rows = match(wanted_order, prov_items[,1])
other = which(!prov_items[,1] %in% wanted_order)
o = c(rows,other)
o = o[!is.na(o)]
prov_items = prov_items[o,]
names(prov_items) = c('Provenance', 'Items')
# prov_items$Provenance  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(prov_items$Provenance, c('G', 'U', 'W', 'Z'))]  
prov_items$percent = paste0(round(prov_items$Items / sum(prov_items$Items)*100,digits = 1),'%')
prov_items <- prov_items %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Items))), 
                pos = Items/2 + dplyr::lead(csum, 1),
                pos = dplyr::if_else(is.na(pos), Items/2, pos))


prov_accessions = data.frame(table(threatened_existing$ProvenanceCode[match(unique(threatened_existing$AccNoFull), threatened_existing$AccNoFull)]))
rows = match(wanted_order, prov_accessions[,1])
other = which(!prov_accessions[,1] %in% wanted_order)
o = c(rows,other)
o = o[!is.na(o)]
prov_accessions = prov_accessions[o,]
names(prov_accessions) = c('Provenance', 'Accessions')
# prov_accessions$Provenance  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(prov_accessions$Provenance, c('G', 'U', 'W', 'Z'))]
prov_accessions$percent = paste0(round(prov_accessions$Accessions / sum(prov_accessions$Accessions)*100,digits = 1),'%')
prov_accessions <- prov_accessions %>% 
  dplyr::mutate(csum = rev(cumsum(rev(Accessions))), 
                pos = Accessions/2 + dplyr::lead(csum, 1),
                pos = dplyr::if_else(is.na(pos), Accessions/2, pos))

flex_data = data.frame(prov_items$Provenance,
                       prov_items$Items, 
                       paste0(round(prov_items$Items/sum(prov_items$Items)*100,digits=2),'%'),
                       prov_accessions$Accessions,
                       paste0(round(prov_accessions$Accessions/sum(prov_accessions$Accessions)*100,digits=2),'%'))
names(flex_data) = c('Provenance', 'Number of items', 'Proportion of items', 'Number of accessions', 'Proportion of accessions')

if(export_data){
  exporting_data_store$Provenance_of_existing_threatened_plants_in_the_LC = flex_data
}

### Create provenance by taxa to see if any have multiple provenance codes. 
taxa_prov_og = threatened_existing |>
  dplyr::group_by(best_name) |>
  summarise(wild_items = sum(.data$ProvenanceCode == 'Wild'),
            z_items = sum(.data$ProvenanceCode == "Wild-derived"),
            garden_items = sum(.data$ProvenanceCode == 'Garden'),
            unknown_items = sum(.data$ProvenanceCode == 'Unknown'),
            original_name = toString(.data$TaxonNameFull |> unique())
            )

taxa_prov = taxa_prov_og[,c(1:5)] |> as.data.frame()
for(i in 2:5){taxa_prov[,i] = taxa_prov[,i] |> as.numeric()}

wild_taxa = taxa_prov$best_name[taxa_prov$wild_items > 0]
z_taxa = taxa_prov$best_name[taxa_prov$z_items > 0]
g_taxa = taxa_prov$best_name[taxa_prov$garden_items > 0]
u_taxa = taxa_prov$best_name[taxa_prov$unknown_items > 0]
x = list() ; category.names = NULL ; colours = NULL
if(length(wild_taxa) > 0 ){x$w = wild_taxa ; category.names = c(category.names, "Wild") ; colours = c(colours, interactive_colour(4)[c(1)])}
if(length(u_taxa) > 0 ){x$u = u_taxa ; category.names = c(category.names, "Unknown") ; colours = c(colours, interactive_colour(4)[c(4)])}
if(length(z_taxa) > 0 ){x$z = z_taxa ; category.names = c(category.names, "Wild-derived"); colours = c(colours, interactive_colour(4)[c(2)])}
if(length(g_taxa) > 0 ){x$g = g_taxa ; category.names = c(category.names, "Garden"); colours = c(colours, interactive_colour(4)[c(3)])}
library(VennDiagram)
venn_diag = venn.diagram(x, filename = NULL,
                         category.names = category.names,
                         fill =colours,
                         disable.logging = TRUE,
                         main.fontfamily = "arial",
                         main.fontface = "sans")


# Make a data.frame of the overlapping taxa for a download.
taxa_prov_og$no_val =  lapply(taxa_prov_og[,-6] |> as.data.frame() |> t() |> as.data.frame(),function(x){
  sum(as.numeric(x[-1]) != 0)
}) |> unlist()

taxa_prov = taxa_prov_og[taxa_prov_og$no_val > 1,]
taxa_prov = taxa_prov[,-7]
taxa_prov =taxa_prov[,c(1,6,2,3,4,5)]
names(taxa_prov) = c("Best name found by BGSmartR",
                     "Original Taxonomic name/s",
                     "Number of Wild Origin Items",
                     "Number of Wild-derived Origin Items",
                     "Number of Garden Origin Items",
                     "Number of Unknown Origin Items")


### (END) Create provenance by taxa to see if any have multiple provenance codes. 

```

```{r Provenance of threatened plants in the LC - Table}
if(report_kind == 'static'){
  flex_data |>
    flextable() |>
    set_caption(caption = paste0('Provenance of existing threatened plants in the living collection')) |> 
    font(fontname = "Calibri (Body)", part = "all") |> 
    fontsize(size = table_font_size, part = "body") |> 
    # add footer if you want
    # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
    #                colwidths = 4) |> 
    theme_booktabs() |> # default theme
    separate_header() |> 
    align(align = "center", part = "all") |>
    autofit()
}
# if(report_kind == 'interactive'){
#  DT::datatable(flex_data, rownames = FALSE,
#                 # caption = paste0('Provenance of existing threatened plants in the LC'),
#                 options = list(dom = 't'))
# }

```


```{r, Provenance of threatened plants in the LC pie charts, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'static'}
# Basic piechart
pie1 = ggplot(prov_items, aes(x="", y=Items, fill=Provenance)) +
  geom_bar(stat="identity", width=1, color="white") +
  ggtheme(base_size = 18) +
  coord_polar("y", start=0) +
  ggtheme(base_size = 18) +
  ggtitle('Items') +
  scale_fill_discrete() +
  theme(legend.position = "none",
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5))

pie2 = ggplot(prov_accessions, aes(x="", y=Accessions, fill=Provenance)) +
  geom_bar(stat="identity", width=1, color="white") +
  ggtheme(base_size = 18) +
  coord_polar("y", start=0) +
  ggtitle('Accessions') + 
  scale_fill_discrete() +
  theme(legend.position = "none",
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
        plot.title = element_text(hjust = 0.5))

if(value_on_fig){
  pie1 = pie1 + geom_label_repel(aes(y = pos,label = percent),
                                 size = 4.5, nudge_x = 1, show.legend = FALSE, fill = "white")   
  
  pie2 = pie2 + geom_label_repel(aes(y = pos,label = percent),
                                 size = 4.5, nudge_x = 1, show.legend = FALSE, fill = "white") 
}

p = ggpubr::ggarrange(pie1, pie2, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")

if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Provenance_of_threatened_plants.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}


p

```

`r if(report_kind == 'interactive'){"##### {.tabset}"}`

`r if(report_kind == 'interactive' & do_over_time){"###### Chart"}`
```{r  Provenance of existing threatened items in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 5), eval = report_kind == 'interactive'}
figs = list()
colo = interactive_colour(4)
names(colo) =c('Wild', 'Wild-derived', 'Garden', 'Unknown')
colo = colo[match(prov_items$Provenance,c('Wild', 'Wild-derived', 'Garden', 'Unknown'))]

# Items
text = paste0(prov_items$Items, ' ',prov_items$Provenance,' origin threatened items')
fig1 <- plot_ly(prov_items, labels = ~Provenance, values = ~Items, hoverinfo = "text",  hovertext = text, sort = FALSE, marker = list(colors = colo))
fig1 <- fig1 |> add_pie(hole = 0.6)
fig1 <- fig1 |> layout(title = "Items",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
figs$items = fig1
#Accessions
text = paste0(prov_accessions$Accessions, ' ',prov_accessions$Provenance,' origin threatened items')
fig2 <- plot_ly(prov_accessions, labels = ~Provenance, values = ~Accessions, hoverinfo = "text",  hovertext = text, sort = FALSE, marker = list(colors = colo))
fig2 <- fig2 |> add_pie(hole = 0.6)
fig2 <- fig2 |> layout(title = "Accessions",
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
figs$accessions = fig2

# Together
p <- plot_ly() |>
  add_pie(data = prov_items, labels = ~Provenance, values = ~Items, hole = 0.6,
          title = "Items", name ='', sort = FALSE, marker = list(colors = colo),
          domain = list(row = 0, column = 0)) |>
  add_pie(data = prov_accessions, labels = ~Provenance, values = ~Accessions, hole = 0.6,
          title = "Accessions", name ='', sort = FALSE, marker = list(colors = colo),
         domain = list(row = 0, column = 1))  |>
  layout(title = "", showlegend = TRUE,
          grid=list(rows=1, columns=2),
         plot_bgcolor="#f0f0f0",
         legend = list(orientation = 'h', xanchor = "center", x = 0.5),
         xaxis = list(title=paste0('Items', 'Accessions'), showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         margin = c(0,0,0,0))


if(separate_figure_folder){
  htmlwidgets::saveWidget(figs$items, file = paste0(figures_dir, '/','threatened_provenance_existing_items.html'))
  htmlwidgets::saveWidget(figs$accessions, file = paste0(figures_dir, '/','threatened_provenance_existing_accessions.html'))
}

p
  
```

`r if(report_kind == 'interactive' & do_over_time){"###### Table"}`
```{r Provenance of existing threatened items in the LC - table, eval = report_kind == 'interactive'}
flex_data |>  DT::datatable(extensions = 'Buttons',
                  rownames = FALSE,
                  options = list(scrollX = TRUE,
                                 pageLength = nrow(flex_data),
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))

```

`r if(report_kind == 'interactive'& do_over_time){"##### {-}"}`

`r if(do_over_time){"We will now look at how the provenance of threatened plants in the living collection has changed over time."}`

`r if(!do_over_time){"We do not have the information to explore how provenance of threatened plants in the living collection has changed over time."}`

```{r, Provenance of threatened plants in the LC- change over time data, eval = do_over_time}
# For each year get the number in each group (species and genera)
over_time_info = lapply(threatened_existing_each_year, function(x){
  garden_current = threatened[x,]

  prov_items =  data.frame(table(garden_current$ProvenanceCode))
  if(nrow(prov_items) == 0){
    return(list(prov_items = prov_items,
                prov_accessions = prov_items))
  }
  names(prov_items) = c('Provenance', 'Items')

  prov_accessions = data.frame(table(garden_current$ProvenanceCode[match(unique(garden_current$AccNoFull), garden_current$AccNoFull)]))
  names(prov_accessions) = c('Provenance', 'Accessions')

  return(list(prov_items = prov_items,
              prov_accessions = prov_accessions))
})
prov_threatened_over_time = over_time_info

```

```{r, Provenance of threatened plants in the LC- change over time figures, fig.fullwidth=FALSE, fig.dim = c(10, 6), , eval = report_kind == 'static' & do_over_time}
to_do = c('items', 'accessions')
for(i in 1:length(to_do)){
  # Get the data from over_time_info.
  data_wanted = NULL
  wanted_item = which(grepl('prov',names(prov_threatened_over_time[[1]])) & grepl(to_do[i],names(prov_threatened_over_time[[1]])) )
  for(jj in 1:length(prov_threatened_over_time)){
    info_year = prov_threatened_over_time[[jj]][[wanted_item]]
    if(nrow(info_year) == 0){
      next
    }
    prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
    info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)),
                           info_year = info_year,
                           prop = prop)
    data_wanted = rbind(data_wanted, info_year)
    
  }
  names(data_wanted) = c('year', 'Provenance', 'Items', 'prop')
  data_wanted[,2] = as.character(data_wanted[,2])
  # data_wanted[,2]  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(data_wanted[,2], c('G', 'U', 'W', 'Z') )]
  data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))
  
  
  if(export_data){
    item_name = paste0('Provenance_of_threatened_plants_in_the_LC_proportion_over_time_',to_do[i])
    exporting_data_store$AA = data_wanted
    names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
  }
  
  p = ggplot(data_wanted, aes(fill=Provenance, y=prop, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) +
    # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    labs(title=paste0("Provenance of ", to_do[i] , " that are threatened in the living collection over time"),
         x ="Year",
         y = "Percentage (%)") +
    guides(fill=guide_legend(title=paste0('Provenance')))# +
  # theme(text = element_text(size=table_font_size))
  
  if(separate_figure_folder){
    ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Provenance_of_threatened_over_time', to_do[i] ,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
  }
  print(p)
  
}
```

`r if(report_kind == 'interactive'& do_over_time){"#####  {.tabset}"}`

`r if(report_kind == 'interactive'& do_over_time){"###### Items"}`
```{r  Provenance of threatened items over time in the living collection - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive'& do_over_time}
wanted_order = c('Wild', 'Wild-derived', 'Garden', 'Unknown')
to_do = c('items', 'accessions')
i=1
colo = interactive_colour(4)
names(colo) =c('Wild', 'Wild-derived', 'Garden', 'Unknown')

# Get the data from over_time_info.
data_wanted = NULL
wanted_item = which(grepl('prov',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
for(jj in 1:length(over_time_info)){
  info_year = over_time_info[[jj]][[wanted_item]]
  if(nrow(info_year) == 0){
    next
  }
  rows = match(wanted_order, info_year[,1])
  other = which(!info_year[,1] %in% wanted_order)
  o = c(rows,other)
  o = o[!is.na(o)]
  info_year = info_year[o,]
  prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
  info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)),
                         info_year = info_year,
                         prop = prop)
  data_wanted = rbind(data_wanted, info_year)
  
}
names(data_wanted) = c('year', 'Provenance', 'Items', 'prop')
data_wanted[,2] = as.character(data_wanted[,2])
# data_wanted[,2]  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(data_wanted[,2], c('G', 'U', 'W', 'Z') )]
data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))


if(export_data){
  item_name = paste0('Provenance_of_threatened_plants_in_the_LC_proportion_over_time_',to_do[i])
  exporting_data_store$AA = data_wanted
  names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
}


text = paste0(round(data_wanted$prop,digits=2), '% (',data_wanted$Items,' items) ')
data_wanted$text = text
data_wanted$Provenance = factor(data_wanted$Provenance, levels= unique(data_wanted$Provenance))
fig <- plot_ly(data_wanted, x = ~year, y = ~prop, color = ~Provenance, type = 'bar', text = ~text,
               hovertemplate = '%{text}',  colors = colo,
               sort = FALSE)
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text='')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-threatened_provenance_over_time_by_item.html'))
}
```

`r if(report_kind == 'interactive'& do_over_time){"###### Accessions"}`
```{r  Provenance of threatened accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive'& do_over_time}
wanted_order = c('Wild', 'Wild-derived', 'Garden', 'Unknown')
to_do = c('items', 'accessions')
i=2
colo = interactive_colour(4)
names(colo) =c('Wild', 'Wild-derived', 'Garden', 'Unknown')

# Get the data from over_time_info.
data_wanted = NULL
wanted_item = which(grepl('prov',names(over_time_info[[1]])) & grepl(to_do[i],names(over_time_info[[1]])) )
for(jj in 1:length(over_time_info)){
  info_year = over_time_info[[jj]][[wanted_item]]
  if(nrow(info_year) == 0){
    next
  }
  rows = match(wanted_order, info_year[,1])
  other = which(!info_year[,1] %in% wanted_order)
  o = c(rows,other)
  o = o[!is.na(o)]
  info_year = info_year[o,]
  prop = round(info_year[,2] / sum(info_year[,2]) *100,digits = 3)
  info_year = data.frame(year = rep(names(over_time_info)[jj], nrow(info_year)),
                         info_year = info_year,
                         prop = prop)
  data_wanted = rbind(data_wanted, info_year)
  
}
names(data_wanted) = c('year', 'Provenance', 'Items', 'prop')
data_wanted[,2] = as.character(data_wanted[,2])
# data_wanted[,2]  = c('Garden', 'Unknown', 'Wild', 'Wild-dervived')[match(data_wanted[,2], c('G', 'U', 'W', 'Z') )]
data_wanted$year = as.numeric(stringr::str_extract(data_wanted$year,'[0-9]{4}'))


if(export_data){
  item_name = paste0('Provenance_of_threatened_plants_in_the_LC_proportion_over_time_',to_do[i])
  exporting_data_store$AA = data_wanted
  names(exporting_data_store)[names(exporting_data_store) == 'AA'] = item_name
}


text = paste0(round(data_wanted$prop,digits=2), '% (',data_wanted$Items,' accessions) ')
data_wanted$text = text
data_wanted$Provenance = factor(data_wanted$Provenance, levels= unique(data_wanted$Provenance))
fig <- plot_ly(data_wanted, x = ~year, y = ~prop, color = ~Provenance, type = 'bar', text = ~text,
               hovertemplate = '%{text}', colors = colo,
               sort = FALSE)
fig <- fig |> layout(yaxis = list(title = 'Percentage', ticksuffix  ='%'),
                     xaxis = list(title = "Year", hoverformat = 'Year: '),
                     legend=list(title=list(text='')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)
fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-threatened_provenance_over_time_by_accession.html'))
}
```

`r if(report_kind == 'interactive'& do_over_time){"##### {-}"}`

In the above analysis we have considered threatened items and accessions where provenance is unique (i.e each accession can only be in one provenance group). Next we look at provenance across taxa. In this case a taxa can be contained in the garden by multiple provenance codes (i.e. one accession of wild origin and one accession of garden origin). To visualise the provenance by taxa we use a Venn diagram visualise how taxa can be expressed by multiple provenances. 

```{r, draw provenace by taxa (venn), results='hide',fig.keep='all', message=FALSE, fig.dim=c(4,4), fig.align = 'center', eval = report_kind == 'interactive' }
grid.newpage()
grid.draw(venn_diag)
```


:::: {.blackbox data-latex=""}

Taxa with multiple provenance types can be downloaded below

```{r}
taxa_prov |>
  download_this(
    output_name = "Taxa with multiple provenances",
    output_extension = ".xlsx",
    button_label = "Download taxa with multiple provenances (xlsx)",
    button_type = "default",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "hvr-sweep-to-left",
    self_contained =  T
  )
```
::::


***
  
### Turnover analysis
  
This section looks at the turnover (gain/loss) of accessions of threatened plants in the living collection.

`r if(!do_over_time){"As we do not have information to explore the data over time we cannot perform turnover analysis."}`

`r if(do_over_time){"We can also view the gain/loss of threatened accessions as a proportion of all gained/lost threatened each year. This will show if the incoming proportion of threatened plants was more or less than the proportion of threatened plants in the collection."}`


```{r, threatened Turnover - Accessions, eval = do_over_time}
# Start from the original_report
max_year = as.numeric(format(Sys.Date(),'%Y'))
years = min_year:max_year
report = report_original

# Set old unknown records to have a date ( = earliest_allowable_record) and remove all records prior to this time.
if(!is.null(old_accession_year_codes)){
  has_old_accessions_code = which(report$AccYear %in% old_accession_year_codes)
  no_old_accessions = length(has_old_accessions_code)
  report$AccYear[has_old_accessions_code] = earliest_allowable_record
}
to_keep = which(as.numeric(report$AccYear) >= earliest_allowable_record  & as.numeric(report$AccYear) <= max_year)
still_issue = nrow(report) - length(to_keep)
report = report[to_keep,]


# report = report[report$AccYear > 1650 & report$AccYear<=max_year,]
report = report[!is.na(report$ItemStatusDate),]

## Convert report into accessions
unique_accessions = unique(report$AccNoFull)
LossYear = rep(NA,nrow(report))
ItemStatusYear = as.numeric(stringr::str_extract(report$ItemStatusDate,pattern = '[0-9]{4}'))
LossYear[report$ItemStatusType %in% c('NotExisting', 'Not Existing')] = ItemStatusYear[report$ItemStatusType %in% c('NotExisting', 'Not Existing')]
report$LossYear = LossYear
# Order the items where the lower indices have the most recent death date.
LossYear_dummy = LossYear
LossYear_dummy[is.na(LossYear_dummy)] = 4000
ordered_items = order(LossYear_dummy,decreasing = T)
report = report[ordered_items, ]
# Match to the ordered accessions.
match_to_best = match(unique_accessions,report$AccNoFull)
#Reduce the data to unique accessions with most recent death date.
report = report[match_to_best,]


#Get the number of accessions existing each year for whole collection/
existing_each_year_LC = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                                                                   AccessionYear = as.character(report$AccYear),
                                                                   ItemStatusDate = as.character(report$ItemStatusDate),
                                                                   ItemStatusType = as.character(report$ItemStatusType),
                                                                   post_date = '3000-01-01')))
## Get the turnover for the whole collection.
overall_turnover = turnover_items(report)
exporting_data_store$overall_turnover = overall_turnover

## Get the trends for threatened accessions.
date = paste0(years, '-12-31')
plant_existing = BGSmartR::exist_at_date(date, AccessionYear = report$AccYear,
                                         ItemStatusDate = report$ItemStatusDate,
                                         ItemStatusType = report$ItemStatusType,
                                         post_date = '3000-01-01')
names(plant_existing) = date
trends = pbapply::pblapply(plant_existing, function(x){
  garden_current = report[x,]
  
  accessions = nrow(garden_current)
  
  breakdown_threatened = table(garden_current$threatened)
  
  return(list(accessions = accessions, breakdown_threatened = breakdown_threatened))
})
names(trends) = names(plant_existing)
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_threatened[!grepl(' ', names(x$breakdown_threatened))])
})))


do_type = TRUE
report_cur = report[which(report$threatened == 'Threatened'),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  
  
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_native = turnover
  
  plots = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'threatened', data_type = 'Accessions')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                                                                  AccessionYear = report_cur$AccYear,
                                                                  ItemStatusDate = report_cur$ItemStatusDate,
                                                                  ItemStatusType = report_cur$ItemStatusType,
                                                                  post_date = '3000-01-01')))
  collection_proportion = existing_each_year/ existing_each_year_LC
  
  plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'Threatened',
                                                   data_type = 'Accessions')
  exporting_data_store$proportional_turnover_native_gain = plots_proportional$data_gain
  exporting_data_store$proportional_turnover_native_loss = plots_proportional$data_loss
}
```


`r if(!do_type){"There are no records that are threatened."}`

```{r, Turnover of threatened items - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6)}

# plots$gain_loss
# 
# plots$net
# 
# plots_accessions$gain_loss
# 
# plots_accessions$net


```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of threatened {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Net"}`
```{r  Gain/Loss of threatened items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$trend_and_net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss"}`
```{r  Net turnover of threatened items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(!do_type){"We now explore how the gain and loss of threatened species compare proportionally to the total number of gained and lost items/accessions and the overall proportion of threatened species each year. To calculate the proportion of the collection that are threatened each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of threatened - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
# plots_proportional$gain
# 
# plots_proportional$loss
# 
# plots_proportional_accessions$gain
# 
# plots_proportional_accessions$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional turnover of threatened {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain"}`
```{r  Proportional Gain of threatened items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss"}`
```{r  Proportional Loss of threatened items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

***

### Survival analysis

```{r, survival analysis, echo = FALSE}
# Remove those without accession year.
do_survival = TRUE
if(!do_over_time){
  do_survival = FALSE
}

if(!is.null(old_accession_year_codes)){
  has_old_accessions_code = which(enriched_report$AccYear %in% old_accession_year_codes)
  no_old_accessions = length(has_old_accessions_code)
  enriched_report$AccYear[has_old_accessions_code] = earliest_allowable_record
}

enriched_report_wanted = enriched_report[as.numeric(enriched_report$AccYear) > 1650 & as.numeric(enriched_report$AccYear) < as.numeric(format(Sys.Date(),'%Y')),]
enriched_report_wanted = enriched_report_wanted[!is.na(enriched_report_wanted$ItemStatusDate),]
enriched_report_wanted = enriched_report_wanted[!is.na(enriched_report_wanted$AccYear),]

# Create a data-frame of the information we want.
wanted_data = data.frame(enriched_report_wanted$TaxonNameFull, enriched_report_wanted$AccYear, enriched_report_wanted$ItemStatusDate, enriched_report_wanted$ItemStatusType)
names(wanted_data) = c('TaxonNameFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType')


if(nrow(wanted_data) == 0){
  do_survival = FALSE
}
if(do_survival){
  #Set accession date (just chose the first of June)
pre_date = rep(NA, length(wanted_data$AccYear))
pre_date = paste(wanted_data$AccYear, "06", "01", sep = "-")

# Convert ItemStatusDate into all dates (i.e if month or day missing add one)
post_date = rep(as.character(Sys.Date()), length(wanted_data$AccYear))
post_date[which(wanted_data$ItemStatusType == "NotExisting")] = as.character(wanted_data$ItemStatusDate[which(wanted_data$ItemStatusType == 
  "NotExisting")])

post_date[which(stringr::str_length(post_date) == 4)] = paste(post_date[which(stringr::str_length(post_date) == 
  4)], "12", "31", sep = "-")
post_date[which(stringr::str_length(post_date) == 7)] = paste(post_date[which(stringr::str_length(post_date) == 
  7)], "28", sep = "-")

# Create the time difference between accession and ItemStatusDate.
time = as.numeric(as.Date(post_date) - as.Date(pre_date)) /365.25

# Create status to tell whether the last item status date was item death or just a check.
status = rep(0,length(wanted_data$AccYear))
status[wanted_data$ItemStatusType == 'NotExisting'] = 1

wanted_data = data.frame(wanted_data, AccessionDate = pre_date,
                         StatusDate = post_date,
                         time = time,
                         Alive_dead = status)


# We haven't removed those with missing ItemStatusDate, these will most likely have negative time so remove these.
missing_itemstatusdate_index = which(wanted_data$time > 0)
status_year = wanted_data$ItemStatusDate |> stringr::str_extract('^[0-9]{4}') |> unlist() |> as.numeric()
status_before_current_year = which(status_year <= format(Sys.Date(),'%Y') |> as.numeric() )
to_keep = intersect(missing_itemstatusdate_index,status_before_current_year)
wanted_data = wanted_data[to_keep,]

enriched_report_wanted = enriched_report_wanted[to_keep,]


## 1) Native or non-native (or doesn't have geography info.)
#Get which records are native.
native_non_native = enriched_report_wanted$threatened

data_to_analyse = data.frame(wanted_data$TaxonNameFull,
                             wanted_data$time,
                             wanted_data$Alive_dead,
                             native_non_native)
names(data_to_analyse) = c("TaxonNameFull", "time", "Alive_dead", "threatened_status")

if(export_data){
  exporting_data_store$Data_for_survival = data_to_analyse
}
}

```

In this section, we look at whether the survival period of threatened plants is longer than non-threatened plants.

Below we present a survival plot (Kaplan-Meier) to show how survival probability changes over time.

<!-- Note that plants can be long-lived and most collections will have plants that are older than the collection itself. Therefore, we need to lean into tools from statistics to analyse the lifespan of plants in a collection. Herein, we use survival analysis to calculate the probability that plants will survive for any number of years.  -->

`r if(FALSE){"Note that \"lifespan\" may be a preferable way to describe this section, however we use survival because accessioned plants may arrive in the garden already a few years old. As we will be using the accession date and item status date to determine the length of time plants have been in the collection this might not align perfectly with the age of the plant. Moreover, it is important to be aware that the data used in the analysis is not \"perfect\" in the sense that plants might have already died a large period of time prior to an audit, moreover this analysis doesn't account for circumstances of plant death which could be natural, or for renovation of a section of the garden, etc. "}`

<!-- In the following analysis, we assume that existing items are still existing to the present day. Therefore, even if the last item status date was a year ago if the plant was found to be existing then we assume it is still existing now. (I could change this to just use alive at Item status date, it won't make much difference. ) -->

<!-- Since we only have the accession year we set the accession date to be the middle of the year (1st June). -->
 
<!-- `r if(do_survival){"Below we show the survival probability (Kaplan-Meier plots) for threatened and non-threatened plants."}` -->

`r if(!do_survival){"Survival analysis cannot be performed. (This could be due to lack of accession year or item status date."}`

```{r, survival analysis - figure, echo  = FALSE, fig.fullwidth=TRUE, fig.dim = c(10, 6), results='hide',fig.keep='all', message=FALSE, eval = do_survival}
# Create Kaplan-Meier plots comparing threatened and not threatened
data_to_analyse$threatened_status[data_to_analyse$threatened_status == 'Not Threatened'] = 'Non-threatened'
data_to_analyse$threatened_status = factor(data_to_analyse$threatened_status, levels = c('Threatened', 'Non-threatened'))
fit = ggsurvfit::survfit2(survival::Surv(time, Alive_dead) ~ threatened_status, data = data_to_analyse)

p = fit |> 
  ggsurvfit::ggsurvfit() +
  ggplot2::labs(
    x = "Years",
    y = "Overall survival probability"
  ) + 
  ggsurvfit::add_confidence_interval()+
    # theme(text = element_text(size=table_font_size))+
  # labs(title=paste0("Survival probability of native and non native plants in the LC")) +
  labs(caption = "Shaded region represents 95% confidence interval.") +
  scale_colour_manual(values = c('#f46d43', '#848484')) +
  scale_fill_manual(values = c('#f46d43', '#848484')) +
   theme(
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text=element_text(size=10))

if(report_kind == 'static'){
  if(separate_figure_folder){
  ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','threatened_widespread_survival.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
  }
  p
} 
```

```{r survival_analysis - interactive figure, eval = do_survival,  fig.fullwidth=TRUE, fig.dim = c(10, 4)}
if(report_kind == 'interactive'){
  fig = ggplotly(p,dynamicTicks = TRUE)
  fig$x$layout$title$text =''
  fig$x$layout$legend$title$text = ''
  fig$x$layout$hovermode = 'x unified'
  fig$x$data[[2]]$name = 'Non-threatened'
  fig$x$data[[2]]$legendgroup = 'Non-threatened'
  fig$x$data[[1]]$name = 'Threatened'
  fig$x$data[[1]]$legendgroup = 'Threatened'
  fig$x$data[[4]]$hoverinfo = 'none'
  fig$x$data[[3]]$hoverinfo = 'none'
  
  time_region = round(fig$x$data[[4]]$text |>
    stringr::str_extract(pattern = 'time:[ ]+[0-9\\.]+') |>
    stringr::str_extract(pattern = '[0-9]+\\.[0-9]+') |>
    as.numeric(), digits = 3) 
  lower_region = round(fig$x$data[[4]]$text |>
    stringr::str_extract(pattern = 'conf.low: [0-9\\.]+') |>
    stringr::str_extract(pattern = '[01]\\.[0-9]+') |>
    as.numeric()*100,digits = 2)
  upper_region = round(fig$x$data[[4]]$text |>
    stringr::str_extract(pattern = 'conf.high: [0-9\\.]+') |>
    stringr::str_extract(pattern = '[01]\\.[0-9]+') |>
    as.numeric()*100, digits = 2)
  time = round(fig$x$data[[2]]$x,digits = 3)
  lower = lower_region[match(time, time_region)]
  upper = upper_region[match(time, time_region)]
  
  fig$x$data[[2]]$text = paste0(#'Age (years): ',round(fig$x$data[[2]]$x,digits = ), '<br>',
                                'Survival probability: ', round(fig$x$data[[2]]$y*100,digits = 2),'% <br>',
                                '95% confidence Interval: (', lower,'%, ', upper, '%)')
  
    time_region = round(fig$x$data[[3]]$text |>
    stringr::str_extract(pattern = 'time:[ ]+[0-9\\.]+') |>
    stringr::str_extract(pattern = '[0-9]+\\.[0-9]+') |>
    as.numeric(), digits = 3) 
  lower_region = round(fig$x$data[[3]]$text |>
    stringr::str_extract(pattern = 'conf.low: [0-9\\.]+') |>
    stringr::str_extract(pattern = '[01]\\.[0-9]+') |>
    as.numeric()*100,digits = 2)
  upper_region = round(fig$x$data[[3]]$text |>
    stringr::str_extract(pattern = 'conf.high: [0-9\\.]+') |>
    stringr::str_extract(pattern = '[01]\\.[0-9]+') |>
    as.numeric()*100, digits = 2)
  time = round(fig$x$data[[1]]$x,digits = 3)
  lower = lower_region[match(time, time_region)]
  upper = upper_region[match(time, time_region)]
  
  fig$x$data[[1]]$text = paste0('Age (years): ',round(fig$x$data[[1]]$x,digits = ), '<br>',
                                'Survival probability: ', round(fig$x$data[[1]]$y*100,digits = 2),'% <br>',
                                '95% confidence Interval: (', lower,'%, ', upper, '%)')
  
  #Set initial y-axis tick values 
  fig$x$layout$yaxis$ticktext = c("0.25", "0.50", "0.75", "1.00")
  fig$x$layout$yaxis$tickvals = c(0.25, 0.50, 0.75, 1.00)
  fig$x$layout$font$size = 10 # font size
  fig$x$layout$yaxis$title$font$size = 14
  fig$x$layout$xaxis$title$font$size = 14
  fig$x$layout$xaxis$title$font$color = NULL
   fig$x$layout$yaxis$title$font$color = NULL
  fig$x$layout$yaxis$tickfont$size = 12
  fig$x$layout$xaxis$tickfont$size = 12
  
  if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Threatened_survival_analysis.html'))
}
# fig |> layout(yaxis = list(dtick = 0.25, 
#       tick0 = 0, 
#       tickmode = "linear"))
fig
}


```

`r if(do_survival){"Below we give the median surivival times for threatened and non-threatened plants in the collection, and the survival probability for a given number of years."}`

```{r, survival analysis - median survival tables, echo = FALSE, message=FALSE, eval = do_survival}
# We can also get survival probabilities in a table
# 

table_data2 = fit |> 
  gtsummary::tbl_survfit(
    probs = c(0.5),
    label_header = "**Median survival (95% CI)**")

table_data2 = table_data2$table_body
table_data2 = table_data2[,-c(1:4)]
table_data2 = table_data2[-1,]
names(table_data2) = c('Characteristic', 'Median survival + 95% CI (Years)')

if(report_kind == 'interactive'){
  table_data2 |>  DT::datatable(extensions = 'Buttons',
                             rownames = FALSE,
                             options = list(scrollX = TRUE,
                                            pageLength = nrow(table_data2),
                                            dom = 'Bt',
                                            buttons = c('copy', 'csv', 'excel', 'pdf'),
                                            paging=FALSE,
                                            ordering=F))
}



```


```{r, survival analysis - tables, echo = FALSE, message=FALSE, eval = do_survival}
# We can also get survival probabilities in a table
# 
table_data = fit |> 
  gtsummary::tbl_survfit(
    times = c(2,5,10,25,50,100),
    label_header = "{time}-year survival (95% CI)")

table_data = table_data$table_body
  table_data = table_data[,-c(1:4)]
  table_data = table_data[-1,]
  names(table_data) = c('Status', '2-year survival (95% CI)',
                        '5-year survival (95% CI)',
                        '10-year survival (95% CI)',
                        '25-year survival (95% CI)',
                        '50-year survival (95% CI)',
                        '100-year survival (95% CI)')
  table_data = data.frame(t(table_data))
  table_data = table_data[-1,]
  names(table_data) = c('Threatened', 'Non-threatened')
  
  
  table_data2 = fit |> 
  gtsummary::tbl_survfit(
    probs = c(0.5),
    label_header = "**Median survival (95% CI)**")

table_data2 = table_data2$table_body
  table_data2 = table_data2[,-c(1:4)]
  table_data2 = table_data2[-1,]
  names(table_data2) = c('Characteristic', 'Median survival (95% CI)')
  
  
  
if(report_kind == 'static'){
  table_data |>
  flextable() |>
  set_caption(caption = paste0('Survival probability table')) |> 
  font(fontname = "Calibri (Body)", part = "all") |> 
  fontsize(size = table_font_size, part = "body") |> 
  # add footer if you want
  # add_footer_row(values = "* p < 0.05. ** p < 0.01. *** p < 0.001.", 
  #                colwidths = 4) |> 
  theme_booktabs() |> # default theme
  autofit() |>
  set_table_properties(width = 1, layout = "autofit")
}
if(report_kind == 'interactive'){
  table_data$year  =paste0(c(2,5,10,25,50,100), '-year survival')
  names(table_data)[3] = 'Survival time (95% CI)'
  table_data = table_data[,c(3,1,2)]
  table_data |>  DT::datatable(extensions = 'Buttons',
                  rownames = FALSE,
                  options = list(scrollX = TRUE,
                                 pageLength = nrow(table_data),
                                 dom = 'Bt',
                                 buttons = c('copy', 'csv', 'excel', 'pdf'),
                                 paging=FALSE,
                                 ordering=F))

}
```


***

###### {-} 
<!-- This section is for exporting the data -->

```{r, export data}
if(export_data){
  #remove all data except wanted to be exported.
  # rm(list=setdiff(ls(), c("exporting_data_store", "output_dir", "collection")))
  # Do some changes (which we do in smaller chunks in the actual code.)

  threatened_data  = exporting_data_store
  # save as .rda.
  save(threatened_data, file = paste0(output_dir,'/',collection, '_threatened_data.rda'))

}
```

```{r, save excel version of data, eval = export_data}
library(xlsx)
file = paste0(output_dir,'/',collection, '_threatened_data.xlsx')
wb <- createWorkbook()
datas <- exporting_data_store

data_titles = names(datas) |> stringr::str_replace_all('_', ' ')
sheetnames <- paste0('Sheet ', 1:length(datas))
sheets <- lapply(sheetnames,  createSheet, wb = wb)
for(i in 1:length(datas)){
  # First we'll create a new row, let's put the title on row 2
  titleRow <-createRow(sheets[[i]], rowIndex=1)
  # We'll add a cell to that row (in the first column) for the title
  sheetTitle <-createCell(titleRow, colIndex=1)
  # We'll add the sheet title to the cell
  setCellValue(sheetTitle[[1,1]], data_titles[i])
  # We'll choose the formatting for the cell
  xlsx::addDataFrame(datas[[i]], sheet = sheets[[i]],row.names = FALSE, startRow = 3)
}
saveWorkbook(wb, file = file)

```

