---
title: "Turnover of plants and groups of plants in the LC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(htmltools)
library(sf)
library(here)
library(flextable)
library(ggrepel)
```

```{r, imported parameters}
# #Inputs
# load('/Users/jakepowell/Cambridge/Enriched_reports_Jake/Copenhagen_enriched_report.rda')
# collection = 'Chicago'
# coordinates = c(52.19378853629289,0.1277234065606053)
# load('/Users/jakepowell/Cambridge/Geography Module/wgsrpd3.rda')
# 
# native = 'Naturally occurring only'
# extinct = TRUE
# doubtful_locations = FALSE
# 
# separate_figure_folder = T
# output_dir = getwd()
# report_kind = 'interactive'
# ggtheme = NULL
# value_on_fig = TRUE
# min_year = 1970
# data_type = 'Accessions'
# earliest_allowable_record = 1650
# old_accession_year_codes = c(1000)
```

```{r, add logo, eval = report_kind == 'interactive'}
htmltools::img(src = knitr::image_uri(paste0(system.file(package = "BGSmartR"), "/logo.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width:100px; height:100px')
```

```{r theme_ggplot2, echo = FALSE}
library(ggplot2)
# Changing the default theme
if(is.null(ggtheme)){
   ggtheme <- function(base_size = 16) {
      ggplot2::theme_bw(base_size = base_size) %+replace%
        ggplot2::theme(
          plot.title = ggplot2::element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
          panel.grid.minor = ggplot2::element_blank(),
          panel.background = element_rect(fill = 'transparent', color = NA),
          panel.border = ggplot2::element_blank(),
          axis.title = ggplot2::element_text(size = rel(0.85), face = "bold"),
          axis.text = ggplot2::element_text(size = rel(0.70), face = "bold"),
          axis.line = ggplot2::element_line(color = "black"),
          legend.title = ggplot2::element_text(size = rel(0.85), face = "bold"),
          legend.text = ggplot2::element_text(size = rel(0.70), face = "bold"),
          legend.key = ggplot2::element_rect(fill = "transparent", colour = NA),
          legend.key.size = ggplot2::unit(1.5, "lines"),
          legend.background = ggplot2::element_rect(fill = "transparent", colour = NA),
          strip.background = ggplot2::element_rect(fill = "#17252D", color = "#17252D"),
          strip.text = ggplot2::element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
        )
    }
}
theme_set(ggtheme())
update_geom_defaults("line", list(size = 2))

# Set the colour scale for the plots
scale_colour_continuous <- scale_colour_continuous
scale_colour_discrete   <- scale_colour_discrete
scale_colour_binned     <- scale_colour_binned
scale_fill_continuous <- scale_fill_continuous
scale_fill_discrete <- scale_fill_discrete
scale_fill_binned <- scale_fill_binned
```

```{r, Add combined geography column to enriched report}
geography_data = enriched_report[,grepl('_area_code_l3', names(enriched_report))]
want = c('000','010','001','011','100','110','101')

#Depending on the values of the options reduce `want` to only the satisfactory values.
##A) introduced / naturally occurring only.
if(native == 'Introduced only'){
 want = want[grepl('^1',want)]
}else if(native == 'Naturally occurring only'){
 want = want[grepl('^0',want)]
}
## B) Extinct
if(!extinct){
 want = want[!grepl('010|011|110',want)]
}
## C) Doubtful
if(!doubtful_locations){
 want = want[!grepl('001|011|101',want)]
}
# Get the columns in wanted info that contain the geography information we want.
geography_data = geography_data[,grepl(paste0(want,collapse='|'), names(geography_data))]

if(length(want) > 1){
 level3codes =do.call("paste", c(geography_data, sep = ", "))
 level3codes = stringr::str_remove(level3codes, ', NA$')
 level3codes = stringr::str_remove(level3codes, 'NA, ')
}else{
 level3codes = geography_data
}
level3codes[level3codes == "NA"] = NA

enriched_report$geography_codes = level3codes
```

```{r, Get location_type_text}
location_type_text = ''
if(native == 'Naturally occurring only'){
 location_type_text = paste0(location_type_text, 'naturally occuring only,', collapse = ' ')  
}else if(native == 'Introduced only'){
  location_type_text = paste0(location_type_text, 'introduced only,', collapse = ' ')  
 
}else{
  location_type_text = paste0(location_type_text, 'Either naturally occurring or introduced,', collapse = ' ')  
}
if(extinct){
   location_type_text = paste0(location_type_text, ' allowing extinct regions and', collapse = ' ')  
}else{
  location_type_text = paste0(location_type_text, ' not including extinct regions and', collapse = ' ')  
}
if(doubtful_locations){
   location_type_text = paste0(location_type_text, ' allowing doubtful regions.', collapse = ' ')  
 
}else{
  location_type_text = paste0(location_type_text, ' not allowing doubtful regions.', collapse = ' ')  
}
```

```{r, Find LC location, echo = FALSE}
coord_contained = TRUE
# Extract the location of the collection.
pnts <- data.frame( x = coordinates[2], y = coordinates[1])

# create a points collection
pnts_sf <- do.call("st_sfc",c(lapply(1:nrow(pnts), 
                                     function(i) {sf::st_point(as.numeric(pnts[i, ]))}), list("crs" = 4326))) 

pnts_trans <- st_transform(pnts_sf, 2163) # apply transformation to pnts sf
tt1_trans <- st_transform(wgsrpd3, 2163)      # apply transformation to polygons sf

intersect = as.vector(st_intersects(tt1_trans, pnts_trans, sparse = FALSE))
collection_geog_details = wgsrpd3[which(intersect),-5]
location_name = collection_geog_details$LEVEL3_NAM
location_code = collection_geog_details$LEVEL3_COD

if(length(location_name) == 0){
  coord_contained = FALSE
  # Try and find the nearest polygon.
  collection_geog_details  =  wgsrpd3[which.min(as.vector(sf::st_distance(tt1_trans, pnts_trans))),-5]
  location_name = collection_geog_details$LEVEL3_NAM
  location_code = collection_geog_details$LEVEL3_COD
}
```

`r if(any(!c('AccNoFull') %in% names(enriched_report))){"*** \n **Warning!**\n\n This enriched report contains missing 'original' columns needed to perform the analyses in this report. Default values have been created. In particular, the following were missing "}`

`r if(!'AccNoFull' %in% names(enriched_report)){"- AccNoFull: assume each item is its own accession.\n"}`

`r if(any(!c('no_gardens','AccNoFull') %in% names(enriched_report))){"***"}`

```{r, Extracting required information from enriched report, echo = F}

needed_columns = c('AccNoFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType')
if(!'AccNoFull' %in% names(enriched_report)){
  enriched_report$AccNoFull = 1:nrow(enriched_report)
}
if(!'AccYear' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$AccYear = rep(0, nrow(enriched_report))
}
if(!'ItemStatusDate' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$ItemStatusDate = rep(NA, nrow(enriched_report))
}
if(!'ItemStatusType' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$ItemStatusType = rep('Existing', nrow(enriched_report))
}
if(all(is.na(enriched_report$ItemStatusDate))){
  do_over_time = FALSE
}
if(all(is.na(enriched_report$ItemStatusType))){
  do_over_time = FALSE
}
if(all(is.na(enriched_report$AccYear))){
  do_over_time = FALSE
}
#Extract endemic information from enriched_report
endemic = rep('Not Endemic', nrow(enriched_report))
endemic_index = which(stringr::str_length(enriched_report$geography_codes) == 3)
endemic[endemic_index] = 'Endemic'
enriched_report$endemic = endemic
rm(endemic)

# Extract threatened. 
threat_cat = c('VU','EN','CR','EW','EX')
threatened = rep('Not Threatened', nrow(enriched_report))
threatened[which(enriched_report$redList_category %in% threat_cat)] = 'Threatened'
threatened[which(enriched_report$POWO_Red_category %in% threat_cat)] = 'Threatened'
enriched_report$threatened = threatened
rm(threatened)

# Extract threatened category. 
threatened_category = rep(NA, nrow(enriched_report))
for(i in 1:length(threat_cat)){
 threatened_category[which(enriched_report$redList_category == threat_cat[i])] = threat_cat[i]
 threatened_category[which(enriched_report$POWO_Red_category == threat_cat[i])] = threat_cat[i]
}
enriched_report$threatened_category = threatened_category
rm(threatened_category)

# Convert Provenance code to text.
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'G'] = 'Garden'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'U'] = 'Unknown'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'W'] = 'Wild'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'Z'] = 'Wild-derived'

# Extract tree.
tree =rep('Not Tree', nrow(enriched_report))
tree[enriched_report$Enrich_is_tree] = 'Tree'
enriched_report$tree = tree
rm(tree)

# Extract native.
native = rep('Not Native', nrow(enriched_report))
native_index = grepl(location_code,enriched_report$geography_codes)
native[native_index] = 'Native'
enriched_report$native = native
rm(native)

# Create a sanitised taxonomic name.
enriched_report$good_name = paste0(enriched_report$sanitised_taxon, ' ', enriched_report$extracted_author)

best_name = paste0(enriched_report$POWO_taxon_name, ' ', enriched_report$POWO_taxon_authors)
best_name[which(best_name == 'NA NA')] = enriched_report$good_name[which(best_name == 'NA NA')]
enriched_report$best_name = best_name

report_original = enriched_report

report = enriched_report[match(c('AccNoFull','AccYear', 'ItemStatusDate', 'ItemStatusType','good_name', 'best_name', 'taxon_type',
                                 'ProvenanceCode', 'endemic', 'threatened', 'threatened_category',
                                 'tree', 'native'),names(enriched_report))]
LossYear = rep(NA,nrow(report))
ItemStatusYear = as.numeric(stringr::str_extract(report$ItemStatusDate,pattern = '[0-9]{4}'))
LossYear[report$ItemStatusType == 'NotExisting'] = ItemStatusYear[report$ItemStatusType == 'NotExisting']
report$LossYear = LossYear
  
if(separate_figure_folder){
  figures_dir = paste0(output_dir,'/',collection,'_Figures')
  dir.create(figures_dir,showWarnings = FALSE)
}

exporting_data_store = list()

## Data for turnovers
max_year = as.numeric(format(Sys.Date(),'%Y'))
years = min_year:max_year

# Set old unknown records to have a date ( = earliest_allowable_record) and remove all records prior to this time.
if(!is.null(old_accession_year_codes)){
  has_old_accessions_code = which(report$AccYear %in% old_accession_year_codes)
  no_old_accessions = length(has_old_accessions_code)
  report$AccYear[has_old_accessions_code] = earliest_allowable_record
}
to_keep = which(as.numeric(report$AccYear) >= earliest_allowable_record  & as.numeric(report$AccYear) <= max_year)
still_issue = nrow(report) - length(to_keep)
report = report[to_keep,]

# Find the number of items in the collection each year.
existing_items_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                        AccessionYear = as.character(report$AccYear),
                        ItemStatusDate = as.character(report$ItemStatusDate),
                        ItemStatusType = as.character(report$ItemStatusType))))

#######################################
# Report restricted to accessions.
#######################################
unique_accessions = unique(report$AccNoFull)

# Order the items where the lower indices have the most recent death date.
LossYear_dummy = LossYear
LossYear_dummy[is.na(LossYear_dummy)] = 4000
ordered_items = order(LossYear_dummy,decreasing = T)

#Match to the ordered accessions.
match_to_best = match(unique_accessions,report$AccNoFull[ordered_items])

#Reduce the data to unique accessions with most recent death date.
report_accessions = report[match_to_best,]

existing_accessions_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                        AccessionYear = report_accessions$AccYear,
                        ItemStatusDate = report_accessions$ItemStatusDate,
                        ItemStatusType = report_accessions$ItemStatusType)))
```

```{r, functions for turnover}
turnover_items <- function(report){
  combined = data.frame(Year = years)

# Items
gain = data.frame(table(report$AccYear))
if(nrow(gain) == 0){
  gained = rep(0,nrow(combined))
  combined$gain_items = gained
}else{
  names(gain) = c('Year', 'Items')
  gain$Year = as.numeric(as.character(gain$Year))
  gain = gain[gain$Year >=min_year & gain$Year <= max_year,]
  gained = rep(0,nrow(combined))
  gained[match(gain$Year, combined$Year)] = gain$Items
  combined$gain_items = gained
}


loss = data.frame(table(report$LossYear))
if(nrow(loss) == 0){
  lost = rep(0,nrow(combined))
  combined$loss_items = lost
}else{
  names(loss) = c('Year', 'Items')
  loss$Year = as.numeric(as.character(loss$Year))
  loss = loss[loss$Year >=min_year & loss$Year <= max_year,]
  lost = rep(0,nrow(combined))
  lost[match(loss$Year, combined$Year)] = loss$Items
  combined$loss_items = lost
}


combined$net_items = combined$gain_items - combined$loss_items
return(combined)
}

plots_turnover <- function(turnover, report_kind, text = '', separate_figure_folder, data_type = 'items'){
  if(report_kind == 'static'){
    dataA = data.frame(turnover[,c(1,3)], rep('Lost', nrow(turnover)))
    names(dataA) = LETTERS[1:3]
    dataB = data.frame(turnover[,c(1,2)], rep('Gain', nrow(turnover)))
    names(dataB) = LETTERS[1:3]
    items_data = rbind(dataA, dataB)
    
    p = ggplot(data=items_data, aes(x=A, y=B, group=C)) +
      geom_line(aes(color=C)) +
        # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
        labs(title=paste0("Turnover of ",text," in the LC (By ",data_type,")"),
            x ="Year",
            y = paste0('Number of ',data_type,'')) +
        guides(color=guide_legend(title=paste0('Gain/Loss'))) +
        # theme(text = element_text(size=table_font_size)) +
      scale_color_manual(values=c("blue", "red"))+
        theme(legend.direction = "horizontal", legend.position = "top", legend.justification = "right")
      
      
      if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','turnover_',stringr::str_replace_all(text, ' ', '_'),'_gain_loss_',data_type,'.pdf'),device = 'pdf', scale = 1, width = 20, height = 12, limitsize = FALSE)
    }
    p1=p
      
    # Color based on value
    color <- ifelse(turnover$net_items < 0, "pink", "lightblue")
    
    p = ggplot(turnover, aes(x = Year, y = net_items)) +
      geom_bar(stat = "identity",
               show.legend = FALSE,
               fill = color,      # Background color
               color = "white") + # Border color
      # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
      labs(title=paste0("Net turnover of ",text," in the LC (By ",data_type,")"),
          x ="Year",
          y = paste0('Number of ',data_type,'')) #+
      # theme(text = element_text(size=table_font_size))

      if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Net_turnover_',stringr::str_replace_all(text, ' ', '_'),'_',data_type,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
    p2=p
    
    return(list(gain_loss = p1, net = p2))
  }
  if(report_kind == 'interactive'){
    dataA = data.frame(turnover[,c(1,3)], rep('Lost', nrow(turnover)))
    names(dataA) = LETTERS[1:3]
    dataB = data.frame(turnover[,c(1,2)], rep('Gain', nrow(turnover)))
    names(dataB) = LETTERS[1:3]
    items_data = rbind(dataA, dataB)
    
    colors = c("lightblue", "pink")
    fig <- plot_ly(items_data, x = ~A, y = ~B, color = ~C, colors = colors, type = 'scatter', mode = 'lines+markers')
    fig <- fig |> layout(yaxis = list(title = paste0("Number of ",data_type,"")),
                       xaxis = list(title = "Year"))
    fig <- fig |> layout(hovermode = 'x unified')
    
    if(separate_figure_folder){
     htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','line-turnover_',stringr::str_replace_all(text, ' ', '_'),'_gain_loss_',data_type,'.html'))
    }
    fig1 = fig
    
    colors = c("pink", "lightblue")
    color <- ifelse(turnover$net_items < 0, "pink", "lightblue")
    texto = paste0('Net: ', turnover$net_items, '<br>',
                  'Gain: ', turnover$gain_items, '<br>',
                  'Loss: ', turnover$loss_items, '<br>'
                  )
    fig = plot_ly(turnover, x = ~Year, y = ~net_items, type = 'bar', hoverinfo = "text",  hovertext = texto, color = color, colors = rev(colors), showlegend = FALSE)
    fig <- fig |> layout(yaxis = list(title = paste0("Net number of ",data_type,"")),
                       xaxis = list(title = "Year"))
    fig <- fig |> layout(hovermode = 'x unified')
    
    if(separate_figure_folder){
     htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Bar-Net_',stringr::str_replace_all(text, ' ', '_'),'turnover_of_',data_type,'.html'))
    }
    
    return(list(gain_loss = fig1, net = fig))

  }
}

proportional_plots_turnover <- function(turnover, overall_turnover, collection_proportion, separate_figure_folder, 
                                        report_kind, text = '', data_type = 'items'){
  if(report_kind == 'static'){
     # Gain
  plot_data = data.frame(year = turnover[,1],
                         specific = turnover[,2], 
                         all = overall_turnover[,2])
  plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
  plot_data$prop_specific_collection = collection_proportion*100
  plot_data_gain = plot_data
   p = ggplot(plot_data, aes(y=prop_specific, x=year)) + 
      geom_bar(stat="identity", width = 1, fill = 'lightblue', col = 'black') +
     geom_point(aes(x = year, y = prop_specific_collection))+
      labs(title=paste0("Proportion of gained ",data_type," that are ",text," in the LC"),
          x ="Year",
          y = "Percentage (%)") +
      # theme(text = element_text(size=table_font_size)) +
        labs(caption = paste0("Black points are proportion of ",data_type," in the LC that are ",text," each year."))
    p1 = p
    
        if(separate_figure_folder){
    ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/',stringr::str_replace_all(text,' ','_'),'_gain_',data_type,'_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
      }
    
  ### Plot of proportion of lost items being native.
  
  plot_data = data.frame(year =  turnover[,1],
                         specific = turnover[,3], 
                         all = overall_turnover[,3])
  plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
  plot_data$prop_specific_collection = collection_proportion*100
  plot_data_loss = plot_data

   p = ggplot(plot_data, aes(y=prop_specific, x=year)) + 
      geom_bar(stat="identity", width = 1, fill = 'pink', col = 'black') +
     geom_point(aes(x = year, y = prop_specific_collection))+
      # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
      labs(title=paste0("Proportion of lost ",data_type," that are ",text," in the LC"),
          x ="Year",
          y = "Percentage (%)") +
      # theme(text = element_text(size=table_font_size)) +
        labs(caption = paste0("Black points are proportion of ",data_type," in the LC that are ",text," each year."))
  
          if(separate_figure_folder){
    ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/',stringr::str_replace_all(text,' ','_'),'_loss_',data_type,'_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
          }
   
   return(list(gain = p1, loss = p, data_gain = plot_data_gain, data_loss = plot_data_loss))
  }
  if(report_kind =='interactive'){
     # Gain
  plot_data = data.frame(year = turnover[,1],
                         specific = turnover[,2], 
                         all = overall_turnover[,2])
  plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
  plot_data$prop_specific_collection = collection_proportion*100
  plot_data_gain = plot_data
   fig = plot_ly(plot_data, x = ~year, y = ~prop_specific, type = 'bar', name = paste0('% ',text,' ',data_type,' gained'), marker = list(color = 'lightblue'))
fig <- fig |> layout(yaxis = list(title = paste0("Percentage")),
                   xaxis = list(title = "Year"))
fig <- fig |> add_trace(x = ~year, y = ~prop_specific_collection, type = 'scatter', mode = 'markers', name = paste0('% ',text,' ',data_type,' in LC'), marker = list(color = 'black'))
fig <- fig |> layout(hovermode = 'x unified')

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_gained_',stringr::str_replace_all(text, ' ', '_'),'_',data_type,'.html'))
}

fig1 = fig
    
  ### Plot of proportion of lost items being native.
  
  plot_data = data.frame(year =  turnover[,1],
                         specific = turnover[,3], 
                         all = overall_turnover[,3])
  plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
  plot_data$prop_specific_collection = collection_proportion*100
  plot_data_loss = plot_data

     fig = plot_ly(plot_data, x = ~year, y = ~prop_specific, type = 'bar', name = paste0('% ',text,' ',data_type,' lost'), marker = list(color = 'pink'))
fig <- fig |> layout(yaxis = list(title = paste0("Percentage")),
                   xaxis = list(title = "Year"))
fig <- fig |> add_trace(x = ~year, y = ~prop_specific_collection, type = 'scatter', mode = 'markers', name = paste0('% ',text,' ',data_type,' in LC'), marker = list(color = 'black'))
fig <- fig |> layout(hovermode = 'x unified')

if(separate_figure_folder){
 htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_lost_',stringr::str_replace_all(text, ' ', '_'),'_',data_type,'.html'))
}

return(list(gain = fig1, loss = fig,  data_gain = plot_data_gain, data_loss = plot_data_loss))
   
  }
}
```

This report uses data from **`r collection`** and analyses turnover in the collection from **`r min_year`** to **`r max_year`**.

`r if(!is.null(old_accession_year_codes)){paste0("Within the collection the accession years '", paste0(old_accession_year_codes, collapse = ', '),"' are used to denote items which have unknown accession year but have being in the collection since near the beginning. There are ", no_old_accessions," items in the collection with these accession years. In this analysis we set the accession year of such plants to ", earliest_allowable_record, ".")}`

`r if(!is.null(old_accession_year_codes)){"Afterwards the"}else{'The'}` collection contains **`r still_issue`** items whose accession year is not within `r earliest_allowable_record` and `r max_year`. These records are excluded in the following analysis.

Turnover can be split into two parts; when a plant arrives into the collection and when the item is removed. We refer to these two parts as *Gain* and *Loss*. From these values we can calculate the net number of plants gained/lost each year. 

For *Gain* we require items to have an accession date associated with their record. 

For *Loss* we require items to have an item status date together with item status type notifying that the item is now 'Not Existing'. Keep in mind this means a plant is recorded to have being lost on the date the record is updated and this might not align completely with when the item was removed or died.

In the analysis presented within we refer to turnover in terms of `r data_type`. 
`r if(data_type == 'Accessions}{"For accessions, An accession is lost only when all items in the accession have been removed."}`. 


### Overall Turnover

```{r, overall turnover}
overall_turnover_items = turnover_items(report)
exporting_data_store$overall_turnover_items = overall_turnover_items
plots = plots_turnover(overall_turnover_items, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'items')
if(report_kind == 'static'){
  plots$gain_loss = plots$gain_loss + ggtitle('Turnover of items in the LC')
  plots$net = plots$gain_loss + ggtitle('Net turnover of items in the LC')
}

overall_turnover_accessions = turnover_items(report_accessions)
exporting_data_store$overall_turnover_accessions = overall_turnover_accessions

plots_accessions = plots_turnover(overall_turnover_accessions, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'accessions', data_type = 'accessions')
if(report_kind == 'static'){
  plots_accessions$gain_loss = plots$gain_loss + ggtitle('Turnover of accessions in the LC')
  plots_accessions$net = plots$gain_loss + ggtitle('Net turnover of accessions in the LC')
}
``` 

In this section we shall consider the turnover of plants in the whole of the LC. 

#### Items

First we will consider the gain and loss of items in the LC.

```{r, Overall turnover of items - static, eval = report_kind == 'static', fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots$gain_loss

plots$net
```

`r if(report_kind == 'interactive'){"#####  Turnover of items {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Gain/Loss"}`
```{r  Gain/Loss of items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
plots$gain_loss

```

`r if(report_kind == 'interactive'){"###### Net"}`
```{r  Net turnover of items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
plots$net

```

`r if(report_kind == 'interactive'){"##### {-}"}`


#### Accessions

First we will consider the gain and accessions of items in the LC.

```{r, Overall turnover of accessions - static, eval = report_kind == 'static', fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_accessions$gain_loss

plots_accessions$net
```

`r if(report_kind == 'interactive'){"#####  Turnover of accessions {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Gain/Loss"}`
```{r  Gain/Loss of accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
plots_accessions$gain_loss

```

`r if(report_kind == 'interactive'){"###### Net"}`
```{r  Net turnover of accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
plots_accessions$net

```

`r if(report_kind == 'interactive'){"##### {-}"}`



### Turnover of new/unique plants

In this section, we look at the gain of plants that were not in the collection in the previous year and the loss of plants that causes the species to no longer be represented in the LC. 

```{r, gain new/unique plants}
# First need to create a list of taxonomic names that were in the LC each previous year.
existing_items = BGSmartR::exist_at_date(date = paste0(years-1,'-12-31'),
                        AccessionYear = report$AccYear,
                        ItemStatusDate = report$ItemStatusDate,
                        ItemStatusType = report$ItemStatusType)

taxonomic_names_in_LC_years = lapply(existing_items, function(x){
  data_cur = report$best_name[x]
  return(unique(data_cur))
})

names(taxonomic_names_in_LC_years) = paste0('plants_previous_year_to_',years)
new_to_collection = unlist(lapply(years, function(x){
  plants_accessioned = unique(report$best_name[report$AccYear == x])
  previous_plants = taxonomic_names_in_LC_years[[which(grepl(x,names(taxonomic_names_in_LC_years)))]]
  sum(!plants_accessioned %in% previous_plants)
}))
total_taxonomic_names = unlist(lapply(years, function(x){
  plants_accessioned = report$best_name[report$AccYear == x]
  length(unique(plants_accessioned))
}))

prop_new_collection = new_to_collection*100 /total_taxonomic_names
prop_new_collection[is.nan(prop_new_collection)] = NA

plot_data = data.frame(years, new_to_collection, total_taxonomic_names, prop_new_collection)
exporting_data_store$gained_taxa_with_new_to_LC = plot_data


plot_data_formatted = rbind(data.frame(year = plot_data$years, taxa = plot_data$new_to_collection, prop = plot_data$prop_new_collection, fill = rep('Yes',nrow(plot_data)) ), 
                            data.frame(year = plot_data$years, taxa = plot_data$total_taxonomic_names -  plot_data$new_to_collection, prop = 100 - plot_data$prop_new_collection, fill = rep('No',nrow(plot_data)) )
)

if(report_kind == 'static'){
  p = ggplot(plot_data_formatted, aes(fill=fill, y=taxa, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) + labs(x = 'Year', y = 'Number of Taxa', fill = 'New to LC')
p = p + scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) + scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))+
  ggtitle("Number of gained taxa to the LC \n by if the taxa is already present")

p1= p

if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Stacked_Bar-Turnover_new_to_LC_gained_number.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

p = ggplot(plot_data_formatted, aes(fill=fill, y=prop, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) + labs(x = 'Year', y = 'Percentage', fill = 'New to LC')
p = p + scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) + scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  ggtitle("Percentage of gained taxa that are new to the LC")

p2 = p

if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Stacked_Bar-Turnover_new_to_LC_gained_percentage.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

gained_new_taxa = list(number = p1, percentage = p2)
}
if(report_kind == 'interactive'){
  # plot_data_formatted$text = paste0(round(stacked_bar_data$no_accessions,digits = 2), ' accessions have ', stacked_bar_data$total_items, ' items')

fig <- plot_ly(plot_data_formatted, x = ~year, y = ~taxa, color = ~fill, type = 'bar')
fig <- fig |> layout(yaxis = list(title = 'Number of Taxa'),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text='New to LC')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)
fig1 = fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-Turnover_new_to_LC_gained_number.html'))
}

fig <- plot_ly(plot_data_formatted, x = ~year, y = ~prop, color = ~fill, type = 'bar')
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text='New to LC')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig2 = fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-Turnover_new_to_LC_gained_percentage.html'))
}
gained_new_taxa = list(number = fig1, percentage = fig2)

}
```

```{r, gain new/unique plants excluding indeterminate and cultivar}
# First need to create a list of taxonomic names that were in the LC each previous year
report_cur = report[!grepl('0|5',report$taxon_type),]
existing_items = BGSmartR::exist_at_date(date = paste0(years-1,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType)

taxonomic_names_in_LC_years = lapply(existing_items, function(x){
  data_cur = report_cur$best_name[x]
  return(unique(data_cur))
})

names(taxonomic_names_in_LC_years) = paste0('plants_previous_year_to_',years)
new_to_collection = unlist(lapply(years, function(x){
  plants_accessioned = unique(report_cur$best_name[report_cur$AccYear == x])
  previous_plants = taxonomic_names_in_LC_years[[which(grepl(x,names(taxonomic_names_in_LC_years)))]]
  sum(!plants_accessioned %in% previous_plants)
}))
total_taxonomic_names = unlist(lapply(years, function(x){
  plants_accessioned = report_cur$best_name[report_cur$AccYear == x]
  length(unique(plants_accessioned))
}))

prop_new_collection = new_to_collection*100 /total_taxonomic_names
prop_new_collection[is.nan(prop_new_collection)] = NA

plot_data = data.frame(years, new_to_collection, total_taxonomic_names, prop_new_collection)
exporting_data_store$gained_taxa_with_new_to_LC_without_cult_or_indet = plot_data


plot_data_formatted = rbind(data.frame(year = plot_data$years, taxa = plot_data$new_to_collection, prop = plot_data$prop_new_collection, fill = rep('Yes',nrow(plot_data)) ), 
                            data.frame(year = plot_data$years, taxa = plot_data$total_taxonomic_names -  plot_data$new_to_collection, prop = 100 - plot_data$prop_new_collection, fill = rep('No',nrow(plot_data)) )
)

if(report_kind == 'static'){
  p = ggplot(plot_data_formatted, aes(fill=fill, y=taxa, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) + labs(x = 'Year', y = 'Number of Taxa', fill = 'New to LC')
p = p + scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) + scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))+
  ggtitle("Number of gained taxa to the LC \n by if the taxa is already present \n after removing cultivars and indeterminate taxa")

p1= p

if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Stacked_Bar-Turnover_new_to_LC_gained_number_without_cult_indet.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

p = ggplot(plot_data_formatted, aes(fill=fill, y=prop, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) + labs(x = 'Year', y = 'Percentage', fill = 'New to LC')
p = p + scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) + scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  ggtitle("Percentage of gained taxa that are new to the LC\n after removing cultivars and indeterminate taxa")

p2 = p

if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Stacked_Bar-Turnover_new_to_LC_gained_percentage_without_cult_indet.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

gained_new_taxa_without_cult_indet = list(number = p1, percentage = p2)
}
if(report_kind == 'interactive'){
  # plot_data_formatted$text = paste0(round(stacked_bar_data$no_accessions,digits = 2), ' accessions have ', stacked_bar_data$total_items, ' items')

fig <- plot_ly(plot_data_formatted, x = ~year, y = ~taxa, color = ~fill, type = 'bar')
fig <- fig |> layout(yaxis = list(title = 'Number of Taxa'),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text='New to LC')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)
fig1 = fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-Turnover_new_to_LC_gained_number_without_cult_indet.html'))
}

fig <- plot_ly(plot_data_formatted, x = ~year, y = ~prop, color = ~fill, type = 'bar')
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text='New to LC')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig2 = fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-Turnover_new_to_LC_gained_percentage_without_cult_indet.html'))
}
gained_new_taxa_without_cult_indet = list(number = fig1, percentage = fig2)

}
```

Below we show the number of gained taxa each year (where we count taxa by unique taxonomic name and author), and show how many of the gained taxa are new to the LC.

```{r, number and percentage gained taxa are new - static, eval = report_kind == 'static', fig.fullwidth=TRUE, fig.dim = c(10, 6)}
gained_new_taxa$number

gained_new_taxa$percentage
```

`r if(report_kind == 'interactive'){"#####  Gained Taxa and whether the taxa is new to the LC {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Number of Taxa"}`
```{r  number gained taxa are new - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
gained_new_taxa$number
```

`r if(report_kind == 'interactive'){"###### Proportion"}`
```{r  percentage gained taxa are new - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
gained_new_taxa$percentage
```

`r if(report_kind == 'interactive'){"##### {-}"}`

Next we perform the same analysis as above but removing new indeterminate taxa and cultivars.

```{r, number and percentage gained taxa are new without cult or indet - static, eval = report_kind == 'static', fig.fullwidth=TRUE, fig.dim = c(10, 6)}
gained_new_taxa_without_cult_indet$number

gained_new_taxa_without_cult_indet$percentage
```

`r if(report_kind == 'interactive'){"#####  Gained Taxa and whether the taxa is new to the LC (removed cultivars and indeterminate taxa) {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Number of Taxa"}`
```{r  number gained taxa are new without cult or indet - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
gained_new_taxa_without_cult_indet$number
```

`r if(report_kind == 'interactive'){"###### Proportion"}`
```{r  percentage gained taxa are new without cult or indet - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
gained_new_taxa_without_cult_indet$percentage
```

`r if(report_kind == 'interactive'){"##### {-}"}`


```{r, loss unique plants}
# First need to create a list of taxonomic names that were in the LC at the end of the year.
existing_items = BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report$AccYear,
                        ItemStatusDate = report$ItemStatusDate,
                        ItemStatusType = report$ItemStatusType)

taxonomic_names_in_LC_years = lapply(existing_items, function(x){
  data_cur = report$best_name[x]
  return(unique(data_cur))
})

names(taxonomic_names_in_LC_years) = paste0('plants_at_end_of_year_',years)

lost_to_the_collection = unlist(lapply(years, function(x){
  plants_lost = unique(report$best_name[which(report$LossYear == x)])
  in_LC_end_year = taxonomic_names_in_LC_years[[which(grepl(x,names(taxonomic_names_in_LC_years)))]]
  sum(!plants_lost %in% in_LC_end_year)
}))

total_taxonomic_names = unlist(lapply(years, function(x){
  plants_lost = report$best_name[which(report$LossYear == x)]
  length(unique(plants_lost))
}))

prop_lost_unique = lost_to_the_collection*100 /total_taxonomic_names
prop_lost_unique[is.nan(prop_lost_unique)] = NA

plot_data = data.frame(years, lost_to_the_collection, total_taxonomic_names, prop_lost_unique)
exporting_data_store$lost_taxa_with_lost_to_LC = plot_data

plot_data_formatted = rbind(data.frame(year = plot_data$years, taxa = plot_data$lost_to_the_collection, prop = plot_data$prop_lost_unique, fill = rep('Yes',nrow(plot_data)) ), 
                            data.frame(year = plot_data$years, taxa = plot_data$total_taxonomic_names -  plot_data$lost_to_the_collection, prop = 100 - plot_data$prop_lost_unique, fill = rep('No',nrow(plot_data)) )
)

if(report_kind == 'static'){
  p = ggplot(plot_data_formatted, aes(fill=fill, y=taxa, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) + labs(x = 'Year', y = 'Number of Taxa', fill = 'No longer represented\nin the LC')
p = p + scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) + scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))+
  ggtitle("Number of lost taxa to the LC \n by if the taxa is no longer represented in th LC")

p1= p

if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Stacked_Bar-Turnover_loss_to_LC_unique_number.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

p = ggplot(plot_data_formatted, aes(fill=fill, y=prop, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) + labs(x = 'Year', y = 'Percentage', fill = 'New to LC')
p = p + scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) + scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  ggtitle("Percentage of lost taxa that are no longer represented in the LC")

p2 = p

if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Stacked_Bar-Turnover_loss_to_LC_unique_percentage.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

loss_unique_taxa = list(number = p1, percentage = p2)
}
if(report_kind == 'interactive'){
  # plot_data_formatted$text = paste0(round(stacked_bar_data$no_accessions,digits = 2), ' accessions have ', stacked_bar_data$total_items, ' items')

fig <- plot_ly(plot_data_formatted, x = ~year, y = ~taxa, color = ~fill, type = 'bar')
fig <- fig |> layout(yaxis = list(title = 'Number of Taxa'),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text='No longer represented<br>in the LC')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)
fig1 = fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-Turnover_loss_to_LC_unique_number.html'))
}

fig <- plot_ly(plot_data_formatted, x = ~year, y = ~prop, color = ~fill, type = 'bar')
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text='No longer represented<br>in the LC')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig2 = fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-Turnover_loss_to_LC_unique_percentage.html'))
}
loss_unique_taxa = list(number = fig1, percentage = fig2)

}
```

```{r, loss unique plants excluding indeterminate and cultivar}
report_cur = report[!grepl('0|5',report$taxon_type),]

# First need to create a list of taxonomic names that were in the LC at the end of the year.
existing_items = BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType)

taxonomic_names_in_LC_years = lapply(existing_items, function(x){
  data_cur = report_cur$best_name[x]
  return(unique(data_cur))
})

names(taxonomic_names_in_LC_years) = paste0('plants_at_end_of_year_',years)

lost_to_the_collection = unlist(lapply(years, function(x){
  plants_lost = unique(report_cur$best_name[which(report_cur$LossYear == x)])
  in_LC_end_year = taxonomic_names_in_LC_years[[which(grepl(x,names(taxonomic_names_in_LC_years)))]]
  sum(!plants_lost %in% in_LC_end_year)
}))

total_taxonomic_names = unlist(lapply(years, function(x){
  plants_lost = report_cur$best_name[which(report_cur$LossYear == x)]
  length(unique(plants_lost))
}))

prop_lost_unique = lost_to_the_collection*100 /total_taxonomic_names
prop_lost_unique[is.nan(prop_lost_unique)] = NA

plot_data = data.frame(years, lost_to_the_collection, total_taxonomic_names, prop_lost_unique)
exporting_data_store$lost_taxa_with_lost_to_LC_without_cult_or_indet = plot_data

plot_data_formatted = rbind(data.frame(year = plot_data$years, taxa = plot_data$lost_to_the_collection, prop = plot_data$prop_lost_unique, fill = rep('Yes',nrow(plot_data)) ), 
                            data.frame(year = plot_data$years, taxa = plot_data$total_taxonomic_names -  plot_data$lost_to_the_collection, prop = 100 - plot_data$prop_lost_unique, fill = rep('No',nrow(plot_data)) )
)

if(report_kind == 'static'){
  p = ggplot(plot_data_formatted, aes(fill=fill, y=taxa, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) + labs(x = 'Year', y = 'Number of Taxa', fill = 'No longer represented\nin the LC')
p = p + scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) + scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))+
  ggtitle("Number of lost taxa to the LC \n by if the taxa is no longer represented in th LC")

p1= p

if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Stacked_Bar-Turnover_loss_to_LC_unique_number_without_cult_indet.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

p = ggplot(plot_data_formatted, aes(fill=fill, y=prop, x=year)) + 
    geom_bar(position="stack", stat="identity", width = 1) + labs(x = 'Year', y = 'Percentage', fill = 'New to LC')
p = p + scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) + scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  ggtitle("Percentage of lost taxa that are no longer represented in the LC")

p2 = p

if(separate_figure_folder){
ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Stacked_Bar-Turnover_loss_to_LC_unique_percentage_without_cult_indet.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
}

loss_unique_taxa_without_cult_indet = list(number = p1, percentage = p2)
}
if(report_kind == 'interactive'){
  # plot_data_formatted$text = paste0(round(stacked_bar_data$no_accessions,digits = 2), ' accessions have ', stacked_bar_data$total_items, ' items')

fig <- plot_ly(plot_data_formatted, x = ~year, y = ~taxa, color = ~fill, type = 'bar')
fig <- fig |> layout(yaxis = list(title = 'Number of Taxa'),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text='No longer represented<br>in the LC')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)
fig1 = fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-Turnover_loss_to_LC_unique_number_without_cult_indet.html'))
}

fig <- plot_ly(plot_data_formatted, x = ~year, y = ~prop, color = ~fill, type = 'bar')
fig <- fig |> layout(yaxis = list(title = 'Percentage'),
                     xaxis = list(title = "Year"),
                     legend=list(title=list(text='No longer represented<br>in the LC')),
                     barmode = 'stack')
fig <- fig |> layout(hovermode = 'x unified', bargap =0)

fig2 = fig

if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Stacked_Bar-Turnover_loss_to_LC_unique_percentage_without_cult_indet.html'))
}
loss_unique_taxa_without_cult_indet = list(number = fig1, percentage = fig2)

}
```

Next up we consider the loss of taxa from the LC each year. In the following graphs we show the total number of taxa which have loss. Moreover, we group the taxa into those that are no longer represented in the LC and those that are at the end of the year. For example a taxa will be classed as 'No longer represented in the LC' if during the year the taxa has experienced loss (plant removal/death) and on the 31st of December that year the taxa is no longer represented in the garden. This method does therefore class taxa which are not in th LC for a period of the year to be still represented in the LC as long as the taxa is reintroduced by the 31st of December.  

```{r, number and percentage lost taxa are new - static, eval = report_kind == 'static', fig.fullwidth=TRUE, fig.dim = c(10, 6)}
loss_unique_taxa$number

loss_unique_taxa$percentage
```

`r if(report_kind == 'interactive'){"#####  Lost taxa and whether they are still represented in the LC {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Number of Taxa"}`
```{r  number loss taxa are new - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
loss_unique_taxa$number
```

`r if(report_kind == 'interactive'){"###### Proportion"}`
```{r  percentage loss taxa are new - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
loss_unique_taxa$percentage
```

`r if(report_kind == 'interactive'){"##### {-}"}`

Next we perform the same analysis as above but removing lost indeterminate taxa and cultivars.

```{r, number and percentage lost taxa are new without cult or indet - static, eval = report_kind == 'static', fig.fullwidth=TRUE, fig.dim = c(10, 6)}
loss_unique_taxa_without_cult_indet$number

loss_unique_taxa_without_cult_indet$percentage
```

`r if(report_kind == 'interactive'){"#####  Lost taxa and whether they are still represented in the LC (removed cultivars and indeterminate taxa) {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Number of Taxa"}`
```{r  number lost taxa are new without cult or indet - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
loss_unique_taxa_without_cult_indet$number
```

`r if(report_kind == 'interactive'){"###### Proportion"}`
```{r  percentage lost taxa are new without cult or indet - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive'}
loss_unique_taxa_without_cult_indet$percentage
```

`r if(report_kind == 'interactive'){"##### {-}"}`


### Turnover of types of taxa

This section explores the turnover of types of taxa. In particular, we use the taxonomic name of plants in the LC to determine whether plants classified as:

- Indeterminate
- Species
- Subspecies
- Varieties
- Forma
- Cultivars
- Hybrids

Not that a record can have multiple of these classifications, such as `Tilia cordata × dasystyla subsp. caucasica`. Since, are not interested in the level of species (i.e if a plant is a subspecies or variety) we are going to explore the turnover of: indeterminate taxa, cultivars, and non-horticultural (those that are not indeterminate, hybrids or cultivars).

#### Indeterminate Taxa

```{r, turnover of indeterminates}
do_indet = TRUE
report_cur = report[grepl('0',report$taxon_type),]
if(nrow(report_cur) == 0){
  do_indet = FALSE
}else{
  turnover_indetermintes = turnover_items(report_cur)
exporting_data_store$turnover_indeterminate_taxa = turnover_indetermintes

plots = plots_turnover(turnover_indetermintes, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'indeterminate plants')

existing_indeterminates_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType)))
collection_proportion = existing_indeterminates_each_year/ existing_items_each_year

plots_proportional = proportional_plots_turnover(turnover = turnover_indetermintes,
                                                 overall_turnover = overall_turnover_items,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'indeterminate')
exporting_data_store$proportional_turnover_indeterminate_taxa_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_indeterminate_taxa_loss = plots_proportional$data_loss

############################################################################
report_cur = report_accessions[grepl('0',report_accessions$taxon_type),]
turnover_indetermintes_accessions = turnover_items(report_cur)
plots_accessions = plots_turnover(turnover_indetermintes_accessions, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'indeterminate plants', data_type = 'accessions')

existing_indeterminates_each_year_accessions = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType)))
collection_proportion = existing_indeterminates_each_year/ existing_accessions_each_year

plots_proportional_accessions = proportional_plots_turnover(turnover = turnover_indetermintes_accessions,
                                                 overall_turnover = overall_turnover_accessions,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'indeterminate',data_type = 'accessions')
exporting_data_store$proportional_turnover_indeterminate_taxa_accessions_gain = plots_proportional_accessions$data_gain
exporting_data_store$proportional_turnover_indeterminate_taxa_accessions_loss = plots_proportional_accessions$data_loss

}

```

`r if(!do_indet){"There are no records that are indeteminate."}`

```{r, Turnover of indeterminates items - static, eval = report_kind == 'static' & do_indet, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
plots$gain_loss

plots$net

plots_accessions$gain_loss

plots_accessions$net
```

`r if(report_kind == 'interactive' & do_indet){"#####  Turnover of indeterminate taxa {.tabset}"}`

`r if(report_kind == 'interactive' & do_indet){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of indeterminates items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_indet}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_indet){"###### Net (Items)"}`
```{r  Net turnover of indeterminates items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_indet}
plots$net
```

`r if(report_kind == 'interactive' & do_indet){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of indeterminates accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_indet}
plots_accessions$gain_loss
```

`r if(report_kind == 'interactive' & do_indet){"###### Net (Accessions)"}`
```{r  Net turnover of indeterminates accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_indet}
plots_accessions$net
```

`r if(report_kind == 'interactive' & do_indet){"##### {-}"}`

`r if(do_indet){"We now explore how the gain and loss of indeterminate items compare proportionally to the total number of gained and lost items and the overall proportion of indeterminate items each year. To calculate the proportion of the collection that is indeteminate each year we only consider records that have an accession year and item status date and extract which items were existing on the first of January each year."}`

```{r, Proportional Turnover of indeterminates items - static, eval = report_kind == 'static' & do_indet, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

plots_proportional_accessions$gain

plots_proportional_accessions$loss

```

`r if(report_kind == 'interactive' & do_indet){"#####  Proportional Turnover of indeterminate taxa {.tabset}"}`

`r if(report_kind == 'interactive' & do_indet){"###### Gain (Items)"}`
```{r  Proportional Gain of indeterminates items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_indet}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_indet){"###### Loss (Items)"}`
```{r  Proportional Loss of indeterminates items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_indet}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_indet){"###### Gain (Accessions)"}`
```{r  Proportional Gain of indeterminates accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_indet}
plots_proportional_accessions$gain
```

`r if(report_kind == 'interactive' & do_indet){"###### Loss (Accessions)"}`
```{r  Proportional Loss of indeterminates accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_indet}
plots_proportional_accessions$loss
```

`r if(report_kind == 'interactive' & do_indet){"##### {-}"}`

#### Cutivars

```{r, turnover of cultivars}
do_type = TRUE
report_cur = report[grepl('5',report$taxon_type),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_cutivars = turnover
  
  plots = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'cultivars', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'cultivars',
                                                   data_type = 'items')
  exporting_data_store$proportional_turnover_cutivars_gain = plots_proportional$data_gain
  exporting_data_store$proportional_turnover_cutivars_loss = plots_proportional$data_loss
  ##############################################################################################
  report_cur = report_accessions[grepl('5',report_accessions$taxon_type),]
  turnover = turnover_items(report_cur)
  plots_accessions = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'cultivars',data_type = 'accessions')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_accessions = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'cultivars',
                                                   data_type = 'accessions')
  exporting_data_store$proportional_turnover_cutivars_gain_accessions = plots_proportional_accessions$data_gain
  exporting_data_store$proportional_turnover_cutivars_loss_accessions = plots_proportional_accessions$data_loss
}
```

`r if(!do_type){"There are no records that are cultivars."}`

```{r, Turnover of cultivars items - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots$gain_loss

plots$net

plots_accessions$gain_loss

plots_accessions$net


```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of cultivars {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of cultivars items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Items)"}`
```{r  Net turnover of cultivars items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of cultivars accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Accessions)"}`
```{r  Net turnover of cultivars accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$net
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(do_type){"We now explore how the gain and loss of cultivars compare proportionally to the total number of gained and lost items/accessions and the overall proportion of cultivars each year. To calculate the proportion of the collection that are cultivars each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of cultivars - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

plots_proportional_accessions$gain

plots_proportional_accessions$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of cultivars {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain (Items)"}`
```{r  Proportional Gain of cultivars items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Items)"}`
```{r  Proportional Loss of cultivars items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"###### Gain (Accessions)"}`
```{r  Proportional Gain of cultivars accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Accessions)"}`
```{r  Proportional Loss of cultivars accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`


#### Non-horticultural

```{r, turnover of Non-horticultural}
do_type = TRUE
report_cur = report[!grepl('0|5|6',report$taxon_type),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_non_horticultural = turnover
  
  plots = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'Non-horticultural', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'Non-horticultural',
                                                   data_type = 'items')
  exporting_data_store$proportional_turnover_non_horticultural_gain = plots_proportional$data_gain
  exporting_data_store$proportional_turnover_non_horticultural_loss = plots_proportional$data_loss
  ##############################################################################################
  report_cur = report_accessions[!grepl('0|5|6',report_accessions$taxon_type),]
  turnover = turnover_items(report_cur)
  plots_accessions = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'Non-horticultural',data_type = 'accessions')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_accessions = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'Non-horticultural',
                                                   data_type = 'accessions')
  exporting_data_store$proportional_turnover_non_horticultural_gain_accessions = plots_proportional_accessions$data_gain
  exporting_data_store$proportional_turnover_non_horticultural_loss_accessions = plots_proportional_accessions$data_loss
}
```

`r if(!do_type){"There are no records that are non-horticultural."}`

```{r, Turnover of Non-horticultural items - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots$gain_loss

plots$net

plots_accessions$gain_loss

plots_accessions$net


```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of Non-horticultural {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of Non-horticultural items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Items)"}`
```{r  Net turnover of Non-horticultural items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of Non-horticultural accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Accessions)"}`
```{r  Net turnover of Non-horticultural accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$net
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(!do_type){"We now explore how the gain and loss of Non-horticultural taxa compare proportionally to the total number of gained and lost items/accessions and the overall proportion of Non-horticultural each year. To calculate the proportion of the collection that are Non-horticultural each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of Non-horticultural - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

plots_proportional_accessions$gain

plots_proportional_accessions$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of Non-horticultural {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain (Items)"}`
```{r  Proportional Gain of Non-horticultural items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Items)"}`
```{r  Proportional Loss of Non-horticultural items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"###### Gain (Accessions)"}`
```{r  Proportional Gain of Non-horticultural accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Accessions)"}`
```{r  Proportional Loss of Non-horticultural accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`


### Provenance of Turnover

This section looks at the provenance of gained and lost items to the collection. The provenance of an item can be in one of four groups, namely: Wild, Wild-derived, Garden or Unknown. We shall consider each provenance type individually. 

```{r, provenance of turnover individual by items}
do_unknown = TRUE
report_cur = report[which(report$ProvenanceCode == 'Unknown'),]
if(nrow(report_cur) == 0){
  do_unknown = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_provenance_unknown = turnover
  
  plots_unknown = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'unknown provenance', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_unknown = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'unknown provenance',
                                                   data_type = 'items')
}

###################################################################################################
do_garden = TRUE
report_cur = report[which(report$ProvenanceCode == 'Garden'),]
if(nrow(report_cur) == 0){
  do_garden = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_provenance_garden = turnover
  
  plots_garden = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'garden provenance', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_garden = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'garden provenance',
                                                   data_type = 'items')
}

###################################################################################################
do_wild_derived = TRUE
report_cur = report[which(report$ProvenanceCode == 'Wild-derived'),]
if(nrow(report_cur) == 0){
  do_wild_derived = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_provenance_wild_derived = turnover
  
  plots_wild_derived = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'wild-derived provenance', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_wild_derived = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'wild-derived provenance',
                                                   data_type = 'items')
}

###################################################################################################
do_wild = TRUE
report_cur = report[which(report$ProvenanceCode == 'Wild'),]
if(nrow(report_cur) == 0){
  do_wild = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_provenance_wild = turnover
  
  plots_wild = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'wild provenance', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_wild = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'wild provenance',
                                                   data_type = 'items')
}

##############################################################################################
```

```{r, provenance of turnover individual by accessions}
if(do_unknown){
  report_cur = report_accessions[which(report_accessions$ProvenanceCode == 'Unknown'),]
turnover = turnover_items(report_cur)
exporting_data_store$turnover_provenance_unknown_accessions = turnover

plots_accessions_unknown = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'unknown provenance',data_type = 'accessions')

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType)))
collection_proportion = existing_each_year/ existing_items_each_year

plots_proportional_accessions_unknown = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover_items,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'unknown provenance',
                                                 data_type = 'accessions')
}

if(do_garden){
  report_cur = report_accessions[which(report_accessions$ProvenanceCode == 'Garden'),]
turnover = turnover_items(report_cur)
exporting_data_store$turnover_provenance_garden_accessions = turnover

plots_accessions_garden = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'garden provenance',data_type = 'accessions')

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType)))
collection_proportion = existing_each_year/ existing_items_each_year

plots_proportional_accessions_garden = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover_items,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'garden provenance',
                                                 data_type = 'accessions')
}

if(do_wild_derived){
  report_cur = report_accessions[which(report_accessions$ProvenanceCode == 'Wild-derived'),]
turnover = turnover_items(report_cur)
exporting_data_store$turnover_provenance_wild_derived_accessions = turnover

plots_accessions_wild_derived = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'wild-derived provenance',data_type = 'accessions')

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType)))
collection_proportion = existing_each_year/ existing_items_each_year

plots_proportional_accessions_wild_derived = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover_items,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'wild-derived provenance',
                                                 data_type = 'accessions')
}

if(do_wild){
  report_cur = report_accessions[which(report_accessions$ProvenanceCode == 'Wild'),]
turnover = turnover_items(report_cur)
exporting_data_store$turnover_provenance_wild_accessions = turnover

plots_accessions_wild = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'wild provenance',data_type = 'accessions')

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType)))
collection_proportion = existing_each_year/ existing_items_each_year

plots_proportional_accessions_wild = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover_items,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'wild provenance',
                                                 data_type = 'accessions')
}


```

#### Garden

In this section we explore the turnover of garden origin plants.

`r if(!do_garden){"There are no records with garden origin."}`

```{r, Turnover of garden provenance items - static, eval = report_kind == 'static' & do_garden, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_garden$gain_loss

plots_garden$net

plots_accessions_garden$gain_loss

plots_accessions_garden$net


```

`r if(report_kind == 'interactive' & do_garden){"#####  Turnover of garden origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_garden){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of garden origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_garden}
plots_garden$gain_loss
```

`r if(report_kind == 'interactive' & do_garden){"###### Net (Items)"}`
```{r  Net turnover of garden origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_garden}
plots_garden$net
```

`r if(report_kind == 'interactive' & do_garden){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of garden origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_garden}
plots_accessions_garden$gain_loss
```

`r if(report_kind == 'interactive' & do_garden){"###### Net (Accessions)"}`
```{r  Net turnover of garden origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_garden}
plots_accessions_garden$net
```

`r if(report_kind == 'interactive' & do_garden){"##### {-}"}`

`r if(do_garden){"We now explore how the gain and loss of garden origin plants compare proportionally to the total number of gained and lost items/accessions and the overall proportion of garden origin each year. To calculate the proportion of the collection that are of garden origin each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of garden origin - static, eval = report_kind == 'static' & do_garden, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional_garden$gain

plots_proportional_garden$loss

plots_proportional_accessions_garden$gain

plots_proportional_accessions_garden$loss

```

`r if(report_kind == 'interactive' & do_garden){"#####  Proportional Turnover of garden origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_garden){"###### Gain (Items)"}`
```{r  Proportional Gain of garden origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_garden}
plots_proportional_garden$gain
```

`r if(report_kind == 'interactive' & do_garden){"###### Loss (Items)"}`
```{r  Proportional Loss of garden origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_garden}
plots_proportional_garden$loss
```

`r if(report_kind == 'interactive' & do_garden){"###### Gain (Accessions)"}`
```{r  Proportional Gain of garden origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_garden}
plots_proportional_accessions_garden$gain
```

`r if(report_kind == 'interactive' & do_garden){"###### Loss (Accessions)"}`
```{r  Proportional Loss of garden origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_garden}
plots_proportional_accessions_garden$loss
```

`r if(report_kind == 'interactive' & do_garden){"##### {-}"}`



#### Wild

In this section we explore the turnover of wild origin plants.

`r if(!do_wild){"There are no records with wild origin."}`

```{r, Turnover of wild provenance items - static, eval = report_kind == 'static' & do_wild, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_wild$gain_loss

plots_wild$net

plots_accessions_wild$gain_loss

plots_accessions_wild$net


```

`r if(report_kind == 'interactive' & do_wild){"#####  Turnover of wild origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_wild){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of wild origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild}
plots_wild$gain_loss
```

`r if(report_kind == 'interactive' & do_wild){"###### Net (Items)"}`
```{r  Net turnover of wild origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild}
plots_wild$net
```

`r if(report_kind == 'interactive' & do_wild){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of wild origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild}
plots_accessions_wild$gain_loss
```

`r if(report_kind == 'interactive' & do_wild){"###### Net (Accessions)"}`
```{r  Net turnover of wild origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild}
plots_accessions_wild$net
```

`r if(report_kind == 'interactive' & do_wild){"##### {-}"}`

`r if(do_wild){"We now explore how the gain and loss of wild origin plants compare proportionally to the total number of gained and lost items/accessions and the overall proportion of wild origin each year. To calculate the proportion of the collection that are of wild origin each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of wild origin - static, eval = report_kind == 'static' & do_wild, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional_wild$gain

plots_proportional_wild$loss

plots_proportional_accessions_wild$gain

plots_proportional_accessions_wild$loss

```

`r if(report_kind == 'interactive' & do_wild){"#####  Proportional Turnover of wild origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_wild){"###### Gain (Items)"}`
```{r  Proportional Gain of wild origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild}
plots_proportional_wild$gain
```

`r if(report_kind == 'interactive' & do_wild){"###### Loss (Items)"}`
```{r  Proportional Loss of wild origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild}
plots_proportional_wild$loss
```

`r if(report_kind == 'interactive' & do_wild){"###### Gain (Accessions)"}`
```{r  Proportional Gain of wild origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild}
plots_proportional_accessions_wild$gain
```

`r if(report_kind == 'interactive' & do_wild){"###### Loss (Accessions)"}`
```{r  Proportional Loss of wild origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild}
plots_proportional_accessions_wild$loss
```

`r if(report_kind == 'interactive' & do_wild){"##### {-}"}`



#### Wild-derived

In this section we explore the turnover of wild-derived origin plants.

`r if(!do_wild_derived){"There are no records with wild-derived origin."}`

```{r, Turnover of wild_derived provenance items - static, eval = report_kind == 'static' & do_wild_derived, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_wild_derived$gain_loss

plots_wild_derived$net

plots_accessions_wild_derived$gain_loss

plots_accessions_wild_derived$net


```

`r if(report_kind == 'interactive' & do_wild_derived){"#####  Turnover of wild-derived origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_wild_derived){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of wild_derived origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild_derived}
plots_wild_derived$gain_loss
```

`r if(report_kind == 'interactive' & do_wild_derived){"###### Net (Items)"}`
```{r  Net turnover of wild_derived origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild_derived}
plots_wild_derived$net
```

`r if(report_kind == 'interactive' & do_wild_derived){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of wild_derived origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild_derived}
plots_accessions_wild_derived$gain_loss
```

`r if(report_kind == 'interactive' & do_wild_derived){"###### Net (Accessions)"}`
```{r  Net turnover of wild_derived origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild_derived}
plots_accessions_wild_derived$net
```

`r if(report_kind == 'interactive' & do_wild_derived){"##### {-}"}`

`r if(do_wild_derived){"We now explore how the gain and loss of wild-derived origin plants compare proportionally to the total number of gained and lost items/accessions and the overall proportion of wild-derived origin each year. To calculate the proportion of the collection that are of wild-derived origin each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of wild_derived origin - static, eval = report_kind == 'static' & do_wild_derived, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional_wild_derived$gain

plots_proportional_wild_derived$loss

plots_proportional_accessions_wild_derived$gain

plots_proportional_accessions_wild_derived$loss

```

`r if(report_kind == 'interactive' & do_wild_derived){"#####  Proportional Turnover of wild-derived origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_wild_derived){"###### Gain (Items)"}`
```{r  Proportional Gain of wild_derived origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild_derived}
plots_proportional_wild_derived$gain
```

`r if(report_kind == 'interactive' & do_wild_derived){"###### Loss (Items)"}`
```{r  Proportional Loss of wild_derived origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild_derived}
plots_proportional_wild_derived$loss
```

`r if(report_kind == 'interactive' & do_wild_derived){"###### Gain (Accessions)"}`
```{r  Proportional Gain of wild_derived origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild_derived}
plots_proportional_accessions_wild_derived$gain
```

`r if(report_kind == 'interactive' & do_wild_derived){"###### Loss (Accessions)"}`
```{r  Proportional Loss of wild_derived origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_wild_derived}
plots_proportional_accessions_wild_derived$loss
```

`r if(report_kind == 'interactive' & do_wild_derived){"##### {-}"}`




#### Unknown

In this section we explore the turnover of unknown origin plants.

`r if(!do_unknown){"There are no records with unknown origin."}`


```{r, Turnover of unknown provenance items - static, eval = report_kind == 'static' & do_unknown, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_unknown$gain_loss

plots_unknown$net

plots_accessions_unknown$gain_loss

plots_accessions_unknown$net


```

`r if(report_kind == 'interactive' & do_unknown){"#####  Turnover of unknown origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_unknown){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of unknown origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_unknown}
plots_unknown$gain_loss
```

`r if(report_kind == 'interactive' & do_unknown){"###### Net (Items)"}`
```{r  Net turnover of unknown origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_unknown}
plots_unknown$net
```

`r if(report_kind == 'interactive' & do_unknown){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of unknown origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_unknown}
plots_accessions_unknown$gain_loss
```

`r if(report_kind == 'interactive' & do_unknown){"###### Net (Accessions)"}`
```{r  Net turnover of unknown origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_unknown}
plots_accessions_unknown$net
```

`r if(report_kind == 'interactive' & do_unknown){"##### {-}"}`

`r if(do_unknown){"We now explore how the gain and loss of unknown origin plants compare proportionally to the total number of gained and lost items/accessions and the overall proportion of unknown origin each year. To calculate the proportion of the collection that are of unknown origin each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of unknown origin - static, eval = report_kind == 'static' & do_unknown, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional_unknown$gain

plots_proportional_unknown$loss

plots_proportional_accessions_unknown$gain

plots_proportional_accessions_unknown$loss

```

`r if(report_kind == 'interactive' & do_unknown){"#####  Proportional Turnover of unknown origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_unknown){"###### Gain (Items)"}`
```{r  Proportional Gain of unknown origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_unknown}
plots_proportional_unknown$gain
```

`r if(report_kind == 'interactive' & do_unknown){"###### Loss (Items)"}`
```{r  Proportional Loss of unknown origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_unknown}
plots_proportional_unknown$loss
```

`r if(report_kind == 'interactive' & do_unknown){"###### Gain (Accessions)"}`
```{r  Proportional Gain of unknown origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_unknown}
plots_proportional_accessions_unknown$gain
```

`r if(report_kind == 'interactive' & do_unknown){"###### Loss (Accessions)"}`
```{r  Proportional Loss of unknown origin accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_unknown}
plots_proportional_accessions_unknown$loss
```

`r if(report_kind == 'interactive' & do_unknown){"##### {-}"}`

### Turnover of Native Species

This section looks at the turnover of plants that are native to the location of `r collection` In other words plants that are native to `r location_name`. Here we use Plants of the World Online (POWO) to determine the geography distribution of individual plants, where we only consider regions where the plant is `r location_type_text`.

```{r, turnover of native}
do_type = TRUE
report_cur = report[which(report$native == 'Native'),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_native = turnover
  
  plots = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'native', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'native',
                                                   data_type = 'items')
  ##############################################################################################
  report_cur = report_accessions[which(report_accessions$native == 'Native'),]
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_native_accessions = turnover
  
  plots_accessions = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'native',data_type = 'accessions')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_accessions = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'native',
                                                   data_type = 'accessions')
}
```

`r if(!do_type){"There are no records that are native."}`

```{r, Turnover of native items - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots$gain_loss

plots$net

plots_accessions$gain_loss

plots_accessions$net


```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of native {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Items)"}`
```{r  Net turnover of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Accessions)"}`
```{r  Net turnover of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$net
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(!do_type){"We now explore how the gain and loss of native species compare proportionally to the total number of gained and lost items/accessions and the overall proportion of native species each year. To calculate the proportion of the collection that are native each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of native - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

plots_proportional_accessions$gain

plots_proportional_accessions$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of native {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain (Items)"}`
```{r  Proportional Gain of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Items)"}`
```{r  Proportional Loss of native items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"###### Gain (Accessions)"}`
```{r  Proportional Gain of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Accessions)"}`
```{r  Proportional Loss of native accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`


### Turnover of Endemic Species

This section looks at the turnover of plants that are endemic, this means plants that are native to a single geography region (not only native plants). Here we use Plants of the World Online (POWO) to determine the geography distribution of individual plants, where we only consider regions where the plant is `r location_type_text`. 

```{r, turnover of endemic}
do_type = TRUE
report_cur = report[report$endemic == 'Endemic',]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_endemic = turnover
  
  plots = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'endemic', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'endemic',
                                                   data_type = 'items')
  ##############################################################################################
  report_cur = report_accessions[report_accessions$endemic == 'Endemic',]
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_endemic_accessions = turnover
  
  plots_accessions = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'endemic',data_type = 'accessions')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_accessions = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'endemic',
                                                   data_type = 'accessions')
}
```

`r if(!do_type){"There are no records that are endemic."}`

```{r, Turnover of endemic items - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots$gain_loss

plots$net

plots_accessions$gain_loss

plots_accessions$net


```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of endemic {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of endemic items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Items)"}`
```{r  Net turnover of endemic items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of endemic accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Accessions)"}`
```{r  Net turnover of endemic accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$net
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(!do_type){"We now explore how the gain and loss of endemic species compare proportionally to the total number of gained and lost items/accessions and the overall proportion of endemic species each year. To calculate the proportion of the collection that are endemic each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of endemic - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

plots_proportional_accessions$gain

plots_proportional_accessions$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of endemic {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain (Items)"}`
```{r  Proportional Gain of endemic items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Items)"}`
```{r  Proportional Loss of endemic items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"###### Gain (Accessions)"}`
```{r  Proportional Gain of endemic accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Accessions)"}`
```{r  Proportional Loss of endemic accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`


### Turnover of Threatened Species

This section looks at the turnover of plants that are threatened. 

```{r, turnover of threatened}
do_type = TRUE
report_cur = report[report$threatened == 'Threatened',]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_threatened = turnover
  
  plots = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'threatened', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'threatened',
                                                   data_type = 'items')
  ##############################################################################################
  report_cur = report_accessions[report_accessions$threatened == 'Threatened',]
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_threatened_accessions = turnover
  
  plots_accessions = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'threatened',data_type = 'accessions')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_accessions = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'threatened',
                                                   data_type = 'accessions')
}
```

`r if(!do_type){"There are no records that are threatened."}`

```{r, Turnover of threatened items - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots$gain_loss

plots$net

plots_accessions$gain_loss

plots_accessions$net


```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of threatened {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of threatened items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Items)"}`
```{r  Net turnover of threatened items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of threatened accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Accessions)"}`
```{r  Net turnover of threatened accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$net
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(!do_type){"We now explore how the gain and loss of threatened species compare proportionally to the total number of gained and lost items/accessions and the overall proportion of threatened species each year. To calculate the proportion of the collection that are threatened each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of threatened - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

plots_proportional_accessions$gain

plots_proportional_accessions$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of threatened {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain (Items)"}`
```{r  Proportional Gain of threatened items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Items)"}`
```{r  Proportional Loss of threatened items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"###### Gain (Accessions)"}`
```{r  Proportional Gain of threatened accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Accessions)"}`
```{r  Proportional Loss of threatened accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`


### Turnover of Trees

This section looks at the turnover of plants that are trees. Here we use BGCI's TreeSearch combined with Plants of the World Online (POWO) (to look at synonymy) to determine which plants in the LC are trees. Note that we use the Genus and Species base to determine trees therefore, this will also include cultivated trees if any exist.


```{r, turnover of tree}
do_type = TRUE
report_cur = report[report$tree == 'Tree',]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_trees = turnover
  
  plots = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'tree', data_type = 'items')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'tree',
                                                   data_type = 'items')
  ##############################################################################################
  report_cur = report_accessions[report_accessions$tree == 'Tree',]
  turnover = turnover_items(report_cur)
  exporting_data_store$turnover_trees_accessions = turnover
  
  plots_accessions = plots_turnover(turnover, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'tree',data_type = 'accessions')
  
  existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-01-01'),
                          AccessionYear = report_cur$AccYear,
                          ItemStatusDate = report_cur$ItemStatusDate,
                          ItemStatusType = report_cur$ItemStatusType)))
  collection_proportion = existing_each_year/ existing_items_each_year
  
  plots_proportional_accessions = proportional_plots_turnover(turnover = turnover,
                                                   overall_turnover = overall_turnover_items,
                                                   collection_proportion = collection_proportion,
                                                   report_kind = report_kind,
                                                   separate_figure_folder = separate_figure_folder,
                                                   text = 'tree',
                                                   data_type = 'accessions')
}
```

`r if(!do_type){"There are no records that are trees."}`

```{r, Turnover of tree items - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots$gain_loss

plots$net

plots_accessions$gain_loss

plots_accessions$net


```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of tree {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Items)"}`
```{r  Gain/Loss of tree items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Items)"}`
```{r  Net turnover of tree items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots$net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss (Accessions)"}`
```{r  Gain/Loss of tree accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"###### Net (Accessions)"}`
```{r  Net turnover of tree accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_accessions$net
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(!do_type){"We now explore how the gain and loss of trees compare proportionally to the total number of gained and lost items/accessions and the overall proportion of trees each year. To calculate the proportion of the collection that are trees each year we only consider records that have an accession year and item status date and extract which items/accessions were existing on the first of January each year."}`

```{r, Proportional Turnover of tree - static, eval = report_kind == 'static'& do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

plots_proportional_accessions$gain

plots_proportional_accessions$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of tree {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain (Items)"}`
```{r  Proportional Gain of tree items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Items)"}`
```{r  Proportional Loss of tree items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"###### Gain (Accessions)"}`
```{r  Proportional Gain of tree accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss (Accessions)"}`
```{r  Proportional Loss of tree accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 6), eval = report_kind == 'interactive' & do_type}
plots_proportional_accessions$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`




```{r, export data, echo = FALSE}
if(export_data){
 turnover_data = exporting_data_store
  save(turnover_data, file = paste0(output_dir,'/',collection, '_turnover_data.rda'))
  
}
```

```{r, save excel version of data, eval = save_excel & export_data}
library(xlsx)
file = paste0(output_dir,'/',collection, '_turnover_data.xlsx')
wb <- createWorkbook()

datas <- exporting_data_store[!grepl('proportional',names(exporting_data_store))]
data_titles = names(datas) |> stringr::str_replace_all('_', ' ')
sheetnames <- paste0('Sheet ', 1:length(datas))
sheets <- lapply(sheetnames,  createSheet, wb = wb)
for(i in 1:length(datas)){
  # First we'll create a new row, let's put the title on row 2
  titleRow <-createRow(sheets[[i]], rowIndex=1)
  # We'll add a cell to that row (in the first column) for the title
  sheetTitle <-createCell(titleRow, colIndex=1)
  # We'll add the sheet title to the cell
  setCellValue(sheetTitle[[1,1]], data_titles[i])
  # We'll choose the formatting for the cell
  xlsx::addDataFrame(datas[[i]], sheet = sheets[[i]],row.names = FALSE, startRow = 3)
}
saveWorkbook(wb, file = file)


file = paste0(output_dir,'/',collection, '_proportional_turnover_data.xlsx')
wb <- createWorkbook()

datas <- exporting_data_store[grepl('proportional',names(exporting_data_store))]
data_titles = names(datas) |> stringr::str_replace_all('_', ' ')
sheetnames <- paste0('Sheet ', 1:length(datas))
sheets <- lapply(sheetnames,  createSheet, wb = wb)
for(i in 1:length(datas)){
  # First we'll create a new row, let's put the title on row 2
  titleRow <-createRow(sheets[[i]], rowIndex=1)
  # We'll add a cell to that row (in the first column) for the title
  sheetTitle <-createCell(titleRow, colIndex=1)
  # We'll add the sheet title to the cell
  setCellValue(sheetTitle[[1,1]], data_titles[i])
  # We'll choose the formatting for the cell
  xlsx::addDataFrame(datas[[i]], sheet = sheets[[i]],row.names = FALSE, startRow = 3)
}
saveWorkbook(wb, file = file)

```
