---
title: "Turnover of plants and groups of plants <br>in the LC"
output: html_document
---

##### {-}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(htmltools)
library(sf)
library(here)
library(flextable)
library(ggrepel)
```

```{r, imported parameters}
# #Inputs
# load('/Users/jakepowell/Cambridge/Enriched_reports_Jake/Copenhagen_enriched_report.rda')
# collection = 'Chicago'
# coordinates = c(52.19378853629289,0.1277234065606053)
# load('/Users/jakepowell/Cambridge/Geography Module/wgsrpd3.rda')
# 
# native = 'Naturally occurring only'
# extinct = TRUE
# doubtful_locations = FALSE
# 
# separate_figure_folder = T
# output_dir = getwd()
# report_kind = 'interactive'
# ggtheme = NULL
# value_on_fig = TRUE
# min_year = 1970
# data_type = 'Accessions'
# earliest_allowable_record = 1650
# old_accession_year_codes = c(1000)
```

```{r, add logo, eval = report_kind == 'interactive'}
htmltools::img(src = knitr::image_uri(paste0(system.file(package = "BGSmartR"), "/logo.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width:100px; height:100px')
```

```{r theme_ggplot2, echo = FALSE}
library(ggplot2)
# Changing the default theme
if(is.null(ggtheme)){
   ggtheme <- function(base_size = 16) {
      ggplot2::theme_bw(base_size = base_size) %+replace%
        ggplot2::theme(
          plot.title = ggplot2::element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
          panel.grid.minor = ggplot2::element_blank(),
          panel.background = element_rect(fill = 'transparent', color = NA),
          panel.border = ggplot2::element_blank(),
          axis.title = ggplot2::element_text(size = rel(0.85), face = "bold"),
          axis.text = ggplot2::element_text(size = rel(0.70), face = "bold"),
          axis.line = ggplot2::element_line(color = "black"),
          legend.title = ggplot2::element_text(size = rel(0.85), face = "bold"),
          legend.text = ggplot2::element_text(size = rel(0.70), face = "bold"),
          legend.key = ggplot2::element_rect(fill = "transparent", colour = NA),
          legend.key.size = ggplot2::unit(1.5, "lines"),
          legend.background = ggplot2::element_rect(fill = "transparent", colour = NA),
          strip.background = ggplot2::element_rect(fill = "#17252D", color = "#17252D"),
          strip.text = ggplot2::element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
        )
    }
}
theme_set(ggtheme())
update_geom_defaults("line", list(size = 2))

# Set the colour scale for the plots
scale_colour_continuous <- scale_colour_continuous
scale_colour_discrete   <- scale_colour_discrete
scale_colour_binned     <- scale_colour_binned
scale_fill_continuous <- scale_fill_continuous
scale_fill_discrete <- scale_fill_discrete
scale_fill_binned <- scale_fill_binned
```

```{r, Add combined geography column to enriched report}
geography_data = enriched_report[,grepl('_area_code_l3', names(enriched_report))]
want = c('000','010','001','011','100','110','101')

#Depending on the values of the options reduce `want` to only the satisfactory values.
##A) introduced / naturally occurring only.
if(native == 'Introduced only'){
 want = want[grepl('^1',want)]
}else if(native == 'Naturally occurring only'){
 want = want[grepl('^0',want)]
}
## B) Extinct
if(!extinct){
 want = want[!grepl('010|011|110',want)]
}
## C) Doubtful
if(!doubtful_locations){
 want = want[!grepl('001|011|101',want)]
}
# Get the columns in wanted info that contain the geography information we want.
geography_data = geography_data[,grepl(paste0(want,collapse='|'), names(geography_data))]

if(length(want) > 1){
 level3codes =do.call("paste", c(geography_data, sep = ", "))
 level3codes = stringr::str_remove(level3codes, ', NA$')
 level3codes = stringr::str_remove(level3codes, 'NA, ')
}else{
 level3codes = geography_data
}
level3codes[level3codes == "NA"] = NA

enriched_report$geography_codes = level3codes
```

```{r, Get location_type_text}
location_type_text = ''
if(native == 'Naturally occurring only'){
 location_type_text = paste0(location_type_text, 'naturally occuring only,', collapse = ' ')  
}else if(native == 'Introduced only'){
  location_type_text = paste0(location_type_text, 'introduced only,', collapse = ' ')  
 
}else{
  location_type_text = paste0(location_type_text, 'Either naturally occurring or introduced,', collapse = ' ')  
}
if(extinct){
   location_type_text = paste0(location_type_text, ' allowing extinct regions and', collapse = ' ')  
}else{
  location_type_text = paste0(location_type_text, ' not including extinct regions and', collapse = ' ')  
}
if(doubtful_locations){
   location_type_text = paste0(location_type_text, ' allowing doubtful regions.', collapse = ' ')  
 
}else{
  location_type_text = paste0(location_type_text, ' not allowing doubtful regions.', collapse = ' ')  
}
```

```{r, Find LC location, echo = FALSE}
coord_contained = TRUE
# Extract the location of the collection.
pnts <- data.frame( x = coordinates[2], y = coordinates[1])

# create a points collection
pnts_sf <- do.call("st_sfc",c(lapply(1:nrow(pnts), 
                                     function(i) {sf::st_point(as.numeric(pnts[i, ]))}), list("crs" = 4326))) 

pnts_trans <- st_transform(pnts_sf, 2163) # apply transformation to pnts sf
tt1_trans <- st_transform(wgsrpd3, 2163)      # apply transformation to polygons sf

intersect = as.vector(st_intersects(tt1_trans, pnts_trans, sparse = FALSE))
collection_geog_details = wgsrpd3[which(intersect),-5]
location_name = collection_geog_details$LEVEL3_NAM
location_code = collection_geog_details$LEVEL3_COD

if(length(location_name) == 0){
  coord_contained = FALSE
  # Try and find the nearest polygon.
  collection_geog_details  =  wgsrpd3[which.min(as.vector(sf::st_distance(tt1_trans, pnts_trans))),-5]
  location_name = collection_geog_details$LEVEL3_NAM
  location_code = collection_geog_details$LEVEL3_COD
}
```

`r if(any(!c('AccNoFull') %in% names(enriched_report))){"*** \n **Warning!**\n\n This enriched report contains missing 'original' columns needed to perform the analyses in this report. Default values have been created. In particular, the following were missing "}`

`r if(!'AccNoFull' %in% names(enriched_report)){"- AccNoFull: assume each item is its own accession.\n"}`

`r if(any(!c('no_gardens','AccNoFull') %in% names(enriched_report))){"***"}`

```{r, Extracting required information from enriched report, echo = F}

needed_columns = c('AccNoFull', 'AccYear', 'ItemStatusDate', 'ItemStatusType')
if(!'AccNoFull' %in% names(enriched_report)){
  enriched_report$AccNoFull = 1:nrow(enriched_report)
}
if(!'AccYear' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$AccYear = rep(0, nrow(enriched_report))
}
if(!'ItemStatusDate' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$ItemStatusDate = rep(NA, nrow(enriched_report))
}
if(!'ItemStatusType' %in% names(enriched_report)){
  do_over_time = FALSE
  enriched_report$ItemStatusType = rep('Existing', nrow(enriched_report))
}
if(all(is.na(enriched_report$ItemStatusDate))){
  do_over_time = FALSE
}
if(all(is.na(enriched_report$ItemStatusType))){
  do_over_time = FALSE
}
if(all(is.na(enriched_report$AccYear))){
  do_over_time = FALSE
}
#Extract endemic information from enriched_report
endemic = rep('Not Endemic', nrow(enriched_report))
endemic_index = which(stringr::str_length(enriched_report$geography_codes) == 3)
endemic[endemic_index] = 'Endemic'
enriched_report$endemic = endemic
rm(endemic)

# Extract threatened. 
threat_cat = c('VU','EN','CR','EW','EX')
threatened = rep('Not Threatened', nrow(enriched_report))
threatened[which(enriched_report$redList_category %in% threat_cat)] = 'Threatened'
threatened[which(enriched_report$POWO_Red_category %in% threat_cat)] = 'Threatened'
enriched_report$threatened = threatened
rm(threatened)

# Extract threatened category. 
threatened_category = rep(NA, nrow(enriched_report))
for(i in 1:length(threat_cat)){
 threatened_category[which(enriched_report$redList_category == threat_cat[i])] = threat_cat[i]
 threatened_category[which(enriched_report$POWO_Red_category == threat_cat[i])] = threat_cat[i]
}
enriched_report$threatened_category = threatened_category
rm(threatened_category)

# Convert Provenance code to text.
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'G'] = 'Garden'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'U'] = 'Unknown'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'W'] = 'Wild'
enriched_report$ProvenanceCode[enriched_report$ProvenanceCode == 'Z'] = 'Wild-derived'

# Extract tree.
tree =rep('Not Tree', nrow(enriched_report))
tree[enriched_report$Enrich_is_tree] = 'Tree'
enriched_report$tree = tree
rm(tree)

# Extract native.
native = rep('Not Native', nrow(enriched_report))
native_index = grepl(location_code,enriched_report$geography_codes)
native[native_index] = 'Native'
enriched_report$native = native
rm(native)

# Create a sanitised taxonomic name.
enriched_report$good_name = paste0(enriched_report$sanitised_taxon, ' ', enriched_report$extracted_author)

best_name = paste0(enriched_report$POWO_taxon_name, ' ', enriched_report$POWO_taxon_authors)
best_name[which(best_name == 'NA NA')] = enriched_report$good_name[which(best_name == 'NA NA')]
enriched_report$best_name = best_name

report_original = enriched_report

report = enriched_report[match(c('AccNoFull','AccYear', 'ItemStatusDate', 'ItemStatusType','good_name', 'best_name', 'taxon_type',
                                 'ProvenanceCode', 'endemic', 'threatened', 'threatened_category',
                                 'tree', 'native', 'geography_codes'),names(enriched_report))]
LossYear = rep(NA,nrow(report))
ItemStatusYear = as.numeric(stringr::str_extract(report$ItemStatusDate,pattern = '[0-9]{4}'))
LossYear[report$ItemStatusType == 'NotExisting'] = ItemStatusYear[report$ItemStatusType == 'NotExisting']
report$LossYear = LossYear
  
if(separate_figure_folder){
  figures_dir = paste0(output_dir,'/',collection,'_Figures')
  dir.create(figures_dir,showWarnings = FALSE)
}

exporting_data_store = list()

## Data for turnovers
max_year = as.numeric(format(Sys.Date(),'%Y'))
years = min_year:max_year

# Set old unknown records to have a date ( = earliest_allowable_record) and remove all records prior to this time.
if(!is.null(old_accession_year_codes)){
  has_old_accessions_code = which(report$AccYear %in% old_accession_year_codes)
  no_old_accessions = length(has_old_accessions_code)
  report$AccYear[has_old_accessions_code] = earliest_allowable_record
}
to_keep = which(as.numeric(report$AccYear) >= earliest_allowable_record  & as.numeric(report$AccYear) <= max_year)
still_issue = nrow(report) - length(to_keep)
report = report[to_keep,]

# Swap to accessions only for breakdown of endemic, native, etc if data_type == 'Accessions'
if(data_type == 'Accessions'){
  unique_accessions = unique(report$AccNoFull)
  
  LossYear = rep(NA,nrow(report))
  ItemStatusYear = as.numeric(stringr::str_extract(report$ItemStatusDate,pattern = '[0-9]{4}'))
  LossYear[report$ItemStatusType %in% c('NotExisting', 'Not Existing')] = ItemStatusYear[report$ItemStatusType %in% c('NotExisting', 'Not Existing')]
  report$LossYear = LossYear
  # Order the items where the lower indices have the most recent death date.
  LossYear_dummy = LossYear
  LossYear_dummy[is.na(LossYear_dummy)] = 4000
  ordered_items = order(LossYear_dummy,decreasing = T)
  report = report[ordered_items, ]
  
  #Match to the ordered accessions.
  match_to_best = match(unique_accessions,report$AccNoFull)
  
  #Reduce the data to unique accessions with most recent death date.
  report = report[match_to_best,]
}

# Find the number of items in the collection each year.
existing_each_year_LC = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = as.character(report$AccYear),
                        ItemStatusDate = as.character(report$ItemStatusDate),
                        ItemStatusType = as.character(report$ItemStatusType),
                        post_date = '3000-01-01')))
```

```{r, Get Trends, echo = F}
date = paste0(years, '-12-31')
plant_existing = BGSmartR::exist_at_date(date, AccessionYear = report$AccYear,
                       ItemStatusDate = report$ItemStatusDate,
                       ItemStatusType = report$ItemStatusType,
                       post_date = '3000-01-01')
names(plant_existing) = date
trends = pbapply::pblapply(plant_existing, function(x){
  garden_current = report[x,]
  
  accessions = nrow(garden_current)
  
  # Breakdown of infrageneric diversity.
  breakdown_infrageneric = table(garden_current$taxon_type)
  
  # Breakdown of provenance.
  breakdown_provenance = table(garden_current$ProvenanceCode)
  
  # Breakdown of endemic
  breakdown_endemic = table(garden_current$endemic)
  
  # Number of native
  breakdown_native = table(garden_current$native)

  # Breakdown of threatened
  breakdown_threatened = table(garden_current$threatened)
  
  # Breakdown of threatened categories
  breakdown_threatened_cateogory = table(garden_current$threatened_category)

  # Breakdown by tree
  breakdown_tree= table(garden_current$tree)

  
  return(list(accessions = accessions, breakdown_infrageneric = breakdown_infrageneric, breakdown_provenance = breakdown_provenance, breakdown_endemic = breakdown_endemic, breakdown_native = breakdown_native, breakdown_threatened = breakdown_threatened, breakdown_threatened_cateogory = breakdown_threatened_cateogory, breakdown_tree = breakdown_tree))
})
names(trends) = names(plant_existing)
```

```{r, functions for turnover}
turnover_items <- function(report){
  report = report[(report$LossYear >= min_year & report$LossYear <= max_year) | is.na(report$LossYear),]
  combined = data.frame(Year = years)
  
  # Items
  gain = data.frame(table(report$AccYear))
  if(nrow(gain) == 0){
    gained = rep(0,nrow(combined))
    combined$gain_items = gained
  }else{
    names(gain) = c('Year', 'Items')
    gain$Year = as.numeric(as.character(gain$Year))
    gain = gain[gain$Year >=min_year & gain$Year <= max_year,]
    gained = rep(0,nrow(combined))
    gained[match(gain$Year, combined$Year)] = gain$Items
    combined$gain_items = gained
  }
  
  
  loss = data.frame(table(report$LossYear))
  if(nrow(loss) == 0){
    lost = rep(0,nrow(combined))
    combined$loss_items = lost
  }else{
    names(loss) = c('Year', 'Items')
    loss$Year = as.numeric(as.character(loss$Year))
    loss = loss[loss$Year >=min_year & loss$Year <= max_year,]
    lost = rep(0,nrow(combined))
    lost[match(loss$Year, combined$Year)] = loss$Items
    combined$loss_items = lost
  }
  
  
  combined$net_items = combined$gain_items - combined$loss_items
  return(combined)
}

plots_turnover <- function(turnover, trend, report_kind, text = '', separate_figure_folder, data_type = 'items'){
    dataA = data.frame(turnover[,c(1,3)], rep('Lost', nrow(turnover)))
    names(dataA) = LETTERS[1:3]
    dataB = data.frame(turnover[,c(1,2)], rep('Gain', nrow(turnover)))
    names(dataB) = LETTERS[1:3]
    items_data = rbind(dataA, dataB)
  if(report_kind == 'static'){

    p = ggplot(data=items_data, aes(x=A, y=B, group=C)) +
      geom_line(aes(color=C)) +
      # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
      labs(title=paste0("Turnover of ",text," in the LC (By ",data_type,")"),
           x ="Year",
           y = paste0('Number of ',data_type,'')) +
      guides(color=guide_legend(title=paste0('Gain/Loss'))) +
      # theme(text = element_text(size=table_font_size)) +
      scale_color_manual(values=c("blue", "red"))+
      theme(legend.direction = "horizontal", legend.position = "top", legend.justification = "right")
    
    
    if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','turnover_',stringr::str_replace_all(text, ' ', '_'),'_gain_loss_',data_type,'.pdf'),device = 'pdf', scale = 1, width = 20, height = 12, limitsize = FALSE)
    }
    p1=p
    
    # Color based on value
    color <- ifelse(turnover$net_items < 0, "pink", "lightblue")
    p = ggplot(turnover, aes(x = Year, y = net_items)) +
      geom_bar(stat = "identity",
               show.legend = FALSE,
               fill = color,      # Background color
               color = "white") + # Border color
      # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
      labs(title=paste0("Net turnover of ",text," in the LC (By ",data_type,")"),
           x ="Year",
           y = paste0('Number of ',data_type,''))
    # theme(text = element_text(size=table_font_size))
    
    if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/','Net_turnover_',stringr::str_replace_all(text, ' ', '_'),'_',data_type,'.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
    p2=p
    
    
    # Trend plot.
    trend_data = data.frame(year = turnover$Year, trend = trend)
    p = ggplot(trend_data, aes(x = year, y = trend)) +
      geom_line(width = 1.1) +
      labs(title="",
           x ="Year",
           y = paste0('Number of ',data_type,''))
    p3 = p
    return(list(gain_loss = p1, net = p2, trend = p3))
  }
  if(report_kind == 'interactive'){
    # gain_loss plot
    colors = c("lightblue", "pink")
    fig <- plot_ly(items_data, x = ~A, y = ~B, color = ~C, colors = colors, type = 'scatter', mode = 'lines+markers')
    fig <- fig |> layout(yaxis = list(title = paste0("Number of ",data_type,"")),
                         xaxis = list(title = "Year"))
    fig <- fig |> layout(hovermode = 'x unified')
    
    if(separate_figure_folder){
      htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','line-turnover_',stringr::str_replace_all(text, ' ', '_'),'_gain_loss_',data_type,'.html'))
    }
    fig1 = fig
    
    # net plot
    colors = c("pink", "lightblue")
    color <- ifelse(turnover$net_items < 0, "pink", "lightblue")
    texto = paste0('Net: ', turnover$net_items, '<br>',
                   'Gain: ', turnover$gain_items, '<br>',
                   'Loss: ', turnover$loss_items, '<br>'
    )
    fig = plot_ly(turnover, x = ~Year, y = ~net_items, type = 'bar', hoverinfo = "x+text",  hovertext = texto, color = color, colors = rev(colors), showlegend = FALSE)
    fig <- fig |> layout(yaxis = list(title = paste0("Net number of ",data_type,"")),
                         xaxis = list(title = "Year"))
    fig <- fig |> layout(hovermode = 'x unified')
    
    if(separate_figure_folder){
      htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Bar-Net_',stringr::str_replace_all(text, ' ', '_'),'turnover_of_',data_type,'.html'))
    }
    fig2 = fig
    
    # Trend plot.
    trend_data = data.frame(year = turnover$Year, trend = trend)
    fig = plot_ly(trend_data, x = ~year, y = ~trend, type = 'scatter', mode = 'line', hoverinfo = "x+text", hovertext = trend)
    fig <- fig |> layout(yaxis = list(title = paste0("Number of ",data_type,"")),
                         xaxis = list(title = "Year"))
    fig <- fig |> layout(hovermode = 'x unified')
    fig3= fig
    
    # Combined trend and net.
    fig <- plot_ly()
    # Add traces
    ay <- list(
      tickfont = list(color = "black"),
      overlaying = "y",
      side = "right",
      title = paste0("Total ", tolower(data_type)))
    
      fig <- fig %>% add_trace(x = years, y = turnover$net_items, name = "Net", type = 'bar', hoverinfo = "x+text",  hovertext = texto, color = color, colors = rev(colors), showlegend = FALSE)
     
    fig <- fig %>% add_trace(x = years, y = trend, name = "Total", mode = "lines+markers", type = "scatter", line = list(color = 'black'), marker = list(color = 'black'),yaxis = "y2")
    
  
    # Set figure title, x and y-axes titles
    fig <- fig |> layout(
      title = "", yaxis2 = ay,
      xaxis = list(title="Year"),
      yaxis = list(title = paste0("Net turnover of ", tolower(data_type)))
    ) |>
      layout(plot_bgcolor='white',
             xaxis = list(
               zerolinecolor = '#ffff',
               zerolinewidth = 2,
               gridcolor = 'ffff'),
             yaxis2 = list(
               zeroline = TRUE,
               zerolinecolor = 'ffff',
               zerolinewidth = 2,
               gridcolor = 'ffff',
               showgrid = FALSE),
              yaxis = list(
                zeroline = TRUE,
                zerolinecolor = '#f0f0f0',
                zerolinewidth = 1,
                gridcolor = '#f0f0f0',
                griddash = 'solid',
                gridwidth = 1,showgrid = T),
             margin = list(t = 10, l = 20, r = 70, b = 20, pad = 4),
             autosize = T
      )
    fig <- fig |> layout(hovermode = 'x unified')
    fig4 = fig
    
    return(list(gain_loss = fig1, net = fig2, trend = fig3, trend_and_net = fig4))
    
  }
}

proportional_plots_turnover <- function(turnover, overall_turnover, collection_proportion, separate_figure_folder, 
                                        report_kind, text = '', data_type = 'items'){
  if(report_kind == 'static'){
    # Gain
    plot_data = data.frame(year = turnover[,1],
                           specific = turnover[,2], 
                           all = overall_turnover[,2])
    plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
    plot_data$prop_specific_collection = collection_proportion*100
    plot_data_gain = plot_data
    p = ggplot(plot_data, aes(y=prop_specific, x=year)) + 
      geom_bar(stat="identity", width = 1, fill = 'lightblue', col = 'black') +
      geom_point(aes(x = year, y = prop_specific_collection))+
      labs(title=paste0("Proportion of gained ",data_type," that are ",text," in the LC"),
           x ="Year",
           y = "Percentage (%)") +
      # theme(text = element_text(size=table_font_size)) +
      labs(caption = paste0("Black points are proportion of ",data_type," in the LC that are ",text," each year."))
    p1 = p
    
    if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/',stringr::str_replace_all(text,' ','_'),'_gain_',data_type,'_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
    
    ### Plot of proportion of lost items being native.
    
    plot_data = data.frame(year =  turnover[,1],
                           specific = turnover[,3], 
                           all = overall_turnover[,3])
    plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
    plot_data$prop_specific_collection = collection_proportion*100
    plot_data_loss = plot_data
    
    p = ggplot(plot_data, aes(y=prop_specific, x=year)) + 
      geom_bar(stat="identity", width = 1, fill = 'pink', col = 'black') +
      geom_point(aes(x = year, y = prop_specific_collection))+
      # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
      labs(title=paste0("Proportion of lost ",data_type," that are ",text," in the LC"),
           x ="Year",
           y = "Percentage (%)") +
      # theme(text = element_text(size=table_font_size)) +
      labs(caption = paste0("Black points are proportion of ",data_type," in the LC that are ",text," each year."))
    
    if(separate_figure_folder){
      ggplot2::ggsave(plot = p, filename = paste0(figures_dir, '/',stringr::str_replace_all(text,' ','_'),'_loss_',data_type,'_proportion.pdf'),device = 'pdf', scale = 1, limitsize = FALSE)
    }
    
    return(list(gain = p1, loss = p, data_gain = plot_data_gain, data_loss = plot_data_loss))
  }
  if(report_kind =='interactive'){
    # Gain
    plot_data = data.frame(year = turnover[,1],
                           specific = turnover[,2], 
                           all = overall_turnover[,2])
    plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
    plot_data$prop_specific_collection = collection_proportion*100
    plot_data_gain = plot_data
    
    fig = plot_ly(plot_data, x = ~year, y = ~prop_specific, type = 'bar', name = paste0('Turnover'), marker = list(color = 'lightblue'), hovertemplate = paste("Turnover: %{y:.2f}%<extra></extra>") )
    fig <- fig |> layout(yaxis = list(title = paste0("Percentage")),
                         xaxis = list(title = "Year"))
    fig <- fig |> add_trace(x = ~year, y = ~prop_specific_collection, type = 'scatter', mode = 'markers', name = paste0('Total'), marker = list(color = 'black'), hovertemplate = paste("Collection: %{y:.2f}%<extra></extra>"))
    fig <- fig |> layout(hovermode = 'x unified', yaxis = list(ticksuffix = '%'))
    
    if(separate_figure_folder){
      htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_gained_',stringr::str_replace_all(text, ' ', '_'),'_',data_type,'.html'))
    }
    
    fig1 = fig
    
    ### Plot of proportion of lost items being native.
    
    plot_data = data.frame(year =  turnover[,1],
                           specific = turnover[,3], 
                           all = overall_turnover[,3])
    plot_data$prop_specific = round(plot_data$specific / plot_data$all *100,digits=2)
    plot_data$prop_specific_collection = collection_proportion*100
    plot_data_loss = plot_data
    
    fig = plot_ly(plot_data, x = ~year, y = ~prop_specific, type = 'bar', name = paste0('Turnover'), marker = list(color = 'pink'))
    fig <- fig |> layout(yaxis = list(title = paste0("Percentage")),
                         xaxis = list(title = "Year"))
    fig <- fig |> add_trace(x = ~year, y = ~prop_specific_collection, type = 'scatter', mode = 'markers', name = paste0('Total'), marker = list(color = 'black'))
    fig <- fig |> layout(hovermode = 'x unified', yaxis = list(ticksuffix = '%'))
    
    if(separate_figure_folder){
      htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/','Line_Bar-Proportion_of_lost_',stringr::str_replace_all(text, ' ', '_'),'_',data_type,'.html'))
    }
    
    return(list(gain = fig1, loss = fig,  data_gain = plot_data_gain, data_loss = plot_data_loss))
    
  }
}
```

This report uses data from **`r collection`** and analyses turnover in the collection from **`r min_year`** to **`r max_year`**.

`r if(!is.null(old_accession_year_codes)){paste0("Within the collection the accession years '", paste0(old_accession_year_codes, collapse = ', '),"' are used to denote items which have unknown accession year but have being in the collection since near the beginning. There are ", no_old_accessions," items in the collection with these accession years. In this analysis we set the accession year of such plants to ", earliest_allowable_record, ".")}`

`r if(!is.null(old_accession_year_codes)){"Afterwards the"}else{'The'}` collection contains **`r still_issue`** items whose accession year is not within `r earliest_allowable_record` and `r max_year`. These records are excluded in the following analysis.

Turnover can be split into two parts; when a plant arrives into the collection and when the item is removed. We refer to these two parts as *Gain* and *Loss*. From these values we can calculate the *Net* number of plants gained/lost each year. 

For *Gain* we require items to have an accession year associated with their record. 

For *Loss* we require items to have an item status date together with item status type notifying that the item is now 'Not Existing'. Keep in mind this means a plant is recorded to have being lost on the date the record is updated and this might not align completely with when the item was removed or died.

In the analysis presented within we refer to turnover in terms of `r tolower(data_type)`. 

In the Net graphs in this report we also overlap the graphs with the overall trend to be able to visualise how the turnover affects the total number of `r tolower(data_type)` over time. The trends are calculated by looking at the `r tolower(data_type)` that are existing on the 31st of December each year. 



### Overall Turnover

```{r, overall turnover}
overall_trend = existing_each_year_LC
overall_turnover = turnover_items(report)
exporting_data_store$overall_turnover = overall_turnover
plots = plots_turnover(turnover = overall_turnover, trend = overall_trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, data_type = tolower(data_type))
if(report_kind == 'static'){
  plots$gain_loss = plots$gain_loss + ggtitle(paste0('Turnover of ',tolower(data_type),' in the LC'))
  plots$net = plots$gain_loss + ggtitle(paste0('Net turnover of ',tolower(data_type),' in the LC'))
}

``` 

In this section we shall consider the turnover of plants in the whole of the LC. 

```{r, Overall turnover of accessions - static, eval = report_kind == 'static', fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots$gain_loss

plots$net
```

`r if(report_kind == 'interactive'){"#####  Turnover of the whole collection {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Net"}`
```{r  Net turnover of accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive'}
plots$trend_and_net

# fig <- subplot(plots$trend, plots$net, plots$gain_loss,
#                nrows = 3,
#                shareX = TRUE,
#                heights = c(0.2, 0.4,0.4)) |> layout(showlegend = FALSE)
# htmlwidgets::saveWidget(widget = fig,file = 'stacked_turnover_overall.html')
```

`r if(report_kind == 'interactive'){"###### Gain/Loss"}`
```{r  Gain/Loss of accessions over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive'}
plots$gain_loss

```

`r if(report_kind == 'interactive'){"##### {-}"}`

***

### Turnover by Category

This section explores the turnover of types of taxa. In particular, we use the taxonomic name of plants in the LC to determine whether plants classified as:

- Indeterminate
- Species
- Subspecies
- Varieties
- Forma
- Cultivars
- Hybrids

Not that a record can have multiple of these classifications, such as `Tilia cordata × dasystyla subsp. caucasica`. Since, are not interested in the level of species (i.e if a plant is a subspecies or variety) we are going to explore the turnover of: indeterminate taxa, horticultural taxa (those that are either cultivars or hybrids), and biological taxa (those that are not indeterminate or horticultural).

#### Indeterminate Taxa

```{r, turnover of indeterminates}
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_infrageneric[grepl('0', names(x$breakdown_infrageneric))])
})))
do_indet = TRUE
report_cur = report[grepl('0',report$taxon_type),]
if(nrow(report_cur) == 0){
  do_indet = FALSE
}else{
  turnover_indetermintes = turnover_items(report_cur)
exporting_data_store$turnover_indeterminate_taxa = turnover_indetermintes

plots = plots_turnover(turnover_indetermintes, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'indeterminate plants', data_type = data_type)

existing_indeterminates_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_indeterminates_each_year/ existing_each_year_LC

plots_proportional = proportional_plots_turnover(turnover = turnover_indetermintes,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'indeterminate',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_indeterminate_taxa_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_indeterminate_taxa_loss = plots_proportional$data_loss
}

```

`r if(!do_indet){"There are no records that are indeteminate."}`

```{r, Turnover of indeterminates items - static, eval = report_kind == 'static' & do_indet, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
plots$net

plots$gain_loss

```

`r if(report_kind == 'interactive' & do_indet){"#####  Turnover of indeterminate taxa {.tabset}"}`

`r if(report_kind == 'interactive' & do_indet){"###### Net"}`
```{r  Net turnover of indeterminates over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_indet}
plots$trend_and_net
```

`r if(report_kind == 'interactive' & do_indet){"###### Gain/Loss"}`
```{r  Gain/Loss of indeterminates over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_indet}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_indet){"##### {-}"}`

`r if(do_indet){paste0("We now explore how the gain and loss of indeterminate ", tolower(data_type)," compare proportionally to the total number of gained and lost ", tolower(data_type)," and the overall proportion of indeterminate ", tolower(data_type)," each year. To calculate the proportion of the collection that is indeteminate each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type)," were existing on the first of January each year.")}`

```{r, Proportional Turnover of indeterminates - static, eval = report_kind == 'static' & do_indet, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

```

`r if(report_kind == 'interactive' & do_indet){"#####  Proportional Turnover of indeterminate taxa {.tabset}"}`

`r if(report_kind == 'interactive' & do_indet){"###### Gain"}`
```{r  Proportional Gain of indeterminates items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_indet}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_indet){"###### Loss"}`
```{r  Proportional Loss of indeterminates over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_indet}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_indet){"##### {-}"}`

#### Horticultural Taxa

We define horticultural taxa to be any taxa that is classed as either a cultivar or a hybrid.

```{r, turnover of horticultural}
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_infrageneric[grepl('5|6', names(x$breakdown_infrageneric))])
})))
do_type = TRUE
report_cur = report[grepl('5|6',report$taxon_type),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_horticultural = turnover

plots = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'horticultural', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'horticultural',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_horticultural_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_horticultural_loss = plots_proportional$data_loss
}

```

`r if(!do_type){"There are no records that are horticultural."}`

```{r, Turnover of horticultural - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
plots$net

plots$gain_loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of horticultural taxa {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Net"}`
```{r  Net turnover of horticultural over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$trend_and_net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss"}`
```{r  Gain/Loss of horticultural over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(do_type){paste0("We now explore how the gain and loss of horticultural ", tolower(data_type)," compare proportionally to the total number of gained and lost ", tolower(data_type)," and the overall proportion of horticultural ", tolower(data_type)," each year. To calculate the proportion of the collection that is horticultural each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type)," were existing on the 31st of December each year.")}`

```{r, Proportional Turnover of horticultural - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of horticultural taxa {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain"}`
```{r  Proportional Gain of horticultural over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss"}`
```{r  Proportional Loss of horticultural over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

#### Biological Taxa

We define biological taxa to be those that are not indeterminate or horticultural (i.e those that are only tagged as species, subspecies, variety or forma.)

```{r, turnover of biological}
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_infrageneric[!grepl('0|5|6', names(x$breakdown_infrageneric))])
})))
do_type = TRUE
report_cur = report[!grepl('0|5|6',report$taxon_type),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_biological = turnover

plots = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'biological', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'biological',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_biological_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_biological_loss = plots_proportional$data_loss
}

```

`r if(!do_type){"There are no records that are biological."}`

```{r, Turnover of biological - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
plots$net

plots$gain_loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of biological taxa {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Net"}`
```{r  Net turnover of biological over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$trend_and_net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss"}`
```{r  Gain/Loss of biological over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(do_type){paste0("We now explore how the gain and loss of biological ", tolower(data_type)," compare proportionally to the total number of gained and lost ", tolower(data_type)," and the overall proportion of biological ", tolower(data_type)," each year. To calculate the proportion of the collection that is biological each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type)," were existing on the 31st of December each year.")}`

```{r, Proportional Turnover of biological - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of biological taxa {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain"}`
```{r  Proportional Gain of biological over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss"}`
```{r  Proportional Loss of biological over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`


***

### Turnover by Provenance

In this section we explore the turnover in provenance. Each `r tolower(data_type)` in the garden has one of the following provenance types:

- Wild: plants that were collected directly from the wild,
- Wild-derived: cultivated plants descended from plants of known wild origin,
- Garden: cultivated plants descended from plants not of known wild origin,
- Unknown: plants with unknown origin.

```{r, turnover of provenance}
##### UNKNOWN
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_provenance[grepl('Unknown', names(x$breakdown_provenance))])
})))
do_unknown = TRUE
report_cur = report[which(report$ProvenanceCode == 'Unknown'),]
if(nrow(report_cur) == 0){
  do_unknown = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_provenance_unknown = turnover

plots_unknown = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'unknown provenance', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional_unknown = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'unknown provenance',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_unknown_provenance_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_unknown_provenance_loss = plots_proportional$data_loss
}

##### Garden
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_provenance[grepl('Garden', names(x$breakdown_provenance))])
})))
do_garden = TRUE
report_cur = report[which(report$ProvenanceCode == 'Garden'),]
if(nrow(report_cur) == 0){
  do_garden = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_provenance_garden = turnover

plots_garden = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'garden provenance', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional_garden = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'garden provenance',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_garden_provenance_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_garden_provenance_loss = plots_proportional$data_loss
}

##### Wild-derived
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_provenance[grepl('Wild-derived', names(x$breakdown_provenance))])
})))
do_wild_derived = TRUE
report_cur = report[which(report$ProvenanceCode == 'Wild-derived'),]
if(nrow(report_cur) == 0){
  do_wild_derived = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_provenance_wild_derived = turnover

plots_wild_derived = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'wild-derived provenance', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional_wild_derived = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'wild-derived provenance',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_wild_derived_provenance_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_wild_derived_provenance_loss = plots_proportional$data_loss
}

##### Wild-derived
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_provenance[names(x$breakdown_provenance) == 'Wild'])
})))
do_wild = TRUE
report_cur = report[which(report$ProvenanceCode == 'Wild'),]
if(nrow(report_cur) == 0){
  do_wild = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_provenance_wild = turnover

plots_wild = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'wild provenance', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional_wild = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'wild provenance',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_wild_provenance_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_wild_provenance_loss = plots_proportional$data_loss
}
```

```{r, provenance together}
ordered_provenance = c('Wild','Wild-derived', 'Garden', 'Unknown')
gain_all = data.frame(year = exporting_data_store$turnover_provenance_unknown$Year,
                      unknown = exporting_data_store$turnover_provenance_unknown$gain_items,
                      garden = exporting_data_store$turnover_provenance_garden$gain_items,
                      wild_derived = exporting_data_store$turnover_provenance_wild_derived$gain_items,
                      wild = exporting_data_store$turnover_provenance_wild$gain_items)

loss_all = data.frame(year = exporting_data_store$turnover_provenance_unknown$Year,
                      unknown = exporting_data_store$turnover_provenance_unknown$loss_items,
                      garden = exporting_data_store$turnover_provenance_garden$loss_items,
                      wild_derived = exporting_data_store$turnover_provenance_wild_derived$loss_items,
                      wild = exporting_data_store$turnover_provenance_wild$loss_items)

 dd = reshape(gain_all, idvar = "year", varying = list(2:ncol(gain_all)),
      v.names = "count", timevar = "group", times = names(gain_all)[-1], direction = "long")
 dd$group = stringr::str_to_title(dd$group)
 dd$group[dd$group == 'Wild_derived'] = 'Wild-derived'
 dd$group = factor(dd$group, levels = ordered_provenance)
 
 gain_all_prop = gain_all[,-1]
 gain_all_prop = (gain_all_prop) / rowSums(gain_all_prop)*100
 gain_all_prop = data.frame(year = gain_all$year, gain_all_prop)
 gain_all_prop = gain_all_prop[which(rowSums(gain_all[,-1])!= 0),]

 dd_prop = reshape(gain_all_prop, idvar = "year", varying = list(2:ncol(gain_all_prop)),
      v.names = "count", timevar = "group", times = names(gain_all_prop)[-1], direction = "long")
 dd_prop$group = stringr::str_to_title(dd_prop$group)
 dd_prop$group[dd_prop$group == 'Wild_derived'] = 'Wild-derived'
 dd_prop$group = factor(dd_prop$group, levels =ordered_provenance)
 
 #loss
  ff = reshape(loss_all, idvar = "year", varying = list(2:ncol(loss_all)),
      v.names = "count", timevar = "group", times = names(loss_all)[-1], direction = "long")
 ff$group = stringr::str_to_title(ff$group)
 ff$group[ff$group == 'Wild_derived'] = 'Wild-derived'
 ff$group = factor(ff$group, levels = ordered_provenance)
 
 loss_all_prop = loss_all[,-1]
 loss_all_prop = (loss_all_prop) / rowSums(loss_all_prop)*100
 loss_all_prop = data.frame(year = gain_all$year, loss_all_prop)
 loss_all_prop = loss_all_prop[which(rowSums(loss_all[,-1])!= 0),]

 ff_prop = reshape(loss_all_prop, idvar = "year", varying = list(2:ncol(loss_all_prop)),
      v.names = "count", timevar = "group", times = names(loss_all_prop)[-1], direction = "long")
 ff_prop$group = stringr::str_to_title(ff_prop$group)
 ff_prop$group[ff_prop$group == 'Wild_derived'] = 'Wild-derived'
 ff_prop$group = factor(ff_prop$group, levels = ordered_provenance)
 
if(report_kind == 'interactive'){
  ## gain
   colour_scale = interactive_colour(4)
  colors = colour_scale[dd$group]
 fig = plot_ly(data = dd, x = ~year, y = ~count, color = ~group, type = 'bar',
         marker = list(color = colors,
                       line = list(color = 'black', width = 0))
         )
 fig <- fig |> layout(hovermode = 'x unified',
                         bargap =0,
                         title = "",
                         xaxis = list(title = "Year"),
                         yaxis = list (title = stringr::str_to_title(data_type)),
                         barmode = 'stack',
                         legend = list(traceorder = 'normal'))
 fig1 = fig

 
  colour_scale = interactive_colour(4)
  colors = colour_scale[dd_prop$group]
  dd_prop$colors  = colors
 fig = plot_ly(data = dd_prop, x = ~year, y = ~count, color = ~group, type = 'bar',
         marker = list(color = colors,
                       line = list(color = 'black', width = 0)),
         customdata = ~group,
         hovertemplate = paste("%{customdata}: %{y:.2f}%<extra></extra>")
         )
 fig <- fig |> layout(hovermode = 'x unified',
                         bargap =0,
                         title = "",
                         xaxis = list(title = "Year"),
                         yaxis = list (title = "Gain Percentage"),
                         barmode = 'stack',
                      legend = list(traceorder = 'normal'))
 fig2 = fig

 
  colour_scale = interactive_colour(4)
  colors = colour_scale[ff$group]
 fig = plot_ly(data = ff, x = ~year, y = ~count, color = ~group, type = 'bar',
         marker = list(color = colors,
                       line = list(color = 'black', width = 0))
         )
 fig <- fig |> layout(hovermode = 'x unified',
                         bargap =0,
                         title = "",
                         xaxis = list(title = "Year"),
                         yaxis = list (title = stringr::str_to_title(data_type)),
                         barmode = 'stack', legend = list(traceorder = 'normal'))
 fig3 = fig

 
  colour_scale = interactive_colour(4)
  colors = colour_scale[ff_prop$group]
  ff_prop$colors  = colors
 fig = plot_ly(data = ff_prop, x = ~year, y = ~count, color = ~group, type = 'bar',
         marker = list(color = colors,
                       line = list(color = 'black', width = 0)),
         customdata = ~group,
         hovertemplate = paste("%{customdata}: %{y:.2f}%<extra></extra>")
         )
 fig <- fig |> layout(hovermode = 'x unified',
                         bargap =0,
                         title = "",
                         xaxis = list(title = "Year"),
                         yaxis = list (title = "Loss Percentage"),
                         barmode = 'stack',legend = list(traceorder = 'normal'))
 fig4 = fig
 fig4
 plots_prov_all = list(gain = fig1, gain_prop = fig2,
                       loss = fig3, loss_prop = fig4)
}
```

We begin by considering breaking down gained `r tolower(data_type)` and lost `r tolower(data_type)` by provenance. We show both the total number of `r tolower(data_type)` and the percentage of gained/lost items each year.

`r if(report_kind == 'interactive'){"#####  Gain/loss by provenance {.tabset}"}`

`r if(report_kind == 'interactive'){"###### Gain (Number)"}`
```{r  Gain (Number) by provenance in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_garden}
plots_prov_all$gain
```

`r if(report_kind == 'interactive'){"###### Loss (Number)"}`
```{r  Loss (Number) by provenance in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_garden}
plots_prov_all$loss
```

`r if(report_kind == 'interactive'){"###### Gain (Percentage)"}`
```{r  Gain (Percentage) by provenance in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_garden}
plots_prov_all$gain_prop
```

`r if(report_kind == 'interactive'){"###### Loss (Percentage)"}`
```{r  Loss (Percentage) by provenance in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_garden}
plots_prov_all$loss_prop
```

`r if(report_kind == 'interactive'){"##### {-}"}`

Next we consider each provenance type individually. 

#### Wild

In this section we explore the turnover of wild origin plants.

`r if(!do_wild){"There are no records with wild origin."}`

```{r, Turnover of wild provenance items - static, eval = report_kind == 'static' & do_wild, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_wild$gain_loss

plots_wild$net
```

`r if(report_kind == 'interactive' & do_wild){"#####  Turnover of wild origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_wild){"###### Net"}`
```{r  Gain/Loss of wild origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_wild}
plots_wild$trend_and_net
```

`r if(report_kind == 'interactive' & do_wild){"###### Gain/Loss "}`
```{r  Net turnover of wild origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_wild}
plots_wild$gain_loss
```

`r if(report_kind == 'interactive' & do_wild){"##### {-}"}`

`r if(do_wild){paste0("We now explore how the gain and loss of wild origin plants compare proportionally to the total number of gained and lost ", tolower(data_type), " and the overall proportion of wild origin each year. To calculate the proportion of the collection that are of wild origin each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type), " were existing on the first of January each year.")}`

```{r, Proportional Turnover of wild origin - static, eval = report_kind == 'static' & do_wild, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional_wild$gain

plots_proportional_wild$loss
```

`r if(report_kind == 'interactive' & do_wild){"#####  Proportional Turnover of wild origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_wild){"###### Gain (Items)"}`
```{r  Proportional Gain of wild origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_wild}
plots_proportional_wild$gain
```

`r if(report_kind == 'interactive' & do_wild){"###### Loss (Items)"}`
```{r  Proportional Loss of wild origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_wild}
plots_proportional_wild$loss
```

`r if(report_kind == 'interactive' & do_wild){"##### {-}"}`



#### Wild-derived

In this section we explore the turnover of wild-derived origin plants.

`r if(!do_wild_derived){"There are no records with wild-derived origin."}`

```{r, Turnover of wild_derived provenance items - static, eval = report_kind == 'static' & do_wild_derived, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_wild_derived$gain_loss

plots_wild_derived$net

```

`r if(report_kind == 'interactive' & do_wild_derived){"#####  Turnover of wild-derived origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_wild_derived){"###### Net"}`
```{r  Gain/Loss of wild_derived origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_wild_derived}
plots_wild_derived$trend_and_net
```

`r if(report_kind == 'interactive' & do_wild_derived){"###### Gain/Loss"}`
```{r  Net turnover of wild_derived origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_wild_derived}
plots_wild_derived$gain_loss
```

`r if(report_kind == 'interactive' & do_wild_derived){"##### {-}"}`

`r if(do_wild_derived){paste0("We now explore how the gain and loss of wild-derived origin plants compare proportionally to the total number of gained and lost ", tolower(data_type), " and the overall proportion of wild-derived origin each year. To calculate the proportion of the collection that are of wild-derived origin each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type), " were existing on the first of January each year.")}`

```{r, Proportional Turnover of wild_derived origin - static, eval = report_kind == 'static' & do_wild_derived, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional_wild_derived$gain

plots_proportional_wild_derived$loss

```

`r if(report_kind == 'interactive' & do_wild_derived){"#####  Proportional Turnover of wild-derived origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_wild_derived){"###### Gain"}`
```{r  Proportional Gain of wild_derived origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_wild_derived}
plots_proportional_wild_derived$gain
```

`r if(report_kind == 'interactive' & do_wild_derived){"###### Loss"}`
```{r  Proportional Loss of wild_derived origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_wild_derived}
plots_proportional_wild_derived$loss
```

`r if(report_kind == 'interactive' & do_wild_derived){"##### {-}"}`




#### Garden

In this section we explore the turnover of garden origin plants.

`r if(!do_garden){"There are no records with garden origin."}`

```{r, Turnover of garden provenance - static, eval = report_kind == 'static' & do_garden, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_garden$gain_loss

plots_garden$net
```

`r if(report_kind == 'interactive' & do_garden){"#####  Turnover of garden origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_garden){"###### Net"}`
```{r  Gain/Loss of garden origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_garden}
plots_garden$trend_and_net
```

`r if(report_kind == 'interactive' & do_garden){"###### Gain/Loss"}`
```{r  Net turnover of garden origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_garden}
plots_garden$gain_loss
```

`r if(report_kind == 'interactive' & do_garden){"##### {-}"}`

`r if(do_garden){paste0("We now explore how the gain and loss of garden origin plants compare proportionally to the total number of gained and lost ", tolower(data_type), " and the overall proportion of garden origin each year. To calculate the proportion of the collection that are of garden origin each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type), " were existing on the first of January each year.")}`

```{r, Proportional Turnover of garden origin - static, eval = report_kind == 'static' & do_garden, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional_garden$gain

plots_proportional_garden$loss
```

`r if(report_kind == 'interactive' & do_garden){"#####  Proportional Turnover of garden origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_garden){"###### Gain"}`
```{r  Proportional Gain of garden origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_garden}
plots_proportional_garden$gain
```

`r if(report_kind == 'interactive' & do_garden){"###### Loss"}`
```{r  Proportional Loss of garden origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_garden}
plots_proportional_garden$loss
```

`r if(report_kind == 'interactive' & do_garden){"##### {-}"}`



#### Unknown

In this section we explore the turnover of unknown origin plants.

`r if(!do_unknown){"There are no records with unknown origin."}`


```{r, Turnover of unknown provenance items - static, eval = report_kind == 'static' & do_unknown, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_unknown$gain_loss

plots_unknown$net

```

`r if(report_kind == 'interactive' & do_unknown){"#####  Turnover of unknown origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_unknown){"###### Net"}`
```{r  Gain/Loss of unknown origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_unknown}
plots_unknown$trend_and_net
```

`r if(report_kind == 'interactive' & do_unknown){"###### Gain/Loss"}`
```{r  Net turnover of unknown origin items over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_unknown}
plots_unknown$gain_loss
```

`r if(report_kind == 'interactive' & do_unknown){"##### {-}"}`

`r if(do_unknown){paste0("We now explore how the gain and loss of unknown origin plants compare proportionally to the total number of gained and lost ", tolower(data_type), " and the overall proportion of unknown origin each year. To calculate the proportion of the collection that are of unknown origin each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type), " were existing on the first of January each year.")}`

```{r, Proportional Turnover of unknown origin - static, eval = report_kind == 'static' & do_unknown, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional_unknown$gain

plots_proportional_unknown$loss

```

`r if(report_kind == 'interactive' & do_unknown){"#####  Proportional Turnover of unknown origin {.tabset}"}`

`r if(report_kind == 'interactive' & do_unknown){"###### Gain"}`
```{r  Proportional Gain of unknown origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_unknown}
plots_proportional_unknown$gain
```

`r if(report_kind == 'interactive' & do_unknown){"###### Loss"}`
```{r  Proportional Loss of unknown origin over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_unknown}
plots_proportional_unknown$loss
```

`r if(report_kind == 'interactive' & do_unknown){"##### {-}"}`

***

### Turnover for Key Collections

In this section we look at turnover across the key collections in the LC, this includes:

- Threatened species, those classed as extinct, extinct in the wild, critically endangered, endangered or vulnerable by [IUCN's Red List](https://www.iucnredlist.org).
- Endemic species, those in the collection that belong to a single region according to [TDWG geographical codes](https://web.archive.org/web/20160125135239/http:/www.nhm.ac.uk/hosted_sites/tdwg/TDWG_geo2.pdf) (Brummitt, 2001) expressed to that system's third level.
- Native species, those in the collection that belong to the same region as the collection according to [TDWG geographical codes](https://web.archive.org/web/20160125135239/http:/www.nhm.ac.uk/hosted_sites/tdwg/TDWG_geo2.pdf) (Brummitt, 2001) expressed to that system's third level.
- Trees accessions in the collection that are contained within [BGCI's GlobalTreeSearch](https://tools.bgci.org/global_tree_search.php) (or synonymous names found in [Plants of the World Online](https://powo.science.kew.org))

#### Threatened Species


```{r, turnover of threatened}
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_threatened[!grepl(' ', names(x$breakdown_threatened))])
})))
do_type = TRUE
report_cur = report[which(report$threatened == 'Threatened'),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_threatened = turnover

plots = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'threatened', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'threatened',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_threatened_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_threatened_loss = plots_proportional$data_loss
}

```

`r if(!do_type){"There are no records that are threatened."}`

```{r, Turnover of threatened - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
plots$net

plots$gain_loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of threatened species {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Net"}`
```{r  Net turnover of threatened over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$trend_and_net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss"}`
```{r  Gain/Loss of threatened over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(do_type){paste0("We now explore how the gain and loss of threatened ", tolower(data_type)," compare proportionally to the total number of gained and lost ", tolower(data_type)," and the overall proportion of threatened ", tolower(data_type)," each year. To calculate the proportion of the collection that is threatened each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type)," were existing on the 31st of December each year.")}`

```{r, Proportional Turnover of threatened - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of threatened species {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain"}`
```{r  Proportional Gain of threatened over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss"}`
```{r  Proportional Loss of threatened over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

***

#### Endemic Species

```{r, turnover of endemic}
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_endemic[!grepl(' ', names(x$breakdown_endemic))])
})))
do_type = TRUE
report_cur = report[which(report$endemic == 'Endemic'),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_endemic = turnover

plots = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'endemic', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'endemic',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_endemic_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_endemic_loss = plots_proportional$data_loss
}

```

`r if(!do_type){"There are no records that are endemic."}`

```{r, Turnover of endemic - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
plots$net

plots$gain_loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of endemic species {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Net"}`
```{r  Net turnover of endemic over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$trend_and_net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss"}`
```{r  Gain/Loss of endemic over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(do_type){paste0("We now explore how the gain and loss of endemic ", tolower(data_type)," compare proportionally to the total number of gained and lost ", tolower(data_type)," and the overall proportion of endemic ", tolower(data_type)," each year. To calculate the proportion of the collection that is endemic each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type)," were existing on the 31st of December each year.")}`

```{r, Proportional Turnover of endemic - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of endemic species {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain"}`
```{r  Proportional Gain of endemic over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss"}`
```{r  Proportional Loss of endemic over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

***

#### Native Species

This section looks at the turnover of plants that are native to the location of `r collection` In other words plants that are native to `r location_name`. 

```{r, turnover of native}
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_native[!grepl(' ', names(x$breakdown_native))])
})))
do_type = TRUE
report_cur = report[which(report$native == 'Native'),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_native = turnover

plots = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'native', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'native',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_native_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_native_loss = plots_proportional$data_loss
}

```

`r if(!do_type){"There are no records that are native."}`

```{r, Turnover of native - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
plots$net

plots$gain_loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of native species {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Net"}`
```{r  Net turnover of native over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$trend_and_net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss"}`
```{r  Gain/Loss of native over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(do_type){paste0("We now explore how the gain and loss of native ", tolower(data_type)," compare proportionally to the total number of gained and lost ", tolower(data_type)," and the overall proportion of native ", tolower(data_type)," each year. To calculate the proportion of the collection that is native each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type)," were existing on the 31st of December each year.")}`

```{r, Proportional Turnover of native - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of native species {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain"}`
```{r  Proportional Gain of native over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss"}`
```{r  Proportional Loss of native over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

***



#### Turnover of Trees

```{r, turnover of tree}
trend = as.numeric(unlist(lapply(trends,function(x){
  sum(x$breakdown_tree[!grepl(' ', names(x$breakdown_tree))])
})))
do_type = TRUE
report_cur = report[which(report$tree == 'Tree'),]
if(nrow(report_cur) == 0){
  do_type = FALSE
}else{
  turnover = turnover_items(report_cur)
exporting_data_store$turnover_tree = turnover

plots = plots_turnover(turnover, trend = trend, report_kind = report_kind, separate_figure_folder = separate_figure_folder, text = 'tree', data_type = data_type)

existing_each_year = as.numeric(colSums(BGSmartR::exist_at_date(date = paste0(years,'-12-31'),
                        AccessionYear = report_cur$AccYear,
                        ItemStatusDate = report_cur$ItemStatusDate,
                        ItemStatusType = report_cur$ItemStatusType,
                        post_date = '3000-01-01')))
collection_proportion = existing_each_year/ existing_each_year_LC

plots_proportional = proportional_plots_turnover(turnover = turnover,
                                                 overall_turnover = overall_turnover,
                                                 collection_proportion = collection_proportion,
                                                 report_kind = report_kind,
                                                 separate_figure_folder = separate_figure_folder,
                                                 text = 'tree',
                                                 data_type = data_type)
exporting_data_store$proportional_turnover_tree_gain = plots_proportional$data_gain
exporting_data_store$proportional_turnover_tree_loss = plots_proportional$data_loss
}

```

`r if(!do_type){"There are no records that are trees."}`

```{r, Turnover of tree - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6)}
plots$net

plots$gain_loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Turnover of trees {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Net"}`
```{r  Net turnover of tree over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$trend_and_net
```

`r if(report_kind == 'interactive' & do_type){"###### Gain/Loss"}`
```{r  Gain/Loss of tree over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots$gain_loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

`r if(do_type){paste0("We now explore how the gain and loss of tree ", tolower(data_type)," compare proportionally to the total number of gained and lost ", tolower(data_type)," and the overall proportion of tree ", tolower(data_type)," each year. To calculate the proportion of the collection that is tree each year we only consider records that have an accession year and item status date and extract which ", tolower(data_type)," were existing on the 31st of December each year.")}`

```{r, Proportional Turnover of tree - static, eval = report_kind == 'static' & do_type, fig.fullwidth=TRUE, fig.dim = c(10, 6),}
plots_proportional$gain

plots_proportional$loss

```

`r if(report_kind == 'interactive' & do_type){"#####  Proportional Turnover of trees {.tabset}"}`

`r if(report_kind == 'interactive' & do_type){"###### Gain"}`
```{r  Proportional Gain of tree over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$gain
```

`r if(report_kind == 'interactive' & do_type){"###### Loss"}`
```{r  Proportional Loss of tree over time in the LC - interactive, fig.fullwidth=TRUE, fig.dim = c(10, 4), eval = report_kind == 'interactive' & do_type}
plots_proportional$loss
```

`r if(report_kind == 'interactive' & do_type){"##### {-}"}`

***



`r if(do_geography_trend){"### Trends in Geographic Distribution"}`

`r if(do_geography_trend){"In this section, we investigate turnover in the geographic distribution of the collection. Note that to obtain the distribution of plants we use POWO. Therefore, only plants matched to POWO have geographic distribution."}`

`r if(do_geography_trend){paste0("Furthermore, in the following figures we show the number of ", data_type, " each region has in the collection each year. Since plants can belong to multiple regions their ", data_type, " count will contribution to multiple regions in the maps. For example Abies pinsapo Boiss. is native to Morocco and Spain therefore a single accession of would add one to the number of ", data_type, " for Morocco and Spain. This means the total number of ", data_type, " in the geography plots below is larger than the number of ", data_type, " in the collection.")}`

`r if(do_geography_trend & report_kind == 'static'){"Below we show the geographic distribution of the collection for 4 fixed years. "}`

`r if(do_geography_trend & report_kind == 'interactive'){"Below we show the gain across geographic distribution. "}`

```{r, geography trends - static, eval = do_geography_trend & report_kind == 'static', fig.fullwidth=TRUE, fig.dim = c(10, 15), fig.caption = 'Trends in the geographic distribution'}
# Format and create individual plots.
# Combine the wanted information into location data.
CUBG_geo_over_time = data.frame(lapply(time_series_info_subsets, function(x){x$breakdown_geography}))
rownames(CUBG_geo_over_time) = wgsrpd3$LEVEL3_NAM
CUBG_geo_over_time_in_list = lapply(CUBG_geo_over_time, function(x){
  x[x==0] = NA
  location_data = data.frame(code =  wgsrpd3$LEVEL3_COD,
                             name = wgsrpd3$LEVEL3_NAM,
                             value = x)
  
  plot_info <- wgsrpd3 |>
    #add the location data to the geometry data
    dplyr::left_join(location_data, by=c("LEVEL3_COD"="code"))
  return(plot_info)
})

dir.create(paste0(figures_dir, '/geography_trends/'),showWarnings = FALSE)
# Loop over and create each individual plot and save
show_plots = round(seq(0,1,1/3)*(length(CUBG_geo_over_time_in_list)-1),digits = 0)+1
plots = list() ; counter = 1
for(i in 1:length(CUBG_geo_over_time_in_list)){
fig = ggplot(CUBG_geo_over_time_in_list[[i]])+
  geom_sf(aes(fill=value),  col="transparent")+
  stat_sf_coordinates(aes(col=value))+
  scale_colour_distiller(direction=1,
                         name="Accessions", na.value = 'black') +
  scale_fill_distiller(direction=1,
                       name="Accessions", na.value = 'black')+
  # scale_fill_continuous(name = 'Acessions',na.value = 'black')+
  # scale_color_continuous(name = 'Acessions',na.value = 'black')+
  coord_sf(expand=FALSE)+
  theme(
    plot.title = ggplot2::element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0.5),
    panel.grid.minor = ggplot2::element_blank(),
    panel.border = ggplot2::element_blank(),
    legend.title = ggplot2::element_text(size = rel(0.85), face = "bold"),
    legend.text = ggplot2::element_text(size = rel(0.70), face = "bold"),
    legend.key = ggplot2::element_rect(fill = "transparent", colour = NA),
    legend.key.size = ggplot2::unit(1.5, "lines"),
    legend.background = ggplot2::element_rect(fill = "transparent", colour = NA),
    strip.background = ggplot2::element_rect(fill = "#17252D", color = "#17252D"),
    strip.text = ggplot2::element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0)),
    axis.line=element_blank(),
    axis.text.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks=element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_blank(),
  ) + ggtitle(years[i])+
  theme(legend.position="bottom")

  if(separate_figure_folder){
      ggsave(plot = fig, filename = paste0(figures_dir, '/geography_trends/plot',i,'.pdf'), device = 'pdf', width = 10, height = 6)
  }
  
  if(i %in%show_plots){
    plots[[counter]] = fig
    counter = counter + 1
  }
}

p = ggpubr::ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], labels = c("", "","",""),
          common.legend = FALSE, ncol = 1, nrow = 4) +
    theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"))
p
```


```{r, geography gain - interactive, eval = do_geography_trend & report_kind == 'interactive', fig.fullwidth=TRUE, fig.dim = c(10, 6)}
# Convert geometry into a format accepted by plotly
wgsrpd3_level3_simp = BGSmartR::wgsrpd3_level3_simp 
mapp = sf::st_cast(wgsrpd3_level3_simp , "MULTIPOLYGON")
geo_sf = geojsonsf::sf_geojson(mapp)
data = rjson::fromJSON(geo_sf)
feat = data$features
for(i in 1:length(feat)){
  feat[[i]]$id = feat[[i]]$properties$code
}
data$features = feat

# Format the geography total count per region each year
regions = wgsrpd3_level3_simp$code

turnover_geography = pbapply::pblapply(regions, function(x){
  report_cur = report[which(grepl(x, report$geography_codes)),]
  return(turnover_items(report_cur))
})
names(turnover_geography)  = regions

geography_gain = data.frame(lapply(turnover_geography, function(x){x$gain_items}))
names(geography_gain) = wgsrpd3$LEVEL3_COD

tt = data.frame(year = years, geography_gain)
dd = reshape(tt, idvar = "year", varying = list(2:length(tt)),
             v.names = "count", timevar = "group", times = names(tt)[-1], direction = "long")

name = wgsrpd3_level3_simp$name[match(dd$group, wgsrpd3_level3_simp$code)]
name <- stringr::str_replace_all(name, pattern = 'Is\\.', 'Islands')
name <- stringr::str_replace_all(name, pattern = 'I\\.', 'Island')
dd$name = name
dd$text = paste0(dd$name,': ', dd$count)


# Next repreat the above process but obtain a ratio for each year (value/max(year)) to use for colouring the scatter points.
geography_gain = data.frame(lapply(turnover_geography, function(x){x$gain_items}))
names(geography_gain) = wgsrpd3$LEVEL3_COD
max_count = as.numeric(apply(geography_gain, MARGIN=c(1), max))
geography_gain = geography_gain / max_count
# as.numeric(apply(CUBG_geo_over_time, MARGIN=c(1), max)) # Check the max = 1
tt = data.frame(year = years, geography_gain)
dd_ratio = reshape(tt, idvar = "year", varying = list(2:length(tt)),
             v.names = "count", timevar = "group", times = names(tt)[-1], direction = "long")
dd$ratio = dd_ratio$count
dd$color = viridis::viridis(101)[dd$ratio*100+1]

# Add the long and lat of the centres of the geometries to dd.
data_g = st_geometry(wgsrpd3_level3_simp)
centers = st_centroid(data_g)
AA = unlist(centers)
center_df = data.frame(long = AA[seq(1,length(AA),2)], lat = AA[seq(2,length(AA),2)])
centers_df = center_df[match(dd$group, wgsrpd3_level3_simp$code),]
dd = data.frame(dd, centers_df)

fig = plot_ly() |>
  layout(
    mapbox=list(
      style="white-bg",
      center = list(lon = 0 ,lat= 60),
      zoom =0.46, 
      lataxis = list(range = c(-59, 90)))) |>
  add_trace(fig, type="choroplethmapbox",
            geojson=data,
            locations=dd$group,
            z=dd$count,
            frame = dd$year,
            text = dd$text, 
            hoverinfo = 'text',
            hoverlabel = list(bgcolor = 'white'),
            # colorscale=col_scale,
            #  #  reversescale = FALSE,
            marker=list(line=list( width=0.01,color ='black'))
            )|>
   add_trace(type="scattermapbox",
             mode = 'markers',
             lon = dd$long,
             lat = dd$lat,
             text = dd$text,
             frame = dd$year,
             color = I(dd$color),
             hoverinfo = 'text',
             hoverlabel = list(bgcolor = 'white'),
             inherit = FALSE) 
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/turnover_gain_geographic_distribution_of_',data_type,'.html'))
}

fig
```

`r if(do_geography_trend & report_kind == 'interactive'){"Below we show the loss across geographic distribution. "}`

```{r, geography loss - interactive, eval = do_geography_trend & report_kind == 'interactive', fig.fullwidth=TRUE, fig.dim = c(10, 6)}
# Convert geometry into a format accepted by plotly
mapp = sf::st_cast(wgsrpd3_level3_simp, "MULTIPOLYGON")
geo_sf = geojsonsf::sf_geojson(mapp)
data = rjson::fromJSON(geo_sf)
feat = data$features
for(i in 1:length(feat)){
  feat[[i]]$id = feat[[i]]$properties$code
}
data$features = feat

# Format the geography total count per region each year
regions = wgsrpd3_level3_simp$code

turnover_geography = pbapply::pblapply(regions, function(x){
  report_cur = report[which(grepl(x, report$geography_codes)),]
  return(turnover_items(report_cur))
})
names(turnover_geography)  = regions

geography_gain = data.frame(lapply(turnover_geography, function(x){x$loss_items}))
names(geography_gain) = wgsrpd3$LEVEL3_COD

tt = data.frame(year = years, geography_gain)
dd = reshape(tt, idvar = "year", varying = list(2:length(tt)),
             v.names = "count", timevar = "group", times = names(tt)[-1], direction = "long")

name = wgsrpd3_level3_simp$name[match(dd$group, wgsrpd3_level3_simp$code)]
name <- stringr::str_replace_all(name, pattern = 'Is\\.', 'Islands')
name <- stringr::str_replace_all(name, pattern = 'I\\.', 'Island')
dd$name = name
dd$text = paste0(dd$name,': ', dd$count)


# Next repreat the above process but obtain a ratio for each year (value/max(year)) to use for colouring the scatter points.
geography_gain = data.frame(lapply(turnover_geography, function(x){x$loss_items}))
names(geography_gain) = wgsrpd3$LEVEL3_COD
max_count = as.numeric(apply(geography_gain, MARGIN=c(1), max))
geography_gain = geography_gain / max_count
# as.numeric(apply(CUBG_geo_over_time, MARGIN=c(1), max)) # Check the max = 1
tt = data.frame(year = years, geography_gain)
dd_ratio = reshape(tt, idvar = "year", varying = list(2:length(tt)),
             v.names = "count", timevar = "group", times = names(tt)[-1], direction = "long")
dd$ratio = dd_ratio$count
dd$color = viridis::viridis(101)[dd$ratio*100+1]

# Add the long and lat of the centres of the geometries to dd.
data_g = st_geometry(wgsrpd3_level3_simp)
centers = st_centroid(data_g)
AA = unlist(centers)
center_df = data.frame(long = AA[seq(1,length(AA),2)], lat = AA[seq(2,length(AA),2)])
centers_df = center_df[match(dd$group, wgsrpd3_level3_simp$code),]
dd = data.frame(dd, centers_df)

fig = plot_ly() |>
  layout(
    mapbox=list(
      style="white-bg",
      center = list(lon = 0 ,lat= 60),
      zoom =0.46, 
      lataxis = list(range = c(-59, 90)))) |>
  add_trace(fig, type="choroplethmapbox",
            geojson=data,
            locations=dd$group,
            z=dd$count,
            frame = dd$year,
            text = dd$text, 
            hoverinfo = 'text',
            hoverlabel = list(bgcolor = 'white'),
            # colorscale=col_scale,
            #  #  reversescale = FALSE,
            marker=list(line=list( width=0.01,color ='black'))
            )|>
   add_trace(type="scattermapbox",
             mode = 'markers',
             lon = dd$long,
             lat = dd$lat,
             text = dd$text,
             frame = dd$year,
             color = I(dd$color),
             hoverinfo = 'text',
             hoverlabel = list(bgcolor = 'white'),
             inherit = FALSE) 
if(separate_figure_folder){
  htmlwidgets::saveWidget(fig, file = paste0(figures_dir, '/turnover_loss_geographic_distribution_of_',data_type,'.html'))
}

fig
```

***
